### 2019-03-20 doc/chapters/conclusions.tex
added latex template

\chapter{Conclusion / Results}
\label{chap:conclusions}

But I must explain to you how all this mistaken idea of denouncing pleasure and praising pain was born and I will give you a complete account of the system, and expound the actual teachings of the great explorer of the truth, the master-builder of human happiness. No one rejects, dislikes, or avoids pleasure itself, because it is pleasure, but because those who do not know how to pursue pleasure rationally encounter consequences that are extremely painful. Nor again is there anyone who loves or pursues or desires to obtain pain of itself, because it is pain, but because occasionally circumstances occur in which toil and pain can procure him some great pleasure. To take a trivial example, which of us ever undertakes laborious physical exercise, except to obtain some advantage from it? But who has any right to find fault with a man who chooses to enjoy a pleasure that has no annoying consequences, or one who avoids a pain that produces no resultant pleasure? On the other hand, we denounce with righteous indignation and dislike men who are so beguiled and demoralized by the charms of pleasure of the moment, so blinded by desire, that they cannot foresee the pain and trouble that are bound to ensue; and equal blame belongs to those who fail in their duty through weakness of will, which is the same as saying through shrinking from toil and pain. These cases are perfectly simple and easy to distinguish. In a free hour, when our power of choice is untrammelled and when nothing prevents our being able to do what we like best, every pleasure is to be welcomed and every pain avoided. But in certain circumstances and owing to the claims of duty or the obligations of business it will frequently occur that pleasures have to be repudiated and annoyances accepted. The wise man therefore always holds in these matters to this principle of selection: he rejects pleasures to secure other greater pleasures, or else he endures pains to avoid worse pains.

\newpage

But I must explain to you how all this mistaken idea of denouncing pleasure and praising pain was born and I will give you a complete account of the system, and expound the actual teachings of the great explorer of the truth, the master-builder of human happiness. No one rejects, dislikes, or avoids pleasure itself, because it is pleasure, but because those who do not know how to pursue pleasure rationally encounter consequences that are extremely painful. Nor again is there anyone who loves or pursues or desires to obtain pain of itself, because it is pain, but because occasionally circumstances occur in which toil and pain can procure him some great pleasure. To take a trivial example, which of us ever undertakes laborious physical exercise, except to obtain some advantage from it? But who has any right to find fault with a man who chooses to enjoy a pleasure that has no annoying consequences, or one who avoids a pain that produces no resultant pleasure? On the other hand, we denounce with righteous indignation and dislike men who are so beguiled and demoralized by the charms of pleasure of the moment, so blinded by desire, that they cannot foresee the pain and trouble that are bound to ensue; and equal blame belongs to those who fail in their duty through weakness of will, which is the same as saying through shrinking from toil and pain. These cases are perfectly simple and easy to distinguish. In a free hour, when our power of choice is untrammelled and when nothing prevents our being able to do what we like best, every pleasure is to be welcomed and every pain avoided. But in certain circumstances and owing to the claims of duty or the obligations of business it will frequently occur that pleasures have to be repudiated and annoyances accepted. The wise man therefore always holds in these matters to this principle of selection: he rejects pleasures to secure other greater pleasures, or else he endures pains to avoid worse pains.

### 2019-03-20 doc/leader/declaration_authorship.tex
added latex template

\chapter*{Declaration of primary authorship}
\label{chap:declaration_authorship}

\vspace*{10mm}

I / We hereby confirm that I / we have written this thesis independently and without using other sources and resources than those specified in the bibliography. All text passages which were not written by me are marked as quotations and provided with the exact indication of its origin.

\vspace{15mm}

\begin{tabbing}
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\kill
Place, Date:		\> [Biel/Burgdorf], \versiondate \\ \\ 
Last Name/s, First Name/s:	\> [Test Peter] 	\> [M\"uster R\"os\"a] \\ \\ \\ \\ 
Signature/s:	\> ......................................\> ...................................... \\
\end{tabbing}

### 2019-03-20 doc/leader/frontpage_with_picture.tex
added latex template

%
% Project documentation template
% ===========================================================================
% This is part of the document "Project documentation template".
% Authors: brd3, kaa1
%

\begin{titlepage}

% BFH-Logo absolute placed at (28,12) on A4 and picture (16:9 or 15cm x 8.5cm)
% Actually not a realy satisfactory solution but working.
%---------------------------------------------------------------------------
\setlength{\unitlength}{1mm}
\begin{textblock}{20}[0,0](28,12)
	\includegraphics[scale=1.0]{images/BFH_Logo_B.png}
\end{textblock}

\begin{textblock}{154}(28,48)
	\begin{picture}(150,2)
		\put(0,0){\color{bfhgrey}\rule{150mm}{2mm}}
	\end{picture}
\end{textblock}

\begin{textblock}{154}[0,0](28,50)
	\includegraphics[scale=1.0]{images/placemarker.jpg}			% define cover picture
\end{textblock}

\begin{textblock}{154}(28,135)
	\begin{picture}(150,2)
		\put(0,0){\color{bfhgrey}\rule{150mm}{2mm}}
	\end{picture}
\end{textblock}
\color{black}

% Institution / titel / subtitel / authors / experts:
%---------------------------------------------------------------------------
\begin{flushleft}

\vspace*{115mm}

\fontsize{26pt}{28pt}\selectfont 
\heading				\\							% Read heading from file leader/title.tex
\vspace{2mm}

\fontsize{16pt}{20pt}\selectfont\vspace{0.3em}
Place your subheading here 			\\				% Insert subheading
\vspace{5mm}

\fontsize{10pt}{12pt}\selectfont
\textbf{Description of thesis (semester- / Bachelor thesis / etc.)} \\		% Insert text
\vspace{3mm}

% Abstract (eingeben):
%---------------------------------------------------------------------------
\begin{textblock}{150}(28,190)
\fontsize{10pt}{12pt}\selectfont
[Insert short text (abstract) if desired] \\ 
This document serves as a template for the compilation of reports according to the guidelines of the BFH. The template is written in LATEX and supports the automatic writing of various directories, references, indexing and glossaries. This small text is a summary of this document with a length of 4 to max. 8 lines. \\ 
The cover picture may be turned on or off in the lines 157/158 of the file template.tex.
\end{textblock}

\begin{textblock}{150}(28,225)
\fontsize{10pt}{17pt}\selectfont
\begin{tabbing}
xxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
Degree course:	\> [z.B. Electrical and Communication Engineering]	\\		% insert name of degree course
Authors:		\> [Test Peter, M\"uster R\"os\"a]		\\					% insert names
Tutor:	\> [Dr.~Xxxx Xxxx, Dr.~Yyyy Yyyy]		\\							% insert names
Constituent:	\> [Wwwww AG]					\\							% insert names
Experts:		\> [Dr.~Zzzz Zzzz]				\\							% insert names
Date:			\> \versiondate					\\							% read from file leader/version.tex
\end{tabbing}

\end{textblock}
\end{flushleft}

\begin{textblock}{150}(28,280)
\noindent 
\color{bfhgrey}\fontsize{9pt}{10pt}\selectfont
Berner Fachhochschule | Haute \'ecole sp\'ecialis\'ee bernoise | Bern University of Applied Sciences
\color{black}\selectfont
\end{textblock}

\end{titlepage}

%
% ===========================================================================
% EOF
%

### 2019-03-20 doc/leader/versions.tex
added latex template

% Control of versions :
% -----------------------------------------------

\begin{textblock}{180}(15,150)
\color{black}
\begin{huge}
Versions
\end{huge}
\vspace{10mm}

\fontsize{10pt}{18pt}\selectfont
\begin{tabbing}
xxxxxxxxxxx\=xxxxxxxxxxxxxxx\=xxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
Version	\> Date	\> Status			\> Remarks		\\
0.1	\> 01.08.2013	\> Draft		\> Lorem ipsum dolor sit amet	\\	
0.2	\> 21.08.2013	\> Draft		\> Phasellus scelerisque	\\ 
0.3	\> 02.09.2013	\> Draft		\> Donec eget aliquam urna. Lorem ipsum dolor sit amet	\\ 
1.0	\> 26.01.2014	\> Final		\> Lorem ipsum dolor sit ametPhasellus scelerisque, leo sed iaculis ornare 	\\ 
1.1	\> 31.01.2014	\> Correction	\> Layout changed	\\
1.2	\> 07.02.2014	\> Addition		\> Chapter 1.1 extended	\\
\end{tabbing}

\end{textblock}

### 2019-03-20 doc/chapters/introduction.tex
fixed äöü

\chapter{Introduction}
\label{chap:introduction}

This document serves to illustrate the \LaTeX{} template on the basis of the corporate design\index{Corporate Design} of the Bern University of Applied Sciences\index{Bern University of Applied Sciences} as well as a manual for its use. It is assumed that the user already has some experience with \LaTeX{} or is willing to familiarize with the subject. In the bibliography the user finds some useful information on \LaTeX{} on various books and documents on the Internet.

% Einträge im Verzeichnis erscheinen lassen ohne hier eine Referenz einzufügen
\nocite{kopka:band1}
\nocite{raichle:bibtex_programmierung}
\nocite{MiKTeX}
\nocite{KOMA}
\nocite{TeXnicCenter}
\nocite{Marti06}
\nocite{Erbsland08}
\nocite{juergens:einfuehrung}
\nocite{juergens:fortgeschritten}

\section{Organising Documents}
\label{sec:einleitung_aufbau}

This document is structured according to the documentation of a project work or a thesis\index{thesis}. In Chapter \ref{chap:instructions}, the packages used are briefly explained, and instructions are given, how the bibliography and the glossary are to be used. Chapter \ref{chap:typeareatest} presents a sample chapter to audit the type area.

In figure \ref{fig:file_structure} the file structure is shown for this template.

\begin{figure}[H]
	\centering
		\includegraphics[scale=0.85]{images/file_structure.png}
	\caption{File structure}
	\label{fig:file_structure}
\end{figure}

\section{Contact}
\label{sec:introduction_contact}

The manufacturers of this template welcome any suggestions for improvement. Chapter \ref{sec:introduction_suggestions} shows possible suggestions for improvement\index{suggestions}.

\begin{table}[H]
	\centering
		\begin{tabular}{lll} \toprule
			\textbf{First Name Last Name} & \textbf{E-mail} & \textbf{Function} \\ \midrule
			Alfred Kaufmann & alfred.kaufmann@bfh.ch & Employer, Project Management, \\
			& & Supplements, Improvements \\ \midrule
			Fritz Dellsperger & Retired & Tips on the structure and layout \\ \midrule
			David Burri & Contracted out & First compilation of the Template \\ \bottomrule
		\end{tabular}
	\caption{Contact Persons}
	\label{tab:Contact Persons}
\end{table}

\section{Suggestions for Improvement}
\label{sec:introduction_suggestions}

\begin{itemize}
	\item Create a BFH Style Files
	\item Template for the Compilation of presentations with \LaTeX{}
\end{itemize}

### 2019-04-02 doc/chapters/introduction.tex
Added section Formal verification.

@@
 \label{chap:introduction}
-
-This document serves to illustrate the \LaTeX{} template on the basis of the corporate design\index{Corporate Design} of the Bern University of Applied Sciences\index{Bern University of Applied Sciences} as well as a manual for its use.
-It is assumed that the user already has some experience with \LaTeX{} or is willing to familiarize with the subject.
-In the bibliography the user finds some useful information on \LaTeX{} on various books and documents on the Internet.
 
@@
 
-\section{Organising Documents}
-\label{sec:einleitung_aufbau}
+\section{Formal verification}
+\label{sec:formal_verification}
 
-This document is structured according to the documentation of a project work or a thesis\index{thesis}.
-In Chapter \ref{chap:instructions}, the packages used are briefly explained, and instructions are given, how the bibliography and the glossary are to be used.
-Chapter \ref{chap:typeareatest} presents a sample chapter to audit the type area.
+The longer and complexer a source code is, the more difficult it is to verify its correctness.
+There are two main approaches of the program verification. The commonly used one is the simulation-based verification.
 
-In figure \ref{fig:file_structure} the file structure is shown for this template.
+To simulate the software design some practical scenarios and assertions are created by a developer at the software testing stage. 
+But it not achievable to simulate all possible states and to test all possible inputs keeping the ability to observe the side effects of a specific part of code.
 
-\begin{figure}[H]
-	\centering
-		\includegraphics[scale=0.85]{images/file_structure.png}
-	\caption{File structure}
-	\label{fig:file_structure}
-\end{figure}
+Another method to find out whether a program works correct is formal verification. It is a systematic process based on mathematical modeling.
+The correctness of a program is analyzed relative to its formal specification. The high level requirements are written corresponding to the micro-architecture specification.
+With formal verification all possible input values are explored algorithmically and exhaustively. \cite{sanghavi:formal_verification}
+
+%In figure \ref{fig:file_structure} the file structure is shown for this template.
+
+%\begin{figure}[H]
+%	\centering
+%		\includegraphics[scale=0.85]{images/file_structure.png}
+%	\caption{File structure}
+%	\label{fig:file_structure}
+%\end{figure}
 

### 2019-04-09 doc/chapters/introduction.tex
Added the section Stainless to introduction.tex

@@
 
-\section{Contact}
-\label{sec:introduction_contact}
+\section{Stainless}
+\label{sec:stainless}
 
-The manufacturers of this template welcome any suggestions for improvement. Chapter \ref{sec:introduction_suggestions} shows possible suggestions for improvement\index{suggestions}.
+Stainless is a framework developed by "Lab for Automated Reasoning and Analysis" (LARA) at EPFL's School of Computer and Communication Sciences to verify Scala programs.
+It explores all input values, reports inputs for which a program fails and demonstrates counterexamples which violate a given specification.
+The specification must be provided by a developer through giving a precondition and postcondition for functions.
+Constrains for input values are specified in a precondition, and for output values - in a postcondition.
+While compiling Stainless tries to prove that for the given precondition of input the postcondition of output always holds.
+Three outcomes of the verification are possible: valid, unvalid and unknown when the framework is unable to prove or find a counterexample.
+The code to be verified should be written in Pure Scala belonging to functional programming paradigm.
+To extend this subset of the Scala language Stainless supports also some imperative features translating them into Pure Scala concepts.
+For more details about Stainless its official website can be explored. \cite{Stainless}
 
-\begin{table}[H]
-	\centering
-		\begin{tabular}{lll} \toprule
-			\textbf{First Name Last Name} & \textbf{E-mail} & \textbf{Function} \\ \midrule
-			Alfred Kaufmann & alfred.kaufmann@bfh.ch & Employer, Project Management, \\
-			& & Supplements, Improvements \\ \midrule
-			Fritz Dellsperger & Retired & Tips on the structure and layout \\ \midrule
-			David Burri & Contracted out & First compilation of the Template \\ \bottomrule
-		\end{tabular}
-	\caption{Contact Persons}
-	\label{tab:Contact Persons}
-\end{table}
+%\begin{table}[H]
+%	\centering
+%		\begin{tabular}{lll} \toprule
+%			\textbf{First Name Last Name} & \textbf{E-mail} & \textbf{Function} \\ \midrule
+%			Alfred Kaufmann & alfred.kaufmann@bfh.ch & Employer, Project Management, \\
+%			& & Supplements, Improvements \\ \midrule
+%			Fritz Dellsperger & Retired & Tips on the structure and layout \\ \midrule
+%			David Burri & Contracted out & First compilation of the Template \\ \bottomrule
+%		\end{tabular}
+%	\caption{Contact Persons}
+%	\label{tab:Contact Persons}
+%\end{table}
 

### 2019-04-29 doc/chapters/introduction.tex
Updated Formal verification in intruduction.tex

@@
 The longer and complexer a source code is, the more difficult it is to verify its correctness.
-There are two main approaches of the program verification. The commonly used one is the simulation-based verification.
+There are two main approaches for verification of program correctness.
 
-To simulate the software design some practical scenarios and assertions are created by a developer at the software testing stage. 
-But it not achievable to simulate all possible states and to test all possible inputs keeping the ability to observe the side effects of a specific part of code.
+The commonly used one is the simulation-based verification.
+To simulate a software design a set of some practical scenarios and assertions is created by a developer at the software testing stage. 
+While testing it is tried to identify a design error by generating an input which can activate a bug.
+It is not achievable to simulate all possible states and to test all possible inputs. 
+Thus, a long-term test coverage model should be created to simulate a software design enough.
+Furthermore it is challenging to test a software keeping the ability to observe the side effects of a specific part of code.
+Practically, a developer runs tests, debugs appeared failures, adjustes a code of a software and extands a testbench verifying previously uncovered aspects.\cite{sanghavi:formal_verification}
 
-Another method to find out whether a program works correct is formal verification. It is a systematic process based on mathematical modeling.
-The correctness of a program is analyzed relative to its formal specification. The high level requirements are written corresponding to the micro-architecture specification.
+Another powerfull method to discover whether a program works correct is formal verification. 
+It is a systematic process based on mathematical modeling. 
+It is unnecessary to create a simulation tests. 
+Correctness of a program is analyzed relative to its formal specification.
+Formal specification is a mathematical description of a software behavior that can be provided to formal verification tools for proving it.
 With formal verification all possible input values are explored algorithmically and exhaustively. \cite{sanghavi:formal_verification}

### 2019-04-30 code/currencysnippet/src/main/scala/example/BasicArithmetic.scala
imported full code

@@ -1,0 +1,47 @@
+package example
+
+import scala.util.Try
+
+/**
+  * @define mulSafe
+  * Some classes have restrictions on upper bounds
+  * for it's underlying value. This might cause the `*`
+  * operator to throw. This method wraps it in a `Try`
+  * block.
+  */
+trait BasicArithmetic[N] {
+
+  def +(n: N): N
+
+  /**
+    * Some classes have restrictions on upper bounds
+    * for it's underlying value. This might cause the `+`
+    * operator to throw. This method wraps it in a `Try`
+    * block.
+    */
+  def addSafe(n: N): Try[N] = Try { this + n }
+
+  def -(n: N): N
+
+  /**
+    * Some classes have restrictions on lower bounds
+    * for it's underlying value. This might cause the `-`
+    * operator to throw. This method wraps it in a `Try`
+    * block.
+    */
+  def subtractSafe(n: N): Try[N] = Try { this - n }
+
+  def *(factor: BigInt): N
+
+  /**
+    * $mulSafe
+    */
+  def multiplySafe(factor: BigInt): Try[N] = Try { this * factor }
+
+  def *(factor: N): N
+
+  /**
+    * $mulSafe
+    */
+  def multiplySafe(factor: N): Try[N] = Try { this * factor }
+}

### 2019-04-30 code/currencysnippet/src/main/scala/example/Consensus.scala

@@ -1,0 +1,31 @@
+package example
+
+/**
+  * Created by chris on 5/13/16.
+  */
+sealed abstract class Consensus {
+
+  def maxBlockSize: Long = 1000000
+
+  def weightScalar: Long = 4
+
+  def maxBlockWeight: Long = maxBlockSize * weightScalar
+
+  /**
+    * BIP141 changes this from 20,000 -> 80,000, to see how sigops are counted please see BIP 141
+    * [[https://github.com/bitcoin/bips/blob/master/bip-0141.mediawiki#sigops]]
+    */
+  def maxSigOps = 80000
+
+  def maxMoney: CurrencyUnit = Satoshis(Int64(2100000000000000L))
+
+  /**
+    * A integer representing the maximum number of public keys you can have in a
+    * OP_CHECKMULTISIG or OP_CHECKMULTISIGVERIFY operation
+    * https://github.com/bitcoin/bitcoin/blob/master/src/script/interpreter.cpp#L903
+    * @return
+    */
+  def maxPublicKeysPerMultiSig = 20
+}
+
+object Consensus extends Consensus

### 2019-04-30 code/currencysnippet/src/main/scala/example/CurrencyUnits.scala

@@ -1,0 +1,125 @@
+package example
+
+sealed abstract class CurrencyUnit
+    extends BasicArithmetic[CurrencyUnit] {
+  type A
+
+  def satoshis: Satoshis
+
+  def >=(c: CurrencyUnit): Boolean = {
+    satoshis.underlying >= c.satoshis.underlying
+  }
+
+  def >(c: CurrencyUnit): Boolean = {
+    satoshis.underlying > c.satoshis.underlying
+  }
+
+  def <(c: CurrencyUnit): Boolean = {
+    satoshis.underlying < c.satoshis.underlying
+  }
+
+  def <=(c: CurrencyUnit): Boolean = {
+    satoshis.underlying <= c.satoshis.underlying
+  }
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  override def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
+
+  override def -(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying - c.satoshis.underlying)
+  }
+
+  override def *(factor: BigInt): CurrencyUnit = {
+    Satoshis(satoshis.underlying * factor)
+  }
+
+  override def *(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying * c.satoshis.underlying)
+  }
+
+  def unary_- : CurrencyUnit = {
+    Satoshis(-satoshis.underlying)
+  }
+
+  def toBigDecimal: BigDecimal
+
+  protected def underlying: A
+}
+
+sealed abstract class Satoshis extends CurrencyUnit {
+  override type A = Int64
+
+  override def toString: String = s"$toLong sat"
+
+  override def satoshis: Satoshis = this
+
+  override def toBigDecimal = BigDecimal(toBigInt)
+
+  def toBigInt: BigInt = BigInt(toLong)
+
+  def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}
+
+object Satoshis extends BaseNumbers[Satoshis] {
+
+  val min = Satoshis(Int64.min)
+  val max = Satoshis(Int64.max)
+  val zero = Satoshis(Int64.zero)
[…]

### 2019-04-30 code/currencysnippet/src/main/scala/example/NumberType.scala

@@ -1,0 +1,310 @@
+package example
+
+import scala.util.{Failure, Success, Try}
+
+/**
+  * Created by chris on 6/4/16.
+  */
+/**
+  * This abstract class is meant to represent a signed and unsigned number in C
+  * This is useful for dealing with codebases/protocols that rely on C's
+  * unsigned integer types
+  */
+sealed abstract class Number[T <: Number[T]]
+    extends BasicArithmetic[T] {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toInt: Int = toBigInt.bigInteger.intValueExact()
+  def toLong: Long = toBigInt.bigInteger.longValueExact()
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => T
+
+  override def +(num: T): T = apply(checkResult(underlying + num.underlying))
+  override def -(num: T): T = apply(checkResult(underlying - num.underlying))
+  override def *(factor: BigInt): T = apply(checkResult(underlying * factor))
+  override def *(num: T): T = apply(checkResult(underlying * num.underlying))
+
+  def >(num: T): Boolean = underlying > num.underlying
+  def >=(num: T): Boolean = underlying >= num.underlying
+  def <(num: T): Boolean = underlying < num.underlying
+  def <=(num: T): Boolean = underlying <= num.underlying
+
+  def <<(num: Int): T = this.<<(apply(num))
+  def >>(num: Int): T = this.>>(apply(num))
+
+  def <<(num: T): T = {
+    checkIfInt(num).map { _ =>
+      apply((underlying << num.toInt) & andMask)
+    }.get
+  }
+
+  def >>(num: T): T = {
+    //this check is for weird behavior with the jvm and shift rights
+    //https://stackoverflow.com/questions/47519140/bitwise-shift-right-with-long-not-equaling-zero/47519728#47519728
+    if (num.toLong > 63) apply(0)
+    else {
+      checkIfInt(num).map { _ =>
[…]

### 2019-04-30 code/currencysnippet/src/main/scala/example/BasicArithmetic.scala
removed comments

@@ -2,46 +2,21 @@ package example
 
 import scala.util.Try
 
-/**
-  * @define mulSafe
-  * Some classes have restrictions on upper bounds
-  * for it's underlying value. This might cause the `*`
-  * operator to throw. This method wraps it in a `Try`
-  * block.
-  */
 trait BasicArithmetic[N] {
 
   def +(n: N): N
 
-  /**
-    * Some classes have restrictions on upper bounds
-    * for it's underlying value. This might cause the `+`
-    * operator to throw. This method wraps it in a `Try`
-    * block.
-    */
   def addSafe(n: N): Try[N] = Try { this + n }
 
   def -(n: N): N
 
-  /**
-    * Some classes have restrictions on lower bounds
-    * for it's underlying value. This might cause the `-`
-    * operator to throw. This method wraps it in a `Try`
-    * block.
-    */
   def subtractSafe(n: N): Try[N] = Try { this - n }
 
   def *(factor: BigInt): N
 
-  /**
-    * $mulSafe
-    */
   def multiplySafe(factor: BigInt): Try[N] = Try { this * factor }
 
   def *(factor: N): N
 
-  /**
-    * $mulSafe
-    */
   def multiplySafe(factor: N): Try[N] = Try { this * factor }
 }

### 2019-04-30 code/currencysnippet/src/main/scala/example/NumberType.scala

@@ -2,33 +2,18 @@ package example
 
 import scala.util.{Failure, Success, Try}
 
-/**
-  * Created by chris on 6/4/16.
-  */
-/**
-  * This abstract class is meant to represent a signed and unsigned number in C
-  * This is useful for dealing with codebases/protocols that rely on C's
-  * unsigned integer types
-  */
 sealed abstract class Number[T <: Number[T]]
     extends BasicArithmetic[T] {
   type A = BigInt
 
-  /** The underlying scala number used to to hold the number */
   protected def underlying: A
 
   def toInt: Int = toBigInt.bigInteger.intValueExact()
   def toLong: Long = toBigInt.bigInteger.longValueExact()
   def toBigInt: BigInt = underlying
 
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
   def andMask: BigInt
 
-  /** Factory function to create the underlying T, for instance a UInt32 */
   def apply: A => T
 
   override def +(num: T): T = apply(checkResult(underlying + num.underlying))
@@ -65,17 +50,12 @@ sealed abstract class Number[T <: Number[T]]
   def &(num: T): T = apply(checkResult(underlying & num.underlying))
   def unary_- : T = apply(-underlying)
 
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
   private def checkResult(result: BigInt): A = {
     require((result & andMask) == result,
             "Result was out of bounds, got: " + result)
     result
   }
 
-  /** Checks if the given nubmer is within range of a Int */
   private def checkIfInt(num: T): Try[Unit] = {
     if (num.toBigInt >= Int.MaxValue || num.toBigInt <= Int.MinValue) {
       Failure(
@@ -87,16 +67,8 @@ sealed abstract class Number[T <: Number[T]]
   }
 }
 
-/**
-  * Represents a signed number in our number system
-  * Instances of this are [[Int32]] or [[Int64]]
-  */
 sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
 
-/**
-  * Represents an unsigned number in our number system
-  * Instances of this are [[UInt32]] or [[UInt64]]
-  */
 sealed abstract class UnsignedNumber[T <: Number[T]] extends Number[T]
 
[…]

### 2019-04-30 code/currencysnippet/src/test/scala/example/CurrencyUnitSpec.scala
added test

@@ -1,0 +1,15 @@
+package example
+
+import scala.util.Try
+import org.scalacheck.{Gen, Prop, Properties}
+
+/**
+  * Created by chris on 6/23/16.
+  */
+class CurrencyUnitSpec extends Properties("CurrencyUnitSpec") {
+  // Modified
+  property("additive identity") = Prop.forAll(Satoshis.zero) {
+    satoshis =>
+      satoshis + CurrencyUnits.zero == satoshis
+  }
+}

### 2019-04-30 code/currencysnippet/src/main/scala/example/BasicArithmetic.scala
removed unused operators and number types

@@ -3,20 +3,5 @@ package example
 import scala.util.Try
 
 trait BasicArithmetic[N] {
-
   def +(n: N): N
-
-  def addSafe(n: N): Try[N] = Try { this + n }
-
-  def -(n: N): N
-
-  def subtractSafe(n: N): Try[N] = Try { this - n }
-
-  def *(factor: BigInt): N
-
-  def multiplySafe(factor: BigInt): Try[N] = Try { this * factor }
-
-  def *(factor: N): N
-
-  def multiplySafe(factor: N): Try[N] = Try { this * factor }
 }

### 2019-04-30 code/currencysnippet/src/main/scala/example/CurrencyUnits.scala

@@ -30,22 +30,6 @@ sealed abstract class CurrencyUnit
     Satoshis(satoshis.underlying + c.satoshis.underlying)
   }
 
-  override def -(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying - c.satoshis.underlying)
-  }
-
-  override def *(factor: BigInt): CurrencyUnit = {
-    Satoshis(satoshis.underlying * factor)
-  }
-
-  override def *(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying * c.satoshis.underlying)
-  }
-
-  def unary_- : CurrencyUnit = {
-    Satoshis(-satoshis.underlying)
-  }
-
   def toBigDecimal: BigDecimal
 
   protected def underlying: A
@@ -79,35 +63,6 @@ object Satoshis extends BaseNumbers[Satoshis] {
   private case class SatoshisImpl(underlying: Int64) extends Satoshis
 }
 
-sealed abstract class Bitcoins extends CurrencyUnit {
-  override type A = BigDecimal
-
-  override def toString: String = s"$toBigDecimal BTC"
-
-  override def toBigDecimal: BigDecimal = underlying
-
-  override def satoshis: Satoshis = {
-    val sat = underlying * CurrencyUnits.btcToSatoshiScalar
-    Satoshis(Int64(sat.toLongExact))
-  }
-}
-
-object Bitcoins extends BaseNumbers[Bitcoins] {
-  val min = Bitcoins((-Consensus.maxMoney).satoshis)
-  val max = Bitcoins(Consensus.maxMoney.satoshis)
-  val zero = Bitcoins(Satoshis.zero)
-  val one = Bitcoins(1)
-
-  def apply(satoshis: Satoshis): Bitcoins = {
-    val b: BigDecimal = satoshis.toLong * CurrencyUnits.satoshisToBTCScalar
-    Bitcoins(b)
-  }
-
-  def apply(underlying: BigDecimal): Bitcoins = BitcoinsImpl(underlying)
-
-  private case class BitcoinsImpl(underlying: BigDecimal) extends Bitcoins
-}
-
 object CurrencyUnits {
 
   /** The number you need to multiply BTC by to get it's satoshis */
@@ -119,7 +74,6 @@ object CurrencyUnits {
   val negativeSatoshi = Satoshis(Int64(-1))
 
[…]

### 2019-04-30 code/currencysnippet/src/main/scala/example/NumberType.scala

@@ -17,9 +17,6 @@ sealed abstract class Number[T <: Number[T]]
   def apply: A => T
 
   override def +(num: T): T = apply(checkResult(underlying + num.underlying))
-  override def -(num: T): T = apply(checkResult(underlying - num.underlying))
-  override def *(factor: BigInt): T = apply(checkResult(underlying * factor))
-  override def *(num: T): T = apply(checkResult(underlying * num.underlying))
 
   def >(num: T): Boolean = underlying > num.underlying
   def >=(num: T): Boolean = underlying >= num.underlying
@@ -69,45 +66,6 @@ sealed abstract class Number[T <: Number[T]]
 
 sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
 
-sealed abstract class UnsignedNumber[T <: Number[T]] extends Number[T]
-
-sealed abstract class UInt5 extends UnsignedNumber[UInt5] {
-  override def apply: A => UInt5 = UInt5(_)
-
-  override def andMask: BigInt = 0x1f
-
-  def byte: Byte = toInt.toByte
-
-  def toUInt8: UInt8 = UInt8(toInt)
-}
-
-sealed abstract class UInt8 extends UnsignedNumber[UInt8] {
-  override def apply: A => UInt8 = UInt8(_)
-
-  override def andMask = 0xff
-
-  def toUInt5: UInt5 = {
-    //this will throw if not in range of a UInt5, come back and look later
-    UInt5(toInt)
-  }
-}
-
-sealed abstract class UInt32 extends UnsignedNumber[UInt32] {
-  override def apply: A => UInt32 = UInt32(_)
-
-  override def andMask = 0xffffffffL
-}
-
-sealed abstract class UInt64 extends UnsignedNumber[UInt64] {
-  override def apply: A => UInt64 = UInt64(_)
-  override def andMask = 0xffffffffffffffffL
-}
-
-sealed abstract class Int32 extends SignedNumber[Int32] {
-  override def apply: A => Int32 = Int32(_)
-  override def andMask = 0xffffffff
-}
-
 sealed abstract class Int64 extends SignedNumber[Int64] {
   override def apply: A => Int64 = Int64(_)
   override def andMask = 0xffffffffffffffffL
@@ -118,132 +76,6 @@ trait BaseNumbers[T] {
   def one: T
   def min: T
   def max: T
-}
-
[…]

### 2019-04-30 code/currencysnippet/src/main/scala/example/Consensus.scala
removed more code

@@ -1,18 +1,0 @@
-package example
-
-sealed abstract class Consensus {
-
-  def maxBlockSize: Long = 1000000
-
-  def weightScalar: Long = 4
-
-  def maxBlockWeight: Long = maxBlockSize * weightScalar
-
-  def maxSigOps = 80000
-
-  def maxMoney: CurrencyUnit = Satoshis(Int64(2100000000000000L))
-
-  def maxPublicKeysPerMultiSig = 20
-}
-
-object Consensus extends Consensus

### 2019-04-30 code/currencysnippet/src/main/scala/example/CurrencyUnits.scala

@@ -6,26 +6,6 @@ sealed abstract class CurrencyUnit
 
   def satoshis: Satoshis
 
-  def >=(c: CurrencyUnit): Boolean = {
-    satoshis.underlying >= c.satoshis.underlying
-  }
-
-  def >(c: CurrencyUnit): Boolean = {
-    satoshis.underlying > c.satoshis.underlying
-  }
-
-  def <(c: CurrencyUnit): Boolean = {
-    satoshis.underlying < c.satoshis.underlying
-  }
-
-  def <=(c: CurrencyUnit): Boolean = {
-    satoshis.underlying <= c.satoshis.underlying
-  }
-
-  def !=(c: CurrencyUnit): Boolean = !(this == c)
-
-  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
   override def +(c: CurrencyUnit): CurrencyUnit = {
     Satoshis(satoshis.underlying + c.satoshis.underlying)
   }
@@ -64,8 +44,6 @@ object Satoshis extends BaseNumbers[Satoshis] {
 }
 
 object CurrencyUnits {
-
-  /** The number you need to multiply BTC by to get it's satoshis */
   val btcToSatoshiScalar: Long = 100000000
   val satoshisToBTCScalar: BigDecimal = BigDecimal(1.0) / btcToSatoshiScalar
   val oneBTC: CurrencyUnit = Satoshis(Int64(btcToSatoshiScalar))

### 2019-04-30 code/currencysnippet/src/main/scala/example/NumberType.scala

@@ -18,49 +18,10 @@ sealed abstract class Number[T <: Number[T]]
 
   override def +(num: T): T = apply(checkResult(underlying + num.underlying))
 
-  def >(num: T): Boolean = underlying > num.underlying
-  def >=(num: T): Boolean = underlying >= num.underlying
-  def <(num: T): Boolean = underlying < num.underlying
-  def <=(num: T): Boolean = underlying <= num.underlying
-
-  def <<(num: Int): T = this.<<(apply(num))
-  def >>(num: Int): T = this.>>(apply(num))
-
-  def <<(num: T): T = {
-    checkIfInt(num).map { _ =>
-      apply((underlying << num.toInt) & andMask)
-    }.get
-  }
-
-  def >>(num: T): T = {
-    //this check is for weird behavior with the jvm and shift rights
-    //https://stackoverflow.com/questions/47519140/bitwise-shift-right-with-long-not-equaling-zero/47519728#47519728
-    if (num.toLong > 63) apply(0)
-    else {
-      checkIfInt(num).map { _ =>
-        apply(underlying >> num.toInt)
-      }.get
-    }
-  }
-
-  def |(num: T): T = apply(checkResult(underlying | num.underlying))
-  def &(num: T): T = apply(checkResult(underlying & num.underlying))
-  def unary_- : T = apply(-underlying)
-
   private def checkResult(result: BigInt): A = {
     require((result & andMask) == result,
             "Result was out of bounds, got: " + result)
     result
-  }
-
-  private def checkIfInt(num: T): Try[Unit] = {
-    if (num.toBigInt >= Int.MaxValue || num.toBigInt <= Int.MinValue) {
-      Failure(
-        new IllegalArgumentException(
-          "Num was not in range of int, got: " + num))
-    } else {
-      Success(())
-    }
   }
 }
 

### 2019-05-07 doc/chapters/introduction.tex
Updated Tables of content in .tex

@@
-\chapter{Introduction}
+\chapter{Introduction (Aufgabestellung)}
 \label{chap:introduction}
+
+In this work some experiments in formal verification of Scala code will be accomplished. 
+The framework Stainless is used as a verification tool. The code of Bitcoin-S, Scala implementation of Bintcoin, is taken as an input for Stainless to be verified. 
+In following the main aspects of formal verification, Stainless and Bitcoin-S are described.
 
@@
 
-\section{Formal verification}
+\section{Formal verification (formal verification vs. simulation-based)}
 \label{sec:formal_verification}
@@
 Furthermore it is challenging to test a software keeping the ability to observe the side effects of a specific part of code.
-Practically, a developer runs tests, debugs appeared failures, adjustes a code of a software and extands a testbench verifying previously uncovered aspects.\cite{sanghavi:formal_verification}
+Practically, a developer runs tests, debugs appeared failures, adjustes a code of a software and extands a testbench verifying previously uncovered aspects. \cite{sanghavi:formal_verification}
 
 Another powerfull method to discover whether a program works correct is formal verification. 
-It is a systematic process based on mathematical modeling. 
-It is unnecessary to create a simulation tests. 
+This is a systematic process based on mathematical modeling. 
+It is unnecessary to create a simulation tests with some possible inputs. 
+With formal verification all possible input values are explored algorithmically and exhaustively.
 Correctness of a program is analyzed relative to its formal specification.
-Formal specification is a mathematical description of a software behavior that can be provided to formal verification tools for proving it.
-With formal verification all possible input values are explored algorithmically and exhaustively. \cite{sanghavi:formal_verification}
+Formal specification is a mathematical description of a software behavior that can be provided to formal verification tools for proving it. \cite{sanghavi:formal_verification}
+
+One of such tools is Stainless. It is used in this work to verify a Scala code.
 
@@
 
-\section{Stainless}
+\section{Stainless (allgemeine Beschreibung: Struktur with pre- and postconditions, Solvers, Terminierung und mögliche Outputs)}
 \label{sec:stainless}
@@
 The specification must be provided by a developer through giving a precondition and postcondition for functions.
-Constrains for input values are specified in a precondition, and for output values - in a postcondition.
-While compiling Stainless tries to prove that for the given precondition of input the postcondition of output always holds.
+
+Constrains for input values are specified in a precondition with the keyword \textit{require}. 
+Restrictions for received output values are defined in a postcondition with the keyword \textit{ensuring}. 
+While compiling Stainless tries to prove that for the given input the postcondition of output always holds.
 Three outcomes of the verification are possible: valid, unvalid and unknown when the framework is unable to prove or find a counterexample.
@@
 
-\section{Suggestions for Improvement}
-\label{sec:introduction_suggestions}
+\section{Bitcoin-S (Projekten und Packages, bitcoin-s-core, Eigenschaften zu prüfen)}
+\label{sec:bitcoin_s}
 
-\begin{itemize}
-	\item Create a BFH Style Files
-	\item Template for the Compilation of presentations with \LaTeX{}
-\end{itemize}
+%\begin{itemize}
+%	\item Create a BFH Style Files
+%	\item Template for the Compilation of presentations with \LaTeX{}
+%\end{itemize}

### 2019-05-08 code/currencysnippet/src/main/scala/example/BasicArithmetic.scala
management summary -> abstract

@@ -1,7 +1,0 @@
-package example
-
-import scala.util.Try
-
-trait BasicArithmetic[N] {
-  def +(n: N): N
-}

### 2019-05-08 doc/chapters/towards_verifying.tex
changed top level titles

\chapter{Towards verifyng Bitcoin-S with Stainless}
\label{chap:connecting}

\section{Integration (Versionskonflikte, neues Plugin)}

\section{Error reporting with sbt und jar}

\section{Extraction of code fragment from Bitcoin-S}

\section{Rewriting of code}

\section{Verification of method "+"}

### 2019-05-08 doc/chapters/using_bitcoins.tex
changed top level titles

\chapter{Using Bitcoin-S}
\label{chap:using_bitcoins}

\section{Creating of a transaction (valid, unvalid with TxBuilder)}

\section{Validation of a transaction (with method checkTransaction)}

\section{{Bugfix for noDublicate}}

### 2019-05-08 doc/chapters/using_stainless.tex
changed top level titles

\chapter{Using of Stainless}
\label{chap:using_stainless}

\section{Konfiguration (Set-up, sbt, jar, frontends/../stainless-scalac)}

\section{Scala compatibility (Pure Scala, imperative features, dedicated BigInt, Generics...)}

### 2019-05-14 doc/leader/frontpage_with_picture.tex
titlepage

@@
 \fontsize{10pt}{12pt}\selectfont
-\textbf{Description of thesis (semester- / Bachelor thesis / etc.)} \\		% Insert text
+\textbf{Bachelor thesis} \\		% Insert text
 \vspace{3mm}
@@
 xxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-Degree course:	\> [z.B. Electrical and Communication Engineering]	\\		% insert name of degree course
-Authors:		\> [Test Peter, M\"uster R\"os\"a]		\\					% insert names
-Tutor:	\> [Dr.~Xxxx Xxxx, Dr.~Yyyy Yyyy]		\\							% insert names
-Constituent:	\> [Wwwww AG]					\\							% insert names
-Experts:		\> [Dr.~Zzzz Zzzz]				\\							% insert names
+Degree course:	\> Computer Science	\\		% insert name of degree course
+Authors:		\> Anna Doukmak, Ramon Boss		\\					% insert names
+Tutor:	\> Kai Brünnler @TODO Dr.?		\\							% insert names
+%Constituent:	\> [Wwwww AG]					\\							% insert names
+Experts:		\> Urs Keller				\\							% insert names
 Date:			\> \versiondate					\\							% read from file leader/version.tex

### 2019-05-14 doc/chapters/introduction.tex
fixed some typos
Updated fromal verificaiton vs. testing in introduction.tex

@@
 In this work some experiments in formal verification of Scala code will be accomplished. 
-The framework Stainless is used as a verification tool. The code of Bitcoin-S, Scala implementation of Bintcoin, is taken as an input for Stainless to be verified. 
+The framework Stainless is used as a verification tool. The code of Bitcoin-S-Core, Scala implementation of Bitcoin protocol, is taken as an input for Stainless to be verified. 
 In following the main aspects of formal verification, Stainless and Bitcoin-S are described.
@@
 
-\section{Formal verification (formal verification vs. simulation-based)}
+\section{Formal verification (formal verification vs. testing)}
 \label{sec:formal_verification}
@@
 The longer and complexer a source code is, the more difficult it is to verify its correctness.
-There are two main approaches for verification of program correctness.
+There are different approaches for verification of program correctness.
 
-The commonly used one is the simulation-based verification.
-To simulate a software design a set of some practical scenarios and assertions is created by a developer at the software testing stage. 
-While testing it is tried to identify a design error by generating an input which can activate a bug.
-It is not achievable to simulate all possible states and to test all possible inputs. 
-Thus, a long-term test coverage model should be created to simulate a software design enough.
+The commonly used one is testing. A set of some practical scenarios and assertions is created by a developer to test a software design and implementation. 
+While testing it is checked whether the actual results match the expected results. Tests help to identify bugs and missing requirements.
+But it is not achievable to test all possible inputs. 
+Thus, a long-term test coverage model should be created to test a software design enough.
 Furthermore it is challenging to test a software keeping the ability to observe the side effects of a specific part of code.
@@
 
-Another powerfull method to discover whether a program works correct is formal verification. 
+Another powerfull method to check whether a program works correct is formal verification. 
 This is a systematic process based on mathematical modeling. 
@@
 
-One of such tools is Stainless. It is used in this work to verify a Scala code.
+One of such tools is Stainless. This framework is used in the work to verify a Scala code.
 

### 2019-05-20 doc/chapters/using_stainless.tex
changed subsections
initial configration sbt

@@
-\chapter{Using of Stainless}
+\chapter{Using Stainless}
 \label{chap:using_stainless}
+This chapter describes the setup and integration process of Stainless in a new or already existing project.
+It also shows the compatibility of Stainless with Scala as Stainless supports only a purely functional subset of Scala which they call \emph{Pure Scala}.
 
-\section{Konfiguration (Set-up, sbt, jar, frontends/../stainless-scalac)}
+\section{Configuration (Set-up, sbt, jar, frontends/../stainless-scalac)}
+There are two ways to integrate Stainless in a Scala project, the scala build tool (abt) plugin or the stainless.jar file.
+Both, when run, analyses the passed code and report warnings to the console about the given code.
+Stainless requires and Scala recommends Java SE Development Kit 8.
+Newer Java versions won't compile.
+
+\subsection{sbt}
+sbt for Scala is like gradle or maven for Java.
+It can compile Scala code continously or manual, manage dependencies with support for Maven-formatted repositories, mixing Scala and Java projects and much more.
+
+A simple sbt project has the following structure:
+\dirtree{%
+  .1 ..
+  .2 project.
+  .3 build.properties.
+  .3 plugins.sbt.
+  .2 src.
+  .3 main.
+  .4 scala.
+  .5 Main.scala.
+  .2 build.sbt.
+}
+\emph{build.properties} specifies the abt version used for this project.
+If the version is not available locally, sbt will download it.
+
+In \emph{plugins.sbt} new sbt plugins can be added.
+A plugin extends the build definition.
+Mostly this means adding and overriding settings.
+
+\emph{build.sbt} defines the build definition.
+There can be several projects or subprojects as sbt doc calls it.
+
+Here an example for a single project in \emph{build.sbt}:
+\begin{lstlisting}[language=scala]
+scalaVersion := "2.12.8"
+
+lazy val root = (project in file("."))
+\end{lstlisting}
+The project is called root and its source files are located in the files directory.
+Executing \code{sbt compile} should now compile the code.
+
+The Stainless webpage has a guide on how to integrate Stainless in an existing project.
+The simplifies steps are:
+\begin{itemize}
+  \item Install an external solver.
+  \item Add Stainless sbt plugin to \emph{plugins.sbt}
+  \item Enable the plugin in \emph{build.sbt} for the project.
+\end{itemize}
+After this setup, Stainless will report errors to the console, when running \code{sbt compile}.
+
+\subsection{JAR file}
 

### 2019-05-22 doc/chapters/using_stainless.tex
more text

@@
 \section{Configuration (Set-up, sbt, jar, frontends/../stainless-scalac)}
-There are two ways to integrate Stainless in a Scala project, the scala build tool (abt) plugin or the stainless.jar file.
+There are two ways to integrate Stainless in a Scala project, the scala build tool (abt) plugin or a command line tool.
 Both, when run, analyses the passed code and report warnings to the console about the given code.
@@
 
-\subsection{JAR file}
+\subsection{Command Line Tool}
+There are two ways to use the command line tool.
+
+Either download a prebuilt JAR file from efpl-lara/stainless GitHub repository or built a binary from source.
+The prebuild versions are released by Stainless.
+The latest was released on January 14, when there was no support for Scala 2.12.
+The 'Bump Scala to 2.12.8' branch was merged on March 4.
+
+If latest features are needed, like support for Scala 2.12, the build from source is required.
+Here a short installation description.
+Full description can be found on the Stainless documentation pages.
+\begin{itemize}
+  \item Install sbt.
+  \item Check out GitHub repository.
+  \item Run \code{sbt universal:stage} inside the project.
+\end{itemize}
+This generates \emph{frontends/scalac/target/universal/stage/bin/stainless-scalac}.
+
+To check the source code with one of those either \code{java -jar downloaded.jar source.scala} or \code{stainless-scalac source.scala} must be invoked.
+The file \emph{source.scala} is the file to be checked.
+
+As compiling without a build tool, this command will become really complex for bigger projects.
+All dependencies must be on the classpath and all source files appended.
+Those are added with \code{-classpath Dep1.jar:Dep2.jar:...:DepN.jar src1.scala src2.scala ... srcM.scala}.
 

### 2019-05-22 doc/chapters/using_bitcoins.tex
changed subsections
chapter intro

@@
-\chapter{Using Bitcoin-S}
+\chapter{Using Bitcoin-S-Core}
 \label{chap:using_bitcoins}
+This chapter describes the relevant parts needed to verify that a non-coinbase transaction cannot generate new coins in Bitcoin-S-Core.
+We will see, how to create valid and invalid transactions with Bitcoin-S-Core and how this transactions are validated against the described property.
 
@@
 \section{Validation of a transaction (with method checkTransaction)}
-
-\section{{Bugfix for noDublicate}}

### 2019-05-22 doc/chapters/introduction.tex
Updated chapter Stainless in introduction.tex

@@
 
-Stainless is a framework developed by "Lab for Automated Reasoning and Analysis" (LARA) at EPFL's School of Computer and Communication Sciences to verify Scala programs.
-It explores all input values, reports inputs for which a program fails and demonstrates counterexamples which violate a given specification.
-The specification must be provided by a developer through giving a precondition and postcondition for functions.
+Stainless is a framework developed by "Lab for Automated Reasoning and Analysis" (LARA) at EPFL's School of Computer and Communication Sciences. The framework is used to verify Scala programs.
+It verifies statically that a program satisfies a specification given by a developer and that a program will not crash at runtime.
+Stainless explores all possible input values, reports inputs for which a program fails and demonstrates counterexamples which violate a given specification.
 
-Constrains for input values are specified in a precondition with the keyword \textit{require}. 
-Restrictions for received output values are defined in a postcondition with the keyword \textit{ensuring}. 
-While compiling Stainless tries to prove that for the given input the postcondition of output always holds.
-Three outcomes of the verification are possible: valid, unvalid and unknown when the framework is unable to prove or find a counterexample.
+The main functions used to write a specification are \textit{require} and \textit{ensuring}. 
+A precondition should be written at the beginning of a function body with \textit{require}.
+Its argument is a boolean expression which corresponds to constraint for inputs of a function being verified.
+A postcondition should be written after a function body with  \textit{ensuring}. It is a verification condition on an output of a function.
+While compiling Stainless tries to prove that the postcondition always holds, assuming a given precondition does hold.
+
+3 outcomes of verification with Stainless are possible: valid, unvalid and unknown. 
+If the postcondition is valid, Stainless could prove that for any inputs constrained in a precondition, the postcondition always holds.
+With invalid postcondition the framework could find at least one counterexample which satisfies a precondition but violates a postcondition.
+Also an output unknown is possible when Stainless is unable to prove a postcondition or find a counterexample. In this case a timeout or an internal error occured.
+Furthermore, it will be verified by Stainless that a precondition cannot be violated.
+
+The following example demonstrates a simple structure of formal specification for calculating factorial.
+
+\begin{lstlisting}[caption=Example of verifying the function calculating factorial of an integer number, captionpos=b, language=Scala]
+    def factorial(n: Int): Int = {
+        require(n >= 0)
+        if (n == 0) {
+          1
+        } else {
+          n * factorial(n - 1)
+        }
+    } ensuring(res => res >= 0)
+\end{lstlisting}
+
+The function recursively calculates factorial of an integer number.
+An input to the function is constrained in \textit{require} with positive value.
+A result of calculation should be also positive, what will be verified by Stainless.
+While compiling Stainless disproves the postcondition and gives the number 17 as a counterexample.
+A number of type Int is a 32 bit value.
+That is why the function factorial of 17 causes an overflow and results to a negative value.
+The program would work correct by changing type of a number to BigInt.
+The outcome of Stainless verification of the function factorial is shown bellow.
+
+\begin{figure}[H]
+	\centering
+		\includegraphics[scale=0.5]{images/output1.png}
+	\caption{Output of Stainless verification for calculating factorial of Int number}
+	\label{fig:output1}
+\end{figure}
+
 The code to be verified should be written in Pure Scala belonging to functional programming paradigm.

### 2019-05-22 doc/chapters/introduction.tex
Correct chapter Stainless in introduction.tex

@@
 
-\begin{lstlisting}[caption=Example of verifying the function calculating factorial of an integer number, captionpos=b, language=Scala]
+\begin{lstlisting}[caption=Example of verifying the function calculating factorial of an integer number, label=listing:factorial, captionpos=b, language=Scala]
     def factorial(n: Int): Int = {
@@
 The function recursively calculates factorial of an integer number.
-An input to the function is constrained in \textit{require} with positive value.
-A result of calculation should be also positive, what will be verified by Stainless.
+An input to the function is constrained in \textit{require} with non-negative value.
+A result of calculation should also be non-negative, what will be verified by Stainless.
 While compiling Stainless disproves the postcondition and gives the number 17 as a counterexample.
 A number of type Int is a 32 bit value.
-That is why the function factorial of 17 causes an overflow and results to a negative value.
-The program would work correct by changing type of a number to BigInt.
-The outcome of Stainless verification of the function factorial is shown bellow.
+That is why calculating factorial of 17 causes an overflow and results to a negative value.
+The program would work correct by changing a type of a number to BigInt.
+The outcome of Stainless verification of the function factorial from Listing \ref{listing:factorial} is shown bellow.
 

### 2019-05-22 doc/chapters/using_bitcoins.tex
more text

@@
 
-\section{Creating of a transaction (valid, unvalid with TxBuilder)}
+\section{Creation of a Transaction}
+Bitcoin-S-Core has a bitcoin transcation builder class with the following signature:
+\begin{lstlisting}[language=scala]
+  BitcoinTxBuilder(
+    destinations: Seq[TransactionOutput], // where we send money
+    utxos: BitcoinTxBuilder.UTXOMap,      // unspent transaction outputs
+    feeRate: FeeUnit,                     // fee rate per byte
+    changeSPK: ScriptPubKey,              // public key
+    network: BitcoinNetwork               // bitcoin network information
+  ): Future[BitcoinTxBuilder]             // sign TxBuilder to get tx
+\end{lstlisting}
+Here is how those parameters are generated.
 
-\section{Validation of a transaction (with method checkTransaction)}
+First, a previous transaction with outputs is needed to spend some money from.
+This is created here, to show the process.
+It could also be parsed from a transaction in the bitcoin network.
+A single output is sufficient for this example.
+So, first create a new keypair to sign the next transaction and have a scriptPubKey where the money is.
+Then define the amount to spend (here 10000 Satishis) collect this information in a transaction output and add it to the previous transaction.
+\begin{lstlisting}[language=scala]
+  val privKey = ECPrivateKey.freshPrivateKey
+  val creditingSPK = P2PKHScriptPubKey(pubKey = privKey.publicKey)
+
+  val amount = Satoshis(Int64(10000))
+
+  val utxo = TransactionOutput(currencyUnit = amount, scriptPubKey = creditingSPK)
+
+  val prevTx = BaseTransaction(
+    version = Int32.one,
+    inputs = List.empty,
+    outputs = List(utxo),
+    lockTime = UInt32.zero
+  )
+\end{lstlisting}
+Next, the new transaction should point to an output of the previous transaction.
+Thus, a outpoint is created with the id of the previous transaction and the index pointing to the specific output of it.
+This is collected in an utxo spending info which is then put in the list of all utxos (only one here).
+\begin{lstlisting}  
+  val outPoint = TransactionOutPoint(prevTx.txId, UInt32.zero)
+
+  val utxoSpendingInfo = BitcoinUTXOSpendingInfo(
+    outPoint = outPoint,
+    output = utxo,
+    signers = List(privKey),
+    redeemScriptOpt = None,
+    scriptWitnessOpt = None,
+    hashType = HashType.sigHashAll
+  )
+
+  val utxos = List(utxoSpendingInfo)
+\end{lstlisting}
+Then, the destination, where the money goes, is defined.
+This includes a destination script pub key, as well as a the amount to spend to it.
+\begin{lstlisting}
+  val destinationAmount = Satoshis(Int64(5000))
+
+  val destinationSPK = P2PKHScriptPubKey(pubKey = ECPrivateKey.freshPrivateKey.publicKey)
+
+  val destinations = List(
+    TransactionOutput(currencyUnit = destinationAmount, scriptPubKey = destinationSPK)
+  )
+\end{lstlisting}
+Finally. define a fee rate, the network params and create a transaction builder.
+\begin{lstlisting}
+  val feeRate = SatoshisPerByte(Satoshis.one)
+
+  val networkParams = RegTest // soem static values for testing
+
+  val txBuilder: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
+    destinations = destinations,
+    utxos = utxos,
+    feeRate = feeRate,
+    changeSPK = creditingSPK,
+    network = networkParams
+  )
+\end{lstlisting}
+After calling sign on the transaction builder a valid transaction is returned.
+\begin{lstlisting}
+  val signedTxF: Future[Transaction] = txBuilder
+    .flatMap(_.sign)
+    .map {
+      tx => println(tx.hex) // transaction in hex for the bitcoin network
+    }
+\end{lstlisting}
+There is no need for a transaction input, since a transaction out point is kind of the pointer to a transaction input.
+
+\section{Validation of a Transaction (with method checkTransaction)}

### 2019-05-22 doc/chapters/using_bitcoins.tex
added ref

@@
 \section{Creation of a Transaction}
+Some parts of the code in this section are from Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
 Bitcoin-S-Core has a bitcoin transcation builder class with the following signature:

### 2019-05-22 doc/chapters/using_bitcoins.tex
added type to transaction

@@
     .map {
-      tx => println(tx.hex) // transaction in hex for the bitcoin network
+      (tx: Transaction) => println(tx.hex) // transaction in hex for the bitcoin network
     }

### 2019-05-22 doc/chapters/using_bitcoins.tex
more text

@@
 \section{Validation of a Transaction (with method checkTransaction)}
+Bitcoin-S-Core offers a function called \emph{checkTransaction}.
+This is its type signature.
+\begin{lstlisting}
+  checkTransaction(transaction: Transaction): Boolean
+\end{lstlisting}
+It takes a transaction as input and returns a boolean whether the transaction is valid or not.
+There are several checks in checkTransaction.
+For example it checks if there is either no input or no output.
+In this case it returns false.
+
+The relevant parts for the property to be verified are the following two lines.
+\begin{lstlisting}
+  val prevOutputs = transaction.inputs.map(_.previousOutput)
+  val noDuplicateInputs = prevOutputs.distinct.size == prevOutputs.size
+\end{lstlisting}
+It gathers all inputs output.
+On this previous outputs it calls distinct and checks if the size stays the same.
+Distinct removes duplicates, so if there was two times the same input the size of inputs before calling distinct would be greater.

### 2019-05-22 doc/chapters/using_bitcoins.tex
remover comment from title
fixed text

@@
 
-\section{Validation of a Transaction (with method checkTransaction)}
+\section{Validation of a Transaction}
 Bitcoin-S-Core offers a function called \emph{checkTransaction}.
@@
 It takes a transaction as input and returns a boolean whether the transaction is valid or not.
+So for example when passing the built tx from above to this function the returned value would be true.
 There are several checks in checkTransaction.

### 2019-05-22 doc/chapters/using_stainless.tex
removed comment from title

@@
 
-\section{Configuration (Set-up, sbt, jar, frontends/../stainless-scalac)}
+\section{Configuration}
 There are two ways to integrate Stainless in a Scala project, the scala build tool (abt) plugin or a command line tool.

### 2019-05-22 doc/leader/versions.tex
added some todos and fixed version

@@
 Version	\> Date	\> Status			\> Remarks		\\
-0.1	\> 01.08.2013	\> Draft		\> Lorem ipsum dolor sit amet	\\	
-0.2	\> 21.08.2013	\> Draft		\> Phasellus scelerisque	\\ 
-0.3	\> 02.09.2013	\> Draft		\> Donec eget aliquam urna. Lorem ipsum dolor sit amet	\\ 
-1.0	\> 26.01.2014	\> Final		\> Lorem ipsum dolor sit ametPhasellus scelerisque, leo sed iaculis ornare 	\\ 
-1.1	\> 31.01.2014	\> Correction	\> Layout changed	\\
-1.2	\> 07.02.2014	\> Addition		\> Chapter 1.1 extended	\\
+1.0	\> TODO fix this	\> Final		\> Final Version 	\\
 \end{tabbing}

### 2019-05-22 code/currsniporig/src/main/scala/example/BasicArithmetic.scala
reverted original code

@@ -1,0 +1,48 @@
+package example
+
+import scala.util.Try
+
+/**
+  * @define mulSafe
+  * Some classes have restrictions on upper bounds
+  * for it's underlying value. This might cause the `*`
+  * operator to throw. This method wraps it in a `Try`
+  * block.
+  */
+trait BasicArithmetic[N] {
+
+  def +(n: N): N
+
+  /**
+    * Some classes have restrictions on upper bounds
+    * for it's underlying value. This might cause the `+`
+    * operator to throw. This method wraps it in a `Try`
+    * block.
+    */
+  def addSafe(n: N): Try[N] = Try { this + n }
+
+  def -(n: N): N
+
+  /**
+    * Some classes have restrictions on lower bounds
+    * for it's underlying value. This might cause the `-`
+    * operator to throw. This method wraps it in a `Try`
+    * block.
+    */
+  def subtractSafe(n: N): Try[N] = Try { this - n }
+
+  def *(factor: BigInt): N
+
+  /**
+    * $mulSafe
+    */
+  def multiplySafe(factor: BigInt): Try[N] = Try { this * factor }
+
+  def *(factor: N): N
+
+  /**
+    * $mulSafe
+    */
+  def multiplySafe(factor: N): Try[N] = Try { this * factor }
+}
+

### 2019-05-22 code/currsniporig/src/main/scala/example/Consensus.scala

@@ -1,0 +1,32 @@
+package example
+
+/**
+  * Created by chris on 5/13/16.
+  */
+sealed abstract class Consensus {
+
+  def maxBlockSize: Long = 1000000
+
+  def weightScalar: Long = 4
+
+  def maxBlockWeight: Long = maxBlockSize * weightScalar
+
+  /**
+    * BIP141 changes this from 20,000 -> 80,000, to see how sigops are counted please see BIP 141
+    * [[https://github.com/bitcoin/bips/blob/master/bip-0141.mediawiki#sigops]]
+    */
+  def maxSigOps = 80000
+
+  def maxMoney: CurrencyUnit = Satoshis(Int64(2100000000000000L))
+
+  /**
+    * A integer representing the maximum number of public keys you can have in a
+    * OP_CHECKMULTISIG or OP_CHECKMULTISIGVERIFY operation
+    * https://github.com/bitcoin/bitcoin/blob/master/src/script/interpreter.cpp#L903
+    * @return
+    */
+  def maxPublicKeysPerMultiSig = 20
+}
+
+object Consensus extends Consensus
+

### 2019-05-22 code/currsniporig/src/main/scala/example/CurrencyUnits.scala

@@ -1,0 +1,126 @@
+package example
+
+sealed abstract class CurrencyUnit
+    extends BasicArithmetic[CurrencyUnit] {
+  type A
+
+  def satoshis: Satoshis
+
+  def >=(c: CurrencyUnit): Boolean = {
+    satoshis.underlying >= c.satoshis.underlying
+  }
+
+  def >(c: CurrencyUnit): Boolean = {
+    satoshis.underlying > c.satoshis.underlying
+  }
+
+  def <(c: CurrencyUnit): Boolean = {
+    satoshis.underlying < c.satoshis.underlying
+  }
+
+  def <=(c: CurrencyUnit): Boolean = {
+    satoshis.underlying <= c.satoshis.underlying
+  }
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  override def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
+
+  override def -(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying - c.satoshis.underlying)
+  }
+
+  override def *(factor: BigInt): CurrencyUnit = {
+    Satoshis(satoshis.underlying * factor)
+  }
+
+  override def *(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying * c.satoshis.underlying)
+  }
+
+  def unary_- : CurrencyUnit = {
+    Satoshis(-satoshis.underlying)
+  }
+
+  def toBigDecimal: BigDecimal
+
+  protected def underlying: A
+}
+
+sealed abstract class Satoshis extends CurrencyUnit {
+  override type A = Int64
+
+  override def toString: String = s"$toLong sat"
+
+  override def satoshis: Satoshis = this
+
+  override def toBigDecimal = BigDecimal(toBigInt)
+
+  def toBigInt: BigInt = BigInt(toLong)
+
+  def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}
+
+object Satoshis extends BaseNumbers[Satoshis] {
+
+  val min = Satoshis(Int64.min)
+  val max = Satoshis(Int64.max)
+  val zero = Satoshis(Int64.zero)
[…]

### 2019-05-22 code/currsniporig/src/main/scala/example/NumberType.scala

@@ -1,0 +1,311 @@
+package example
+
+import scala.util.{Failure, Success, Try}
+
+/**
+  * Created by chris on 6/4/16.
+  */
+/**
+  * This abstract class is meant to represent a signed and unsigned number in C
+  * This is useful for dealing with codebases/protocols that rely on C's
+  * unsigned integer types
+  */
+sealed abstract class Number[T <: Number[T]]
+    extends BasicArithmetic[T] {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toInt: Int = toBigInt.bigInteger.intValueExact()
+  def toLong: Long = toBigInt.bigInteger.longValueExact()
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => T
+
+  override def +(num: T): T = apply(checkResult(underlying + num.underlying))
+  override def -(num: T): T = apply(checkResult(underlying - num.underlying))
+  override def *(factor: BigInt): T = apply(checkResult(underlying * factor))
+  override def *(num: T): T = apply(checkResult(underlying * num.underlying))
+
+  def >(num: T): Boolean = underlying > num.underlying
+  def >=(num: T): Boolean = underlying >= num.underlying
+  def <(num: T): Boolean = underlying < num.underlying
+  def <=(num: T): Boolean = underlying <= num.underlying
+
+  def <<(num: Int): T = this.<<(apply(num))
+  def >>(num: Int): T = this.>>(apply(num))
+
+  def <<(num: T): T = {
+    checkIfInt(num).map { _ =>
+      apply((underlying << num.toInt) & andMask)
+    }.get
+  }
+
+  def >>(num: T): T = {
+    //this check is for weird behavior with the jvm and shift rights
+    //https://stackoverflow.com/questions/47519140/bitwise-shift-right-with-long-not-equaling-zero/47519728#47519728
+    if (num.toLong > 63) apply(0)
+    else {
+      checkIfInt(num).map { _ =>
[…]

### 2019-05-22 code/currsniporig/src/main/scala/example/BasicArithmetic.scala
removed unused * and -

@@ -1,7 +1,5 @@
 package example
 
-import scala.util.Try
-
 /**
   * @define mulSafe
   * Some classes have restrictions on upper bounds
@@ -13,36 +11,5 @@ trait BasicArithmetic[N] {
 
   def +(n: N): N
 
-  /**
-    * Some classes have restrictions on upper bounds
-    * for it's underlying value. This might cause the `+`
-    * operator to throw. This method wraps it in a `Try`
-    * block.
-    */
-  def addSafe(n: N): Try[N] = Try { this + n }
-
-  def -(n: N): N
-
-  /**
-    * Some classes have restrictions on lower bounds
-    * for it's underlying value. This might cause the `-`
-    * operator to throw. This method wraps it in a `Try`
-    * block.
-    */
-  def subtractSafe(n: N): Try[N] = Try { this - n }
-
-  def *(factor: BigInt): N
-
-  /**
-    * $mulSafe
-    */
-  def multiplySafe(factor: BigInt): Try[N] = Try { this * factor }
-
-  def *(factor: N): N
-
-  /**
-    * $mulSafe
-    */
-  def multiplySafe(factor: N): Try[N] = Try { this * factor }
 }
 

### 2019-05-22 code/currsniporig/src/main/scala/example/CurrencyUnits.scala

@@ -30,18 +30,6 @@ sealed abstract class CurrencyUnit
     Satoshis(satoshis.underlying + c.satoshis.underlying)
   }
 
-  override def -(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying - c.satoshis.underlying)
-  }
-
-  override def *(factor: BigInt): CurrencyUnit = {
-    Satoshis(satoshis.underlying * factor)
-  }
-
-  override def *(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying * c.satoshis.underlying)
-  }
-
   def unary_- : CurrencyUnit = {
     Satoshis(-satoshis.underlying)
   }

### 2019-05-22 code/currsniporig/src/main/scala/example/BasicArithmetic.scala
removed unused number types

@@ -1,12 +1,5 @@
 package example
 
-/**
-  * @define mulSafe
-  * Some classes have restrictions on upper bounds
-  * for it's underlying value. This might cause the `*`
-  * operator to throw. This method wraps it in a `Try`
-  * block.
-  */
 trait BasicArithmetic[N] {
 
   def +(n: N): N

### 2019-05-22 code/currsniporig/src/main/scala/example/Consensus.scala

@@ -1,8 +1,5 @@
 package example
 
-/**
-  * Created by chris on 5/13/16.
-  */
 sealed abstract class Consensus {
 
   def maxBlockSize: Long = 1000000
@@ -11,20 +8,10 @@ sealed abstract class Consensus {
 
   def maxBlockWeight: Long = maxBlockSize * weightScalar
 
-  /**
-    * BIP141 changes this from 20,000 -> 80,000, to see how sigops are counted please see BIP 141
-    * [[https://github.com/bitcoin/bips/blob/master/bip-0141.mediawiki#sigops]]
-    */
   def maxSigOps = 80000
 
   def maxMoney: CurrencyUnit = Satoshis(Int64(2100000000000000L))
 
-  /**
-    * A integer representing the maximum number of public keys you can have in a
-    * OP_CHECKMULTISIG or OP_CHECKMULTISIGVERIFY operation
-    * https://github.com/bitcoin/bitcoin/blob/master/src/script/interpreter.cpp#L903
-    * @return
-    */
   def maxPublicKeysPerMultiSig = 20
 }
 

### 2019-05-22 code/currsniporig/src/main/scala/example/CurrencyUnits.scala

@@ -97,8 +97,6 @@ object Bitcoins extends BaseNumbers[Bitcoins] {
 }
 
 object CurrencyUnits {
-
-  /** The number you need to multiply BTC by to get it's satoshis */
   val btcToSatoshiScalar: Long = 100000000
   val satoshisToBTCScalar: BigDecimal = BigDecimal(1.0) / btcToSatoshiScalar
   val oneBTC: CurrencyUnit = Satoshis(Int64(btcToSatoshiScalar))

### 2019-05-22 code/currsniporig/src/main/scala/example/NumberType.scala

@@ -2,33 +2,18 @@ package example
 
 import scala.util.{Failure, Success, Try}
 
-/**
-  * Created by chris on 6/4/16.
-  */
-/**
-  * This abstract class is meant to represent a signed and unsigned number in C
-  * This is useful for dealing with codebases/protocols that rely on C's
-  * unsigned integer types
-  */
 sealed abstract class Number[T <: Number[T]]
     extends BasicArithmetic[T] {
   type A = BigInt
 
-  /** The underlying scala number used to to hold the number */
   protected def underlying: A
 
   def toInt: Int = toBigInt.bigInteger.intValueExact()
   def toLong: Long = toBigInt.bigInteger.longValueExact()
   def toBigInt: BigInt = underlying
 
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
   def andMask: BigInt
 
-  /** Factory function to create the underlying T, for instance a UInt32 */
   def apply: A => T
 
   override def +(num: T): T = apply(checkResult(underlying + num.underlying))
@@ -48,8 +33,6 @@ sealed abstract class Number[T <: Number[T]]
   }
 
   def >>(num: T): T = {
-    //this check is for weird behavior with the jvm and shift rights
-    //https://stackoverflow.com/questions/47519140/bitwise-shift-right-with-long-not-equaling-zero/47519728#47519728
     if (num.toLong > 63) apply(0)
     else {
       checkIfInt(num).map { _ =>
@@ -62,17 +45,12 @@ sealed abstract class Number[T <: Number[T]]
   def &(num: T): T = apply(checkResult(underlying & num.underlying))
   def unary_- : T = apply(-underlying)
 
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
   private def checkResult(result: BigInt): A = {
     require((result & andMask) == result,
             "Result was out of bounds, got: " + result)
     result
   }
 
-  /** Checks if the given nubmer is within range of a Int */
   private def checkIfInt(num: T): Try[Unit] = {
[…]

### 2019-05-22 code/currsniporig/src/main/scala/example/CurrencyUnits.scala
updated

@@ -30,10 +30,6 @@ sealed abstract class CurrencyUnit
     Satoshis(satoshis.underlying + c.satoshis.underlying)
   }
 
-  def unary_- : CurrencyUnit = {
-    Satoshis(-satoshis.underlying)
-  }
-
   def toBigDecimal: BigDecimal
 
   protected def underlying: A
@@ -78,22 +74,6 @@ sealed abstract class Bitcoins extends CurrencyUnit {
     val sat = underlying * CurrencyUnits.btcToSatoshiScalar
     Satoshis(Int64(sat.toLongExact))
   }
-}
-
-object Bitcoins extends BaseNumbers[Bitcoins] {
-  val min = Bitcoins((-Consensus.maxMoney).satoshis)
-  val max = Bitcoins(Consensus.maxMoney.satoshis)
-  val zero = Bitcoins(Satoshis.zero)
-  val one = Bitcoins(1)
-
-  def apply(satoshis: Satoshis): Bitcoins = {
-    val b: BigDecimal = satoshis.toLong * CurrencyUnits.satoshisToBTCScalar
-    Bitcoins(b)
-  }
-
-  def apply(underlying: BigDecimal): Bitcoins = BitcoinsImpl(underlying)
-
-  private case class BitcoinsImpl(underlying: BigDecimal) extends Bitcoins
 }
 
 object CurrencyUnits {

### 2019-05-22 code/currsniporig/src/main/scala/example/NumberType.scala

@@ -23,28 +23,6 @@ sealed abstract class Number[T <: Number[T]]
   def <(num: T): Boolean = underlying < num.underlying
   def <=(num: T): Boolean = underlying <= num.underlying
 
-  def <<(num: Int): T = this.<<(apply(num))
-  def >>(num: Int): T = this.>>(apply(num))
-
-  def <<(num: T): T = {
-    checkIfInt(num).map { _ =>
-      apply((underlying << num.toInt) & andMask)
-    }.get
-  }
-
-  def >>(num: T): T = {
-    if (num.toLong > 63) apply(0)
-    else {
-      checkIfInt(num).map { _ =>
-        apply(underlying >> num.toInt)
-      }.get
-    }
-  }
-
-  def |(num: T): T = apply(checkResult(underlying | num.underlying))
-  def &(num: T): T = apply(checkResult(underlying & num.underlying))
-  def unary_- : T = apply(-underlying)
-
   private def checkResult(result: BigInt): A = {
     require((result & andMask) == result,
             "Result was out of bounds, got: " + result)

### 2019-05-22 doc/chapters/towards_verifying.tex
changed subsections
added some space
changed title
more text

@@
 
-\section{Error reporting with sbt und jar}
+\section{Error Reporting with sbt and JAR}
 
-\section{Extraction of code fragment from Bitcoin-S}
+\section{Trying to Verify checkTransaction}
 
-\section{Rewriting of code}
+\subsection{Findings}
 
-\section{Verification of method "+"}
+\subsection{Bugfix}
+
+\section{Verification of Method "+ 0" // Change this}
+After so many failures, we have decided to search for the smallest unit in Bitcoin-S-Core that is worthwhile to verify.
+We found the addition of two Satoshis would be a good candidate.
+To make it even more easy, we decided to verify only the addition, where we add zero to an amount of Satoshis.
+The signature looks like the following.
+\begin{lstlisting}
+  +(c: CurrencyUnit): CurrencyUnit
+\end{lstlisting}
+Where the class Satoshis extends CurrencyUnit.
+The post condition would be, that the parameter c must be zero.
+\begin{lstlisting}
+  require(c.satoshis.underlying == Int64.zero)
+\end{lstlisting}
+\emph{c.satoshis} is an abstract method of CurrencyUnit that must return an instance of the class Satoshis.
+
+\emph{c.satoshis.underlying} is also an abstract method of CurrencyUnit that must return an instance of the abstract type A.
+
+Both are implemented in Satoshis where A is set to Int64.
+So the underlying number of the parameter must be zero.
+
+Next we ensure, that the result is the same value as \emph{this}.
+\begin{lstlisting}
+  ensuring(res => res.satoshis == this.satoshis)
+\end{lstlisting}
+Here, we can directly use equals (==) on Satoshis, because its a case class and Int64, the only parameter of Satoshis is a case class too.
+
+\subsection{Rewriting Abstract Type Member}
+Stainless output:
+\begin{lstlisting}
+[ Error  ] CurrencyUnits.scala:5:3: Stainless doesn't support abstract type members
+           type A
+\end{lstlisting}
+
+\subsection{Rewriting Generics}
+
+\subsection{Rewriting Objects}
+Stainless output:
+\begin{lstlisting}
+[ Error  ] CurrencyUnits.scala:54:1: Objects cannot extend classes or implement
+           traits, use a case object instead
+           object Satoshis extends BaseNumbers[Satoshis] {
+\end{lstlisting}
+
+\subsection{Rewriting BigInt Constructor (only literal argument, no long argument, etc.)}
+Stainless output:
+\begin{lstlisting}
+[ Error  ] CurrencyUnits.scala:47:33: Only literal arguments are allowed for BigInt.
+           def toBigInt: BigInt = BigInt(toLong)
+\end{lstlisting}
+
+\subsection{Rewriting Usage of BigInt \&-Function}
+
+\subsection{Rewriting require (funktioniert evtl?)}

### 2019-05-22 code/currsniporig/src/main/scala/example/Consensus.scala

@@ -1,19 +1,0 @@
-package example
-
-sealed abstract class Consensus {
-
-  def maxBlockSize: Long = 1000000
-
-  def weightScalar: Long = 4
-
-  def maxBlockWeight: Long = maxBlockSize * weightScalar
-
-  def maxSigOps = 80000
-
-  def maxMoney: CurrencyUnit = Satoshis(Int64(2100000000000000L))
-
-  def maxPublicKeysPerMultiSig = 20
-}
-
-object Consensus extends Consensus
-

### 2019-05-23 doc/chapters/towards_verifying.tex
more text

@@
 
+After writing the verification, we ran Stainless on the source files.
+It prints a lot of errors because of code incompatibility.
+So we had to rewrite it as described in the following sections.
+
 \subsection{Rewriting Abstract Type Member}
@@
 \end{lstlisting}
+This should be easy to rewrite by using generics instead of an abstract type, right?
+Unfortunately not.
+The problem is, CurrencyUnit uses its implementing class Satoshis.
+
+Simplified code.
+\begin{lstlisting}
+  sealed abstract class CurrencyUnit {
+    def +(c: CurrencyUnit): CurrencyUnit =
+      Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
+
+  sealed abstract class Satoshis extends CurrencyUnit
+\end{lstlisting}
+Thus, when we use generics, Satoshis is not of type CurrencyUnit anymore and neither of type CurrencyUnit[A], where A is the new generic type.
+Satoshis extends CurrencyUnit with type Int64, so it is of type CurrencyUnit[Int64].
+That's to specific.
+
+Since there is no easy way to fix it and the code should stay as much as possible the original, we decided to remove the abstract type.
+The verification is a bit more limited, to only Satoshis, but that's no problem, because the goal is to verify the addition of Satoshis.
+
+So by removing the abstract type and fix it to Int64 as defined by Satoshis, there was one error less in Stainless.
 
 \subsection{Rewriting Generics}
+Stainless output:
+\begin{lstlisting}
+  [ Error  ] NumberType.scala:3:30: Unknown type parameter type T
+             sealed abstract class Number[T <: Number[T]]
+
+\end{lstlisting}
 
 \subsection{Rewriting Objects}
+// TODO reason? because we can not extend case object?
 Stainless output:
@@
 \end{lstlisting}
+As described before, Stainless supports only a subset of Scala.
+They provide their own dedicated BigInt library.
+This library does not have support for dynamic BigInt construction only for string literals.
+
+Again, a simplified version.
+\begin{lstlisting}
+  sealed abstract class Satoshis extends CurrencyUnit {
+    def toBigInt: BigInt = BigInt(toLong)
+    def toLong: Long = underlying.toLong
+  }
+
+  object Int64 extends BaseNumbers[Int64] {
+    private case class Int64Impl(underlying: BigInt) extends Int64
+  }
+\end{lstlisting}
+This would be really hard to refactor, because Bitcoin-S-Core tries to be as much dynamic as possible so it can be used with other cryptocurrencies too.
+Maybe it would even be impossible, because they need to parse a lot from the bitcoin network.
+
+Restricting the code one more time helps.
+\emph{toBigInt} can directly return underlying, because in case of Int64, which is what Satoshis use, it is a BigInt.
+So it does not be converted from BigInt to Long and back to BigInt.
 

### 2019-05-23 doc/chapters/towards_verifying.tex
rewrite inner class

@@
 
+\subsection{Rewriting Private Inner Classes}
+Stainless output (simplified):
+\begin{lstlisting}
+  [Warning ] CurrencyUnits.scala:36:3: Could not extract tree in class:
+             case private class SatoshisImpl extends Satoshis
+\end{lstlisting}
+Stainless is not able to extract private classes inside other objects.
+Bitcoin-S-core uses it a lot, because they separate the class from its implementation.
+\begin{lstlisting}
+  object Int64 extends BaseNumbers[Int64] {
+    private case class Int64Impl(underlying: BigInt) extends Int64 
+  }
+\end{lstlisting}
+This is an easy one.
+Just extract the inner class out of the object.
+This is not exactly the same code but since the class is still private it can not be extended outside of the file.
+
 \subsection{Rewriting Usage of BigInt \&-Function}

### 2019-05-24 doc/chapters/towards_verifying.tex
rewriting rype member

@@
 \subsection{Rewriting Generics}
+// I dont understand this.. maybe just a missing feature in Stainless?
 Stainless output:
@@
 
+\end{lstlisting}
+
+\begin{lstlisting}
+  sealed abstract class Number[T <: Number[T]] extends BasicArithmetic[T]  
 \end{lstlisting}
@@
 
+\subsection{Rewriting Type Member}
+Stainless output:
+\begin{lstlisting}
+  [Warning ] NumberType.scala:5:3: Could not extract tree in class: type A
+             = BigInt (class scala.reflect.internal.Trees$TypeDef)
+             type A = BigInt
+\end{lstlisting}
+Since the type A is not overridden in any of its subclasses, this can simply be hard coded.
+\begin{lstlisting}
+  sealed abstract class Number {
+    type A = BigInt
+    protected def underlying: A
+    def apply: A => T
+  }
+\end{lstlisting}
+Replace every occurence of A with BigInt.
+This might be a missing feature in Stainless that should not be to hard to fix.
+There is an open issue on Github \mypound511.
+
 \subsection{Rewriting Usage of BigInt \&-Function}

### 2019-05-24 doc/chapters/towards_verifying.tex
added stainless output
rewriting & BigInt

@@
 \subsection{Rewriting Usage of BigInt \&-Function}
+Stainless output:
+\begin{lstlisting}
+  [ Error  ] NumberType.scala:19:14: Unknown call to & on result (BigInt) with
+             arguments List(Number.this.andMask) of type List(BigInt)
+             require((result & andMask) == result,
+\end{lstlisting}
+Due to the restrictions on BigInt, the \& function on it is not supported too.
+\begin{lstlisting}
+  sealed abstract class Number extends BasicArithmetic[Int64] {
+    def andMask: BigInt
+
+    override def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+    private def checkResult(result: BigInt): BigInt = {
+      require((result & andMask) == result, "Result was out of bounds, got: " + result)
+      result
+    }
+  }
+\end{lstlisting}
+This is a bound check.
+It checks, if the result of the addition is in range of the specified type (Int64 here).
+So here the \& mask can be replaced with a bound check whether the result is in range of Long.MinValue and Long.MaxValue.
+Again the code gets a bit more static but for this usecase its ok.
 

### 2019-05-24 doc/chapters/towards_verifying.tex
rewriting require

@@
 
-\subsection{Rewriting require (funktioniert evtl?)}
+\subsection{Rewriting require}
+Stainless output:
+\begin{lstlisting}
+  [Warning ] NumberType.scala:54:3: Could not extract tree in class:
+             scala.this.Predef.require(Int64Impl.this.underlying.<=(
+             math.this.BigInt.long2bigInt(9223372036854775807L)),
+             "Number was too big for a int64, got: ".+(Int64Impl.this.underlying))
+             (class scala.reflect.internal.Trees$Apply)
+             require(underlying <= 9223372036854775807L,
+\end{lstlisting}
+This is, because Stainless does not support require with a second String parameter.
+Removing this String fixes the error.

### 2019-05-24 doc/chapters/towards_verifying.tex
propagate require

@@
 Removing this String fixes the error.
+
+\subsection{Propagate require}
+Finally the code works with Stainless.
+But there is another problem.
+Bitcoin-S-Core uses require like a fail-fast method where as Stainless needs it to verify the code.
+
+Stainless output:
+\begin{lstlisting}
+  [Warning ] Found counter-example:
+  [Warning ]   thiss: { x: Object | @unchecked isNumber(x) } -> Int64Impl(0)
+  [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> 
+                 Int64Impl(9223372036854775808)
+\end{lstlisting}
+Corresponding code:
+\begin{lstlisting}
+  sealed abstract class Number extends BasicArithmetic[Int64] {
+    override def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+    private def checkResult(result: BigInt): BigInt = {
+      require(
+        result <= BigInt("9223372036854775807")
+        && result >= BigInt("-9223372036854775808")
+      )
+      result
+    }
+  }
+\end{lstlisting}
+But how can Stainless find a counter example ignoring the require in checkResult?
+Since Stainless is a static verification tool, it tests every possibility.
+So it can use a number bigger than the maximum Int64 and pass it to the addition.
+The require in checkResult will then fail.
+Thus, the addition need to have the restrictions of checkResult too.
+\begin{lstlisting}
+  override def +(num: Int64): Int64 = {
+    require(
+      num.underlying <= BigInt("9223372036854775807")
+      && num.underlying >= BigInt("-9223372036854775808")
+      && this.underlying <= BigInt("9223372036854775807")
+      && this.underlying >= BigInt("-9223372036854775808")
+    )
+    apply(checkResult(underlying + num.underlying))
+  }
+\end{lstlisting}
+Stainless finds another counter example:
+\begin{lstlisting}
+  [Warning ] Found counter-example:
+  [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(1)
+  [Warning ]   thiss: { x: Object | @unchecked isNumber(x) } ->
+                 Int64Impl(9223372036854775807)
+\end{lstlisting}
+Sure, it can just add one to the maximum Int64 and the require does not hold anymore.
+The goal is to check the addition of zero so lets add a restriction for num to be zero.
+
+Finally, it works and verifies correctly.

### 2019-05-24 doc/chapters/towards_verifying.tex
result image
issue -> pull request
rewriting objects

@@
 \subsection{Rewriting Objects}
-// TODO reason? because we can not extend case object?
+// TODO reason?
+
 Stainless output:
@@
 \end{lstlisting}
+Changing the objects that extends classes to case objects solves this problem.
 
@@
 This might be a missing feature in Stainless that should not be to hard to fix.
-There is an open issue on Github \mypound511.
+There is an open pull request \mypound470 on Github for this issue.
 
@@
 Finally, it works and verifies correctly.
+
+\section{Result}
+\begin{figure}[H]
+	\centering
+		\includegraphics[scale=0.45]{images/final_verify_output.png}
+	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
+	\label{fig:output1}
+\end{figure}

### 2019-05-25 doc/chapters/towards_verifying.tex
fixed grammer

@@
 \end{lstlisting}
-Thus, when we use generics, Satoshis is not of type CurrencyUnit anymore and neither of type CurrencyUnit[A], where A is the new generic type.
+Thus, changeing it to generics, Satoshis is not of type CurrencyUnit anymore and neither of type CurrencyUnit[A], where A is the new generic type.
 Satoshis extends CurrencyUnit with type Int64, so it is of type CurrencyUnit[Int64].
@@
 
-Since there is no easy way to fix it and the code should stay as much as possible the original, we decided to remove the abstract type.
+Since there is no easy way to fix it and the code should stay as much as possible the original, removing the abstract type might be the only choice.
 The verification is a bit more limited, to only Satoshis, but that's no problem, because the goal is to verify the addition of Satoshis.

### 2019-05-25 doc/chapters/towards_verifying.tex
fixed grammer

@@
 
-Next we ensure, that the result is the same value as \emph{this}.
+To ensure that the result is the same value as \emph{this}:
 \begin{lstlisting}
@@
 \end{lstlisting}
-Here, we can directly use equals (==) on Satoshis, because its a case class and Int64, the only parameter of Satoshis is a case class too.
+Here, one can directly use equals (==) on Satoshis, because its a case class and Int64, the only parameter of Satoshis is a case class too.
 
-After writing the verification, we ran Stainless on the source files.
+After writing the verification, running Stainless on the source files was the next step.
 It prints a lot of errors because of code incompatibility.
-So we had to rewrite it as described in the following sections.
+So lets describe the code rewriting in the following sections.
 

### 2019-05-25 doc/chapters/towards_verifying.tex
splitted chapter

--- 2019-05-24 doc/chapters/towards_verifying.tex
@@
-\chapter{Towards verifyng Bitcoin-S with Stainless}
+\chapter{Trying to Verify checkTransaction}
 \label{chap:connecting}
@@
 
-\section{Trying to Verify checkTransaction}
+\section{Findings}
 
-\subsection{Findings}
-
-\subsection{Bugfix}
-
-\section{Verification of Method "+ 0" // Change this}
-After so many failures, we have decided to search for the smallest unit in Bitcoin-S-Core that is worthwhile to verify.
-We found the addition of two Satoshis would be a good candidate.
-To make it even more easy, we decided to verify only the addition, where we add zero to an amount of Satoshis.
-The signature looks like the following.
-\begin{lstlisting}
-  +(c: CurrencyUnit): CurrencyUnit
-\end{lstlisting}
-Where the class Satoshis extends CurrencyUnit.
-The post condition would be, that the parameter c must be zero.
-\begin{lstlisting}
-  require(c.satoshis.underlying == Int64.zero)
-\end{lstlisting}
-\emph{c.satoshis} is an abstract method of CurrencyUnit that must return an instance of the class Satoshis.
-
-\emph{c.satoshis.underlying} is also an abstract method of CurrencyUnit that must return an instance of the abstract type A.
-
-Both are implemented in Satoshis where A is set to Int64.
-So the underlying number of the parameter must be zero.
-
-Next we ensure, that the result is the same value as \emph{this}.
-\begin{lstlisting}
-  ensuring(res => res.satoshis == this.satoshis)
-\end{lstlisting}
-Here, we can directly use equals (==) on Satoshis, because its a case class and Int64, the only parameter of Satoshis is a case class too.
-
-After writing the verification, we ran Stainless on the source files.
-It prints a lot of errors because of code incompatibility.
-So we had to rewrite it as described in the following sections.
-
-\subsection{Rewriting Abstract Type Member}
-Stainless output:
-\begin{lstlisting}
-[ Error  ] CurrencyUnits.scala:5:3: Stainless doesn't support abstract type members
-           type A
-\end{lstlisting}
-This should be easy to rewrite by using generics instead of an abstract type, right?
-Unfortunately not.
-The problem is, CurrencyUnit uses its implementing class Satoshis.
-
-Simplified code.
-\begin{lstlisting}
-  sealed abstract class CurrencyUnit {
-    def +(c: CurrencyUnit): CurrencyUnit =
-      Satoshis(satoshis.underlying + c.satoshis.underlying)
-  }
-
-  sealed abstract class Satoshis extends CurrencyUnit
-\end{lstlisting}
-Thus, when we use generics, Satoshis is not of type CurrencyUnit anymore and neither of type CurrencyUnit[A], where A is the new generic type.
-Satoshis extends CurrencyUnit with type Int64, so it is of type CurrencyUnit[Int64].
-That's to specific.
-
-Since there is no easy way to fix it and the code should stay as much as possible the original, we decided to remove the abstract type.
-The verification is a bit more limited, to only Satoshis, but that's no problem, because the goal is to verify the addition of Satoshis.
-
-So by removing the abstract type and fix it to Int64 as defined by Satoshis, there was one error less in Stainless.
-
-\subsection{Rewriting Generics}
-// I dont understand this.. maybe just a missing feature in Stainless?
-Stainless output:
-\begin{lstlisting}
-  [ Error  ] NumberType.scala:3:30: Unknown type parameter type T
-             sealed abstract class Number[T <: Number[T]]
-
-\end{lstlisting}
-
-\begin{lstlisting}
-  sealed abstract class Number[T <: Number[T]] extends BasicArithmetic[T]  
-\end{lstlisting}
-
-\subsection{Rewriting Objects}
-// TODO reason?
-
-Stainless output:
-\begin{lstlisting}
-[ Error  ] CurrencyUnits.scala:54:1: Objects cannot extend classes or implement
-           traits, use a case object instead
-           object Satoshis extends BaseNumbers[Satoshis] {
-\end{lstlisting}
-Changing the objects that extends classes to case objects solves this problem.
-
-\subsection{Rewriting BigInt Constructor (only literal argument, no long argument, etc.)}
-Stainless output:
-\begin{lstlisting}
-[ Error  ] CurrencyUnits.scala:47:33: Only literal arguments are allowed for BigInt.
-           def toBigInt: BigInt = BigInt(toLong)
-\end{lstlisting}
-As described before, Stainless supports only a subset of Scala.
-They provide their own dedicated BigInt library.
-This library does not have support for dynamic BigInt construction only for string literals.
-
-Again, a simplified version.
-\begin{lstlisting}
-  sealed abstract class Satoshis extends CurrencyUnit {
-    def toBigInt: BigInt = BigInt(toLong)
-    def toLong: Long = underlying.toLong
-  }
-
-  object Int64 extends BaseNumbers[Int64] {
-    private case class Int64Impl(underlying: BigInt) extends Int64
-  }
-\end{lstlisting}
-This would be really hard to refactor, because Bitcoin-S-Core tries to be as much dynamic as possible so it can be used with other cryptocurrencies too.
-Maybe it would even be impossible, because they need to parse a lot from the bitcoin network.
-
-Restricting the code one more time helps.
-\emph{toBigInt} can directly return underlying, because in case of Int64, which is what Satoshis use, it is a BigInt.
-So it does not be converted from BigInt to Long and back to BigInt.
-
-\subsection{Rewriting Private Inner Classes}
-Stainless output (simplified):
-\begin{lstlisting}
-  [Warning ] CurrencyUnits.scala:36:3: Could not extract tree in class:
-             case private class SatoshisImpl extends Satoshis
-\end{lstlisting}
-Stainless is not able to extract private classes inside other objects.
-Bitcoin-S-core uses it a lot, because they separate the class from its implementation.
-\begin{lstlisting}
-  object Int64 extends BaseNumbers[Int64] {
-    private case class Int64Impl(underlying: BigInt) extends Int64 
-  }
-\end{lstlisting}
-This is an easy one.
-Just extract the inner class out of the object.
-This is not exactly the same code but since the class is still private it can not be extended outside of the file.
-
-\subsection{Rewriting Type Member}
-Stainless output:
-\begin{lstlisting}
-  [Warning ] NumberType.scala:5:3: Could not extract tree in class: type A
-             = BigInt (class scala.reflect.internal.Trees$TypeDef)
-             type A = BigInt
-\end{lstlisting}
-Since the type A is not overridden in any of its subclasses, this can simply be hard coded.
-\begin{lstlisting}
-  sealed abstract class Number {
-    type A = BigInt
-    protected def underlying: A
-    def apply: A => T
-  }
-\end{lstlisting}
-Replace every occurence of A with BigInt.
-This might be a missing feature in Stainless that should not be to hard to fix.
-There is an open pull request \mypound470 on Github for this issue.
-
-\subsection{Rewriting Usage of BigInt \&-Function}
-Stainless output:
-\begin{lstlisting}
-  [ Error  ] NumberType.scala:19:14: Unknown call to & on result (BigInt) with
-             arguments List(Number.this.andMask) of type List(BigInt)
-             require((result & andMask) == result,
-\end{lstlisting}
-Due to the restrictions on BigInt, the \& function on it is not supported too.
-\begin{lstlisting}
-  sealed abstract class Number extends BasicArithmetic[Int64] {
-    def andMask: BigInt
-
-    override def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-    private def checkResult(result: BigInt): BigInt = {
-      require((result & andMask) == result, "Result was out of bounds, got: " + result)
-      result
-    }
-  }
-\end{lstlisting}
-This is a bound check.
-It checks, if the result of the addition is in range of the specified type (Int64 here).
-So here the \& mask can be replaced with a bound check whether the result is in range of Long.MinValue and Long.MaxValue.
-Again the code gets a bit more static but for this usecase its ok.
-
-\subsection{Rewriting require}
-Stainless output:
-\begin{lstlisting}
-  [Warning ] NumberType.scala:54:3: Could not extract tree in class:
-             scala.this.Predef.require(Int64Impl.this.underlying.<=(
-             math.this.BigInt.long2bigInt(9223372036854775807L)),
-             "Number was too big for a int64, got: ".+(Int64Impl.this.underlying))
-             (class scala.reflect.internal.Trees$Apply)
-             require(underlying <= 9223372036854775807L,
-\end{lstlisting}
-This is, because Stainless does not support require with a second String parameter.
-Removing this String fixes the error.
-
-\subsection{Propagate require}
-Finally the code works with Stainless.
-But there is another problem.
-Bitcoin-S-Core uses require like a fail-fast method where as Stainless needs it to verify the code.
-
-Stainless output:
-\begin{lstlisting}
-  [Warning ] Found counter-example:
-  [Warning ]   thiss: { x: Object | @unchecked isNumber(x) } -> Int64Impl(0)
-  [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> 
-                 Int64Impl(9223372036854775808)
-\end{lstlisting}
-Corresponding code:
-\begin{lstlisting}
-  sealed abstract class Number extends BasicArithmetic[Int64] {
-    override def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-    private def checkResult(result: BigInt): BigInt = {
-      require(
-        result <= BigInt("9223372036854775807")
-        && result >= BigInt("-9223372036854775808")
-      )
-      result
-    }
-  }
-\end{lstlisting}
-But how can Stainless find a counter example ignoring the require in checkResult?
-Since Stainless is a static verification tool, it tests every possibility.
-So it can use a number bigger than the maximum Int64 and pass it to the addition.
-The require in checkResult will then fail.
-Thus, the addition need to have the restrictions of checkResult too.
-\begin{lstlisting}
-  override def +(num: Int64): Int64 = {
-    require(
-      num.underlying <= BigInt("9223372036854775807")
-      && num.underlying >= BigInt("-9223372036854775808")
-      && this.underlying <= BigInt("9223372036854775807")
-      && this.underlying >= BigInt("-9223372036854775808")
-    )
-    apply(checkResult(underlying + num.underlying))
-  }
-\end{lstlisting}
-Stainless finds another counter example:
-\begin{lstlisting}
-  [Warning ] Found counter-example:
-  [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(1)
-  [Warning ]   thiss: { x: Object | @unchecked isNumber(x) } ->
-                 Int64Impl(9223372036854775807)
-\end{lstlisting}
-Sure, it can just add one to the maximum Int64 and the require does not hold anymore.
-The goal is to check the addition of zero so lets add a restriction for num to be zero.
-
-Finally, it works and verifies correctly.
-
-\section{Result}
-\begin{figure}[H]
-	\centering
-		\includegraphics[scale=0.45]{images/final_verify_output.png}
-	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
-	\label{fig:output1}
-\end{figure}
+\section{Bugfix}

### 2019-05-25 doc/chapters/towards_verifying_addition.tex
splitted chapter

\chapter{Towards Verifying Addition with Zero}
After so many failures, we have decided to search for the smallest unit in Bitcoin-S-Core that is worthwhile to verify.
We found the addition of two Satoshis would be a good candidate.
To make it even more easy, we decided to verify only the addition, where we add zero to an amount of Satoshis.
The signature looks like the following.
\begin{lstlisting}
  +(c: CurrencyUnit): CurrencyUnit
\end{lstlisting}
Where the class Satoshis extends CurrencyUnit.
The post condition would be, that the parameter c must be zero.
\begin{lstlisting}
  require(c.satoshis.underlying == Int64.zero)
\end{lstlisting}
\emph{c.satoshis} is an abstract method of CurrencyUnit that must return an instance of the class Satoshis.

\emph{c.satoshis.underlying} is also an abstract method of CurrencyUnit that must return an instance of the abstract type A.

Both are implemented in Satoshis where A is set to Int64.
So the underlying number of the parameter must be zero.

To ensure that the result is the same value as \emph{this}:
\begin{lstlisting}
  ensuring(res => res.satoshis == this.satoshis)
\end{lstlisting}
Here, one can directly use equals (==) on Satoshis, because its a case class and Int64, the only parameter of Satoshis is a case class too.

After writing the verification, running Stainless on the source files was the next step.
It prints a lot of errors because of code incompatibility.
So lets describe the code rewriting in the following sections.

\section{Rewriting Abstract Type Member}
Stainless output:
\begin{lstlisting}
[ Error  ] CurrencyUnits.scala:5:3: Stainless doesn't support abstract type members
           type A
\end{lstlisting}
This should be easy to rewrite by using generics instead of an abstract type, right?
Unfortunately not.
The problem is, CurrencyUnit uses its implementing class Satoshis.

Simplified code.
\begin{lstlisting}
  sealed abstract class CurrencyUnit {
    def +(c: CurrencyUnit): CurrencyUnit =
      Satoshis(satoshis.underlying + c.satoshis.underlying)
  }

  sealed abstract class Satoshis extends CurrencyUnit
\end{lstlisting}
Thus, changeing it to generics, Satoshis is not of type CurrencyUnit anymore and neither of type CurrencyUnit[A], where A is the new generic type.
Satoshis extends CurrencyUnit with type Int64, so it is of type CurrencyUnit[Int64].
That's to specific.

Since there is no easy way to fix it and the code should stay as much as possible the original, removing the abstract type might be the only choice.
The verification is a bit more limited, to only Satoshis, but that's no problem, because the goal is to verify the addition of Satoshis.

So by removing the abstract type and fix it to Int64 as defined by Satoshis, there was one error less in Stainless.

\section{Rewriting Generics}
// I dont understand this.. maybe just a missing feature in Stainless?
Stainless output:
\begin{lstlisting}
  [ Error  ] NumberType.scala:3:30: Unknown type parameter type T
             sealed abstract class Number[T <: Number[T]]

\end{lstlisting}

\begin{lstlisting}
  sealed abstract class Number[T <: Number[T]] extends BasicArithmetic[T]  
\end{lstlisting}

\section{Rewriting Objects}
// TODO reason?

Stainless output:
\begin{lstlisting}
[ Error  ] CurrencyUnits.scala:54:1: Objects cannot extend classes or implement
           traits, use a case object instead
           object Satoshis extends BaseNumbers[Satoshis] {
\end{lstlisting}
Changing the objects that extends classes to case objects solves this problem.

\section{Rewriting BigInt Constructor (only literal argument, no long argument, etc.)}
Stainless output:
\begin{lstlisting}
[ Error  ] CurrencyUnits.scala:47:33: Only literal arguments are allowed for BigInt.
           def toBigInt: BigInt = BigInt(toLong)
\end{lstlisting}
As described before, Stainless supports only a subset of Scala.
They provide their own dedicated BigInt library.
This library does not have support for dynamic BigInt construction only for string literals.

Again, a simplified version.
\begin{lstlisting}
  sealed abstract class Satoshis extends CurrencyUnit {
    def toBigInt: BigInt = BigInt(toLong)
    def toLong: Long = underlying.toLong
  }

  object Int64 extends BaseNumbers[Int64] {
    private case class Int64Impl(underlying: BigInt) extends Int64
  }
\end{lstlisting}
This would be really hard to refactor, because Bitcoin-S-Core tries to be as much dynamic as possible so it can be used with other cryptocurrencies too.
Maybe it would even be impossible, because they need to parse a lot from the bitcoin network.

Restricting the code one more time helps.
\emph{toBigInt} can directly return underlying, because in case of Int64, which is what Satoshis use, it is a BigInt.
So it does not be converted from BigInt to Long and back to BigInt.

\section{Rewriting Private Inner Classes}
Stainless output (simplified):
\begin{lstlisting}
  [Warning ] CurrencyUnits.scala:36:3: Could not extract tree in class:
             case private class SatoshisImpl extends Satoshis
\end{lstlisting}
Stainless is not able to extract private classes inside other objects.
Bitcoin-S-core uses it a lot, because they separate the class from its implementation.
\begin{lstlisting}
  object Int64 extends BaseNumbers[Int64] {
    private case class Int64Impl(underlying: BigInt) extends Int64 
  }
\end{lstlisting}
This is an easy one.
Just extract the inner class out of the object.
This is not exactly the same code but since the class is still private it can not be extended outside of the file.

\section{Rewriting Type Member}
Stainless output:
\begin{lstlisting}
  [Warning ] NumberType.scala:5:3: Could not extract tree in class: type A
             = BigInt (class scala.reflect.internal.Trees$TypeDef)
             type A = BigInt
\end{lstlisting}
Since the type A is not overridden in any of its subclasses, this can simply be hard coded.
\begin{lstlisting}
  sealed abstract class Number {
    type A = BigInt
    protected def underlying: A
    def apply: A => T
  }
\end{lstlisting}
Replace every occurence of A with BigInt.
This might be a missing feature in Stainless that should not be to hard to fix.
There is an open pull request \mypound470 on Github for this issue.

\section{Rewriting Usage of BigInt \&-Function}
Stainless output:
\begin{lstlisting}
  [ Error  ] NumberType.scala:19:14: Unknown call to & on result (BigInt) with
             arguments List(Number.this.andMask) of type List(BigInt)
             require((result & andMask) == result,
\end{lstlisting}
Due to the restrictions on BigInt, the \& function on it is not supported too.
\begin{lstlisting}
  sealed abstract class Number extends BasicArithmetic[Int64] {
    def andMask: BigInt

    override def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))

    private def checkResult(result: BigInt): BigInt = {
      require((result & andMask) == result, "Result was out of bounds, got: " + result)
      result
    }
  }
\end{lstlisting}
This is a bound check.
It checks, if the result of the addition is in range of the specified type (Int64 here).
So here the \& mask can be replaced with a bound check whether the result is in range of Long.MinValue and Long.MaxValue.
Again the code gets a bit more static but for this usecase its ok.

\section{Rewriting require}
Stainless output:
\begin{lstlisting}
  [Warning ] NumberType.scala:54:3: Could not extract tree in class:
             scala.this.Predef.require(Int64Impl.this.underlying.<=(
             math.this.BigInt.long2bigInt(9223372036854775807L)),
             "Number was too big for a int64, got: ".+(Int64Impl.this.underlying))
             (class scala.reflect.internal.Trees$Apply)
             require(underlying <= 9223372036854775807L,
\end{lstlisting}
This is, because Stainless does not support require with a second String parameter.
Removing this String fixes the error.

\section{Propagate require}
Finally the code works with Stainless.
But there is another problem.
Bitcoin-S-Core uses require like a fail-fast method where as Stainless needs it to verify the code.

Stainless output:
\begin{lstlisting}
  [Warning ] Found counter-example:
  [Warning ]   thiss: { x: Object | @unchecked isNumber(x) } -> Int64Impl(0)
  [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> 
                 Int64Impl(9223372036854775808)
\end{lstlisting}
Corresponding code:
\begin{lstlisting}
  sealed abstract class Number extends BasicArithmetic[Int64] {
    override def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))

    private def checkResult(result: BigInt): BigInt = {
      require(
        result <= BigInt("9223372036854775807")
        && result >= BigInt("-9223372036854775808")
      )
      result
    }
  }
\end{lstlisting}
But how can Stainless find a counter example ignoring the require in checkResult?
Since Stainless is a static verification tool, it tests every possibility.
So it can use a number bigger than the maximum Int64 and pass it to the addition.
The require in checkResult will then fail.
Thus, the addition need to have the restrictions of checkResult too.
\begin{lstlisting}
  override def +(num: Int64): Int64 = {
    require(
      num.underlying <= BigInt("9223372036854775807")
      && num.underlying >= BigInt("-9223372036854775808")
      && this.underlying <= BigInt("9223372036854775807")
      && this.underlying >= BigInt("-9223372036854775808")
    )
    apply(checkResult(underlying + num.underlying))
  }
\end{lstlisting}
Stainless finds another counter example:
\begin{lstlisting}
  [Warning ] Found counter-example:
  [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(1)
  [Warning ]   thiss: { x: Object | @unchecked isNumber(x) } ->
                 Int64Impl(9223372036854775807)
\end{lstlisting}
Sure, it can just add one to the maximum Int64 and the require does not hold anymore.
The goal is to check the addition of zero so lets add a restriction for num to be zero.

Finally, it works and verifies correctly.

\section{Result}
\begin{figure}[H]
	\centering
		\includegraphics[scale=0.45]{images/final_verify_output.png}
	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
	\label{fig:output1}
\end{figure}

### 2019-05-25 doc/chapters/using_stainless.tex
splitted chapter

@@
 \section{Scala compatibility (Pure Scala, imperative features, dedicated BigInt, Generics...)}
+Stainless supports only a subset of Scala.
+They call it Pure Scala.
+We will see some of this restrictions we have detected during this work in the next chapter.
+There is a list of what's supported in the Stainless documentation.

### 2019-05-25 doc/chapters/using_bitcoins.tex
fixed use of old code

@@
 \begin{lstlisting}
-  val prevOutputs = transaction.inputs.map(_.previousOutput)
-  val noDuplicateInputs = prevOutputs.distinct.size == prevOutputs.size
+  val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
+  val noDuplicateInputs = prevOutputTxIds.distinct.size == prevOutputTxIds.size
 \end{lstlisting}
-It gathers all inputs output.
-On this previous outputs it calls distinct and checks if the size stays the same.
-Distinct removes duplicates, so if there was two times the same input the size of inputs before calling distinct would be greater.
+It gathers all inputs output transaction ids.
+On this previous outputs ids it calls distinct and checks if the size stays the same.
+Distinct removes duplicates, so if there was two times the same input transaction id the size of those ids would be greater before calling distinct.
+
+There is a bug in those two lines, as described in the next chapter.

### 2019-05-25 doc/chapters/trying_verify_checktx.tex
text

\chapter{Trying to Verify checkTransaction}
\label{chap:connecting}

\section{Integration (Versionskonflikte, neues Plugin)}
During this work, Stainless bumped the sbt plugin from version 0.13.17 to 1.2.8 and Scala from 2.11.12 to 2.12.8.
So this section might be out of date now.

Bitcoin-S-Core uses sbt 1.2.8 and Scala 2.12.8.

//TODO

\section{Error Reporting with sbt and JAR}

\section{Findings}

\section{Bugfix}

### 2019-05-25 doc/chapters/introduction.tex
End of Stainless in introduction.tex

@@
 It verifies statically that a program satisfies a specification given by a developer and that a program will not crash at runtime.
-Stainless explores all possible input values, reports inputs for which a program fails and demonstrates counterexamples which violate a given specification.
+Stainless explores all possible input values, reports inputs for which a program fails and demonstrates counterexamples which violate a given specification.\cite{Stainless:introduction}
 
@@
 A postcondition should be written after a function body with  \textit{ensuring}. It is a verification condition on an output of a function.
-While compiling Stainless tries to prove that the postcondition always holds, assuming a given precondition does hold.
+While compiling Stainless tries to prove that the postcondition always holds, assuming a given precondition does hold.\cite{Stainless:introduction}
 
@@
 Also an output unknown is possible when Stainless is unable to prove a postcondition or find a counterexample. In this case a timeout or an internal error occured.
-Furthermore, it will be verified by Stainless that a precondition cannot be violated.
+Furthermore, it will be verified by Stainless that a precondition cannot be violated.\cite{Stainless:introduction}
 
@@
 
-The code to be verified should be written in Pure Scala belonging to functional programming paradigm.
-To extend this subset of the Scala language Stainless supports also some imperative features translating them into Pure Scala concepts.
-For more details about Stainless its official website can be explored. \cite{Stainless}
+Stainless verifies scala programs using SMT (satisfiability modulo theories) solvers. 
+A program code will be transformed by Stainless to the Pure Scala fragment supported by Inox solver. 
+This solver was also developed by LARA and allows usage of SMT solvers such as Z3, CVC4 and Princess, adding support of other higher-level features.
+Preconditions, postconditions and assertions will be translated into SMT formulas, so a SMT solver can determine if all properties hold.
 

### 2019-05-25 doc/chapters/introduction.tex
Added partially the description of Bitcoin-S to introduction.tex

@@
 In following the main aspects of formal verification, Stainless and Bitcoin-S are described.
-
-% Einträge im Verzeichnis erscheinen lassen ohne hier eine Referenz einzufügen
-\nocite{kopka:band1}
-\nocite{raichle:bibtex_programmierung}
-\nocite{MiKTeX}
-\nocite{KOMA}
-\nocite{TeXnicCenter}
-\nocite{Marti06}
-\nocite{Erbsland08}
-\nocite{juergens:einfuehrung}
-\nocite{juergens:fortgeschritten}
 
@@
 One of such tools is Stainless. This framework is used in the work to verify a Scala code.
-
-%In figure \ref{fig:file_structure} the file structure is shown for this template.
-
-%\begin{figure}[H]
-%	\centering
-%		\includegraphics[scale=0.85]{images/file_structure.png}
-%	\caption{File structure}
-%	\label{fig:file_structure}
-%\end{figure}
 
@@
 
-Stainless verifies scala programs using SMT (satisfiability modulo theories) solvers. 
-A program code will be transformed by Stainless to the Pure Scala fragment supported by Inox solver. 
+Stainless verifies scala programs using SMT (satisfiability modulo theories) solvers. A program code will be transformed by Stainless to the Pure Scala fragment supported by Inox solver. 
 This solver was also developed by LARA and allows usage of SMT solvers such as Z3, CVC4 and Princess, adding support of other higher-level features.
@@
 
-%\begin{table}[H]
-%	\centering
-%		\begin{tabular}{lll} \toprule
-%			\textbf{First Name Last Name} & \textbf{E-mail} & \textbf{Function} \\ \midrule
-%			Alfred Kaufmann & alfred.kaufmann@bfh.ch & Employer, Project Management, \\
-%			& & Supplements, Improvements \\ \midrule
-%			Fritz Dellsperger & Retired & Tips on the structure and layout \\ \midrule
-%			David Burri & Contracted out & First compilation of the Template \\ \bottomrule
-%		\end{tabular}
-%	\caption{Contact Persons}
-%	\label{tab:Contact Persons}
-%\end{table}
-
 \section{Bitcoin-S (Projekten und Packages, bitcoin-s-core, Eigenschaften zu prüfen)}
 \label{sec:bitcoin_s}
+
+In the work some fragments of the code of Bitcoin-S will be verified. Bitcoin-S is an open source Scala implementation of the Bitcoin protocol. 
+There are different projects belonging to Bitcoin-S.
+Protocol data structures are defined in the package \textbf{bitcoin-s-core}.
+This package is the main target to be verified because it specifies the base of Bitcoin protocol, and it is crucial for correct protocol functionality.
+There is also the package \textbf{bitcoind-rpc} which is a RPC client implementation to communicate with \textit{bitcoind}, daemon of Bitcoin Core running on a machine.
+Developers of Bitcoin-S are working on the implementation of a lightweight bitcoin client in the package \textbf{bitcoin-s-spv-node}.
 

### 2019-05-26 doc/chapters/introduction.tex
End of Bincoin-S in introduction.tex

@@
 
+The code being verified with Stainless should be written in Pure Scala, that corresponds to the functional programming. But also some imperative features are supported by Stainless.
+
 Stainless verifies scala programs using SMT (satisfiability modulo theories) solvers. A program code will be transformed by Stainless to the Pure Scala fragment supported by Inox solver. 
@@
 
-In the work some fragments of the code of Bitcoin-S will be verified. Bitcoin-S is an open source Scala implementation of the Bitcoin protocol. 
+In the work some fragments of the code of Bitcoin-S has to be verified. Bitcoin-S is an open source Scala implementation of the Bitcoin protocol. 
 There are different projects belonging to Bitcoin-S.
@@
 Developers of Bitcoin-S are working on the implementation of a lightweight bitcoin client in the package \textbf{bitcoin-s-spv-node}.
+The whole list of packages can be found on the github \cite{BitcoinSProject}.
 
-%\begin{itemize}
-%	\item Create a BFH Style Files
-%	\item Template for the Compilation of presentations with \LaTeX{}
-%\end{itemize}
+Bitcoin-S is a large project implementing many functionalities.
+In this work one of them is going to be examinated and worked on, namely the property of the Bitcoin that a non-coinbase transaction can not generate new coins.
+Thus, the code of Bitcoin-S should be analyzied, and fragments implementing this feature should be identified.
+Afterwards, the fragments should be rewritten, so that Stainless can accept it.
+After that a formal specification for functions should be defined and annotated according Stainless libraries, so that Stainless can verify corrrectness of the code.

### 2019-05-28 doc/chapters/towards_verifying_addition.tex
typo

@@
 We found the addition of two Satoshis would be a good candidate.
-To make it even more easy, we decided to verify only the addition, where we add zero to an amount of Satoshis.
+To make it even easier, we decided to verify only the addition, where we add zero to an amount of Satoshis.
 The signature looks like the following.
@@
 \end{lstlisting}
-Here, one can directly use equals (==) on Satoshis, because its a case class and Int64, the only parameter of Satoshis is a case class too.
+Here, one can directly use equals (==) on Satoshis, because it is a case class and Int64, the only parameter of Satoshis is a case class too.
 
@@
 It prints a lot of errors because of code incompatibility.
-So lets describe the code rewriting in the following sections.
+So let's describe the code rewriting in the following sections.
 
@@
 \end{lstlisting}
-Thus, changeing it to generics, Satoshis is not of type CurrencyUnit anymore and neither of type CurrencyUnit[A], where A is the new generic type.
+Thus, changing it to generics, Satoshis is not of type CurrencyUnit anymore and neither of type CurrencyUnit[A], where A is the new generic type.
 Satoshis extends CurrencyUnit with type Int64, so it is of type CurrencyUnit[Int64].
-That's to specific.
+That's too specific.
 
@@
 \end{lstlisting}
-Replace every occurence of A with BigInt.
-This might be a missing feature in Stainless that should not be to hard to fix.
+Replace every occurrence of A with BigInt.
+This might be a missing feature in Stainless that should not be too hard to fix.
 There is an open pull request \mypound470 on Github for this issue.
@@
 So here the \& mask can be replaced with a bound check whether the result is in range of Long.MinValue and Long.MaxValue.
-Again the code gets a bit more static but for this usecase its ok.
+Again the code gets a bit more static but for this use case it is OK.
 
@@
 \section{Propagate require}
-Finally the code works with Stainless.
+Finally, the code works with Stainless.
 But there is another problem.
-Bitcoin-S-Core uses require like a fail-fast method where as Stainless needs it to verify the code.
+Bitcoin-S-Core uses require like a fail-fast method whereas Stainless needs it to verify the code.
 
@@
 Sure, it can just add one to the maximum Int64 and the require does not hold anymore.
-The goal is to check the addition of zero so lets add a restriction for num to be zero.
+The goal is to check the addition of zero so let's add a restriction for num to be zero.
 

### 2019-05-28 doc/chapters/trying_verify_checktx.tex
more text

@@
 
-Bitcoin-S-Core uses sbt 1.2.8 and Scala 2.12.8.
+Bitcoin-S-Core uses sbt 1.2.8 and Scala 2.12.8, while Stainless sbt plugin is on sbt 0.13.17 and Scala 2.11.12.
 
-//TODO
+Sbt introduced new features in the 1.x releade used by Bitcoin-S.
+
+For example, they introduced the new slash syntax.
+\begin{lstlisting}
+  lazy val root = (project in file("."))
+    .settings(
+      name := "hello",
+      version in ThisBuild := "1.0.0"
+    )
+\end{lstlisting}
+The version line written without slash becomes now one with slash.
+The right and left parameter of in-operator reverse and instad of the in-operator is now a slash.
+\begin{lstlisting}
+  ThisBuild / version := "1.0.0",
+\end{lstlisting}
+This is just one example of many others.
+
+The bigger problem is, due to the different Scala and sbt versions, the error message, after trying to go in a sbt shell.
+\begin{lstlisting}
+  [warn] There may be incompatibilities among your library dependencies; run 'evicted'
+         to see detailed eviction warnings.
+  [error] java.lang.NoClassDefFoundError: sbt/SourcePosition
+  ...
+  Project loading failed: (r)etry, (q)uit, (l)ast, or (i)gnore?
+\end{lstlisting}
+Downgrading Bitcoin-S sbt version to 0.13.17 fixes the error but then it can not load some libraries only compiled for newer versions.
+So this would take too much time to fix and changes the Bitcoin-S code unwantingly.
+
+The next approach is to use stainless cli instead of sbt.
+Running stainless on all source files does not work, because the dependencies are missing.
+The parameter \emph{-classpath} can resolve it but the value of this parameter must be the paths to all the dependencies separated by a ':'.
+Finally, \emph{core} depends on \emph{secp256k1jni}, another package of Bitcoin-S written in Java.
+So this needs to be in the source files to.
+
+The final command looks like this in \emph{core} folder of Bitcoin-S:
+\begin{lstlisting}[language=bash]
+  $ stainless
+    -classpath ".:$(find ~/.ivy2/ -type f -name *.jar | tr '\n' ':')"
+    $(find . -type f -name *.scala | tr '\n' ' ')
+    $(find ../secp256k1jni -type f -name *.java | tr '\n' ' ')
+\end{lstlisting}
+The output was some Stainless error.
+\begin{lstlisting}
+  [Internal] Error: object scala.reflect.macros.internal.macroImpl in compiler mirror
+             not found.. Trace:
+  [Internal] - scala.reflect.internal.MissingRequirementError$.signal
+             (MissingRequirementError.scala:17)
+  ...
+  [Internal] object scala.reflect.macros.internal.macroImpl in compiler mirror not found.
+  [Internal] Please inform the authors of Inox about this message
+\end{lstlisting}
+After some discussion, we decided to go another way, because the errors may take too much time and it might lead to a next error.
+Let's extract the code needed to verify a transaction mainly the class Transaction and ScriptInterpreter with many other classes their depending on.
+
+After this extraction Stainless was successfully integrated with both, sbt and JAR.
+This leads us to another finding as described in the next section.
 

### 2019-05-28 doc/leader/frontpage_with_picture.tex
added comment

@@
 Authors:		\> Anna Doukmak, Ramon Boss		\\					% insert names
-Tutor:	\> Kai Brünnler @TODO Dr.?		\\							% insert names
+Tutor:	\> Kai Brünnler // TODO Dr.?		\\							% insert names
 %Constituent:	\> [Wwwww AG]					\\							% insert names
-Experts:		\> Urs Keller				\\							% insert names
+Experts:		\> Urs Keller // TODO remove?				\\							% insert names
 Date:			\> \versiondate					\\							% read from file leader/version.tex

### 2019-05-28 doc/chapters/using_bitcoins.tex
added language=scala

@@
 The relevant parts for the property to be verified are the following two lines.
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)

### 2019-05-28 doc/chapters/trying_verify_checktx.tex
more text

@@
 
-\section{Findings}
+\section{Bugfix}
+After so much digging and trying to understand the checkTransaction code, we found a bug even without Stainless.
 
-\section{Bugfix}
+Here the relevant code of checkTransaction again:
+\begin{lstlisting}[language=scala]
+  val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
+  val noDuplicateInputs = prevOutputTxIds.distinct.size == prevOutputTxIds.size
+\end{lstlisting}
+What happens, if a TransactionOutPoint (previousOutput) references the same Transaction ID (txId) another does?
+
+According to the Bitcoin protocol this is possible.
+A transaction can have multiple outputs that should be referenceable by the next transaction.
+What should not be possible is a transaction referencing the same output twice.
+This was a bug in Bitcoin Core known as CVE-2018–17144 which was patched on September 18 2018.
+
+Here, Bitcoin-S did a bit too much and marked all transaction as invalid, if they referenced a transaction twice in the next transaction.
+The fix is, to check on TransactionOutPoint instead of TransactionOutPoint.txId, because TransactionOutPoint contains the txId as well as the output index it references.
+\begin{lstlisting}[language=scala]
+  val prevOutputs = transaction.inputs.map(_.previousOutput)
+  val noDuplicateInputs = prevOutputs.distinct.size == prevOutputs.size
+\end{lstlisting}
+Since TransactionOutPoint is a case class and Scala has a built in == for case classes there is no need to implement TransactionOutPoint.==.
+
+This was fixed in pull request \mypound435 on GitHub at April 23 2019 by us along a unit test to prevent this bug from apperaing again in the future.

### 2019-05-28 doc/chapters/trying_verify_checktx.tex
text

@@
 \label{chap:connecting}
+This chapter describes the process we've made on trying to integrate Stainless in Bitcoin-S.
+It shows the differences between the error reporting by the sbt plugin and the JAR.
+Finally, during our work on Bitcoin-S we found a bug that we explain along its corresponding bug fix.
 
-\section{Integration (Versionskonflikte, neues Plugin)}
+\section{Integration}
 During this work, Stainless bumped the sbt plugin from version 0.13.17 to 1.2.8 and Scala from 2.11.12 to 2.12.8.

### 2019-05-28 doc/chapters/conclusion.tex
changed top level titles
removed placeholder

--- 2019-03-20 doc/chapters/conclusions.tex
@@
-\chapter{Conclusion / Results}
-\label{chap:conclusions}
+\chapter{Conclusion}
+\label{chap:conclusion}
 
-But I must explain to you how all this mistaken idea of denouncing pleasure and praising pain was born and I will give you a complete account of the system, and expound the actual teachings of the great explorer of the truth, the master-builder of human happiness.
-No one rejects, dislikes, or avoids pleasure itself, because it is pleasure, but because those who do not know how to pursue pleasure rationally encounter consequences that are extremely painful.
-Nor again is there anyone who loves or pursues or desires to obtain pain of itself, because it is pain, but because occasionally circumstances occur in which toil and pain can procure him some great pleasure.
-To take a trivial example, which of us ever undertakes laborious physical exercise, except to obtain some advantage from it?
-But who has any right to find fault with a man who chooses to enjoy a pleasure that has no annoying consequences, or one who avoids a pain that produces no resultant pleasure?
-On the other hand, we denounce with righteous indignation and dislike men who are so beguiled and demoralized by the charms of pleasure of the moment, so blinded by desire, that they cannot foresee the pain and trouble that are bound to ensue; and equal blame belongs to those who fail in their duty through weakness of will, which is the same as saying through shrinking from toil and pain.
-These cases are perfectly simple and easy to distinguish.
-In a free hour, when our power of choice is untrammelled and when nothing prevents our being able to do what we like best, every pleasure is to be welcomed and every pain avoided.
-But in certain circumstances and owing to the claims of duty or the obligations of business it will frequently occur that pleasures have to be repudiated and annoyances accepted.
-The wise man therefore always holds in these matters to this principle of selection: he rejects pleasures to secure other greater pleasures, or else he endures pains to avoid worse pains.
-
-\newpage
-
-But I must explain to you how all this mistaken idea of denouncing pleasure and praising pain was born and I will give you a complete account of the system, and expound the actual teachings of the great explorer of the truth, the master-builder of human happiness.
-No one rejects, dislikes, or avoids pleasure itself, because it is pleasure, but because those who do not know how to pursue pleasure rationally encounter consequences that are extremely painful.
-Nor again is there anyone who loves or pursues or desires to obtain pain of itself, because it is pain, but because occasionally circumstances occur in which toil and pain can procure him some great pleasure.
-To take a trivial example, which of us ever undertakes laborious physical exercise, except to obtain some advantage from it?
-But who has any right to find fault with a man who chooses to enjoy a pleasure that has no annoying consequences, or one who avoids a pain that produces no resultant pleasure?
-On the other hand, we denounce with righteous indignation and dislike men who are so beguiled and demoralized by the charms of pleasure of the moment, so blinded by desire, that they cannot foresee the pain and trouble that are bound to ensue; and equal blame belongs to those who fail in their duty through weakness of will, which is the same as saying through shrinking from toil and pain.
-These cases are perfectly simple and easy to distinguish.
-In a free hour, when our power of choice is untrammelled and when nothing prevents our being able to do what we like best, every pleasure is to be welcomed and every pain avoided.
-But in certain circumstances and owing to the claims of duty or the obligations of business it will frequently occur that pleasures have to be repudiated and annoyances accepted.
-The wise man therefore always holds in these matters to this principle of selection: he rejects pleasures to secure other greater pleasures, or else he endures pains to avoid worse pains.
+\section{Future Work}

### 2019-05-28 doc/leader/declaration_authorship.tex
added names

@@
 
-I / We hereby confirm that I / we have written this thesis independently and without using other sources and resources than those specified in the bibliography.
+We hereby confirm that we have written this thesis independently and without using other sources and resources than those specified in the bibliography.
 All text passages which were not written by me are marked as quotations and provided with the exact indication of its origin.
@@
 xxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\kill
-Place, Date:		\> [Biel/Burgdorf], \versiondate \\ \\ 
-Last Name/s, First Name/s:	\> [Test Peter] 	\> [M\"uster R\"os\"a] \\ \\ \\ \\ 
-Signature/s:	\> ......................................\> ...................................... \\
+Place, Date:		\> Biel, \versiondate \\ \\ 
+Last Names, First Names:	\> Doukmak Anna 	\> Boss Ramon \\ \\ \\ \\ 
+Signatures:	\> ......................................\> ...................................... \\
 \end{tabbing}

### 2019-05-28 doc/chapters/using_stainless.tex
fixed typos

@@
 \section{Configuration}
-There are two ways to integrate Stainless in a Scala project, the scala build tool (abt) plugin or a command line tool.
-Both, when run, analyses the passed code and report warnings to the console about the given code.
+There are two ways to integrate Stainless in a Scala project, the Scala build tool (sbt) plugin or a command line tool.
+Both, when run, analyses the code passed and report warnings to the console about the given code.
 Stainless requires and Scala recommends Java SE Development Kit 8.
@@
 sbt for Scala is like gradle or maven for Java.
-It can compile Scala code continously or manual, manage dependencies with support for Maven-formatted repositories, mixing Scala and Java projects and much more.
+It can compile Scala code continuously or manual, manage dependencies with support for Maven-formatted repositories, mixing Scala and Java projects and much more.
 
@@
 }
-\emph{build.properties} specifies the abt version used for this project.
+\emph{build.properties} specifies the sbt version used for this project.
 If the version is not available locally, sbt will download it.
@@
 \end{lstlisting}
-The project is called root and its source files are located in the files directory.
+The project is called root and its source files are located in the files' directory.
 Executing \code{sbt compile} should now compile the code.
 
-The Stainless webpage has a guide on how to integrate Stainless in an existing project.
+The Stainless web page has a guide on how to integrate Stainless in an existing project.
 The simplifies steps are:
@@
 Either download a prebuilt JAR file from efpl-lara/stainless GitHub repository or built a binary from source.
-The prebuild versions are released by Stainless.
+Prebuild versions are released by Stainless.
 The latest was released on January 14, when there was no support for Scala 2.12.

### 2019-05-28 doc/chapters/trying_verify_checktx.tex
added link
fixed typos

@@
 
-Sbt introduced new features in the 1.x releade used by Bitcoin-S.
+Sbt introduced new features in the 1.x release used by Bitcoin-S.
 
 For example, they introduced the new slash syntax.
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   lazy val root = (project in file("."))
@@
 \end{lstlisting}
+
 The version line written without slash becomes now one with slash.
-The right and left parameter of in-operator reverse and instad of the in-operator is now a slash.
-\begin{lstlisting}
+The right and left parameter of in-operator reverse and instead of the in-operator is now a slash.
+\begin{lstlisting}[language=scala]
   ThisBuild / version := "1.0.0",
 \end{lstlisting}
+
 This is just one example of many others.
@@
 \end{lstlisting}
+
 Downgrading Bitcoin-S sbt version to 0.13.17 fixes the error but then it can not load some libraries only compiled for newer versions.
@@
 \end{lstlisting}
+
 The output was some Stainless error.
@@
 \end{lstlisting}
+
 After some discussion, we decided to go another way, because the errors may take too much time and it might lead to a next error.
@@
 \section{Error Reporting with sbt and JAR}
+// TODO https://github.com/epfl-lara/stainless/issues/484?
 
@@
 \end{lstlisting}
+
 What happens, if a TransactionOutPoint (previousOutput) references the same Transaction ID (txId) another does?
@@
 What should not be possible is a transaction referencing the same output twice.
-This was a bug in Bitcoin Core known as CVE-2018–17144 which was patched on September 18 2018.
+This was a bug in Bitcoin Core known as CVE-2018–17144 which was patched on September 18, 2018.
 
@@
 
-This was fixed in pull request \mypound435 on GitHub at April 23 2019 by us along a unit test to prevent this bug from apperaing again in the future.
+This was fixed in pull request \mypound435 on GitHub at April 23, 2019 by us along a unit test to prevent this bug from appearing again in the future.

### 2019-05-28 doc/chapters/towards_verifying_addition.tex
fixed typos

@@
 The signature looks like the following.
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   +(c: CurrencyUnit): CurrencyUnit
 \end{lstlisting}
+
 Where the class Satoshis extends CurrencyUnit.
 The post condition would be, that the parameter c must be zero.
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   require(c.satoshis.underlying == Int64.zero)
 \end{lstlisting}
+
 \emph{c.satoshis} is an abstract method of CurrencyUnit that must return an instance of the class Satoshis.
@@
 To ensure that the result is the same value as \emph{this}:
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   ensuring(res => res.satoshis == this.satoshis)
 \end{lstlisting}
-Here, one can directly use equals (==) on Satoshis, because it is a case class and Int64, the only parameter of Satoshis is a case class too.
+
+Here, equals (==) can be used directly on Satoshis, because it is a case class and Int64, the only parameter of Satoshis is a case class too.
 
@@
 Simplified code.
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   sealed abstract class CurrencyUnit {
@@
 \end{lstlisting}
+
 Thus, changing it to generics, Satoshis is not of type CurrencyUnit anymore and neither of type CurrencyUnit[A], where A is the new generic type.
@@
 
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   sealed abstract class Number[T <: Number[T]] extends BasicArithmetic[T]  
@@
 \end{lstlisting}
+
 Changing the objects that extends classes to case objects solves this problem.
@@
 \end{lstlisting}
+
 As described before, Stainless supports only a subset of Scala.
@@
 Again, a simplified version.
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   sealed abstract class Satoshis extends CurrencyUnit {
@@
 \end{lstlisting}
+
 This would be really hard to refactor, because Bitcoin-S-Core tries to be as much dynamic as possible so it can be used with other cryptocurrencies too.
@@
 \end{lstlisting}
+
 Stainless is not able to extract private classes inside other objects.
 Bitcoin-S-core uses it a lot, because they separate the class from its implementation.
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   object Int64 extends BaseNumbers[Int64] {
@@
 \end{lstlisting}
+
 This is an easy one.
@@
 \end{lstlisting}
+
 Since the type A is not overridden in any of its subclasses, this can simply be hard coded.
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   sealed abstract class Number {
@@
 \end{lstlisting}
+
 Replace every occurrence of A with BigInt.
 This might be a missing feature in Stainless that should not be too hard to fix.
-There is an open pull request \mypound470 on Github for this issue.
+There is an open pull request \mypound470 on GitHub for this issue.
 
@@
 \end{lstlisting}
+
 Due to the restrictions on BigInt, the \& function on it is not supported too.
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   sealed abstract class Number extends BasicArithmetic[Int64] {
@@
 \end{lstlisting}
+
 This is a bound check.
@@
 \end{lstlisting}
+
 This is, because Stainless does not support require with a second String parameter.
@@
 \end{lstlisting}
+
 Corresponding code:
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   sealed abstract class Number extends BasicArithmetic[Int64] {
@@
 \end{lstlisting}
+
 But how can Stainless find a counter example ignoring the require in checkResult?
@@
 Thus, the addition need to have the restrictions of checkResult too.
-\begin{lstlisting}
+\begin{lstlisting}[language=scala]
   override def +(num: Int64): Int64 = {
@@
 \end{lstlisting}
+
 Stainless finds another counter example:
@@
 \end{lstlisting}
+
 Sure, it can just add one to the maximum Int64 and the require does not hold anymore.

### 2019-05-28 doc/chapters/using_stainless.tex
Added partially Scala compatibility.

--- 2019-05-22 doc/chapters/using_stainless.tex
@@
 \section{Scala compatibility (Pure Scala, imperative features, dedicated BigInt, Generics...)}
+
+To be verified the Scala code extended with Stainless annotations is passed to the framework as input.
+The source code should be written in so called Pure Scala, a functional subset of Scala, so that Stainless can perform the verification.
+Additionally the framework supports some imperative features. We will see some of this restrictions we have detected during this work in the next chapter.
+
+\subsection{Functional Pure Scala}
+The following functional elements are a part of Pure Scala:
+\begin{itemize}
+  \item \textbf{Algebraic Data Types:} abstract class, case class, case object. 
+  The root of ADT is an abstarct class, unless only one case class or case object is defined.
+  An abstract class cannot define fields and constructor arguments. 
+  Case class and case object can extend an abstract class.
+  An example of the ADT definition:
+  \begin{lstlisting}[language=Scala]
+    abstract class MyList
+    case object MyEmpty extends MyList
+    case class MyCons(elem: BigInt, rest: MyList) extends MyList
+  \end{lstlisting}
+  \item \textbf{Objects} for grouping classes and functions.
+  An example of object definition:
+  \begin{lstlisting}[language=Scala]
+    object Specs {
+      def increment(a: BigInt): BigInt = {
+        a + 1
+      }
+    case class Identifier(id: BigInt)
+    }
+  \end{lstlisting}
+  \item \textbf{Booleans:} logical operators  \textit{\&\&}, \textit{||}, \textit{!}, relational operator \textit{==}, Stainless syntax for boolean implication \textit{==>}. 
+  The short-circuit interpretation is used, when the second argument of a boolean expression is evaluated only when needed.  
+  \item \textbf{Generics.} Only invariant type parameters are supported, what means if S is subtype of T then List[S] and List[T] do not have sub-typing relationship.
+  Covariant and contravariant type parameters, such as List[+T] and List[-T], are not supported. 
+  Furthermore, only simple hierarchie is aloowed, when subclass must define the same type parameters in the same order as a parent class.
+  An example using Generics:
+  \begin{lstlisting}[language=Scala]
+    object Test {
+      abstract class List[T]
+      case class Cons[T](hd: T, tl: List[T]) extends List[T]
+      case class Nil[T]() extends List[T]
+      def contains[T](l: List[T], el: T) = { ... }
+    }
+  \end{lstlisting}
+  \item \textbf{Methods} can be defined in an abstract class and implemented or overridden in case classes. 
+  The call of superclass methods with the keyword \textit{super} is allowed. 
+  The method parameters can have default values. 
+  An example of method definition in case of inheritance:
+  \begin{lstlisting}[language=Scala]
+    object Test {
+      sealed abstract class Base {
+        def double(x: BigInt): BigInt = x * 2
+      }
+
+      case class Override() extends Base {
+        override def double(x: BigInt): BigInt = {
+          super.double(x + 1) + 42
+        }
+      }
+    }
+  \end{lstlisting}
+  \item \textbf{Pattern matching.} 
+  Simple nested pattern matching, matching on type only, pattern quards and custom pattern matching with \textit{unapply} method are supported.
+  \item \textbf{Values}. 
+  Only immutable variables are allowed. Thus, they must be defined with the keyword \textit{val} (and not with \textit{var}).
+  \item \textbf{Inner functions.}
+  \item \textbf{Anonymous functions.}
+  \item \textbf{Anonymous classes} with an explicit parent. 
+  An anonymous class should have the same number of public members as its parent.
+  An example of anonymous class implementation:
+  \begin{lstlisting}[language=Scala]
+    abstract class Foo {
+      def bar: Int
+    }
+
+    def makeFoo(x: Int): Foo = new Foo {
+      def bar: Int = x
+    }
+  \end{lstlisting}
+  \item \textbf{Local classes}. The creation of local classes within a function are allowed.
+\end{itemize}
+\subsection{Predefined types}
+\subsection{Imperative features}

### 2019-05-28 doc/appendix/appendixexemple.tex

\chapter{Arbitrary Appendix}
\label{chap:appendix_arb}
\endgroup

The European languages are members of the same family. Their separate existence is a myth. For science, music, sport, etc, Europe uses the same vocabulary. The languages only differ in their grammar, their pronunciation and their most common words. Everyone realizes why a new common language would be desirable: one could refuse to pay expensive translators. To achieve this, it would be necessary to have uniform grammar, pronunciation and more common words. If several languages coalesce, the grammar of the resulting language is more simple and regular than that of the individual languages. The new common language will be more simple and regular than the existing European languages. It will be as simple as Occidental; in fact, it will be Occidental.

### 2019-05-28 doc/chapters/introduction.tex

@@
 \label{chap:introduction}
-
 In this work some experiments in formal verification of Scala code will be accomplished. 
@@
 
-The commonly used one is testing. A set of some practical scenarios and assertions is created by a developer to test a software design and implementation. 
-While testing it is checked whether the actual results match the expected results. Tests help to identify bugs and missing requirements.
+The commonly used one is testing.
+A set of some practical scenarios and assertions is created by a developer to test a software design and implementation. 
+While testing, it is checked whether the actual results match the expected results.
+Tests help to identify bugs and missing requirements.
 But it is not achievable to test all possible inputs. 
 Thus, a long-term test coverage model should be created to test a software design enough.
-Furthermore it is challenging to test a software keeping the ability to observe the side effects of a specific part of code.
-Practically, a developer runs tests, debugs appeared failures, adjustes a code of a software and extands a testbench verifying previously uncovered aspects. \cite{sanghavi:formal_verification}
+Furthermore, it is challenging to test a software keeping the ability to observe the side effects of a specific part of code.
+Practically, a developer runs tests, debugs appeared failures, adjusts a code of a software and extends a test bench verifying previously uncovered aspects. \cite{sanghavi:formal_verification}
 
-Another powerfull method to check whether a program works correct is formal verification. 
+Another powerful method to check whether a program works correct is formal verification. 
 This is a systematic process based on mathematical modeling. 
@@
 
-One of such tools is Stainless. This framework is used in the work to verify a Scala code.
+One of such tools is Stainless.
+This framework is used in the work to verify a Scala code.
 
@@
 
-Stainless is a framework developed by "Lab for Automated Reasoning and Analysis" (LARA) at EPFL's School of Computer and Communication Sciences. The framework is used to verify Scala programs.
+Stainless is a framework developed by "Lab for Automated Reasoning and Analysis" (LARA) at EPFL's School of Computer and Communication Sciences.
+The framework is used to verify Scala programs.
 It verifies statically that a program satisfies a specification given by a developer and that a program will not crash at runtime.
@@
 A precondition should be written at the beginning of a function body with \textit{require}.
-Its argument is a boolean expression which corresponds to constraint for inputs of a function being verified.
-A postcondition should be written after a function body with  \textit{ensuring}. It is a verification condition on an output of a function.
+Its argument is a Boolean expression which corresponds to constraint for inputs of a function being verified.
+A postcondition should be written after a function body with \textit{ensuring}.
+It is a verification condition on an output of a function.
 While compiling Stainless tries to prove that the postcondition always holds, assuming a given precondition does hold.\cite{Stainless:introduction}
 
-3 outcomes of verification with Stainless are possible: valid, unvalid and unknown. 
+3 outcomes of verification with Stainless are possible: valid, invalid and unknown. 
 If the postcondition is valid, Stainless could prove that for any inputs constrained in a precondition, the postcondition always holds.
 With invalid postcondition the framework could find at least one counterexample which satisfies a precondition but violates a postcondition.
-Also an output unknown is possible when Stainless is unable to prove a postcondition or find a counterexample. In this case a timeout or an internal error occured.
+Also, an output unknown is possible when Stainless is unable to prove a postcondition or find a counterexample.
+In this case a timeout or an internal error occurred.
 Furthermore, it will be verified by Stainless that a precondition cannot be violated.\cite{Stainless:introduction}
@@
 That is why calculating factorial of 17 causes an overflow and results to a negative value.
-The program would work correct by changing a type of a number to BigInt.
+The program would work correct by changing a type of number to BigInt.
 The outcome of Stainless verification of the function factorial from Listing \ref{listing:factorial} is shown bellow.
@@
 
-Stainless verifies scala programs using SMT (satisfiability modulo theories) solvers. A program code will be transformed by Stainless to the Pure Scala fragment supported by Inox solver. 
+Stainless verifies Scala programs using SMT (satisfiability modulo theories) solvers. A program code will be transformed by Stainless to the Pure Scala fragment supported by Inox solver. 
 This solver was also developed by LARA and allows usage of SMT solvers such as Z3, CVC4 and Princess, adding support of other higher-level features.
-Preconditions, postconditions and assertions will be translated into SMT formulas, so a SMT solver can determine if all properties hold.
+Preconditions, postconditions and assertions will be translated into SMT formulas, so an SMT solver can determine if all properties hold.
 
@@
 
-In the work some fragments of the code of Bitcoin-S has to be verified. Bitcoin-S is an open source Scala implementation of the Bitcoin protocol. 
+In the work some fragments of the code of Bitcoin-S has to be verified.
+Bitcoin-S is an open source Scala implementation of the Bitcoin protocol. 
 There are different projects belonging to Bitcoin-S.
@@
 This package is the main target to be verified because it specifies the base of Bitcoin protocol, and it is crucial for correct protocol functionality.
-There is also the package \textbf{bitcoind-rpc} which is a RPC client implementation to communicate with \textit{bitcoind}, daemon of Bitcoin Core running on a machine.
+There is also the package \textbf{bitcoind-rpc} which is an RPC client implementation to communicate with \textit{bitcoind}, daemon of Bitcoin Core running on a machine.
 Developers of Bitcoin-S are working on the implementation of a lightweight bitcoin client in the package \textbf{bitcoin-s-spv-node}.
-The whole list of packages can be found on the github \cite{BitcoinSProject}.
+The whole list of packages can be found on GitHub \cite{BitcoinSProject}.
 
 Bitcoin-S is a large project implementing many functionalities.
-In this work one of them is going to be examinated and worked on, namely the property of the Bitcoin that a non-coinbase transaction can not generate new coins.
-Thus, the code of Bitcoin-S should be analyzied, and fragments implementing this feature should be identified.
+In this work one of them is going to be examined and worked on, namely the property of the Bitcoin that a non-coinbase transaction can not generate new coins.
+Thus, the code of Bitcoin-S should be analyzed, and fragments implementing this feature should be identified.
 Afterwards, the fragments should be rewritten, so that Stainless can accept it.
-After that a formal specification for functions should be defined and annotated according Stainless libraries, so that Stainless can verify corrrectness of the code.
+After that a formal specification for functions should be defined and annotated according Stainless libraries, so that Stainless can verify correctness of the code.

### 2019-05-28 doc/chapters/using_bitcoins.tex

@@
 This chapter describes the relevant parts needed to verify that a non-coinbase transaction cannot generate new coins in Bitcoin-S-Core.
-We will see, how to create valid and invalid transactions with Bitcoin-S-Core and how this transactions are validated against the described property.
+We will see, how to create valid and invalid transactions with Bitcoin-S-Core and how these transactions are validated against the described property.
 
@@
 Some parts of the code in this section are from Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
-Bitcoin-S-Core has a bitcoin transcation builder class with the following signature:
+Bitcoin-S-Core has a bitcoin transaction builder class with the following signature:
 \begin{lstlisting}[language=scala]
@@
 A single output is sufficient for this example.
-So, first create a new keypair to sign the next transaction and have a scriptPubKey where the money is.
-Then define the amount to spend (here 10000 Satishis) collect this information in a transaction output and add it to the previous transaction.
+So, first create a new key pair to sign the next transaction and have a scriptPubKey where the money is.
+Then define the amount to spend (here 10000 Satoshis) collect this information in a transaction output and add it to the previous transaction.
 \begin{lstlisting}[language=scala]
@@
 Next, the new transaction should point to an output of the previous transaction.
-Thus, a outpoint is created with the id of the previous transaction and the index pointing to the specific output of it.
+Thus, an outpoint is created with the id of the previous transaction and the index pointing to the specific output of it.
 This is collected in an utxo spending info which is then put in the list of all utxos (only one here).
@@
 Then, the destination, where the money goes, is defined.
-This includes a destination script pub key, as well as a the amount to spend to it.
+This includes a destination script pub key, as well as the amount to spend to it.
 \begin{lstlisting}
@@
 \end{lstlisting}
-Finally. define a fee rate, the network params and create a transaction builder.
+Finally, define a fee rate, the network params and create a transaction builder.
 \begin{lstlisting}
@@
 \end{lstlisting}
-After calling sign on the transaction builder a valid transaction is returned.
+After calling sign on the transaction builder, a valid transaction is returned.
 \begin{lstlisting}
@@
 \end{lstlisting}
-It takes a transaction as input and returns a boolean whether the transaction is valid or not.
+It takes a transaction as input and returns a Boolean whether the transaction is valid or not.
 So for example when passing the built tx from above to this function the returned value would be true.
 There are several checks in checkTransaction.
-For example it checks if there is either no input or no output.
+For example, it checks if there is either no input or no output.
 In this case it returns false.

### 2019-05-28 doc/chapters/using_stainless.tex
Updated Bibiliography
fixed typos

@@
 The source code should be written in so called Pure Scala, a functional subset of Scala, so that Stainless can perform the verification.
-Additionally the framework supports some imperative features. We will see some of this restrictions we have detected during this work in the next chapter.
+Additionally, the framework supports some imperative features.
+We will see some of these restrictions we have detected during this work in the next chapter.
 
@@
   \item \textbf{Algebraic Data Types:} abstract class, case class, case object. 
-  The root of ADT is an abstarct class, unless only one case class or case object is defined.
+  The root of ADT is an abstract class, unless only one case class or case object is defined.
   An abstract class cannot define fields and constructor arguments. 
@@
   \item \textbf{Booleans:} logical operators  \textit{\&\&}, \textit{||}, \textit{!}, relational operator \textit{==}, Stainless syntax for boolean implication \textit{==>}. 
-  The short-circuit interpretation is used, when the second argument of a boolean expression is evaluated only when needed.  
+  The short-circuit interpretation is used, when the second argument of a Boolean expression is evaluated only when needed.
   \item \textbf{Generics.} Only invariant type parameters are supported, what means if S is subtype of T then List[S] and List[T] do not have sub-typing relationship.
   Covariant and contravariant type parameters, such as List[+T] and List[-T], are not supported. 
-  Furthermore, only simple hierarchie is aloowed, when subclass must define the same type parameters in the same order as a parent class.
+  Furthermore, only simple hierarchies are allowed, when subclass must define the same type parameters in the same order as a parent class.
   An example using Generics:
@@
   \item \textbf{Pattern matching.} 
-  Simple nested pattern matching, matching on type only, pattern quards and custom pattern matching with \textit{unapply} method are supported.
+  Simple nested pattern matching, matching on type only, pattern guards and custom pattern matching with \textit{unapply} method are supported.
   \item \textbf{Values}. 

### 2019-05-28 doc/chapters/using_stainless.tex
Added Imperative features.

--- 2019-05-28 doc/chapters/using_stainless.tex
@@
   \end{lstlisting}
-  \item \textbf{Local classes}. The creation of local classes within a function are allowed.
+  \item \textbf{Local classes}. 
+  The creation of local classes within a function are allowed. \cite{Stainless:pure_scala}
+  \item \textbf{Functional arrays} which use the operation that do not modify an array.
 \end{itemize}
+
 \subsection{Predefined types}
+
 \subsection{Imperative features}
+
+Stainless supports extensions to Pure Scala implemeting some imperative features.
+They will be translated by Stainless into Pure Scala during a preprocessing phase.
+
+The following imperative components are supported:
+
+\begin{itemize}
+  \item \textbf{Mutable variables} which can be declared with the keyword \textit{var}.
+  Declaration of mutable variables is allowed only as a local variable within a function or in a case classe constuctor.
+  \item  \textbf{While loops}. 
+  Because besides main calculations in a loop the state of a loop counter changes at every iteration as a side effect, this kind of programming component does not belong to the functional paradigm and cannot be handled by solvers.
+  Thus, Stainless generates automatically a postcondition with the negation of the loop condition as an argument. 
+  Moreover, to make it stronger, the automatically generated postcondition can be customised with the keyword \textit{invariant}.
+  An example of while loop implementation:
+  \begin{lstlisting}[language=Scala]
+    var res = 0
+    var i = 0
+    (while(i < 10) {
+      res = res + i
+      i = i + 1
+    }) invariant(i >= 0 && res >= i)
+  \end{lstlisting}
+  \item \textbf{Arrays.} Stainless extends functional arrays allowing the reassignment of arrays elements.
+  \item ...
+\end{itemize}

### 2019-05-28 doc/chapters/using_stainless.tex
Added Stainless library.

@@
 
-\subsection{Predefined types}
-
 \subsection{Imperative features}
@@
   \item \textbf{Mutable variables} which can be declared with the keyword \textit{var}.
-  Declaration of mutable variables is allowed only as a local variable within a function or in a case classe constuctor.
+  Declaration of mutable variables is allowed only as a local variable within a function, or in a case classe constuctor, or in trait if a default value is not assigned.
   \item  \textbf{While loops}. 
@@
   \end{lstlisting}
-  \item \textbf{Arrays.} Stainless extends functional arrays allowing the reassignment of arrays elements.
-  \item ...
+  \item \textbf{Function \textit{old}} can be used in postconditions. 
+  It allows to get value of a variable before the execution of the block.
+  This feature is useful dealing with mutable objects.
+  An example of using the function \textit{old}:
+  \begin{lstlisting}[language=Scala]
+    case class A(var x: Int)
+    def inc(a: A): Unit = {
+      a.x = a.x + 1
+    } ensuring(_ => a.x == old(a).x + 1)
+  \end{lstlisting}
+  \item \textbf{Arrays.} Stainless extends functional arrays allowing the reassignment of arrays elements. \cite{Stainless:imperative}
 \end{itemize}
+
+\subsection{Stainless library}
+
+Besides annotations and functions helping for verification  some core data types and operation on them is defined in the Stainless library.
+The own implementation and dedicated mapping of data types allows Stainless to transform it correctly into mathematical expressions of SMT solvers.
+Thus, the functionality of some data types can be restricted.
+
+For example, BigInt implemented in Stainless has constructors only with a parameter of the type Int or String, while other types of a parameter is allowed in Scala.
+Furthermore, Stainless defines its own collection List[T] with the standard and some additional functions because SMT solvers do not support it.
+Moreover, the own Sets and Maps are used in Stainless. \cite{Stainless:library}

### 2019-05-29 doc/chapters/towards_verifying_addition.tex
case object

@@
              sealed abstract class Number[T <: Number[T]]
-
 \end{lstlisting}
 
+This is a missing feature in Stainless.
 \begin{lstlisting}[language=scala]
@@
 \section{Rewriting Objects}
-// TODO reason?
-
 Stainless output:
@@
 Changing the objects that extends classes to case objects solves this problem.
+This is, due to the internal design of Scala and Java.
+It's possible to reason about a case object but not about an object.
+This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
+The issue \mypound520 on Stainless GitHub gives some thoughts about this.
 

### 2019-05-29 doc/chapters/towards_verifying_addition.tex
generics
fixed console outputs

@@
 Stainless output:
-\begin{lstlisting}
+\begin{verbatim}
 [ Error  ] CurrencyUnits.scala:5:3: Stainless doesn't support abstract type members
            type A
-\end{lstlisting}
+\end{verbatim}
 This should be easy to rewrite by using generics instead of an abstract type, right?
@@
 \section{Rewriting Generics}
-// I dont understand this.. maybe just a missing feature in Stainless?
 Stainless output:
-\begin{lstlisting}
+\begin{verbatim}
   [ Error  ] NumberType.scala:3:30: Unknown type parameter type T
              sealed abstract class Number[T <: Number[T]]
-\end{lstlisting}
+\end{verbatim}
 
 This is a missing feature in Stainless.
+To track this, there is the issue \mypound519 on GitHub.
 \begin{lstlisting}[language=scala]
@@
 \end{lstlisting}
+For the sake of convenience this is also made static by parameterize T with Int64.
 
@@
 Stainless output:
-\begin{lstlisting}
+\begin{verbatim}
 [ Error  ] CurrencyUnits.scala:54:1: Objects cannot extend classes or implement
@@
            object Satoshis extends BaseNumbers[Satoshis] {
-\end{lstlisting}
+\end{verbatim}
 
@@
 Stainless output:
-\begin{lstlisting}
+\begin{verbatim}
 [ Error  ] CurrencyUnits.scala:47:33: Only literal arguments are allowed for BigInt.
            def toBigInt: BigInt = BigInt(toLong)
-\end{lstlisting}
+\end{verbatim}
 
@@
 Stainless output (simplified):
-\begin{lstlisting}
+\begin{verbatim}
   [Warning ] CurrencyUnits.scala:36:3: Could not extract tree in class:
              case private class SatoshisImpl extends Satoshis
-\end{lstlisting}
+\end{verbatim}
 
@@
 Stainless output:
-\begin{lstlisting}
+\begin{verbatim}
   [Warning ] NumberType.scala:5:3: Could not extract tree in class: type A
@@
              type A = BigInt
-\end{lstlisting}
+\end{verbatim}
 
@@
 Stainless output:
-\begin{lstlisting}
+\begin{verbatim}
   [ Error  ] NumberType.scala:19:14: Unknown call to & on result (BigInt) with
@@
              require((result & andMask) == result,
-\end{lstlisting}
+\end{verbatim}
 
@@
 Stainless output:
-\begin{lstlisting}
+\begin{verbatim}
   [Warning ] NumberType.scala:54:3: Could not extract tree in class:
@@
              require(underlying <= 9223372036854775807L,
-\end{lstlisting}
+\end{verbatim}
 
@@
 Stainless output:
-\begin{lstlisting}
+\begin{verbatim}
   [Warning ] Found counter-example:
@@
                  Int64Impl(9223372036854775808)
-\end{lstlisting}
+\end{verbatim}
 
@@
 Stainless finds another counter example:
-\begin{lstlisting}
+\begin{verbatim}
   [Warning ] Found counter-example:
@@
                  Int64Impl(9223372036854775807)
-\end{lstlisting}
+\end{verbatim}
 

### 2019-05-29 doc/chapters/towards_verifying_addition.tex
small font size for verbatim

@@
 Stainless output:
-\begin{verbatim}
+{\footnotesize\begin{verbatim}
 [ Error  ] CurrencyUnits.scala:5:3: Stainless doesn't support abstract type members
            type A
-\end{verbatim}
+\end{verbatim}}
 This should be easy to rewrite by using generics instead of an abstract type, right?
@@
 Stainless output:
-\begin{verbatim}
+{\footnotesize\begin{verbatim}
   [ Error  ] NumberType.scala:3:30: Unknown type parameter type T
              sealed abstract class Number[T <: Number[T]]
-\end{verbatim}
+\end{verbatim}}
 
@@
 Stainless output:
-\begin{verbatim}
+{\footnotesize\begin{verbatim}
 [ Error  ] CurrencyUnits.scala:54:1: Objects cannot extend classes or implement
@@
            object Satoshis extends BaseNumbers[Satoshis] {
-\end{verbatim}
+\end{verbatim}}
 
@@
 Stainless output:
-\begin{verbatim}
+{\footnotesize\begin{verbatim}
 [ Error  ] CurrencyUnits.scala:47:33: Only literal arguments are allowed for BigInt.
            def toBigInt: BigInt = BigInt(toLong)
-\end{verbatim}
+\end{verbatim}}
 
@@
 Stainless output (simplified):
-\begin{verbatim}
+{\footnotesize\begin{verbatim}
   [Warning ] CurrencyUnits.scala:36:3: Could not extract tree in class:
              case private class SatoshisImpl extends Satoshis
-\end{verbatim}
+\end{verbatim}}
 
@@
 Stainless output:
-\begin{verbatim}
+{\footnotesize\begin{verbatim}
   [Warning ] NumberType.scala:5:3: Could not extract tree in class: type A
@@
              type A = BigInt
-\end{verbatim}
+\end{verbatim}}
 
@@
 Stainless output:
-\begin{verbatim}
+{\footnotesize\begin{verbatim}
   [ Error  ] NumberType.scala:19:14: Unknown call to & on result (BigInt) with
@@
              require((result & andMask) == result,
-\end{verbatim}
+\end{verbatim}}
 
@@
 Stainless output:
-\begin{verbatim}
+{\footnotesize\begin{verbatim}
   [Warning ] NumberType.scala:54:3: Could not extract tree in class:
@@
              require(underlying <= 9223372036854775807L,
-\end{verbatim}
+\end{verbatim}}
 
@@
 Stainless output:
-\begin{verbatim}
+{\footnotesize\begin{verbatim}
   [Warning ] Found counter-example:
@@
                  Int64Impl(9223372036854775808)
-\end{verbatim}
+\end{verbatim}}
 
@@
 Stainless finds another counter example:
-\begin{verbatim}
+{\footnotesize\begin{verbatim}
   [Warning ] Found counter-example:
@@
                  Int64Impl(9223372036854775807)
-\end{verbatim}
+\end{verbatim}}
 

### 2019-05-29 doc/leader/frontpage_with_picture.tex
Updated title page.

@@
 Authors:		\> Anna Doukmak, Ramon Boss		\\					% insert names
-Tutor:	\> Kai Brünnler // TODO Dr.?		\\							% insert names
+Tutor:	\> Prof. Dr. Kai Brünnler // TODO Dr.?		\\							% insert names
 %Constituent:	\> [Wwwww AG]					\\							% insert names
-Experts:		\> Urs Keller // TODO remove?				\\							% insert names
+External referee:		\> Urs Keller // TODO remove?				\\							% insert names
 Date:			\> \versiondate					\\							% read from file leader/version.tex

### 2019-05-29 doc/chapters/towards_verifying_addition.tex
updated titles

@@
 
-\section{Rewriting Abstract Type Member}
+\section{Get Rid of Abstract Type Member}
 Stainless output:
@@
 
-\section{Rewriting Generics}
+\section{Get Rid of Special Generics}
 Stainless output:
@@
 
-\section{Rewriting Objects}
+\section{Turning object Into case object}
 Stainless output:
@@
 
-\section{Rewriting BigInt Constructor (only literal argument, no long argument, etc.)}
+\section{Replace BigInt Constructor Argument With String Literal}
 Stainless output:
@@
 
-\section{Rewriting Private Inner Classes}
+\section{Extracting Private Inner Classes}
 Stainless output (simplified):
@@
 
-\section{Rewriting Type Member}
+\section{Get Rid of Concrete Type Member}
 Stainless output:
@@
 
-\section{Rewriting Usage of BigInt \&-Function}
+\section{Replace BigInt \&-Function With Bound Check}
 Stainless output:
@@
 
-\section{Rewriting require}
+\section{Remove Second Parameter of require}
 Stainless output:

### 2019-05-29 doc/chapters/trying_verify_checktx.tex
moved using_bitcoins to trying

@@
 \label{chap:connecting}
-This chapter describes the process we've made on trying to integrate Stainless in Bitcoin-S.
-It shows the differences between the error reporting by the sbt plugin and the JAR.
-Finally, during our work on Bitcoin-S we found a bug that we explain along its corresponding bug fix.
+// TODO
+
+\section{Creation of a Transaction}
+Some parts of the code in this section are from Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
+Bitcoin-S-Core has a bitcoin transaction builder class with the following signature:
+\begin{lstlisting}[language=scala]
+  BitcoinTxBuilder(
+    destinations: Seq[TransactionOutput], // where we send money
+    utxos: BitcoinTxBuilder.UTXOMap,      // unspent transaction outputs
+    feeRate: FeeUnit,                     // fee rate per byte
+    changeSPK: ScriptPubKey,              // public key
+    network: BitcoinNetwork               // bitcoin network information
+  ): Future[BitcoinTxBuilder]             // sign TxBuilder to get tx
+\end{lstlisting}
+Here is how those parameters are generated.
+
+First, a previous transaction with outputs is needed to spend some money from.
+This is created here, to show the process.
+It could also be parsed from a transaction in the bitcoin network.
+A single output is sufficient for this example.
+So, first create a new key pair to sign the next transaction and have a scriptPubKey where the money is.
+Then define the amount to spend (here 10000 Satoshis) collect this information in a transaction output and add it to the previous transaction.
+\begin{lstlisting}[language=scala]
+  val privKey = ECPrivateKey.freshPrivateKey
+  val creditingSPK = P2PKHScriptPubKey(pubKey = privKey.publicKey)
+
+  val amount = Satoshis(Int64(10000))
+
+  val utxo = TransactionOutput(currencyUnit = amount, scriptPubKey = creditingSPK)
+
+  val prevTx = BaseTransaction(
+    version = Int32.one,
+    inputs = List.empty,
+    outputs = List(utxo),
+    lockTime = UInt32.zero
+  )
+\end{lstlisting}
+Next, the new transaction should point to an output of the previous transaction.
+Thus, an outpoint is created with the id of the previous transaction and the index pointing to the specific output of it.
+This is collected in an utxo spending info which is then put in the list of all utxos (only one here).
+\begin{lstlisting}  
+  val outPoint = TransactionOutPoint(prevTx.txId, UInt32.zero)
+
+  val utxoSpendingInfo = BitcoinUTXOSpendingInfo(
+    outPoint = outPoint,
+    output = utxo,
+    signers = List(privKey),
+    redeemScriptOpt = None,
+    scriptWitnessOpt = None,
+    hashType = HashType.sigHashAll
+  )
+
+  val utxos = List(utxoSpendingInfo)
+\end{lstlisting}
+Then, the destination, where the money goes, is defined.
+This includes a destination script pub key, as well as the amount to spend to it.
+\begin{lstlisting}
+  val destinationAmount = Satoshis(Int64(5000))
+
+  val destinationSPK = P2PKHScriptPubKey(pubKey = ECPrivateKey.freshPrivateKey.publicKey)
+
+  val destinations = List(
+    TransactionOutput(currencyUnit = destinationAmount, scriptPubKey = destinationSPK)
+  )
+\end{lstlisting}
+Finally, define a fee rate, the network params and create a transaction builder.
+\begin{lstlisting}
+  val feeRate = SatoshisPerByte(Satoshis.one)
+
+  val networkParams = RegTest // soem static values for testing
+
+  val txBuilder: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
+    destinations = destinations,
+    utxos = utxos,
+    feeRate = feeRate,
+    changeSPK = creditingSPK,
+    network = networkParams
+  )
+\end{lstlisting}
+After calling sign on the transaction builder, a valid transaction is returned.
+\begin{lstlisting}
+  val signedTxF: Future[Transaction] = txBuilder
+    .flatMap(_.sign)
+    .map {
+      (tx: Transaction) => println(tx.hex) // transaction in hex for the bitcoin network
+    }
+\end{lstlisting}
+There is no need for a transaction input, since a transaction out point is kind of the pointer to a transaction input.
+
+\section{Validation of a Transaction}
+Bitcoin-S-Core offers a function called \emph{checkTransaction}.
+This is its type signature.
+\begin{lstlisting}
+  checkTransaction(transaction: Transaction): Boolean
+\end{lstlisting}
+It takes a transaction as input and returns a Boolean whether the transaction is valid or not.
+So for example when passing the built tx from above to this function the returned value would be true.
+There are several checks in checkTransaction.
+For example, it checks if there is either no input or no output.
+In this case it returns false.
+
+The relevant parts for the property to be verified are the following two lines.
+\begin{lstlisting}[language=scala]
+  val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
+  val noDuplicateInputs = prevOutputTxIds.distinct.size == prevOutputTxIds.size
+\end{lstlisting}
+It gathers all inputs output transaction ids.
+On this previous outputs ids it calls distinct and checks if the size stays the same.
+Distinct removes duplicates, so if there was two times the same input transaction id the size of those ids would be greater before calling distinct.
+
+There is a bug in those two lines, as described in the next chapter.
 

### 2019-05-29 doc/chapters/trying_verify_checktx.tex
changed title
rewrite

--- 2019-05-28 doc/chapters/trying_verify_checktx.tex
@@
 \label{chap:connecting}
-This chapter describes the process we've made on trying to integrate Stainless in Bitcoin-S.
-It shows the differences between the error reporting by the sbt plugin and the JAR.
-Finally, during our work on Bitcoin-S we found a bug that we explain along its corresponding bug fix.
+// TODO
 
-\section{Integration}
-During this work, Stainless bumped the sbt plugin from version 0.13.17 to 1.2.8 and Scala from 2.11.12 to 2.12.8.
-So this section might be out of date now.
+\section{Creation of a Transaction}
 
-Bitcoin-S-Core uses sbt 1.2.8 and Scala 2.12.8, while Stainless sbt plugin is on sbt 0.13.17 and Scala 2.11.12.
+Some of the code in this section are copied or modified parts of Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
+Bitcoin-S-Core has a bitcoin transaction builder class with the following signature:
+\begin{lstlisting}[language=scala]
+  BitcoinTxBuilder(
+    destinations: Seq[TransactionOutput], // where we send money
+    utxos: BitcoinTxBuilder.UTXOMap,      // unspent transaction outputs
+    feeRate: FeeUnit,                     // fee rate per byte
+    changeSPK: ScriptPubKey,              // public key
+    network: BitcoinNetwork               // bitcoin network information
+  ): Future[BitcoinTxBuilder]             // sign TxBuilder to get tx
+\end{lstlisting}
 
-Sbt introduced new features in the 1.x release used by Bitcoin-S.
+Here is how those parameters are generated.
 
-For example, they introduced the new slash syntax.
+First, we need a some bitcoins/satoshis.
+Thus, we create a transaction with one single output.
+This transaction can also be parsed from the real bitcoin network but to show the progression we create one.
 \begin{lstlisting}[language=scala]
-  lazy val root = (project in file("."))
-    .settings(
-      name := "hello",
-      version in ThisBuild := "1.0.0"
+  val privKey = ECPrivateKey.freshPrivateKey
+  val creditingSPK = P2PKHScriptPubKey(pubKey = privKey.publicKey)
+
+  val amount = Satoshis(Int64(10000))
+
+  val utxo = TransactionOutput(currencyUnit = amount, scriptPubKey = creditingSPK)
+
+  val prevTx = BaseTransaction(
+    version = Int32.one,
+    inputs = List.empty,
+    outputs = List(utxo),
+    lockTime = UInt32.zero
   )
@@
 
-The version line written without slash becomes now one with slash.
-The right and left parameter of in-operator reverse and instead of the in-operator is now a slash.
+On line one and two we are creating a new keypair to sign the next transaction and have a scriptPubKey where the bitcoins are.
+Line four specifies the amount of satoshis we are having in the transaction.
+Then we create the actual transaction from line 6 to 13.
+
+Now that we have some bitcoins, we create the new transaction to spend them.
+
+First, we need some out points.
+This are pointers to previous transactions' outputs.
+We use the index zero, because the previous transaction has only one output that becomes the first index zero.
 \begin{lstlisting}[language=scala]
-  ThisBuild / version := "1.0.0",
+  val outPoint = TransactionOutPoint(prevTx.txId, UInt32.zero)
+
+  val utxoSpendingInfo = BitcoinUTXOSpendingInfo(
+    outPoint = outPoint,
+    output = utxo,
+    signers = List(privKey),
+    redeemScriptOpt = None,
+    scriptWitnessOpt = None,
+    hashType = HashType.sigHashAll
+  )
+
+  val utxos = List(utxoSpendingInfo)
 \end{lstlisting}
 
-This is just one example of many others.
+Second, we need destinations to spend the bitcoins to.
+For the sake of convenience we create only one.
+\begin{lstlisting}[language=scala]
+  val destinationAmount = Satoshis(Int64(5000))
 
-The bigger problem is, due to the different Scala and sbt versions, the error message, after trying to go in a sbt shell.
-\begin{lstlisting}
-  [warn] There may be incompatibilities among your library dependencies; run 'evicted'
-         to see detailed eviction warnings.
-  [error] java.lang.NoClassDefFoundError: sbt/SourcePosition
-  ...
-  Project loading failed: (r)etry, (q)uit, (l)ast, or (i)gnore?
+  val destinationSPK = P2PKHScriptPubKey(pubKey = ECPrivateKey.freshPrivateKey.publicKey)
+
+  val destinations = List(
+    TransactionOutput(currencyUnit = destinationAmount, scriptPubKey = destinationSPK)
+  )
 \end{lstlisting}
 
-Downgrading Bitcoin-S sbt version to 0.13.17 fixes the error but then it can not load some libraries only compiled for newer versions.
-So this would take too much time to fix and changes the Bitcoin-S code unwantingly.
+We spend 5000 satoshis to the newly created script public key.
 
-The next approach is to use stainless cli instead of sbt.
-Running stainless on all source files does not work, because the dependencies are missing.
-The parameter \emph{-classpath} can resolve it but the value of this parameter must be the paths to all the dependencies separated by a ':'.
-Finally, \emph{core} depends on \emph{secp256k1jni}, another package of Bitcoin-S written in Java.
-So this needs to be in the source files to.
+Finally, we define the fee rate in satoshis per one byte transaction size as well as some bitcoin network parameters.
+\begin{lstlisting}[language=scala]
+  val feeRate = SatoshisPerByte(Satoshis.one)
 
-The final command looks like this in \emph{core} folder of Bitcoin-S:
-\begin{lstlisting}[language=bash]
-  $ stainless
-    -classpath ".:$(find ~/.ivy2/ -type f -name *.jar | tr '\n' ':')"
-    $(find . -type f -name *.scala | tr '\n' ' ')
-    $(find ../secp256k1jni -type f -name *.java | tr '\n' ' ')
+  val networkParams = RegTest // soem static values for testing
 \end{lstlisting}
 
-The output was some Stainless error.
-\begin{lstlisting}
-  [Internal] Error: object scala.reflect.macros.internal.macroImpl in compiler mirror
-             not found.. Trace:
-  [Internal] - scala.reflect.internal.MissingRequirementError$.signal
-             (MissingRequirementError.scala:17)
-  ...
-  [Internal] object scala.reflect.macros.internal.macroImpl in compiler mirror not found.
-  [Internal] Please inform the authors of Inox about this message
+Now lets build the transaction with those data.
+\begin{lstlisting}[language=scala]
+  val txBuilder: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
+    destinations = destinations,
+    utxos = utxos,
+    feeRate = feeRate,
+    changeSPK = creditingSPK,
+    network = networkParams
+  )
+
+  val signedTxF: Future[Transaction] = txBuilder
+    .flatMap(_.sign)
+    .map {
+      (tx: Transaction) => println(tx.hex) // transaction in hex for the bitcoin network
+    }
 \end{lstlisting}
 
-After some discussion, we decided to go another way, because the errors may take too much time and it might lead to a next error.
-Let's extract the code needed to verify a transaction mainly the class Transaction and ScriptInterpreter with many other classes their depending on.
+Line one to seven creates a transaction builder which is then signed on line nine to thirteen.
+On line twelve we can now use our transaction object.
+For example, after calling the \emph{hex} function on it, we can send the returned string to the real bitcoin network.
 
-After this extraction Stainless was successfully integrated with both, sbt and JAR.
-This leads us to another finding as described in the next section.
+\section{Validation of a Transaction}
 
-\section{Error Reporting with sbt and JAR}
-// TODO https://github.com/epfl-lara/stainless/issues/484?
+Bitcoin-S-Core offers a function called \emph{checkTransaction}.
+This is its type signature.
+\begin{lstlisting}[language=scala]
+  checkTransaction(transaction: Transaction): Boolean
+\end{lstlisting}
+We can pass a transaction and it returns a Boolean whether the transaction is valid or not.
+So for example when we pass the built transaction from above to it the returned value would be true, because it's a valid transaction.
 
-\section{Bugfix}
+There are several checks in checkTransaction.
+For example, it checks if there is either no input or no output.
+In this case it returns false.
+
+The relevant part for the property we are trying to verify is the following.
+\begin{lstlisting}[language=scala]
+  val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
+  val noDuplicateInputs = prevOutputTxIds.distinct.size == prevOutputTxIds.size
+\end{lstlisting}
+
+It gathers all transaction ids referenced by the out points.
+When calling \emph{distinct} on the returned list, it returns a list where duplicated are removed.
+Is the size of the new list is the same as the size of the old, we know that there was no duplicate transaction id.
+
+There is a bug, because checkTransaction checks on the transaction id that we are going to explain a next section.
+
+\section{Process of the Verification // TODO change?}
+
+Trying to integrate Stainless in Bitcoin-S caused a lot of troubles, because of version conflicts.
+If you are interested in it you can read \hl{TODO} in the appendix.
+
+It takes too much time to do it and after some discussion, we decided to extract only the parts of the code that are needed for checkTransaction.
+
+The extracted code has more than 1500 lines.
+After running Stainless on it, it throws a really huge bunch of errors about what Stainless can not reason about.
+This would require to change nearly all the extracted code.
+
+We arrive at the decision that we need a smaller part to verify.
+We have chosen to verify the addition with zero of Bitcoin-Ss' CurrencyUnit class as described in the next chapter.
+
+During the examination of the checkTransaction function, we found a bug in Bitcoin-S.
+You can read more about it in the next section.
+
+\section{Fixing a Bug in Bitcoin-S}
+
 After so much digging and trying to understand the checkTransaction code, we found a bug even without Stainless.

### 2019-05-29 doc/chapters/trying_verify_checktx.tex
changed some sentences

@@
 We arrive at the decision that we need a smaller part to verify.
-We have chosen to verify the addition with zero of Bitcoin-Ss' CurrencyUnit class as described in the next chapter.
+We have chosen to verify the addition with zero of Bitcoin-Ss CurrencyUnit class as described in the next chapter.
 
-During the examination of the checkTransaction function, we found a bug in Bitcoin-S.
-You can read more about it in the next section.
+But before that you can read about the bug we found and fixed in the next section.
 
@@
 
-After so much digging and trying to understand the checkTransaction code, we found a bug even without Stainless.
+During the examination of the checkTransaction function, we found a bug in Bitcoin-S.
 

### 2019-05-30 code/currsniporig/src/main/scala/code/initial/InitialCode.scala
initial code

@@ -1,0 +1,103 @@
+package code.initial
+
+trait BasicArithmetic[N] {
+  def +(n: N): N
+}
+
+sealed abstract class Number[T <: Number[T]]
+  extends BasicArithmetic[T] {
+  type A = BigInt
+
+  protected def underlying: A
+
+  def toLong: Long = toBigInt.bigInteger.longValueExact()
+  def toBigInt: BigInt = underlying
+
+  def andMask: BigInt
+
+  def apply: A => T
+
+  override def +(num: T): T = apply(checkResult(underlying + num.underlying))
+
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+
+sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
+
+sealed abstract class Int64 extends SignedNumber[Int64] {
+  override def apply: A => Int64 = Int64(_)
+  override def andMask = 0xffffffffffffffffL
+}
+
+trait BaseNumbers[T] {
+  def zero: T
+  def one: T
+  def min: T
+  def max: T
+}
+
+object Int64 extends BaseNumbers[Int64] {
+  private case class Int64Impl(underlying: BigInt) extends Int64 {
+    require(underlying >= -9223372036854775808L,
+      "Number was too small for a int64, got: " + underlying)
+    require(underlying <= 9223372036854775807L,
+      "Number was too big for a int64, got: " + underlying)
+  }
+
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(-9223372036854775808L)
+  lazy val max = Int64(9223372036854775807L)
+
+  def apply(long: Long): Int64 = Int64(BigInt(long))
+
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+
+
+sealed abstract class CurrencyUnit
+  extends BasicArithmetic[CurrencyUnit] {
+  type A
+
+  def satoshis: Satoshis
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  override def +(c: CurrencyUnit): CurrencyUnit =
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+
+  def toBigDecimal: BigDecimal
+
+  protected def underlying: A
+}
+
[…]

### 2019-05-30 code/currsniporig/src/main/scala/code/initial/InitialCode.scala
indent

@@ -1,103 +1,102 @@
-package code.initial
+  package code.initial
 
-trait BasicArithmetic[N] {
-  def +(n: N): N
-}
-
-sealed abstract class Number[T <: Number[T]]
-  extends BasicArithmetic[T] {
-  type A = BigInt
-
-  protected def underlying: A
-
-  def toLong: Long = toBigInt.bigInteger.longValueExact()
-  def toBigInt: BigInt = underlying
-
-  def andMask: BigInt
-
-  def apply: A => T
-
-  override def +(num: T): T = apply(checkResult(underlying + num.underlying))
-
-  private def checkResult(result: BigInt): A = {
-    require((result & andMask) == result,
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}
-
-sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
-
-sealed abstract class Int64 extends SignedNumber[Int64] {
-  override def apply: A => Int64 = Int64(_)
-  override def andMask = 0xffffffffffffffffL
-}
-
-trait BaseNumbers[T] {
-  def zero: T
-  def one: T
-  def min: T
-  def max: T
-}
-
-object Int64 extends BaseNumbers[Int64] {
-  private case class Int64Impl(underlying: BigInt) extends Int64 {
-    require(underlying >= -9223372036854775808L,
-      "Number was too small for a int64, got: " + underlying)
-    require(underlying <= 9223372036854775807L,
-      "Number was too big for a int64, got: " + underlying)
+  trait BasicArithmetic[N] {
+    def +(n: N): N
   }
 
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
+  sealed abstract class Number[T <: Number[T]]
+    extends BasicArithmetic[T] {
+    type A = BigInt
 
-  lazy val min = Int64(-9223372036854775808L)
-  lazy val max = Int64(9223372036854775807L)
+    protected def underlying: A
 
-  def apply(long: Long): Int64 = Int64(BigInt(long))
+    def toLong: Long = toBigInt.bigInteger.longValueExact()
+    def toBigInt: BigInt = underlying
 
-  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
-}
+    def andMask: BigInt
+
+    def apply: A => T
+
+    override def +(num: T): T = apply(checkResult(underlying + num.underlying))
+
[…]

### 2019-05-30 doc/chapters/towards_verifying_addition.tex
fixed typo

@@
 \chapter{Towards Verifying Addition with Zero}
+After realizing, that it would consume to much time to rewrite the whole Bitcoin-S code
 After so many failures, we have decided to search for the smallest unit in Bitcoin-S-Core that is worthwhile to verify.
@@
 To make it even easier, we decided to verify only the addition, where we add zero to an amount of Satoshis.
+
+The initial code is:
+\lstinputlisting[language=scala]{../code/currsniporig/src/main/scala/code/initial/Code.scala}
+
+When we run Stainless on it it throws the following errors:
+{\footnotesize\begin{verbatim}
+  [ Error  ] Code.scala:43:1: Objects cannot extend classes or implement traits,
+             use a case object instead
+           object Int64 extends BaseNumbers[Int64] {
+           ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+  [ Error  ] Code.scala:65:3: Stainless doesn't support abstract type members
+              type A
+              ^^^^^^
+  [ Error  ] Code.scala:86:33: Only literal arguments are allowed for BigInt.
+              def toBigInt: BigInt = BigInt(toLong)
+                                            ^^^^^^
+  [ Error  ] Code.scala:93:1: Objects cannot extend classes or implement traits,
+             use a case object instead
+            object Satoshis extends BaseNumbers[Satoshis] {
+            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+  [  Info  ] Shutting down executor service.
+\end{verbatim}}
+
 The signature looks like the following.
@@
   sealed abstract class CurrencyUnit {
+    type A
+
     def +(c: CurrencyUnit): CurrencyUnit =
       Satoshis(satoshis.underlying + c.satoshis.underlying)
+
+    protected def underlying: A
   }

### 2019-05-30 code/currsniporig/src/main/scala/code/initial/Code.scala

@@ -36,8 +36,8 @@
   trait BaseNumbers[T] {
     def zero: T
     def one: T
-    def min: T
-    def max: T
+    def min: T //maybe remove
+    def max: T //maybe remove
   }
 
   object Int64 extends BaseNumbers[Int64] {
@@ -51,8 +51,8 @@
     lazy val zero = Int64(0)
     lazy val one = Int64(1)
 
-    lazy val min = Int64(-9223372036854775808L)
-    lazy val max = Int64(9223372036854775807L)
+    lazy val min = Int64(-9223372036854775808L) //maybe remove
+    lazy val max = Int64(9223372036854775807L) //maybe remove
 
     def apply(long: Long): Int64 = Int64(BigInt(long))
 
@@ -91,8 +91,8 @@
   }
 
   object Satoshis extends BaseNumbers[Satoshis] {
-    val min = Satoshis(Int64.min)
-    val max = Satoshis(Int64.max)
+    val min = Satoshis(Int64.min) //maybe remove
+    val max = Satoshis(Int64.max) //maybe remove
     val zero = Satoshis(Int64.zero)
     val one = Satoshis(Int64.one)
 

### 2019-05-30 code/currsniporig/src/main/scala/code/initial/Code.scala
removed more

@@ -5,7 +5,7 @@
   }
 
   sealed abstract class Number[T <: Number[T]]
-    extends BasicArithmetic[T] {
+      extends BasicArithmetic[T] {
     type A = BigInt
 
     protected def underlying: A
@@ -36,8 +36,8 @@
   trait BaseNumbers[T] {
     def zero: T
     def one: T
-    def min: T //maybe remove
-    def max: T //maybe remove
+//    def min: T //maybe remove
+//    def max: T //maybe remove
   }
 
   object Int64 extends BaseNumbers[Int64] {
@@ -51,8 +51,8 @@
     lazy val zero = Int64(0)
     lazy val one = Int64(1)
 
-    lazy val min = Int64(-9223372036854775808L) //maybe remove
-    lazy val max = Int64(9223372036854775807L) //maybe remove
+//    lazy val min = Int64(-9223372036854775808L) //maybe remove
+//    lazy val max = Int64(9223372036854775807L) //maybe remove
 
     def apply(long: Long): Int64 = Int64(BigInt(long))
 
@@ -71,8 +71,6 @@
     override def +(c: CurrencyUnit): CurrencyUnit =
       Satoshis(satoshis.underlying + c.satoshis.underlying)
 
-    def toBigDecimal: BigDecimal
-
     protected def underlying: A
   }
 
@@ -80,8 +78,6 @@
     override type A = Int64
 
     override def satoshis: Satoshis = this
-
-    override def toBigDecimal = BigDecimal(toBigInt)
 
     def toBigInt: BigInt = BigInt(toLong)
 
@@ -91,8 +87,8 @@
   }
 
   object Satoshis extends BaseNumbers[Satoshis] {
-    val min = Satoshis(Int64.min) //maybe remove
-    val max = Satoshis(Int64.max) //maybe remove
+//    val min = Satoshis(Int64.min) //maybe remove
+//    val max = Satoshis(Int64.max) //maybe remove
     val zero = Satoshis(Int64.zero)
     val one = Satoshis(Int64.one)
 

### 2019-05-30 doc/chapters/towards_verifying_addition.tex
text

@@
 \chapter{Towards Verifying Addition with Zero}
-After realizing, that it would consume to much time to rewrite the whole Bitcoin-S code
-After so many failures, we have decided to search for the smallest unit in Bitcoin-S-Core that is worthwhile to verify.
+After realizing, that it would consume to much time to rewrite the Bitcoin-S code and even the extracted part with checkTransaction, we decided to extract the smallest unit in Bitcoin-S-Core that is worthwhile to verify.
 We found the addition of two Satoshis would be a good candidate.
-To make it even easier, we decided to verify only the addition, where we add zero to an amount of Satoshis.
+To make it even easier, we decided to verify only the addition of some satoshis with zero.
 
-The initial code is:
+The initial code is this:
 \lstinputlisting[language=scala]{../code/currsniporig/src/main/scala/code/initial/Code.scala}
 
-When we run Stainless on it it throws the following errors:
+The additions signature looks like this:
+\begin{lstlisting}[language=scala]
+  +(c: CurrencyUnit): CurrencyUnit
+\end{lstlisting}
+
+When we run Stainless on it, it throws the following errors:
 {\footnotesize\begin{verbatim}
@@
 
-The signature looks like the following.
-\begin{lstlisting}[language=scala]
-  +(c: CurrencyUnit): CurrencyUnit
-\end{lstlisting}
+So let's see how we can fix those errors.
 
-Where the class Satoshis extends CurrencyUnit.
-The post condition would be, that the parameter c must be zero.
-\begin{lstlisting}[language=scala]
-  require(c.satoshis.underlying == Int64.zero)
-\end{lstlisting}
+\section{Turning object Into case object}
 
-\emph{c.satoshis} is an abstract method of CurrencyUnit that must return an instance of the class Satoshis.
+Stainless output:
+{\footnotesize\begin{verbatim}
+  [ Error  ] Code.scala:43:1: Objects cannot extend classes or implement traits,
+             use a case object instead
+           object Int64 extends BaseNumbers[Int64] {
+           ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+  [ Error  ] Code.scala:93:1: Objects cannot extend classes or implement traits,
+            use a case object instead
+            object Satoshis extends BaseNumbers[Satoshis] {
+            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+\end{verbatim}}
 
-\emph{c.satoshis.underlying} is also an abstract method of CurrencyUnit that must return an instance of the abstract type A.
+Here, we can just change the objects from object to case object.
+Stainless recommendation is to use objects for modules and case objects as algebraic data type.
 
-Both are implemented in Satoshis where A is set to Int64.
-So the underlying number of the parameter must be zero.
-
-To ensure that the result is the same value as \emph{this}:
-\begin{lstlisting}[language=scala]
-  ensuring(res => res.satoshis == this.satoshis)
-\end{lstlisting}
-
-Here, equals (==) can be used directly on Satoshis, because it is a case class and Int64, the only parameter of Satoshis is a case class too.
-
-After writing the verification, running Stainless on the source files was the next step.
-It prints a lot of errors because of code incompatibility.
-So let's describe the code rewriting in the following sections.
+This is, due to the internal design of Scala and Java.
+It's possible to reason about case object but not about object.
+This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
+The issue \mypound520 on Stainless GitHub gives some thoughts if you want to know more.
 
 \section{Get Rid of Abstract Type Member}
+
 Stainless output:
 {\footnotesize\begin{verbatim}
-[ Error  ] CurrencyUnits.scala:5:3: Stainless doesn't support abstract type members
+  [ Error  ] Code.scala:65:3: Stainless doesn't support abstract type members
               type A
+              ^^^^^^
 \end{verbatim}}
+
 This should be easy to rewrite by using generics instead of an abstract type, right?
 Unfortunately not.
-The problem is, CurrencyUnit uses its implementing class Satoshis.
+The problem is, CurrencyUnit uses one of its implementing class Satoshis.
 
@@
 
-Thus, changing it to generics, Satoshis is not of type CurrencyUnit anymore and neither of type CurrencyUnit[A], where A is the new generic type.
+What happens, if we typify CurrencyUnit with A, meaning to make it generic with type A?
+
 Satoshis extends CurrencyUnit with type Int64, so it is of type CurrencyUnit[Int64].
-That's too specific.
+That's too specific, because the addition would require CurrencyUnits of type CurrencyUnit[A] not CurrencyUnit[Int64].
 
-Since there is no easy way to fix it and the code should stay as much as possible the original, removing the abstract type might be the only choice.
-The verification is a bit more limited, to only Satoshis, but that's no problem, because the goal is to verify the addition of Satoshis.
+Since there is no easy way to fix it and the code should stay as much as possible the original, we just remove the abstract type and fixed it to Int64.
+This limits the verification a bit, but as we only want to verify the addition in satoshis, thats OK.
 
-So by removing the abstract type and fix it to Int64 as defined by Satoshis, there was one error less in Stainless.
+\section{Replace BigInt Constructor Argument With String Literal}
+Stainless output:
+{\footnotesize\begin{verbatim}
+  [ Error  ] Code.scala:86:33: Only literal arguments are allowed for BigInt.
+              def toBigInt: BigInt = BigInt(toLong)
+                                            ^^^^^^
+\end{verbatim}}
+
+As described before, Stainless supports only a subset of Scala.
+The BigInt from the Stainless library is a bit restricted.
+It does not support dynamic BigInt construction.
+Only from string literals.
+
+Again, a simplified code version.
+\begin{lstlisting}[language=scala]
+  sealed abstract class Satoshis extends CurrencyUnit { 
+    def toBigInt: BigInt = BigInt(toLong)
+    def toLong: Long = underlying.toLong
+  }
+\end{lstlisting}
+
+This would be really hard to refactor, because Bitcoin-S tries to be as much dynamic as possible so it can be used with other cryptocurrencies than bitcoins too.
+Maybe it would even be impossible, because they need to parse a lot from the bitcoin network.
+
+Luckily, we can use toBigInt on underlying directly instead of going over toLong.
+
+After fixing all Stainless errors, a new error appears.
 
 \section{Get Rid of Special Generics}
+
 Stainless output:
 {\footnotesize\begin{verbatim}
-  [ Error  ] NumberType.scala:3:30: Unknown type parameter type T
+  [ Error  ] Code.scala:7:30: Unknown type parameter type T
              sealed abstract class Number[T <: Number[T]]
+                                         ^^^^^^^^^^^^^^
 \end{verbatim}}
@@
 \end{lstlisting}
-For the sake of convenience this is also made static by parameterize T with Int64.
 
-\section{Turning object Into case object}
+For the sake of convenience we make this a concrete type by replacing T with Int64.
+
+Now, there are two new errors.
+
+\section{Get Rid of Concret Type}
+
 Stainless output:
 {\footnotesize\begin{verbatim}
-[ Error  ] CurrencyUnits.scala:54:1: Objects cannot extend classes or implement
-           traits, use a case object instead
-           object Satoshis extends BaseNumbers[Satoshis] {
+  [Warning ] Code.scala:9:3: Could not extract tree in class: type A =
+             BigInt (class scala.reflect.internal.Trees$TypeDef)
+             type A = BigInt
+             ^^^^^^^^^^^^^^^
 \end{verbatim}}
 
-Changing the objects that extends classes to case objects solves this problem.
-This is, due to the internal design of Scala and Java.
-It's possible to reason about a case object but not about an object.
-This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
-The issue \mypound520 on Stainless GitHub gives some thoughts about this.
+This is easy.
+We just replace all occurence of A with BigInt, since A is not overrided in an implementing class.
+This is not exact the same code, because an implementing class could not override A anymore but thats fine for our verification.
 
-\section{Replace BigInt Constructor Argument With String Literal}
+This was a missing feature of Stainless that was fixed on Mai 28, 2019 with pull request \mypound470.
+
+\section{Replace BigInt \&-Function With Bound Check}
+
 Stainless output:
 {\footnotesize\begin{verbatim}
-[ Error  ] CurrencyUnits.scala:47:33: Only literal arguments are allowed for BigInt.
-           def toBigInt: BigInt = BigInt(toLong)
+  [ Error  ] Code.scala:22:14: Unknown call to & on result (BigInt) with arguments
+             List(Number.this.andMask) of type List(BigInt)
+            require((result & andMask) == result,
+                    ^^^^^^^^^^^^^^^^
 \end{verbatim}}
 
-As described before, Stainless supports only a subset of Scala.
-They provide their own dedicated BigInt library.
-This library does not have support for dynamic BigInt construction only for string literals.
-
-Again, a simplified version.
-\begin{lstlisting}[language=scala]
-  sealed abstract class Satoshis extends CurrencyUnit {
-    def toBigInt: BigInt = BigInt(toLong)
-    def toLong: Long = underlying.toLong
-  }
-
-  object Int64 extends BaseNumbers[Int64] {
-    private case class Int64Impl(underlying: BigInt) extends Int64
-  }
-\end{lstlisting}
-
-This would be really hard to refactor, because Bitcoin-S-Core tries to be as much dynamic as possible so it can be used with other cryptocurrencies too.
-Maybe it would even be impossible, because they need to parse a lot from the bitcoin network.
-
-Restricting the code one more time helps.
-\emph{toBigInt} can directly return underlying, because in case of Int64, which is what Satoshis use, it is a BigInt.
-So it does not be converted from BigInt to Long and back to BigInt.
-
-\section{Extracting Private Inner Classes}
-Stainless output (simplified):
-{\footnotesize\begin{verbatim}
-  [Warning ] CurrencyUnits.scala:36:3: Could not extract tree in class:
-             case private class SatoshisImpl extends Satoshis
-\end{verbatim}}
-
-Stainless is not able to extract private classes inside other objects.
-Bitcoin-S-core uses it a lot, because they separate the class from its implementation.
-\begin{lstlisting}[language=scala]
-  object Int64 extends BaseNumbers[Int64] {
-    private case class Int64Impl(underlying: BigInt) extends Int64 
-  }
-\end{lstlisting}
-
-This is an easy one.
-Just extract the inner class out of the object.
-This is not exactly the same code but since the class is still private it can not be extended outside of the file.
-
-\section{Get Rid of Concrete Type Member}
-Stainless output:
-{\footnotesize\begin{verbatim}
-  [Warning ] NumberType.scala:5:3: Could not extract tree in class: type A
-             = BigInt (class scala.reflect.internal.Trees$TypeDef)
-             type A = BigInt
-\end{verbatim}}
-
-Since the type A is not overridden in any of its subclasses, this can simply be hard coded.
-\begin{lstlisting}[language=scala]
-  sealed abstract class Number {
-    type A = BigInt
-    protected def underlying: A
-    def apply: A => T
-  }
-\end{lstlisting}
-
-Replace every occurrence of A with BigInt.
-This might be a missing feature in Stainless that should not be too hard to fix.
-There is an open pull request \mypound470 on GitHub for this issue.
-
-\section{Replace BigInt \&-Function With Bound Check}
-Stainless output:
-{\footnotesize\begin{verbatim}
-  [ Error  ] NumberType.scala:19:14: Unknown call to & on result (BigInt) with
-             arguments List(Number.this.andMask) of type List(BigInt)
-             require((result & andMask) == result,
-\end{verbatim}}
-
-Due to the restrictions on BigInt, the \& function on it is not supported too.
+Due to the restrictions on BigInt, we can not use the \& function on.
+Simplified code:
 \begin{lstlisting}[language=scala]
@@
 This is a bound check.
-It checks, if the result of the addition is in range of the specified type (Int64 here).
-So here the \& mask can be replaced with a bound check whether the result is in range of Long.MinValue and Long.MaxValue.
-Again the code gets a bit more static but for this use case it is OK.
+It checks, if the result of the addition is in range of the specified type, which is now the hard coded Int64.
 
-\section{Remove Second Parameter of require}
-Stainless output:
-{\footnotesize\begin{verbatim}
-  [Warning ] NumberType.scala:54:3: Could not extract tree in class:
-             scala.this.Predef.require(Int64Impl.this.underlying.<=(
-             math.this.BigInt.long2bigInt(9223372036854775807L)),
-             "Number was too big for a int64, got: ".+(Int64Impl.this.underlying))
-             (class scala.reflect.internal.Trees$Apply)
-             require(underlying <= 9223372036854775807L,
-\end{verbatim}}
+So, we can replace the \& mask with a bound check whether the result is in range of Long.MinValue and Long.MaxValue.
+Again the code gets a bit more static and it's not the exact same code anymore.
 
-This is, because Stainless does not support require with a second String parameter.
-Removing this String fixes the error.
+Running Stainless produces again new errors.
 
-\section{Propagate require}
-Finally, the code works with Stainless.
-But there is another problem.
-Bitcoin-S-Core uses require like a fail-fast method whereas Stainless needs it to verify the code.
+\section{Extracting Private Inner Classes}
 
@@
 {\footnotesize\begin{verbatim}
+  [Warning ] Code.scala:90:3: Could not extract tree in class: case private
+             class SatoshisImpl extends Satoshis with Product with Serializable {
+\end{verbatim}}
+
+Stainless can not extract the private class inside the object.
+Bitcoin-S uses this a lot, because they separate the class from its implementation.
+Simplified:
+\begin{lstlisting}[language=scala]
+  object Int64 extends BaseNumbers[Int64] {
+    private case class Int64Impl(underlying: BigInt) extends Int64 
+  }
+\end{lstlisting}
+
+This is an easy to fix.
+We just extract the private class out of the object.
+This is not exactly the same code but since the class is still private it can not be extended outside of the file.
+
+Now we get som weird warnings about require.
+
+\section{Remove Second Parameter of require}
+
+Stainless output:
+{\footnotesize\begin{verbatim}
+  [Warning ] Code.scala:51:3: Could not extract tree in class: scala.this.Predef
+             .require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(
+             -9223372036854775808L)), "Number was too small for a int64, got: "
+             .+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
+            require(underlying >= -9223372036854775808L,
+            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+  [Warning ] Code.scala:53:3: Could not extract tree in class: scala.this.Predef
+             .require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(
+             9223372036854775807L)), "Number was too big for a int64, got: "
+             .+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
+            require(underlying <= 9223372036854775807L,
+            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+  [ Error  ] checkResult$0 depends on missing dependencies: require$1.
+\end{verbatim}}
+
+Seems like Stainless does not support the second string parameter of require or at least it throws a warning about it.
+We can saftly remove it.
+
+A new error apperas.
+
+\section{Rewrite BigInt Comparison with Long}
+
+Stainless output:
+{\footnotesize\begin{verbatim}
+  [ Error  ] inv$4 depends on missing dependencies: long2bigInt$0.
+\end{verbatim}}
+
+This error is hard to understand but we can see that there is a missing Long to BigInt conversion.
+So we search for all longs in the code.
+\begin{lstlisting}[language=scala]
+  private case class Int64Impl(underlying: BigInt) extends Int64 {
+    require(underlying >= -9223372036854775808L)
+    require(underlying <= 9223372036854775807L)
+  }
+\end{lstlisting}
+
+Looks like it can not compare a BigInt with a Long value.
+We can easily convert this Long value to a BigInt with string literals parameter.
+
+Finally, we get some output from Stainless about the verification in the code.
+{\footnotesize\begin{verbatim}
+[...]
+[Warning ]  => INVALID
 [Warning ] Found counter-example:
 [Warning ]   thiss: { x: Object | @unchecked isNumber(x) } -> Int64Impl(0)
-  [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> 
-                 Int64Impl(9223372036854775808)
+[Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(9223372036854775808)
+[...]
+[  Info  ] +  precond. (call checkResult(thiss, underlying(thiss) + u ...)  invalid  Code.scala:16:45
+[...]
+[  Info  ] total: 5  valid: 4  (4 from cache) invalid: 1  unknown: 0  time:  1.317
+\end{verbatim}}
+
+This show, that there is an invalid specification in checkResult and Stainless prints a counter example for it.
+
+Let's ignore this for a moment and write the verification for our addition with zero.
+
+\section{Write Specification for Verification}
+
+As specified, our verification must only support addition with zero.
+So we restrict the parameter to be zero in the precondition.
+\begin{lstlisting}[language=scala]
+  require(c.satoshis == Satoshis.zero)
+\end{lstlisting}
+
+We ensure the result is the same value as \emph{this} in the postcondition.
+\begin{lstlisting}[language=scala]
+  ensuring(res => res.satoshis == this.satoshis)
+\end{lstlisting}
+
+We can use equals (==) directly on Satoshis, because it is a case class.
+The addition will now look like this:
+\begin{lstlisting}[language=scala]
+  override def +(c: CurrencyUnit): CurrencyUnit = {
+    require(c.satoshis == Satoshis.zero)
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  } ensuring(res => res.satoshis == this.satoshis)
+\end{lstlisting}
+
+That's all we need to verify our addition.
+
+Now we will look into the previous error.
+
+\section{Propagate require}
+
+There is another problem with Bitcoin-S.
+Bitcoin-S-Core uses require as a fail-fast method whereas Stainless needs it to verify the code.
+
+The Stainless output again:
+{\footnotesize\begin{verbatim}
+  [Warning ] Found counter-example:
+  [Warning ]   thiss: { x: Object | @unchecked isNumber(x) } -> Int64Impl(0)
+  [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(9223372036854775808)
 \end{verbatim}}
@@
 
-But how can Stainless find a counter example ignoring the require in checkResult?
+But how does Stainless find a counter example ignoring the require in checkResult?
 Since Stainless is a static verification tool, it tests every possibility.
 So it can use a number bigger than the maximum Int64 and pass it to the addition.
-The require in checkResult will then fail.
-Thus, the addition need to have the restrictions of checkResult too.
+The require in checkResult fails.
+
+Thus, we need to add the restriction from checkResult to the addition too.
 \begin{lstlisting}[language=scala]
@@
 
-Sure, it can just add one to the maximum Int64 and the require does not hold anymore.
-The goal is to check the addition of zero so let's add a restriction for num to be zero.
+Sure, when adding one to the maximum Int64 the require does not hold anymore.
+Since we do only allow zero as a parameter the easiest way is to restrict it to zero here too.
 
-Finally, it works and verifies correctly.
-
-\section{Result}
+Finally, everything is green and correctly verified.
 \begin{figure}[H]
@@
 \end{figure}
+
+But is it really the original code that we verified?
+And why did we have to change that much?
+This and other questions will be answered in the next chapter.

### 2019-05-30 doc/chapters/trying_verify_checktx.tex
added image
added comment image
fixed typo
chapter description

@@
 \label{chap:connecting}
-// TODO
+
+This chapter describes some parts of Bitcoin-S.
+We are going to create a transaction and show you the relevant parts of checkTransaction, where transactions are checked against some properties.
+Then we see some troubles occurred during this work.
+In the end you will see how we fixed a bug in Bitcoin-S.
 
@@
 
-We arrive at the decision that we need a smaller part to verify.
+We arrived at the decision that we need a smaller part to verify.
 We have chosen to verify the addition with zero of Bitcoin-Ss CurrencyUnit class as described in the next chapter.
@@
 Since TransactionOutPoint is a case class and Scala has a built in == for case classes there is no need to implement TransactionOutPoint.==.
+\begin{figure}[H]
+	\centering
+		\includegraphics[scale=0.396]{images/bitcoin-s-pr.png}
+	\caption{Line changes for PR \#435 from GitHub}
+	\label{fig:output1}
+\end{figure}
 
 This was fixed in pull request \mypound435 on GitHub at April 23, 2019 by us along a unit test to prevent this bug from appearing again in the future.
+
+\begin{figure}[H]
+	\centering
+		\includegraphics[scale=0.5]{images/bitcoin-s-pr-comment.png}
+	\caption{Nice catch comment on PR \#435 from GitHub}
+	\label{fig:output1}
+\end{figure}

### 2019-05-30 doc/appendix/appendixexemple.tex
added some labels

@@
-\chapter{Arbitrary Appendix}
+\chapter{Practical Challenges with Stainless}
 \label{chap:appendix_arb}
@@
 
-The European languages are members of the same family.
-Their separate existence is a myth.
-For science, music, sport, etc, Europe uses the same vocabulary.
-The languages only differ in their grammar, their pronunciation and their most common words.
-Everyone realizes why a new common language would be desirable: one could refuse to pay expensive translators.
-To achieve this, it would be necessary to have uniform grammar, pronunciation and more common words.
-If several languages coalesce, the grammar of the resulting language is more simple and regular than that of the individual languages.
-The new common language will be more simple and regular than the existing European languages.
-It will be as simple as Occidental; in fact, it will be Occidental.
+\section{Integration}
+
+During this work, Stainless bumped the sbt plugin from version 0.13.17 to 1.2.8 and Scala from 2.11.12 to 2.12.8.
+So this section might be out of date now.
+
+Bitcoin-S-Core uses sbt 1.2.8 and Scala 2.12.8, while Stainless sbt plugin is on sbt 0.13.17 and Scala 2.11.12.
+
+Sbt introduced new features in the 1.x release used by Bitcoin-S.
+Most of them can be written the sbt 0.13.17 way.
+
+The bigger problem is, due to the different Scala and sbt versions, the error message, after trying to go in a sbt shell.
+\begin{verbatim}
+  [warn] There may be incompatibilities among your library dependencies; run 'evicted'
+         to see detailed eviction warnings.
+  [error] java.lang.NoClassDefFoundError: sbt/SourcePosition
+  ...
+  Project loading failed: (r)etry, (q)uit, (l)ast, or (i)gnore?
+\end{verbatim}
+
+Downgrading Bitcoin-S sbt version to 0.13.17 fixes the error but then it can not load some libraries only compiled for newer versions.
+So this would take too much time to fix and changes the Bitcoin-S code unwantingly.
+
+The next approach is to use stainless cli instead of sbt.
+Running stainless on all source files does not work, because the dependencies are missing.
+The parameter \emph{-classpath} can resolve it but the value of this parameter must be the paths to all the dependencies separated by a ':'.
+Finally, \emph{core} depends on \emph{secp256k1jni}, another package of Bitcoin-S written in Java.
+So this needs to be in the source files to.
+
+The final command looks like this in \emph{core} folder of Bitcoin-S:
+\begin{lstlisting}[language=bash]
+  $ stainless
+    -classpath ".:$(find ~/.ivy2/ -type f -name *.jar | tr '\n' ':')"
+    $(find . -type f -name *.scala | tr '\n' ' ')
+    $(find ../secp256k1jni -type f -name *.java | tr '\n' ' ')
+\end{lstlisting}
+
+The output was some Stainless error.
+\begin{verbatim}
+  [Internal] Error: object scala.reflect.macros.internal.macroImpl in compiler mirror
+             not found.. Trace:
+  [Internal] - scala.reflect.internal.MissingRequirementError$.signal
+             (MissingRequirementError.scala:17)
+  ...
+  [Internal] object scala.reflect.macros.internal.macroImpl in compiler mirror not found.
+  [Internal] Please inform the authors of Inox about this message
+\end{verbatim}
+
+After some discussion, we decided to go another way, because the errors may take too much time and it might lead to a next error.
+Let's extract the code needed to verify a transaction mainly the class Transaction and ScriptInterpreter with many other classes their depending on.
+
+After this extraction Stainless was successfully integrated with both, sbt and JAR.
+This leads us to another finding as described in the next section.
+
+\section{Error Reporting with sbt and JAR}
+
+Running \code{sbt compile} in the project with Stainless ended without error.
+But it also ended with no output.
+So we were not able to change the code so Stainless would accept it.
+
+After running the JAR file on the code there were many errors as described in chapter \ref{chap:verify_add}.
+
+So the sbt plugin does not complain always but the JAR file did.
+The open issue \mypound484 on GitHub might describe this error.

### 2019-05-30 doc/chapters/towards_verifying_addition.tex
added section result
added some labels

@@
 \chapter{Towards Verifying Addition with Zero}
+\label{chap:verify_add}
+
 After realizing, that it would consume to much time to rewrite the Bitcoin-S code and even the extracted part with checkTransaction, we decided to extract the smallest unit in Bitcoin-S-Core that is worthwhile to verify.
@@
 
+\section{Result}
+
 Finally, everything is green and correctly verified.

### 2019-05-30 doc/appendix/appendixexemple.tex
moved text to appendix

@@
 So this section might be out of date now.
+
+There are two ways to integrate Stainless in a Scala project, the Scala build tool (sbt) plugin or a command line tool.
+Both, when run, analyses the code passed and report warnings to the console about the given code.
+Stainless requires and Scala recommends Java SE Development Kit 8.
+Newer Java versions won't work.
+
+To use the latest version of the sbt tool you have to build it locally.
+You can run \code{sbt universal:stage} in the cloned Stainless git repository.
+This generates \emph{frontends/scalac/target/universal/stage/bin/stainless-scalac}.
 

### 2019-05-30 doc/chapters/using_stainless.tex
moved text to appendix

@@
-\chapter{Using Stainless}
+\chapter{Using Stainless // TODO remove this chapter}
 \label{chap:using_stainless}
-This chapter describes the setup and integration process of Stainless in a new or already existing project.
-It also shows the compatibility of Stainless with Scala as Stainless supports only a purely functional subset of Scala which they call \emph{Pure Scala}.
-
-\section{Configuration}
-There are two ways to integrate Stainless in a Scala project, the Scala build tool (sbt) plugin or a command line tool.
-Both, when run, analyses the code passed and report warnings to the console about the given code.
-Stainless requires and Scala recommends Java SE Development Kit 8.
-Newer Java versions won't compile.
-
-\subsection{sbt}
-sbt for Scala is like gradle or maven for Java.
-It can compile Scala code continuously or manual, manage dependencies with support for Maven-formatted repositories, mixing Scala and Java projects and much more.
-
-A simple sbt project has the following structure:
-\dirtree{%
-  .1 ..
-  .2 project.
-  .3 build.properties.
-  .3 plugins.sbt.
-  .2 src.
-  .3 main.
-  .4 scala.
-  .5 Main.scala.
-  .2 build.sbt.
-}
-\emph{build.properties} specifies the sbt version used for this project.
-If the version is not available locally, sbt will download it.
-
-In \emph{plugins.sbt} new sbt plugins can be added.
-A plugin extends the build definition.
-Mostly this means adding and overriding settings.
-
-\emph{build.sbt} defines the build definition.
-There can be several projects or subprojects as sbt doc calls it.
-
-Here an example for a single project in \emph{build.sbt}:
-\begin{lstlisting}[language=scala]
-scalaVersion := "2.12.8"
-
-lazy val root = (project in file("."))
-\end{lstlisting}
-The project is called root and its source files are located in the files' directory.
-Executing \code{sbt compile} should now compile the code.
-
-The Stainless web page has a guide on how to integrate Stainless in an existing project.
-The simplifies steps are:
-\begin{itemize}
-  \item Install an external solver.
-  \item Add Stainless sbt plugin to \emph{plugins.sbt}
-  \item Enable the plugin in \emph{build.sbt} for the project.
-\end{itemize}
-After this setup, Stainless will report errors to the console, when running \code{sbt compile}.
-
-\subsection{Command Line Tool}
-There are two ways to use the command line tool.
-
-Either download a prebuilt JAR file from efpl-lara/stainless GitHub repository or built a binary from source.
-Prebuild versions are released by Stainless.
-The latest was released on January 14, when there was no support for Scala 2.12.
-The 'Bump Scala to 2.12.8' branch was merged on March 4.
-
-If latest features are needed, like support for Scala 2.12, the build from source is required.
-Here a short installation description.
-Full description can be found on the Stainless documentation pages.
-\begin{itemize}
-  \item Install sbt.
-  \item Check out GitHub repository.
-  \item Run \code{sbt universal:stage} inside the project.
-\end{itemize}
-This generates \emph{frontends/scalac/target/universal/stage/bin/stainless-scalac}.
-
-To check the source code with one of those either \code{java -jar downloaded.jar source.scala} or \code{stainless-scalac source.scala} must be invoked.
-The file \emph{source.scala} is the file to be checked.
-
-As compiling without a build tool, this command will become really complex for bigger projects.
-All dependencies must be on the classpath and all source files appended.
-Those are added with \code{-classpath Dep1.jar:Dep2.jar:...:DepN.jar src1.scala src2.scala ... srcM.scala}.
 

### 2019-06-02 doc/chapters/introduction.tex
Updated introdunction.

@@
-\chapter{Introduction (Aufgabestellung)}
+\chapter{Introduction}
 \label{chap:introduction}
-In this work some experiments in formal verification of Scala code will be accomplished. 
-The framework Stainless is used as a verification tool. The code of Bitcoin-S-Core, Scala implementation of Bitcoin protocol, is taken as an input for Stainless to be verified. 
-In following the main aspects of formal verification, Stainless and Bitcoin-S are described.
+In this work we experiment in formal verification of Scala code. 
+We use the verification framework Stainless to verify the code of Bitcoin-S-Core, the Scala implementation of the Bitcoin protocol.
+In the following we describe the main aspects of formal verification, Stainless and the project Bitcoin-S.
 
-\section{Formal verification (formal verification vs. testing)}
+\section{What is formal verification}
 \label{sec:formal_verification}
 
-The longer and complexer a source code is, the more difficult it is to verify its correctness.
-There are different approaches for verification of program correctness.
+The longer and more complex a source code is, the more difficult it is to verify its correctness.
+There are different approaches for verification of the program correctness.
 
 The commonly used one is testing.
-A set of some practical scenarios and assertions is created by a developer to test a software design and implementation. 
-While testing, it is checked whether the actual results match the expected results.
+A developer creates a set of some practical scenarios and assertions to test a software design and implementation. 
+So, the actual results are checked to match the expected results.
 Tests help to identify bugs and missing requirements.
-But it is not achievable to test all possible inputs. 
-Thus, a long-term test coverage model should be created to test a software design enough.
-Furthermore, it is challenging to test a software keeping the ability to observe the side effects of a specific part of code.
-Practically, a developer runs tests, debugs appeared failures, adjusts a code of a software and extends a test bench verifying previously uncovered aspects. \cite{sanghavi:formal_verification}
+However, not all possible inputs can be tested. 
+Therefore, a long-term test coverage model is usually created to test a software design enough.
+Nevertheless, this is still not easy to test a software being able to observe the side effects of a code.
+\textbf{Practically, a developer runs tests, debugs appeared failures, adjusts a code and extends a test bench to verify the previously uncovered aspects.
+(TODO// do we need this sentence?)}\cite{sanghavi:formal_verification}
 
@@
 Correctness of a program is analyzed relative to its formal specification.
-Formal specification is a mathematical description of a software behavior that can be provided to formal verification tools for proving it. \cite{sanghavi:formal_verification}
+Formal specification is a mathematical description of a software behavior that can be provided to the formal verification tool for proving it.\cite{sanghavi:formal_verification}
 
-One of such tools is Stainless.
-This framework is used in the work to verify a Scala code.
+One of such tools is Stainless which we use in our work.
 
-\section{Stainless (allgemeine Beschreibung: Struktur with pre- and postconditions, Solvers, Terminierung und mögliche Outputs)}
+\section{Overview of Stainless}
 \label{sec:stainless}
 
-Stainless is a framework developed by "Lab for Automated Reasoning and Analysis" (LARA) at EPFL's School of Computer and Communication Sciences.
-The framework is used to verify Scala programs.
-It verifies statically that a program satisfies a specification given by a developer and that a program will not crash at runtime.
-Stainless explores all possible input values, reports inputs for which a program fails and demonstrates counterexamples which violate a given specification.\cite{Stainless:introduction}
+Stainless is developed by "Lab for Automated Reasoning and Analysis" (LARA) at EPFL's School of Computer and Communication Sciences.
+Using this framework we can verify the correctness of Scala programs.
+Stainless statically verifies that a program satisfies a given specification and that a program will not crash at runtime.
+The framework explores all possible input values, reports inputs for which a program fails and demonstrates counterexamples which violate a given specification.\cite{Stainless:introduction}
 
 The main functions used to write a specification are \textit{require} and \textit{ensuring}. 
-A precondition should be written at the beginning of a function body with \textit{require}.
-Its argument is a Boolean expression which corresponds to constraint for inputs of a function being verified.
-A postcondition should be written after a function body with \textit{ensuring}.
-It is a verification condition on an output of a function.
-While compiling Stainless tries to prove that the postcondition always holds, assuming a given precondition does hold.\cite{Stainless:introduction}
+With \textit{require} we define a precondition of a function, which we want to verify.
+\textit{require} is called at the beginning of the function body.
+A constraint for inputs to the function is passed as an argument to \textit{require} in the form of a Boolean expression.
+With \textit{ensuring} we provide a postcondition which is a condition on an output of the function.
+\textit{ensuring} is called after the function body.
+While compiling Stainless tries to prove that the postcondition always holds, assuming the given precondition does hold.\cite{Stainless:introduction}
 
-3 outcomes of verification with Stainless are possible: valid, invalid and unknown. 
-If the postcondition is valid, Stainless could prove that for any inputs constrained in a precondition, the postcondition always holds.
-With invalid postcondition the framework could find at least one counterexample which satisfies a precondition but violates a postcondition.
-Also, an output unknown is possible when Stainless is unable to prove a postcondition or find a counterexample.
-In this case a timeout or an internal error occurred.
-Furthermore, it will be verified by Stainless that a precondition cannot be violated.\cite{Stainless:introduction}
+The following example demonstrates a simple formal specification for the function calculating factorial.
+We has taken the example from the Stainless documentation \cite{Stainless:introduction} and modified it, so Stainless reports an error verifying it.
 
-The following example demonstrates a simple structure of formal specification for calculating factorial.
-
-\begin{lstlisting}[caption=Example of verifying the function calculating factorial of an integer number, label=listing:factorial, captionpos=b, language=Scala]
+\begin{lstlisting}[language=Scala]
     def factorial(n: Int): Int = {
@@
 The function recursively calculates factorial of an integer number.
-An input to the function is constrained in \textit{require} with non-negative value.
-A result of calculation should also be non-negative, what will be verified by Stainless.
-While compiling Stainless disproves the postcondition and gives the number 17 as a counterexample.
-A number of type Int is a 32 bit value.
-That is why calculating factorial of 17 causes an overflow and results to a negative value.
+An input to the function is constrained with \textit{require} to a non-negative value.
+A result of the calculation should be also a non-negative.
+Thus, Stainless will verify whether the result of the factorial calculation is non-negative for all non-negative inputs.
+
+Stainless can produce 3 outcomes of postcondition verification: \textit{valid}, \textit{invalid} and \textit{unknown}.
+If the postcondition is \textit{valid} Stainless could prove that for any inputs constrained in the precondition, the postcondition always holds.
+Reporting the postcondition as \textit{invalid} the framework could find at least one counterexample which satisfies the precondition but violates the postcondition.
+If Stainless is unable to prove the postcondition or find a counterexample it reports the outcome \textit{unknown}.
+In this case a timeout or an internal error has occurred.
+Furthermore, Stainless verifies that the precondition cannot be violated through the calling of the function being verified with an argument out of the precondition constraint.
+So, the framework reports also \textit{valid} or \textit{invalid} for the precondition. \cite{Stainless:introduction}
+
+Let's return back to our incorrect example with factorial. While compiling Stainless disproves the postcondition and gives the number 17 as a counterexample.
+The type of an argument and return value is defined as Int, which is 32 bit long.
+Thus, calculating factorial of 17 causes an overflow and results to a negative value.
 The program would work correct by changing a type of number to BigInt.
-The outcome of Stainless verification of the function factorial from Listing \ref{listing:factorial} is shown bellow.
+
+Stainless reports the following result of the factorial verification from our example:
 
@@
 
-The code being verified with Stainless should be written in Pure Scala, that corresponds to the functional programming. But also some imperative features are supported by Stainless.
+The code being verified must be written in Pure Scala, a functional subset of Scala. 
+The whole specification of Pure Scala is given on the website of Stainless \cite{Stainless:pure_scala}.
+Moreover, Stainless supports some imperative features translating them into Pure Scala during a preprocessing phase.
+To summarise them, the following extensions to the functional Scala are available \textbf{(//TODO: may be delete it)}: 
+\begin{itemize}
+  \item using mutable variables in some cases
+  \item using while loops
+  \item gettng an old value of a variable before the execution of a function being verified
+  \item using arrays that allow reassignment of its elements
+\end{itemize}
+The description of the supported imperative features and examples can be found on the website \cite{Stainless:imperative}. 
+In addition, Stainless has its own library with annotations, reimplementaion of some core data types, collections and input-output functions and much more.
+Some of them are described on the website of \cite{Stainless:library}.
+More details to the library can be found in the source code of Stainless on GitHub \cite{Stainless:github} in the folder \textit{/frontends/library/stainless}.
 
-Stainless verifies Scala programs using SMT (satisfiability modulo theories) solvers. A program code will be transformed by Stainless to the Pure Scala fragment supported by Inox solver. 
-This solver was also developed by LARA and allows usage of SMT solvers such as Z3, CVC4 and Princess, adding support of other higher-level features.
-Preconditions, postconditions and assertions will be translated into SMT formulas, so an SMT solver can determine if all properties hold.
+Stainless verifies Scala programs using SMT (satisfiability modulo theories) solvers. 
+The framework transforms a program code to the Pure Scala fragment supported by Inox solver.
+This solver was also developed by LARA-Group and allows usage of different SMT solvers such as Z3, CVC4 and Princess, adding support of other higher-level features.
+Thus, preconditions, postconditions and assertions are translated into SMT formulas, so an SMT solver can determine if all properties hold.\cite{Stainless:introduction}\cite{Stainless}
 
-\section{Bitcoin-S (Projekten und Packages, bitcoin-s-core, Eigenschaften zu prüfen)}
+\section{Overview of Bitcoin-S}
 \label{sec:bitcoin_s}

### 2019-06-04 code/currsniporig/src/main/scala/code/end/Code.scala
intended code

@@ -1,90 +1,90 @@
-package code.end
+  package code.end
 
-trait BasicArithmetic[N] {
-  def +(n: N): N
-}
-
-sealed abstract class Number
-    extends BasicArithmetic[Int64] {
-
-  protected def underlying: BigInt
-
-  def toBigInt: BigInt = underlying
-
-  def apply: BigInt => Int64
-
-  override def +(num: Int64): Int64 = {
-    require(
-          num.underlying <= BigInt(0)
-      && this.underlying <= BigInt("9223372036854775807")
-      && this.underlying >= BigInt("-9223372036854775808")
-    )
-    apply(checkResult(underlying + num.underlying))
+  trait BasicArithmetic[N] {
+    def +(n: N): N
   }
 
-  private def checkResult(result: BigInt): BigInt = {
-    require(
-         result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808")
-    )
-    result
+  sealed abstract class Number
+      extends BasicArithmetic[Int64] {
+
+    protected def underlying: BigInt
+
+    def toBigInt: BigInt = underlying
+
+    def apply: BigInt => Int64
+
+    override def +(num: Int64): Int64 = {
+      require(
+            num.underlying <= BigInt(0)
+        && this.underlying <= BigInt("9223372036854775807")
+        && this.underlying >= BigInt("-9223372036854775808")
+      )
+      apply(checkResult(underlying + num.underlying))
+    }
+
+    private def checkResult(result: BigInt): BigInt = {
+      require(
+           result <= BigInt("9223372036854775807")
+        && result >= BigInt("-9223372036854775808")
+      )
+      result
+    }
   }
-}
 
-sealed abstract class SignedNumber extends Number
+  sealed abstract class SignedNumber extends Number
 
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: BigInt => Int64 = Int64(_)
-}
+  sealed abstract class Int64 extends SignedNumber {
+    override def apply: BigInt => Int64 = Int64(_)
+  }
 
-trait BaseNumbers[T] {
-  def zero: T
-  def one: T
-}
+  trait BaseNumbers[T] {
+    def zero: T
+    def one: T
+  }
 
-case object Int64 extends BaseNumbers[Int64] {
-  lazy val zero = Int64(0)
[…]

### 2019-06-04 code/currsniporig/src/main/scala/code/initial/Code.scala

@@ -36,8 +36,6 @@
   trait BaseNumbers[T] {
     def zero: T
     def one: T
-//    def min: T //maybe remove
-//    def max: T //maybe remove
   }
 
   object Int64 extends BaseNumbers[Int64] {
@@ -50,9 +48,6 @@
 
     lazy val zero = Int64(0)
     lazy val one = Int64(1)
-
-//    lazy val min = Int64(-9223372036854775808L) //maybe remove
-//    lazy val max = Int64(9223372036854775807L) //maybe remove
 
     def apply(long: Long): Int64 = Int64(BigInt(long))
 
@@ -87,8 +82,6 @@
   }
 
   object Satoshis extends BaseNumbers[Satoshis] {
-//    val min = Satoshis(Int64.min) //maybe remove
-//    val max = Satoshis(Int64.max) //maybe remove
     val zero = Satoshis(Int64.zero)
     val one = Satoshis(Int64.one)
 

### 2019-06-04 doc/chapters/introduction.tex
fixed typos

@@
 \label{chap:introduction}
+
 In this work we experiment in formal verification of Scala code. 
-We use the verification framework Stainless to verify the code of Bitcoin-S-Core, the Scala implementation of the Bitcoin protocol.
+We use the verification framework Stainless to verify parts of the code of Bitcoin-S-Core, a Scala implementation of the Bitcoin protocol.
 In the following we describe the main aspects of formal verification, Stainless and the project Bitcoin-S.
@@
 The longer and more complex a source code is, the more difficult it is to verify its correctness.
-There are different approaches for verification of the program correctness.
+There are different approaches to show the correctness of a program.
 
@@
 However, not all possible inputs can be tested. 
-Therefore, a long-term test coverage model is usually created to test a software design enough.
-Nevertheless, this is still not easy to test a software being able to observe the side effects of a code.
+Therefore, a long-term test coverage model is usually created to test a software design.
+Nevertheless, it is still not easy to test a software and its side effects.
 \textbf{Practically, a developer runs tests, debugs appeared failures, adjusts a code and extends a test bench to verify the previously uncovered aspects.
@@
 It is unnecessary to create a simulation tests with some possible inputs.
-With formal verification all possible input values are explored algorithmically and exhaustively.
+With formal verification all possible inputs are explored algorithmically and exhaustively.
 Correctness of a program is analyzed relative to its formal specification.

### 2019-06-04 doc/chapters/trying_verify_checktx.tex
added some labels
consistend titles
moved image
changed sentence

@@
 \chapter{Trying to Verify checkTransaction}
-\label{chap:connecting}
+\label{chap:verify_check}
 
@@
 
-\section{Process of the Verification // TODO change?}
+\section{Process of Verification // TODO change}
 
@@
 After running Stainless on it, it throws a really huge bunch of errors about what Stainless can not reason about.
-This would require to change nearly all the extracted code.
+And after fixing some of those errors there appear new ones as we can see in the next chapter.
+This would require to change nearly everything of the extracted code.
 
@@
 During the examination of the checkTransaction function, we found a bug in Bitcoin-S.
+
+\begin{figure}[H]
+	\centering
+		\includegraphics[scale=0.5]{images/bitcoin-s-pr-comment.png}
+	\caption{Nice catch comment on our PR \#435 on GitHub}
+	\label{fig:output1}
+\end{figure}
 
@@
 This was fixed in pull request \mypound435 on GitHub at April 23, 2019 by us along a unit test to prevent this bug from appearing again in the future.
-
-\begin{figure}[H]
-	\centering
-		\includegraphics[scale=0.5]{images/bitcoin-s-pr-comment.png}
-	\caption{Nice catch comment on PR \#435 from GitHub}
-	\label{fig:output1}
-\end{figure}

### 2019-06-04 doc/chapters/conclusion.tex
added part of conclusion

@@
 
-\section{Future Work}
+As we saw, it is really hard to verify a software not written with verification in mind.
+It needs a loot of changes in the software because verification is mathematical and the current software is written mostly in Object-oriented style.
+A functional programmed software would be much easier to reason about.
+
+Thus either Stainless must find ways to translate more of built-in object-oriented patterns of Scala to their verification tool or developers must invest more in functional programming.
+
+The code we verified in the previous chapter is not the original code from Bitcoin-S.
+So we can not guarantee the correctness of the addition of Satoshis with zero in Bitcoin-S.
+There are changes that stays the same like the change from object to case object but there are other changes where we can not say if they are the same or not like the bound check in section \ref{sec:bound_check}.

### 2019-06-04 doc/chapters/introduction.tex
first part

@@
 In this work we experiment in formal verification of Scala code. 
-We use the verification framework Stainless to verify parts of the code of Bitcoin-S-Core, a Scala implementation of the Bitcoin protocol.
-In the following we describe the main aspects of formal verification, Stainless and the project Bitcoin-S.
+We use the verification framework Stainless to verify parts of the code of Bitcoin-S, a Scala implementation of the Bitcoin protocol.
+In the following we describe the main aspects of formal verification, Stainless and Bitcoin-S.
 
@@
 
-The commonly used one is testing.
-A developer creates a set of some practical scenarios and assertions to test a software design and implementation. 
-So, the actual results are checked to match the expected results.
-Tests help to identify bugs and missing requirements.
-However, not all possible inputs can be tested. 
-Therefore, a long-term test coverage model is usually created to test a software design.
-Nevertheless, it is still not easy to test a software and its side effects.
-\textbf{Practically, a developer runs tests, debugs appeared failures, adjusts a code and extends a test bench to verify the previously uncovered aspects.
-(TODO// do we need this sentence?)}\cite{sanghavi:formal_verification}
-
-Another powerful method to check whether a program works correct is formal verification.
+A powerful method to check whether a program works correctly is formal verification.
 This is a systematic process based on mathematical modeling.
-It is unnecessary to create a simulation tests with some possible inputs.
 With formal verification all possible inputs are explored algorithmically and exhaustively.
-Correctness of a program is analyzed relative to its formal specification.
-Formal specification is a mathematical description of a software behavior that can be provided to the formal verification tool for proving it.\cite{sanghavi:formal_verification}
+This is in contrast to testing, where only predefined inputs are tested.
+The correctness of a program is analyzed relative to its formal specification.
+Formal specification is a mathematical description of a software behavior that can be used by the verification tool to verify the code.\cite{sanghavi:formal_verification}
 
@@
 Stainless statically verifies that a program satisfies a given specification and that a program will not crash at runtime.
-The framework explores all possible input values, reports inputs for which a program fails and demonstrates counterexamples which violate a given specification.\cite{Stainless:introduction}
+The framework explores all possible inputs and finds a counter examples for possible failures in a program which violate the given specification.\cite{Stainless:introduction}
 
@@
 With \textit{require} we define a precondition of a function, which we want to verify.
-\textit{require} is called at the beginning of the function body.
-A constraint for inputs to the function is passed as an argument to \textit{require} in the form of a Boolean expression.
+\textit{require} is placed at the beginning of the function body.
+We pass a constraint for the inputs of a function as an argument to \textit{require} in the form of a Boolean expression.
+
 With \textit{ensuring} we provide a postcondition which is a condition on an output of the function.
 \textit{ensuring} is called after the function body.
-While compiling Stainless tries to prove that the postcondition always holds, assuming the given precondition does hold.\cite{Stainless:introduction}
+
+On invoking, Stainless tries to prove that the postcondition always holds, assuming the given precondition does hold.\cite{Stainless:introduction}
 
 The following example demonstrates a simple formal specification for the function calculating factorial.
-We has taken the example from the Stainless documentation \cite{Stainless:introduction} and modified it, so Stainless reports an error verifying it.
-
+This is a modified example from the Stainless documentation, so Stainless reports an error verifying it.
 \begin{lstlisting}[language=Scala]
@@
 The function recursively calculates factorial of an integer number.
-An input to the function is constrained with \textit{require} to a non-negative value.
-A result of the calculation should be also a non-negative.
+The input of the function is constrained with \textit{require} to a non-negative value.
+The result should also be non-negative.
 Thus, Stainless will verify whether the result of the factorial calculation is non-negative for all non-negative inputs.
@@
 If Stainless is unable to prove the postcondition or find a counterexample it reports the outcome \textit{unknown}.
-In this case a timeout or an internal error has occurred.
-Furthermore, Stainless verifies that the precondition cannot be violated through the calling of the function being verified with an argument out of the precondition constraint.
-So, the framework reports also \textit{valid} or \textit{invalid} for the precondition. \cite{Stainless:introduction}
+In this case a timeout or an internal error occurred.
+Furthermore, Stainless checks for calls in the code to the function with invalid arguments violating the precondition.
+For example if we invoke factorial with a negative number it reports \textit{invalid}.\cite{Stainless:introduction}
 
-Let's return back to our incorrect example with factorial. While compiling Stainless disproves the postcondition and gives the number 17 as a counterexample.
-The type of an argument and return value is defined as Int, which is 32 bit long.
-Thus, calculating factorial of 17 causes an overflow and results to a negative value.
-The program would work correct by changing a type of number to BigInt.
-
-Stainless reports the following result of the factorial verification from our example:
-
+Let's return back to our incorrect example with factorial.
+Stainless reports the following result for the factorial verification from our example:
 \begin{figure}[H]
@@
 
-The code being verified must be written in Pure Scala, a functional subset of Scala. 
-The whole specification of Pure Scala is given on the website of Stainless \cite{Stainless:pure_scala}.
-Moreover, Stainless supports some imperative features translating them into Pure Scala during a preprocessing phase.
-To summarise them, the following extensions to the functional Scala are available \textbf{(//TODO: may be delete it)}: 
-\begin{itemize}
-  \item using mutable variables in some cases
-  \item using while loops
-  \item gettng an old value of a variable before the execution of a function being verified
-  \item using arrays that allow reassignment of its elements
-\end{itemize}
-The description of the supported imperative features and examples can be found on the website \cite{Stainless:imperative}. 
-In addition, Stainless has its own library with annotations, reimplementaion of some core data types, collections and input-output functions and much more.
-Some of them are described on the website of \cite{Stainless:library}.
-More details to the library can be found in the source code of Stainless on GitHub \cite{Stainless:github} in the folder \textit{/frontends/library/stainless}.
+Stainless disproves the postcondition and gives the number 17 as a counterexample.
+Due to an overflow in the 32 bit Int the result becomes a negative value.
+The program would work correct by changing the type from Int to BigInt.
+
+Stainless supports only a subset of Scala.
+They call is Pure Scala.
+You can find the specification of Pure Scala on the Stainless website.\cite{Stainless:pure_scala}
+
+In addition, Stainless has its own library with annotations, reimplementaion of some core data types, collections and input-output functions and more.
+Some of them are described on the website.\cite{Stainless:library}
+There are more details about library in the source code of Stainless on GitHub in the folder \textit{frontends/library/stainless}.\cite{Stainless:github}
 

### 2019-06-04 doc/chapters/towards_verifying_addition.tex
consistend titles
added final code
added part of conclusion
fixed line numbers

@@
 {\footnotesize\begin{verbatim}
-  [ Error  ] Code.scala:43:1: Objects cannot extend classes or implement traits,
+  [ Error  ] Code.scala:41:1: Objects cannot extend classes or implement traits,
              use a case object instead
@@
            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-  [ Error  ] Code.scala:65:3: Stainless doesn't support abstract type members
+  [ Error  ] Code.scala:60:3: Stainless doesn't support abstract type members
               type A
               ^^^^^^
-  [ Error  ] Code.scala:86:33: Only literal arguments are allowed for BigInt.
+  [ Error  ] Code.scala:77:33: Only literal arguments are allowed for BigInt.
               def toBigInt: BigInt = BigInt(toLong)
                                             ^^^^^^
-  [ Error  ] Code.scala:93:1: Objects cannot extend classes or implement traits,
+  [ Error  ] Code.scala:84:1: Objects cannot extend classes or implement traits,
              use a case object instead
@@
 
-\section{Turning object Into case object}
+\section{Turn object Into case object}
 
@@
            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-  [ Error  ] Code.scala:93:1: Objects cannot extend classes or implement traits,
+  [ Error  ] Code.scala:84:1: Objects cannot extend classes or implement traits,
             use a case object instead
@@
 {\footnotesize\begin{verbatim}
-  [ Error  ] Code.scala:65:3: Stainless doesn't support abstract type members
+  [ Error  ] Code.scala:60:3: Stainless doesn't support abstract type members
               type A
@@
 {\footnotesize\begin{verbatim}
-  [ Error  ] Code.scala:86:33: Only literal arguments are allowed for BigInt.
+  [ Error  ] Code.scala:77:33: Only literal arguments are allowed for BigInt.
               def toBigInt: BigInt = BigInt(toLong)
@@
 \section{Replace BigInt \&-Function With Bound Check}
+\label{sec:bound_check}
 
@@
 
-\section{Extracting Private Inner Classes}
+\section{Extract Private Inner Classes}
 
@@
 
+The verified code.
+\lstinputlisting[language=scala]{../code/currsniporig/src/main/scala/code/end/Code.scala}
+
 But is it really the original code that we verified?

### 2019-06-04 doc/chapters/towards_verifying_addition.tex
fixed upper type bound

@@
 \section{Replace BigInt Constructor Argument With String Literal}
+
 Stainless output:
@@
 It does not support dynamic BigInt construction.
-Only from string literals.
+Only from literal arguments.
 
@@
 This is a missing feature in Stainless.
+It does not support upper type bound on the class itself.
 To track this, there is the issue \mypound519 on GitHub.

### 2019-06-04 doc/chapters/trying_verify_checktx.tex
fixed
fixed order

@@
 
-This chapter describes some parts of Bitcoin-S.
+This chapter describes some parts of Bitcoin-S needed to verify our property.
 We are going to create a transaction and show you the relevant parts of checkTransaction, where transactions are checked against some properties.
@@
 
-The relevant part for the property we are trying to verify is the following.
+The relevant part for the bug we found.
 \begin{lstlisting}[language=scala]
@@
 
-There is a bug, because checkTransaction checks on the transaction id that we are going to explain a next section.
-
-\section{Process of Verification // TODO change}
-
-Trying to integrate Stainless in Bitcoin-S caused a lot of troubles, because of version conflicts.
-If you are interested in it you can read \hl{TODO} in the appendix.
-
-It takes too much time to do it and after some discussion, we decided to extract only the parts of the code that are needed for checkTransaction.
-
-The extracted code has more than 1500 lines.
-After running Stainless on it, it throws a really huge bunch of errors about what Stainless can not reason about.
-And after fixing some of those errors there appear new ones as we can see in the next chapter.
-This would require to change nearly everything of the extracted code.
-
-We arrived at the decision that we need a smaller part to verify.
-We have chosen to verify the addition with zero of Bitcoin-Ss CurrencyUnit class as described in the next chapter.
-
-But before that you can read about the bug we found and fixed in the next section.
+There is a bug, because checkTransaction checks on the transaction id that we are going to explain the next section.
 
@@
 This was fixed in pull request \mypound435 on GitHub at April 23, 2019 by us along a unit test to prevent this bug from appearing again in the future.
+
+\section{Process of Verification // TODO change}
+
+Trying to integrate Stainless in Bitcoin-S caused a lot of troubles, because of version conflicts.
+If you are interested in it you can read \hl{TODO} in the appendix.
+
+It takes too much time to do it and after some discussion, we decided to extract only the parts of the code that are needed for checkTransaction.
+
+The extracted code has more than 1500 lines.
+After running Stainless on it, it throws a really huge bunch of errors about what Stainless can not reason about.
+And after fixing some of those errors there appear new ones as we can see in the next chapter.
+This would require to change nearly everything of the extracted code.
+
+We arrived at the decision that we need a smaller part to verify.
+We have chosen to verify the addition with zero of Bitcoin-Ss CurrencyUnit class as described in the next chapter.

### 2019-06-04 doc/chapters/trying_verify_checktx.tex
added foot note

@@
 What should not be possible is a transaction referencing the same output twice.
-This was a bug in Bitcoin Core known as CVE-2018–17144 which was patched on September 18, 2018.
+This was a bug in Bitcoin Core known as CVE-2018–17144 which was patched on September 18, 2018. \cite{cve201817144}
 
@@
 \end{lstlisting}
+
 Since TransactionOutPoint is a case class and Scala has a built in == for case classes there is no need to implement TransactionOutPoint.==.
@@
 
-This was fixed in pull request \mypound435 on GitHub at April 23, 2019 by us along a unit test to prevent this bug from appearing again in the future.
+This was fixed in pull request \mypound435 on GitHub\footnote{https://github.com/bitcoin-s/bitcoin-s/pull/435} at April 23, 2019 by us along a unit test to prevent this bug from appearing again in the future.
 

### 2019-06-04 doc/chapters/towards_verifying_addition.tex
added foot notes

@@
 This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
-The issue \mypound520 on Stainless GitHub gives some thoughts if you want to know more.
+The issue \mypound520 on Stainless GitHub\footnote{https://github.com/epfl-lara/stainless/issues/520} gives some thoughts if you want to know more.
 
@@
 It does not support upper type bound on the class itself.
-To track this, there is the issue \mypound519 on GitHub.
+To track this, there is the issue \mypound519 on GitHub\footnote{https://github.com/epfl-lara/stainless/issues/519}.
 \begin{lstlisting}[language=scala]
@@
 
-This was a missing feature of Stainless that was fixed on Mai 28, 2019 with pull request \mypound470.
+This was a missing feature of Stainless that was fixed on Mai 28, 2019 with pull request \mypound470 on GitHub\footnote{https://github.com/epfl-lara/stainless/pull/470}.
 

### 2019-06-04 doc/chapters/introduction.tex
fixed footnotes

@@
 The correctness of a program is analyzed relative to its formal specification.
-Formal specification is a mathematical description of a software behavior that can be used by the verification tool to verify the code.\cite{sanghavi:formal_verification}
+Formal specification is a mathematical description of a software behavior that can be used by the verification tool to verify the code. \cite{sanghavi:formal_verification}
 
@@
 Stainless statically verifies that a program satisfies a given specification and that a program will not crash at runtime.
-The framework explores all possible inputs and finds a counter examples for possible failures in a program which violate the given specification.\cite{Stainless:introduction}
+The framework explores all possible inputs and finds a counter examples for possible failures in a program which violate the given specification. \cite{Stainless:introduction}
 
@@
 
-On invoking, Stainless tries to prove that the postcondition always holds, assuming the given precondition does hold.\cite{Stainless:introduction}
+On invoking, Stainless tries to prove that the postcondition always holds, assuming the given precondition does hold. \cite{Stainless:introduction}
 
@@
 Furthermore, Stainless checks for calls in the code to the function with invalid arguments violating the precondition.
-For example if we invoke factorial with a negative number it reports \textit{invalid}.\cite{Stainless:introduction}
+For example if we invoke factorial with a negative number it reports \textit{invalid}. \cite{Stainless:introduction}
 
@@
 They call is Pure Scala.
-You can find the specification of Pure Scala on the Stainless website.\cite{Stainless:pure_scala}
+You can find the specification of Pure Scala on the Stainless website\footnote{\url{https://epfl-lara.github.io/stainless/purescala.html}}.
 
 In addition, Stainless has its own library with annotations, reimplementaion of some core data types, collections and input-output functions and more.
-Some of them are described on the website.\cite{Stainless:library}
-There are more details about library in the source code of Stainless on GitHub in the folder \textit{frontends/library/stainless}.\cite{Stainless:github}
+Some of them are described on the website\footnote{\url{https://epfl-lara.github.io/stainless/library.html}}.
+There are more details about library in the source code of Stainless on GitHub in the folder \textit{frontends/library/stainless}\footnote{\url{https://github.com/epfl-lara/stainless}}.
 
@@
 Developers of Bitcoin-S are working on the implementation of a lightweight bitcoin client in the package \textbf{bitcoin-s-spv-node}.
-The whole list of packages can be found on GitHub \cite{BitcoinSProject}.
+The whole list of packages can be found on GitHub\footnote{\url{https://github.com/bitcoin-s}}.
 

### 2019-06-04 doc/chapters/towards_verifying_addition.tex
fixed footnotes

--- 2019-06-04 doc/chapters/towards_verifying_addition.tex
@@
 This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
-The issue \mypound520 on Stainless GitHub gives some thoughts if you want to know more.
+The issue \mypound520 on Stainless GitHub\footnote{\url{https://github.com/epfl-lara/stainless/issues/520}} gives some thoughts if you want to know more.
 
@@
 It does not support upper type bound on the class itself.
-To track this, there is the issue \mypound519 on GitHub.
+To track this, there is the issue \mypound519 on GitHub\footnote{\url{https://github.com/epfl-lara/stainless/issues/519}}.
 \begin{lstlisting}[language=scala]
@@
 
-This was a missing feature of Stainless that was fixed on Mai 28, 2019 with pull request \mypound470.
+This was a missing feature of Stainless that was fixed on Mai 28, 2019 with pull request \mypound470 on GitHub\footnote{\url{https://github.com/epfl-lara/stainless/pull/470}}.
 

### 2019-06-05 doc/chapters/trying_verify_checktx.tex
added ref
changed title
fixed foot note

@@
 
-This was fixed in pull request \mypound435 on GitHub\footnote{https://github.com/bitcoin-s/bitcoin-s/pull/435} at April 23, 2019 by us along a unit test to prevent this bug from appearing again in the future.
+This was fixed in pull request \mypound435 on GitHub\footnote{\url{https://github.com/bitcoin-s/bitcoin-s/pull/435}} at April 23, 2019 by us along a unit test to prevent this bug from appearing again in the future.
 
-\section{Process of Verification // TODO change}
+\section{Adjusting Property}
 
 Trying to integrate Stainless in Bitcoin-S caused a lot of troubles, because of version conflicts.
-If you are interested in it you can read \hl{TODO} in the appendix.
+If you are interested in it you can read more in the appendix \ref{chap:appendix_arb}.
 

### 2019-06-05 doc/chapters/towards_verifying_addition.tex
updated titles

@@
-\chapter{Towards Verifying Addition with Zero}
+\chapter{Towards Verifying the Addition with Zero Property}
 \label{chap:verify_add}
@@
 
-\section{Turn object Into case object}
+\section{Turning Object into Case Object}
 
@@
 
-\section{Get Rid of Abstract Type Member}
+\section{Getting Rid of Abstract Type Member}
 
@@
 
-\section{Replace BigInt Constructor Argument With String Literal}
+\section{Replacing the BigInt Constructor Argument With A String Literal}
 
@@
 
-\section{Get Rid of Special Generics}
+\section{Getting Rid of Special Generics}
 
@@
 
-\section{Get Rid of Concret Type}
+\section{Getting Rid of Concrete Type}
 
@@
 This was a missing feature of Stainless that was fixed on Mai 28, 2019 with pull request \mypound470 on GitHub\footnote{\url{https://github.com/epfl-lara/stainless/pull/470}}.
+Now it should work without this change.
 
-\section{Replace BigInt \&-Function With Bound Check}
+\section{Replacing the BigInt \&-Function With Bound Check}
 \label{sec:bound_check}
@@
 
-\section{Extract Private Inner Classes}
+\section{Extracting Private Inner Classes}
 
@@
 
-\section{Remove Second Parameter of require}
+\section{Removeing Second Parameter of require}
 
@@
 
-\section{Rewrite BigInt Comparison with Long}
+\section{Rewriting BigInt Comparison with Long}
 
@@
 
-\section{Write Specification for Verification}
+\section{Writing Specification for Verification}
 
@@
 
-\section{Propagate require}
+\section{Propagating require}
 

### 2019-06-05 doc/chapters/introduction.tex
added subtitles
updated intro

@@
 
-In this work we experiment in formal verification of Scala code. 
-We use the verification framework Stainless to verify parts of the code of Bitcoin-S, a Scala implementation of the Bitcoin protocol.
+In this work we use the verification framework Stainless to verify parts of the code of Bitcoin-S, a Scala implementation of the Bitcoin protocol.
 In the following we describe the main aspects of formal verification, Stainless and Bitcoin-S.
@@
 After that a formal specification for functions should be defined and annotated according Stainless libraries, so that Stainless can verify correctness of the code.
+
+\subsection{No Inflation Property}
+
+\subsection{Addition with Zero Property}

### 2019-06-05 doc/chapters/introduction.tex
added text

@@
 
+// TODO move to chapter description and change Overview of stainless to formal verification with stainless
+
 The longer and more complex a source code is, the more difficult it is to verify its correctness.
@@
 This is a systematic process based on mathematical modeling.
-With formal verification all possible inputs are explored algorithmically and exhaustively.
+With formal verification all possible inputs are covered algorithmically.
 This is in contrast to testing, where only predefined inputs are tested.
 The correctness of a program is analyzed relative to its formal specification.
-Formal specification is a mathematical description of a software behavior that can be used by the verification tool to verify the code. \cite{sanghavi:formal_verification}
+Formal specification is a mathematical description of a software behavior that can be used by the verification tool to verify the code.
+We will see an example shortly.
 
@@
 
-In the work some fragments of the code of Bitcoin-S has to be verified.
+In the work we verify some fragments of the code of Bitcoin-S.
 Bitcoin-S is an open source Scala implementation of the Bitcoin protocol. 

### 2019-06-05 doc/appendix/appendixexemple.tex
added foot note
fixed footnotes
fixed appendix

@@
+\addcontentsline{toc}{chapter}{APPENDICES}
+
 \chapter{Practical Challenges with Stainless}
@@
 
-During this work, Stainless bumped the sbt plugin from version 0.13.17 to 1.2.8 and Scala from 2.11.12 to 2.12.8.
+During this work, Stainless bumped the sbt plugin from version 0.13.17 to 1.2.8 and now supports also Scala 2.12.8 and prevoiusly 2.11.12.
 So this section might be out of date now.
@@
 So the sbt plugin does not complain always but the JAR file did.
-The open issue \mypound484 on GitHub might describe this error.
+The open issue \mypound484 on GitHub\footnote{url{https://github.com/epfl-lara/stainless/issues/484}} might describe this error.

### 2019-06-05 doc/chapters/trying_verify_checktx.tex
updated titles
todo

@@
-\chapter{Trying to Verify checkTransaction}
+\chapter{Trying to Verify the no Inflation Property}
 \label{chap:verify_check}
 
-This chapter describes some parts of Bitcoin-S needed to verify our property.
-We are going to create a transaction and show you the relevant parts of checkTransaction, where transactions are checked against some properties.
+This chapter describes the part of Bitcoin-S needed to verify our property.
+We are going to create a transaction and show you the relevant parts of the method checkTransaction, where transactions are checked against some properties.
 Then we see some troubles occurred during this work.
@@
 
-Some of the code in this section are copied or modified parts of Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
-Bitcoin-S-Core has a bitcoin transaction builder class with the following signature:
+Some of the code in this section is copied or adapted parts of Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
+Bitcoin-S-Core has a bitcoin transaction builder class with the following constructor:
 \begin{lstlisting}[language=scala]
@@
     feeRate: FeeUnit,                     // fee rate per byte
-    changeSPK: ScriptPubKey,              // public key
+    changeSPK: ScriptPubKey,              // where to send the change
     network: BitcoinNetwork               // bitcoin network information
@@
 Thus, we create a transaction with one single output.
-This transaction can also be parsed from the real bitcoin network but to show the progression we create one.
+This transaction can be parsed from the bitcoin network but we create a transaction manually in order to see this process.
+
+// TODO explain more. We create a fake transaction blabla
+
 \begin{lstlisting}[language=scala]
@@
 On line one and two we are creating a new keypair to sign the next transaction and have a scriptPubKey where the bitcoins are.
-Line four specifies the amount of satoshis we are having in the transaction.
+Line four specifies the amount of satoshis we have in the transaction.
 Then we create the actual transaction from line 6 to 13.

### 2019-06-07 presentation/solidity-highlighting.tex
added template

% Copyright 2017 Sergei Tikhomirov, MIT License
% https://github.com/s-tikhomirov/solidity-latex-highlighting/

\usepackage{listings, xcolor}

\definecolor{verylightgray}{rgb}{.97,.97,.97}

\lstdefinelanguage{Solidity}{
	keywords=[1]{anonymous, assembly, assert, balance, break, call, callcode, case, catch, class, constant, continue, constructor, contract, debugger, default, delegatecall, delete, do, else, emit, event, experimental, export, external, false, finally, for, function, gas, if, implements, import, in, indexed, instanceof, interface, internal, is, length, library, log0, log1, log2, log3, log4, memory, modifier, new, payable, pragma, private, protected, public, pure, push, require, return, returns, revert, selfdestruct, send, solidity, storage, struct, suicide, super, switch, then, this, throw, transfer, true, try, typeof, using, value, view, while, with, addmod, ecrecover, keccak256, mulmod, ripemd160, sha256, sha3}, % generic keywords including crypto operations
	keywordstyle=[1]\color{blue}\bfseries,
	keywords=[2]{address, bool, byte, bytes, bytes1, bytes2, bytes3, bytes4, bytes5, bytes6, bytes7, bytes8, bytes9, bytes10, bytes11, bytes12, bytes13, bytes14, bytes15, bytes16, bytes17, bytes18, bytes19, bytes20, bytes21, bytes22, bytes23, bytes24, bytes25, bytes26, bytes27, bytes28, bytes29, bytes30, bytes31, bytes32, enum, int, int8, int16, int24, int32, int40, int48, int56, int64, int72, int80, int88, int96, int104, int112, int120, int128, int136, int144, int152, int160, int168, int176, int184, int192, int200, int208, int216, int224, int232, int240, int248, int256, mapping, string, uint, uint8, uint16, uint24, uint32, uint40, uint48, uint56, uint64, uint72, uint80, uint88, uint96, uint104, uint112, uint120, uint128, uint136, uint144, uint152, uint160, uint168, uint176, uint184, uint192, uint200, uint208, uint216, uint224, uint232, uint240, uint248, uint256, var, void, ether, finney, szabo, wei, days, hours, minutes, seconds, weeks, years},	% types; money and time units
	keywordstyle=[2]\color{teal}\bfseries,
	keywords=[3]{block, blockhash, coinbase, difficulty, gaslimit, number, timestamp, msg, data, gas, sender, sig, value, now, tx, gasprice, origin},	% environment variables
	keywordstyle=[3]\color{violet}\bfseries,
	identifierstyle=\color{black},
	sensitive=false,
	comment=[l]{//},
	morecomment=[s]{/*}{*/},
	commentstyle=\color{gray}\ttfamily,
	stringstyle=\color{red}\ttfamily,
	morestring=[b]',
	morestring=[b]"
}

\lstset{
  language=Solidity,
  escapeinside={<@}{@>},
	backgroundcolor=\color{verylightgray},
	extendedchars=true,
	basicstyle=\footnotesize\ttfamily,
	showstringspaces=false,
	showspaces=false,
	numbers=left,
	numberstyle=\footnotesize,
	numbersep=9pt,
	tabsize=2,
	breaklines=true,
	showtabs=false,
	captionpos=b
}

### 2019-06-08 doc/chapters/challenge_stainless.tex
new chapter
added some text

--- 2019-06-05 doc/appendix/appendixexemple.tex
@@
-\addcontentsline{toc}{chapter}{APPENDICES}
-
 \chapter{Practical Challenges with Stainless}
 \label{chap:appendix_arb}
-\endgroup
+
+\section{sbt vs JAR}
+
+We can either use the sbt plugin or a JAR file to check the code with Stainless.
+
+Invoking the JAR on our source code Stainless will verify it.
+If we have a bigger project, this becomes really tricky because we must pass all files needed including the dependencies.
+This is in contrast to the sbt plugin, where we can integrate Stainless in our compilation process.
+When we call compile, Stainless verifies the code and stops the compilation if the verification fails.
+
+Having a fixed version configured in the sbt build file, every developer has the same Stainless features available.
+This should prevent incompatibility with new or deprecated features when we use different plugins.
+
+So the sbt plugin has clear advantages over the JAR file since its integrated directly.
+We do not have to download it manually and find the right version and if we bump the version we can just edit it in the build file and every developer is on the same version again.
+
+However, there are some drawbacks currently.
+For example the sbt plugin does not always report errors.
+We will read mor about that shortly.
 
@@
 
-During this work, Stainless bumped the sbt plugin from version 0.13.17 to 1.2.8 and now supports also Scala 2.12.8 and prevoiusly 2.11.12.
+During this work, Stainless updated the sbt plugin to support sbt 1.2.8 from 0.13.17 and Scala 2.12.8 from 2.11.12.
 So this section might be out of date now.

### 2019-06-08 doc/chapters/challenge_stainless.tex
updated stainless challenges

@@
 \label{chap:appendix_arb}
+
+In this chapter we explore the practical challenges that could occur by using Stainless.
 
@@
 For example the sbt plugin does not always report errors.
-We will read mor about that shortly.
+We will read more about that shortly.
 
@@
 
-There are two ways to integrate Stainless in a Scala project, the Scala build tool (sbt) plugin or a command line tool.
-Both, when run, analyses the code passed and report warnings to the console about the given code.
 Stainless requires and Scala recommends Java SE Development Kit 8.
@@
 
-The output was some Stainless error.
+\emph{.ivy2} is the dependency cache of sbt.
+The \emph{tr} replaces the first char with the second so '\\n' with either ':' or ' '.
+
+With this command, Stainless throws the next error:
 \begin{verbatim}
@@
 
-After some discussion, we decided to go another way, because the errors may take too much time and it might lead to a next error.
-Let's extract the code needed to verify a transaction mainly the class Transaction and ScriptInterpreter with many other classes their depending on.
+So we can not know how many errors will face us.
+Lets go another way, because the errors may take too much time and it might lead to a next error.
+We extract the code needed to verify a transaction mainly the class Transaction and ScriptInterpreter with many other classes their depending on.
 
 After this extraction Stainless was successfully integrated with both, sbt and JAR.
-This leads us to another finding as described in the next section.
-
-\section{Error Reporting with sbt and JAR}
 
@@
 But it also ended with no output.
-So we were not able to change the code so Stainless would accept it.
+So we can not able to change the code so Stainless would accept it since we do not know what to change.
 
-After running the JAR file on the code there were many errors as described in chapter \ref{chap:verify_add}.
+So the sbt plugin does not complain always where the JAR file did.
+The open \href{https://github.com/epfl-lara/stainless/issues/484}{issue \mypound484 on GitHub} might describe exactly this error.
 
-So the sbt plugin does not complain always but the JAR file did.
-The open issue \mypound484 on GitHub\footnote{url{https://github.com/epfl-lara/stainless/issues/484}} might describe this error.
+Now we can finally run Stainless on our code.
+But this leads us to the next findings described in the previous chapters.

### 2019-06-08 doc/chapters/challenge_stainless.tex
typos

@@
 So we can not know how many errors will face us.
-Lets go another way, because the errors may take too much time and it might lead to a next error.
+Let's go another way, because the errors may take too much time and it might lead to a next error.
 We extract the code needed to verify a transaction mainly the class Transaction and ScriptInterpreter with many other classes their depending on.
@@
 
-So the sbt plugin does not complain always where the JAR file did.
+So the sbt plugin does not always complain where the JAR file did.
 The open \href{https://github.com/epfl-lara/stainless/issues/484}{issue \mypound484 on GitHub} might describe exactly this error.

### 2019-06-08 doc/chapters/introduction.tex
added title
changed from foot note to href

@@
 They call is Pure Scala.
-You can find the specification of Pure Scala on the Stainless website\footnote{\url{https://epfl-lara.github.io/stainless/purescala.html}}.
+You can find the specification of Pure Scala on the Stainless website in the section \href{https://epfl-lara.github.io/stainless/purescala.html}{Pure Scala}.
 
 In addition, Stainless has its own library with annotations, reimplementaion of some core data types, collections and input-output functions and more.
-Some of them are described on the website\footnote{\url{https://epfl-lara.github.io/stainless/library.html}}.
-There are more details about library in the source code of Stainless on GitHub in the folder \textit{frontends/library/stainless}\footnote{\url{https://github.com/epfl-lara/stainless}}.
+Some of them are described on the website in the section \href{https://epfl-lara.github.io/stainless/library.html}{Stainless Library}.
+There are more details about library in the source code of Stainless on GitHub in the folder \href{https://github.com/epfl-lara/stainless}{\textit{frontends/library/stainless}}.
 
@@
 Developers of Bitcoin-S are working on the implementation of a lightweight bitcoin client in the package \textbf{bitcoin-s-spv-node}.
-The whole list of packages can be found on GitHub\footnote{\url{https://github.com/bitcoin-s}}.
+The whole list of packages can be found on GitHub in the \href{https://github.com/bitcoin-s}{organization Bitcoin-S}.
 
@@
 
+\section{The Properties to Verify}
+
 \subsection{No Inflation Property}

### 2019-06-08 doc/chapters/towards_verifying_addition.tex
changed from foot note to href

@@
 This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
-The issue \mypound520 on Stainless GitHub\footnote{\url{https://github.com/epfl-lara/stainless/issues/520}} gives some thoughts if you want to know more.
+The \href{https://github.com/epfl-lara/stainless/issues/520}{issue \mypound520} on Stainless GitHub gives some thoughts if you want to know more.
 
@@
 It does not support upper type bound on the class itself.
-To track this, there is the issue \mypound519 on GitHub\footnote{\url{https://github.com/epfl-lara/stainless/issues/519}}.
+To track this, there is the \href{https://github.com/epfl-lara/stainless/issues/519}{issue \mypound519} on GitHub.
 \begin{lstlisting}[language=scala]
@@
 
-This was a missing feature of Stainless that was fixed on Mai 28, 2019 with pull request \mypound470 on GitHub\footnote{\url{https://github.com/epfl-lara/stainless/pull/470}}.
+This was a missing feature of Stainless that was fixed on Mai 28, 2019 with \href{https://github.com/epfl-lara/stainless/pull/470}{pull request \mypound470} on GitHub.
 Now it should work without this change.

### 2019-06-08 doc/chapters/trying_verify_checktx.tex
changed from foot note to href

@@
 
-This was fixed in pull request \mypound435 on GitHub\footnote{\url{https://github.com/bitcoin-s/bitcoin-s/pull/435}} at April 23, 2019 by us along a unit test to prevent this bug from appearing again in the future.
+This was fixed in \href{https://github.com/bitcoin-s/bitcoin-s/pull/435}{pull request \mypound435} on GitHub at April 23, 2019 by us along a unit test to prevent this bug from appearing again in the future.
 

### 2019-06-08 doc/chapters/conclusion.tex
typo

@@
 
-Thus either Stainless must find ways to translate more of built-in object-oriented patterns of Scala to their verification tool or developers must invest more in functional programming.
+Thus, either Stainless must find ways to translate more of built-in object-oriented patterns of Scala to their verification tool or developers must invest more in functional programming.
 
@@
 So we can not guarantee the correctness of the addition of Satoshis with zero in Bitcoin-S.
-There are changes that stays the same like the change from object to case object but there are other changes where we can not say if they are the same or not like the bound check in section \ref{sec:bound_check}.
+There are changes that stays the same as the change from object to case object but there are other changes where we can not say if they are the same or not like the bound check in section \ref{sec:bound_check}.

### 2019-06-08 doc/chapters/introduction.tex
Updated Introduction

@@
 
-In this work we use the verification framework Stainless to verify parts of the code of Bitcoin-S, a Scala implementation of the Bitcoin protocol.
-In the following we describe the main aspects of formal verification, Stainless and Bitcoin-S.
-
-\section{What is formal verification}
-\label{sec:formal_verification}
-
-// TODO move to chapter description and change Overview of stainless to formal verification with stainless
-
 The longer and more complex a source code is, the more difficult it is to verify its correctness.
 There are different approaches to show the correctness of a program.
-
-A powerful method to check whether a program works correctly is formal verification.
-This is a systematic process based on mathematical modeling.
-With formal verification all possible inputs are covered algorithmically.
-This is in contrast to testing, where only predefined inputs are tested.
+One of them is formal verification.
+In contrast to testing, where only predefined inputs are tested, with formal verification all possible inputs are covered.
 The correctness of a program is analyzed relative to its formal specification.
 Formal specification is a mathematical description of a software behavior that can be used by the verification tool to verify the code.
-We will see an example shortly.
+We will see an example of a formal specification shortly.
 
-One of such tools is Stainless which we use in our work.
+In this work we use the verification framework Stainless to verify parts of the code of Bitcoin-S, a Scala implementation of the Bitcoin protocol.
+In the following we look at the main aspects of formal verification with Stainless and properties of Bitcoin-S we want to verify.
 
-\section{Overview of Stainless}
+\section{Formal verification with Stainless}
 \label{sec:stainless}
@@
 Using this framework we can verify the correctness of Scala programs.
+The following overview of the framework is taken from the section \href{https://epfl-lara.github.io/stainless/intro.html}{Introduction} of the Stainless documentation \cite{Stainless:documentation}.
+
 Stainless statically verifies that a program satisfies a given specification and that a program will not crash at runtime.
-The framework explores all possible inputs and finds a counter examples for possible failures in a program which violate the given specification. \cite{Stainless:introduction}
+The framework explores all possible inputs and finds a counter examples for possible failures in a program which violate the given specification.
 
@@
 
-On invoking, Stainless tries to prove that the postcondition always holds, assuming the given precondition does hold. \cite{Stainless:introduction}
+On invoking, Stainless tries to prove that the postcondition always holds, assuming the given precondition does hold.
 
 The following example demonstrates a simple formal specification for the function calculating factorial.
-This is a modified example from the Stainless documentation, so Stainless reports an error verifying it.
+This is a modified example from the section \href{https://epfl-lara.github.io/stainless/intro.html}{Introduction} of the Stainless documentation \cite{Stainless:documentation}, so Stainless reports an error verifying it.
 \begin{lstlisting}[language=Scala]
@@
 Furthermore, Stainless checks for calls in the code to the function with invalid arguments violating the precondition.
-For example if we invoke factorial with a negative number it reports \textit{invalid}. \cite{Stainless:introduction}
+For example if we invoke factorial with a negative number it reports \textit{invalid}.
 
@@
 They call is Pure Scala.
-You can find the specification of Pure Scala on the Stainless website in the section \href{https://epfl-lara.github.io/stainless/purescala.html}{Pure Scala}.
+You can find the specification of Pure Scala on the Stainless documentation website \cite{Stainless:documentation} in the section \href{https://epfl-lara.github.io/stainless/purescala.html}{Pure Scala}.
 
@@
 Some of them are described on the website in the section \href{https://epfl-lara.github.io/stainless/library.html}{Stainless Library}.
-There are more details about library in the source code of Stainless on GitHub in the folder \href{https://github.com/epfl-lara/stainless}{\textit{frontends/library/stainless}}.
+There are more details about library in the source code of Stainless on GitHub \cite{Stainless:github} in the folder \href{https://github.com/epfl-lara/stainless/tree/master/frontends/library/stainless}{\textit{frontends/library/stainless}}.
 
 Stainless verifies Scala programs using SMT (satisfiability modulo theories) solvers.
-The framework transforms a program code to the Pure Scala fragment supported by Inox solver.
-This solver was also developed by LARA-Group and allows usage of different SMT solvers such as Z3, CVC4 and Princess, adding support of other higher-level features.
-Thus, preconditions, postconditions and assertions are translated into SMT formulas, so an SMT solver can determine if all properties hold.\cite{Stainless:introduction}\cite{Stainless}
 

### 2019-06-08 doc/chapters/introduction.tex
text

--- 2019-06-08 doc/chapters/introduction.tex
@@
 Stainless statically verifies that a program satisfies a given specification and that a program will not crash at runtime.
-The framework explores all possible inputs and finds a counter examples for possible failures in a program which violate the given specification. \cite{Stainless:introduction}
+The framework covers all possible inputs and finds a counter examples for possible failures in a program which violate the given specification. \cite{Stainless:introduction}
 
 The main functions used to write a specification are \textit{require} and \textit{ensuring}.
+
 With \textit{require} we define a precondition of a function, which we want to verify.
 \textit{require} is placed at the beginning of the function body.
-We pass a constraint for the inputs of a function as an argument to \textit{require} in the form of a Boolean expression.
+\textit{require} specifies, if the parameter of the function hold a certain condition.
 
-With \textit{ensuring} we provide a postcondition which is a condition on an output of the function.
-\textit{ensuring} is called after the function body.
+With \textit{ensuring}, we define the postcondition of a function.
+\textit{ensuring} is placed at the end of the function after the body.
+\textit{ensuting} specifies, if the return value or the result of the function hold a certain condition.
 
-On invoking, Stainless tries to prove that the postcondition always holds, assuming the given precondition does hold. \cite{Stainless:introduction}
+On invoking Stainless, it tries to prove that the postcondition always holds, assuming the given precondition does hold. \cite{Stainless:introduction}
 

### 2019-06-08 doc/chapters/introduction.tex

--- 2019-06-08 doc/chapters/introduction.tex
@@
 Stainless statically verifies that a program satisfies a given specification and that a program will not crash at runtime.
-The framework explores all possible inputs and finds a counter examples for possible failures in a program which violate the given specification.
+The framework covers all possible inputs and finds a counter examples for possible failures in a program which violate the given specification.
 
 The main functions used to write a specification are \textit{require} and \textit{ensuring}.
+
 With \textit{require} we define a precondition of a function, which we want to verify.
 \textit{require} is placed at the beginning of the function body.
-We pass a constraint for the inputs of a function as an argument to \textit{require} in the form of a Boolean expression.
+\textit{require} specifies, if the parameter of the function hold a certain condition.
 
-With \textit{ensuring} we provide a postcondition which is a condition on an output of the function.
-\textit{ensuring} is called after the function body.
+With \textit{ensuring}, we define the postcondition of a function.
+\textit{ensuring} is placed at the end of the function after the body.
+\textit{ensuting} specifies, if the return value or the result of the function hold a certain condition.
 
-On invoking, Stainless tries to prove that the postcondition always holds, assuming the given precondition does hold.
+On invoking Stainless, it tries to prove that the postcondition always holds, assuming the given precondition does hold.
 

### 2019-06-08 doc/chapters/trying_verify_checktx.tex
changed text

@@
 
-This chapter describes the part of Bitcoin-S needed to verify our property.
+This chapter describes the part of Bitcoin-S needed to verify the No Inflation Property described before.
 We are going to create a transaction and show you the relevant parts of the method checkTransaction, where transactions are checked against some properties.
@@
 
-Some of the code in this section is copied or adapted parts of Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
+Some of the code in this section is copied or adapted from Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
 Bitcoin-S-Core has a bitcoin transaction builder class with the following constructor:
@@
   BitcoinTxBuilder(
-    destinations: Seq[TransactionOutput], // where we send money
+    destinations: Seq[TransactionOutput], // where to send the money
     utxos: BitcoinTxBuilder.UTXOMap,      // unspent transaction outputs
@@
     network: BitcoinNetwork               // bitcoin network information
-  ): Future[BitcoinTxBuilder]             // sign TxBuilder to get tx
+  ): Future[BitcoinTxBuilder]
 \end{lstlisting}
 
-Here is how those parameters are generated.
+To return a Future does not make sense, since the implementing constructor calls either Future.successfull or Future.fromTry which returns an already resolved Future.
+This might be for future purposes.
 
-First, we need a some bitcoins/satoshis.
-Thus, we create a transaction with one single output.
-This transaction can be parsed from the bitcoin network but we create a transaction manually in order to see this process.
+Now we create a transaction.
 
-// TODO explain more. We create a fake transaction blabla
-
+First, we need a some money.
+Thus, we create a fake transaction with one single output.
+This transaction can be parsed from the bitcoin network but we create one manually in order to see this process.
 \begin{lstlisting}[language=scala]
@@
 
-On line one and two we are creating a new keypair to sign the next transaction and have a scriptPubKey where the bitcoins are.
+On line one and two we create a new keypair to sign the next transaction and have a scriptPubKey where the bitcoins are.
+This is our keypair.
+So the money is transferred to our public key.
 Line four specifies the amount of satoshis we have in the transaction.
@@
 
-Now that we have some bitcoins, we create the new transaction to spend them.
+Now that we have some bitcoins, we create the new transaction where we want to spend them.
 
 First, we need some out points.
-This are pointers to previous transactions' outputs.
+They point to outputs of previous transactions.
 We use the index zero, because the previous transaction has only one output that becomes the first index zero.
+If there were two previous outputs, the second output would become the index 1 and so on.
 \begin{lstlisting}[language=scala]
@@
 
+This utxs are the inputs of our transaction.
+
 Second, we need destinations to spend the bitcoins to.
@@
 
-We spend 5000 satoshis to the newly created script public key.
+We spend 5000 satoshis to the newly created random public key.
 
 Finally, we define the fee rate in satoshis per one byte transaction size as well as some bitcoin network parameters.
+The bitcoin network parameters are not important so we use some static values for testing purpose.
 \begin{lstlisting}[language=scala]
@@
 
-  val networkParams = RegTest // soem static values for testing
+  val networkParams = RegTest // some static values for testing
 \end{lstlisting}
@@
 \begin{lstlisting}[language=scala]
-  val txBuilder: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
-    destinations = destinations,
-    utxos = utxos,
-    feeRate = feeRate,
-    changeSPK = creditingSPK,
-    network = networkParams
+  val txBuilderF: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
+    destinations = destinations, // where to send the money
+    utxos = utxos,               // unspent transaction outputs
+    feeRate = feeRate,           // fee rate per byte
+    changeSPK = creditingSPK,    // where to send the change
+    network = networkParams      // bitcoin network information
   )
 
-  val signedTxF: Future[Transaction] = txBuilder
-    .flatMap(_.sign)
+  val signedTxF: Future[Transaction] = txBuilderF
+    .flatMap(_.sign)                       // call sign on the transaction builder
     .map {
@@
 
-Line one to seven creates a transaction builder which is then signed on line nine to thirteen.
-On line twelve we can now use our transaction object.
-For example, after calling the \emph{hex} function on it, we can send the returned string to the real bitcoin network.
+Line one to seven creates a transaction builder which is then signed on line ten.
+We can now use our transaction object on line twelve.
+For example, after calling \emph{hex} on it, we can send the returned string to the bitcoin network.
 
@@
 
-Bitcoin-S-Core offers a function called \emph{checkTransaction}.
-This is its type signature.
+Bitcoin-S offers a function called \emph{checkTransaction} located in the ScriptInterpreter object.
+This is its type signature:
 \begin{lstlisting}[language=scala]
@@
 \end{lstlisting}
+
 We can pass a transaction and it returns a Boolean whether the transaction is valid or not.
-So for example when we pass the built transaction from above to it the returned value would be true, because it's a valid transaction.
+So for example when we pass the transaction built before the returned value would be true, because it's a valid transaction.
+It might not be accepted by the bitcoin network but for a transaction on its own its valid.
+We can not check context with it, because we can only pass one transaction.
 
@@
 For example, it checks if there is either no input or no output.
-In this case it returns false.
+In this case we get false.
 
-The relevant part for the bug we found.
+The relevant part for the bug we found:
 \begin{lstlisting}[language=scala]
@@
 It gathers all transaction ids referenced by the out points.
-When calling \emph{distinct} on the returned list, it returns a list where duplicated are removed.
-Is the size of the new list is the same as the size of the old, we know that there was no duplicate transaction id.
-
-There is a bug, because checkTransaction checks on the transaction id that we are going to explain the next section.
+When we call \emph{distinct} on the returned list, we get a list with duplicate removed.
+If the size of the new list is the same as the size of the old, we know that there was no duplicate transaction id, because, as said, distinct removes the duplicates.
 
@@
 
-During the examination of the checkTransaction function, we found a bug in Bitcoin-S.
+We can see that there is a bug in the checkTransaction function from before, recognized and fixed through this work.
 
@@
 
-What happens, if a TransactionOutPoint (previousOutput) references the same Transaction ID (txId) another does?
+What happens, if we have two TransactionOutPoints (previousOutputs) with a different index but referencing the same Transaction ID (txId)?
 
@@
 A transaction can have multiple outputs that should be referenceable by the next transaction.
+So this is clearly a bug.
+
 What should not be possible is a transaction referencing the same output twice.
-This was a bug in Bitcoin Core known as CVE-2018–17144 which was patched on September 18, 2018. \cite{cve201817144}
+This bug occured in Bitcoin Core known as CVE-2018–17144 which was patched on September 18, 2018. \cite{cve201817144}
 
-Here, Bitcoin-S did a bit too much and marked all transaction as invalid, if they referenced a transaction twice in the next transaction.
+Here, Bitcoin-S did a bit too much and marked all transaction as invalid, if they referenced the same transaction twice.
 The fix is, to check on TransactionOutPoint instead of TransactionOutPoint.txId, because TransactionOutPoint contains the txId as well as the output index it references.
+So in pseudo code, we check on the tuple (tx, index) instead of (tx).
+The fixed code:
 \begin{lstlisting}[language=scala]
@@
 
-This was fixed in \href{https://github.com/bitcoin-s/bitcoin-s/pull/435}{pull request \mypound435} on GitHub at April 23, 2019 by us along a unit test to prevent this bug from appearing again in the future.
+This was fixed in \href{https://github.com/bitcoin-s/bitcoin-s/pull/435}{pull request \mypound435} on GitHub at April 23, 2019 through this work along a unit test to prevent this bug from appearing again in the future.
 

### 2019-06-08 doc/chapters/trying_verify_checktx.tex
more text

@@
 We are going to create a transaction and show you the relevant parts of the method checkTransaction, where transactions are checked against some properties.
-Then we see some troubles occurred during this work.
-In the end you will see how we fixed a bug in Bitcoin-S.
+Then, we will see the bug in Bitcoin-S found during this work and its fix.
+In the end we find reasons why it was needed for this work to change the verification from the No Inflation Property to the Addition with Zero Property.
 
@@
 
-\section{Adjusting Property}
+\section{Adjusting No Inflation Property}
 
-Trying to integrate Stainless in Bitcoin-S caused a lot of troubles, because of version conflicts.
-If you are interested in it you can read more in the appendix \ref{chap:appendix_arb}.
+Trying to integrate Stainless in Bitcoin-S caused a lot of troubles, mainly because of version conflicts.
+You can find more about it in chapter \ref{chap:appendix_arb}.
 

### 2019-06-08 doc/chapters/trying_verify_checktx.tex
added label

@@
 
-This chapter describes the part of Bitcoin-S needed to verify the No Inflation Property described before.
+This chapter describes the part of Bitcoin-S needed to verify the \nameref{property_1} described before.
 We are going to create a transaction and show you the relevant parts of the method checkTransaction, where transactions are checked against some properties.
 Then, we will see the bug in Bitcoin-S found during this work and its fix.
-In the end we find reasons why it was needed for this work to change the verification from the No Inflation Property to the Addition with Zero Property.
+In the end we see why the \nameref{property_1} needed to be changed to the \nameref{property_2}.
 

### 2019-06-08 doc/chapters/trying_verify_checktx.tex
more text

@@
 
-It takes too much time to do it and after some discussion, we decided to extract only the parts of the code that are needed for checkTransaction.
+After integrating Stainless in Bitcoin-S there are many errors.
+It takes too much time to fix them all so it should be easier to extract the classes needed for checkTransaction.
 
 The extracted code has more than 1500 lines.
-After running Stainless on it, it throws a really huge bunch of errors about what Stainless can not reason about.
-And after fixing some of those errors there appear new ones as we can see in the next chapter.
-This would require to change nearly everything of the extracted code.
+After running Stainless on it, it still throws a huge bunch of errors about what Stainless can not reason about.
+After fixing some of those errors there appear new ones.
+So this would require to change nearly everything of the extracted code.
 
-We arrived at the decision that we need a smaller part to verify.
-We have chosen to verify the addition with zero of Bitcoin-Ss CurrencyUnit class as described in the next chapter.
+Lets adjust the property from the \nameref{property_1} to the \nameref{property_2}, because there is not enough time to fix all this errors and write the verification.
+This is way less interesting to verify but gives still many insights in what needs to be done to verify some parts of Bitcoin-S and other source code.
+
+So lets look at the \nameref{property_2}.

### 2019-06-08 doc/chapters/towards_verifying_addition.tex
changed text

@@
 
-After realizing, that it would consume to much time to rewrite the Bitcoin-S code and even the extracted part with checkTransaction, we decided to extract the smallest unit in Bitcoin-S-Core that is worthwhile to verify.
-We found the addition of two Satoshis would be a good candidate.
-To make it even easier, we decided to verify only the addition of some satoshis with zero.
+After realizing, that it would consume to much time to rewrite the Bitcoin-S code and even the extracted part with checkTransaction, the smallest unit in Bitcoin-S-Core that is worthwhile to verify was extracted.
+This could be the addition of two CurrencyUnits.
+To make it even easier, the addition of CurrencyUnits with zero.
+CurrencyUnits is an abstract class in Bitcoin-S, representing currencies like Satoshis.
 
-The initial code is this:
+Here we can see the extracted code needed for the addition of CurrencyUnits:
 \lstinputlisting[language=scala]{../code/currsniporig/src/main/scala/code/initial/Code.scala}

### 2019-06-09 doc/chapters/challenge_stainless.tex
text
changed text

@@
 
-In this chapter we explore the practical challenges that could occur by using Stainless.
+In this chapter we look at the practical challenges that could occur by using Stainless.
+Since Stainless is still in development and there is no 1.x release there might still be breaking changes and improvements fixing problems discribed now.
 
@@
 Invoking the JAR on our source code Stainless will verify it.
-If we have a bigger project, this becomes really tricky because we must pass all files needed including the dependencies.
+If we have a bigger project, this becomes really tricky, because we must pass all files needed including the dependencies.
 This is in contrast to the sbt plugin, where we can integrate Stainless in our compilation process.
@@
 Now we can finally run Stainless on our code.
-But this leads us to the next findings described in the previous chapters.
+But this leads us to the next findings.
+We must rewrite most of the code, as described in the previous chapters.

### 2019-06-09 doc/chapters/introduction.tex
added label
text

@@
 
-\section{Overview of Bitcoin-S}
-\label{sec:bitcoin_s}
-
-In the work we verify some fragments of the code of Bitcoin-S.
-Bitcoin-S is an open source Scala implementation of the Bitcoin protocol. 
-There are different projects belonging to Bitcoin-S.
-Protocol data structures are defined in the package \textbf{bitcoin-s-core}.
-This package is the main target to be verified because it specifies the base of Bitcoin protocol, and it is crucial for correct protocol functionality.
-There is also the package \textbf{bitcoind-rpc} which is an RPC client implementation to communicate with \textit{bitcoind}, daemon of Bitcoin Core running on a machine.
-Developers of Bitcoin-S are working on the implementation of a lightweight bitcoin client in the package \textbf{bitcoin-s-spv-node}.
-The whole list of packages can be found on GitHub in the \href{https://github.com/bitcoin-s}{organization Bitcoin-S}.
-
-Bitcoin-S is a large project implementing many functionalities.
-In this work one of them is going to be examined and worked on, namely the property of the Bitcoin that a non-coinbase transaction can not generate new coins.
-Thus, the code of Bitcoin-S should be analyzed, and fragments implementing this feature should be identified.
-Afterwards, the fragments should be rewritten, so that Stainless can accept it.
-After that a formal specification for functions should be defined and annotated according Stainless libraries, so that Stainless can verify correctness of the code.
-
 \section{The Properties to Verify}
 
+Bitcoin-S is a large project found on GitHub implementing many functionalities.
+In this work two of them is going to be examined and worked on, namely the \nameref{property_1} and the \nameref{property_2} describer shortly.
+Thus, we go through the relevant code parts to verify those properties, look at some challenges with Stainless and see why we should write software with verification in mind.
+We will rewrite the code to Pure Scala, because Bitcoin-S is not written in it.
+
 \subsection{No Inflation Property}
+\label{property_1}
 
 \subsection{Addition with Zero Property}
+\label{property_2}

### 2019-06-10 doc/chapters/introduction.tex
fixed
fixed title
Added No Inflation Property.

@@
 
-\section{Formal verification with Stainless}
+\section{Formal Verification with Stainless}
 \label{sec:stainless}
@@
 Bitcoin-S is a large project found on GitHub implementing many functionalities.
-In this work two of them is going to be examined and worked on, namely the \nameref{property_1} and the \nameref{property_2} describer shortly.
+In this work we look at two of them, namely the \nameref{property_1} and the \nameref{property_2} described shortly.
 Thus, we go through the relevant code parts to verify those properties, look at some challenges with Stainless and see why we should write software with verification in mind.
@@
 \label{property_1}
+The decision to verify this property was inspired by the bug found in Bitcoin Core in September 2018  \cite{cve201817144} which could allow to spend the same unspent transaction output (UTXO) in a transaction multiple times.
+Hence, new coins could be created out of thin air.
+Thus, the property, that a non-coinbase transaction cannot generate new coins, is chosen to be verified in this work.
+
+As we can see later in chapter \ref{chap:verify_check}, we must rewrite the huge fragment of the code implementing this property.
+This reimplementation into Pure Scala needs a lot of time.
+So, we adjust the plan and verify another functionality of Bitcoin-S.
+Nevertheless, during the analysis of the No Inflation Property we have detected and fixed a bug in Bitcoin-S, what is described in section \ref{sec:bugfix}.
 

### 2019-06-10 doc/chapters/introduction.tex
Added Addition with Zero Property in introduction.tex

@@
 Thus, the property, that a non-coinbase transaction cannot generate new coins, is chosen to be verified in this work.
+We call it here as the No Inflation Property.
 
@@
 \label{property_2}
+In Bitcoin-S there is a datatype Satoshis representing an amount of bitcoins.
+We verify the addition of Satoshis with zero Satoshis.
+This operation should result in the same amount of Satoshis.
+We call it in the work as the Addition with Zero Property.
+
+Using Stainless, we successfully verify this property.
+But the process of the verification with the tool requires many changes in the code, so that Stainless can accept it.
+We look at all needed modifications in chapter \ref{chap:verify_add}.

### 2019-06-11 doc/chapters/introduction.tex
changed some lines

--- 2019-06-09 doc/chapters/introduction.tex
@@
 
+The decision to verify this property was inspired by the bug found in Bitcoin Core in September 2018  \cite{cve201817144} which allowed to spend the same unspent transaction output (UTXO) in a transaction multiple times.
+Hence, new coins could be created out of the air.
+Thus, we try to verify the property, that a non-coinbase transaction cannot generate new coins.
+Lets name it the \nameref{property_1}.
+
+As we can see later in chapter \ref{chap:verify_check}, we must rewrite a huge part of the code implementing this property.
+This reimplementation into Pure Scala needs a lot of time.
+So, we adjust the plan and verify another functionality of Bitcoin-S.
+Nevertheless, during the analysis of the \nameref{property_1} we look at a bug in Bitcoin-S found during this work and see the code changes for the bugfix in section \ref{sec:bugfix}.
+
 \subsection{Addition with Zero Property}
 \label{property_2}
+
+In Bitcoin-S there is a class Satoshis representing an amount of bitcoins.
+We look at the verification of the addition of Satoshis with zero Satoshis.
+This operation should result in the same amount of Satoshis.
+Lets call it the \nameref{property_2}.
+
+Using Stainless, we see the successfull verification of this property.
+But the process of the verification with the tool requires many changes in the code, so that Stainless can accept it.
+We look at all needed modifications in chapter \ref{chap:verify_add}.

### 2019-06-11 doc/chapters/trying_verify_checktx.tex
fixed title
Added No Inflation Property.
changed some lines

@@
-\chapter{Trying to Verify the no Inflation Property}
+\chapter{Trying to Verify the No Inflation Property}
 \label{chap:verify_check}
@@
 
-To return a Future does not make sense, since the implementing constructor calls either Future.successfull or Future.fromTry which returns an already resolved Future.
+The return type Future does not make sense here, since the implementing constructor calls either Future.successfull or Future.fromTry which returns an already resolved Future.
 This might be for future purposes.
@@
 \section{Fixing a Bug in Bitcoin-S}
+\label{sec:bugfix}
 

### 2019-06-11 doc/chapters/introduction.tex
typo

@@
 \textit{ensuring} is placed at the end of the function after the body.
-\textit{ensuting} specifies, if the return value or the result of the function hold a certain condition.
+\textit{ensuring} specifies, if the return value or the result of the function hold a certain condition.
 
@@
 
-The function recursively calculates factorial of an integer number.
+It recursively calculates factorial of an integer number.
 The input of the function is constrained with \textit{require} to a non-negative value.
@@
 
-Let's return back to our incorrect example with factorial.
+Let's return to our incorrect example with factorial.
 Stainless reports the following result for the factorial verification from our example:
@@
 Thus, we try to verify the property, that a non-coinbase transaction cannot generate new coins.
-Lets name it the \nameref{property_1}.
+Let's name it the \nameref{property_1}.
 
@@
 This operation should result in the same amount of Satoshis.
-Lets call it the \nameref{property_2}.
+Let's call it the \nameref{property_2}.
 
-Using Stainless, we see the successfull verification of this property.
+Using Stainless, we see the successful verification of this property.
 But the process of the verification with the tool requires many changes in the code, so that Stainless can accept it.

### 2019-06-11 doc/chapters/towards_verifying_addition.tex
typo
fixed title

@@
 
-After realizing, that it would consume to much time to rewrite the Bitcoin-S code and even the extracted part with checkTransaction, the smallest unit in Bitcoin-S-Core that is worthwhile to verify was extracted.
+After realizing, that it would consume too much time to rewrite the Bitcoin-S code and even the extracted part with checkTransaction, the smallest unit in Bitcoin-S-Core that is worthwhile to verify was extracted.
 This could be the addition of two CurrencyUnits.
@@
 
-The additions signature looks like this:
+The additions' signature looks like this:
 \begin{lstlisting}[language=scala]
@@
 Since there is no easy way to fix it and the code should stay as much as possible the original, we just remove the abstract type and fixed it to Int64.
-This limits the verification a bit, but as we only want to verify the addition in satoshis, thats OK.
+This limits the verification a bit, but as we only want to verify the addition in satoshis, that's OK.
 
@@
 This is easy.
-We just replace all occurence of A with BigInt, since A is not overrided in an implementing class.
-This is not exact the same code, because an implementing class could not override A anymore but thats fine for our verification.
+We just replace all occurrence of A with BigInt, since A is not overrode in an implementing class.
+This is not exact the same code, because an implementing class could not override A anymore but that's fine for our verification.
 
@@
 
-Now we get som weird warnings about require.
+Now we get some weird warnings about require.
 
-\section{Removeing Second Parameter of require}
+\section{Removing Second Parameter of Require}
 
@@
 Seems like Stainless does not support the second string parameter of require or at least it throws a warning about it.
-We can saftly remove it.
+We can safely remove it.
 
-A new error apperas.
+A new error appears.
 
@@
 
-This error is hard to understand but we can see that there is a missing Long to BigInt conversion.
+This error is hard to understand, but we can see that there is a missing Long to BigInt conversion.
 So we search for all longs in the code.
@@
 
-\section{Propagating require}
+\section{Propagating Require}
 
@@
 Sure, when adding one to the maximum Int64 the require does not hold anymore.
-Since we do only allow zero as a parameter the easiest way is to restrict it to zero here too.
+Since we do only allow zero as a parameter, the easiest way is to restrict it to zero here too.
 

### 2019-06-11 doc/chapters/trying_verify_checktx.tex
typo

@@
 
-Some of the code in this section is copied or adapted from Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
+Some code in this section is copied or adapted from Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
 Bitcoin-S-Core has a bitcoin transaction builder class with the following constructor:
@@
 
-The return type Future does not make sense here, since the implementing constructor calls either Future.successfull or Future.fromTry which returns an already resolved Future.
+The return type Future does not make sense here, since the implementing constructor calls either Future.successful or Future.fromTry which returns an already resolved Future.
 This might be for future purposes.
@@
 Thus, we create a fake transaction with one single output.
-This transaction can be parsed from the bitcoin network but we create one manually in order to see this process.
+This transaction can be parsed from the bitcoin network, but we create one manually in order to see this process.
 \begin{lstlisting}[language=scala]
@@
 
-This utxs are the inputs of our transaction.
+This utxos are the inputs of our transaction.
 
@@
 Finally, we define the fee rate in satoshis per one byte transaction size as well as some bitcoin network parameters.
-The bitcoin network parameters are not important so we use some static values for testing purpose.
+The bitcoin network parameters are not important, so we use some static values for testing purpose.
 \begin{lstlisting}[language=scala]
@@
 What should not be possible is a transaction referencing the same output twice.
-This bug occured in Bitcoin Core known as CVE-2018–17144 which was patched on September 18, 2018. \cite{cve201817144}
+This bug occurred in Bitcoin Core known as CVE-2018–17144 which was patched on September 18, 2018. \cite{cve201817144}
 
@@
 
-This was fixed in \href{https://github.com/bitcoin-s/bitcoin-s/pull/435}{pull request \mypound435} on GitHub at April 23, 2019 through this work along a unit test to prevent this bug from appearing again in the future.
+This was fixed in \href{https://github.com/bitcoin-s/bitcoin-s/pull/435}{pull request \mypound435} on GitHub at April 23, 2019, through this work along a unit test to prevent this bug from appearing again in the future.
 
@@
 
-After integrating Stainless in Bitcoin-S there are many errors.
+After integrating Stainless in Bitcoin-S, there were many errors.
 It takes too much time to fix them all so it should be easier to extract the classes needed for checkTransaction.
@@
 After running Stainless on it, it still throws a huge bunch of errors about what Stainless can not reason about.
-After fixing some of those errors there appear new ones.
-So this would require to change nearly everything of the extracted code.
+After fixing some of those errors, there appear new ones.
+So this would require changing nearly everything of the extracted code.
 
-Lets adjust the property from the \nameref{property_1} to the \nameref{property_2}, because there is not enough time to fix all this errors and write the verification.
+Let's adjust the property from the \nameref{property_1} to the \nameref{property_2}, because there is not enough time to fix all these errors and write the verification.
 This is way less interesting to verify but gives still many insights in what needs to be done to verify some parts of Bitcoin-S and other source code.
 
-So lets look at the \nameref{property_2}.
+So let's look at the \nameref{property_2}.

### 2019-06-11 doc/chapters/challenge_stainless.tex
typo

@@
 In this chapter we look at the practical challenges that could occur by using Stainless.
-Since Stainless is still in development and there is no 1.x release there might still be breaking changes and improvements fixing problems discribed now.
+Since Stainless is still in development and there is no 1.x release there might still be breaking changes and improvements fixing problems described now.
 
@@
 Downgrading Bitcoin-S sbt version to 0.13.17 fixes the error but then it can not load some libraries only compiled for newer versions.
-So this would take too much time to fix and changes the Bitcoin-S code unwantingly.
+So this would take too much time to fix and changes the Bitcoin-S code inadvertent.
 

### 2019-06-11 code/currsniporig/src/main/scala/code/end/Code.scala
added some comments

@@ -61,13 +61,13 @@
     def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
 
     override def +(c: CurrencyUnit): CurrencyUnit = {
-      require(
+      require(                                       // <--- specification
            c.satoshis == Satoshis.zero
         && this.underlying.toBigInt <= BigInt("9223372036854775807")
         && this.underlying.toBigInt >= BigInt("-9223372036854775808")
       )
       Satoshis(satoshis.underlying + c.satoshis.underlying)
-    } ensuring(res => res.satoshis == this.satoshis)
+    } ensuring(res => res.satoshis == this.satoshis) // <--- specification
 
     protected def underlying: Int64
   }

### 2019-06-12 doc/chapters/towards_verifying_addition.tex
clearer text

@@
 
-This is, due to the internal design of Scala and Java.
+This is, due to the internal design of Scala.
 It's possible to reason about case object but not about object.
@@
 
-  sealed abstract class Satoshis extends CurrencyUnit
+  sealed abstract class Satoshis extends CurrencyUnit {
+    override type A = Int64
+  }
 \end{lstlisting}
@@
 
-Satoshis extends CurrencyUnit with type Int64, so it is of type CurrencyUnit[Int64].
-That's too specific, because the addition would require CurrencyUnits of type CurrencyUnit[A] not CurrencyUnit[Int64].
+Satoshis extends CurrencyUnit with type Int64, so it would be of type CurrencyUnit[Int64].
+That's too specific, because the return type of the addition is then CurrencyUnit[Int64] not CurrencyUnit[A].
+Maybe the Bitcoin-S developers should reimplement this part and not use Satoshis directly.
 
-Since there is no easy way to fix it and the code should stay as much as possible the original, we just remove the abstract type and fixed it to Int64.
+Since there is no easy way to fix it and the code should stay as much as possible the original, we just remove the abstract type and set it to Int64.
 This limits the verification a bit, but as we only want to verify the addition in satoshis, that's OK.

### 2019-06-12 doc/chapters/towards_verifying_addition.tex
ore explanation

@@
 The BigInt from the Stainless library is a bit restricted.
-It does not support dynamic BigInt construction.
-Only from literal arguments.
+One such restriction is, that BigInt does not support dynamic BigInt construction.
+Thus, the constructor parameter of BitInt must be a literal arguments.
 
@@
 
-This would be really hard to refactor, because Bitcoin-S tries to be as much dynamic as possible so it can be used with other cryptocurrencies than bitcoins too.
-Maybe it would even be impossible, because they need to parse a lot from the bitcoin network.
+This would be really hard to refactor, because Bitcoin-S tries to be as much dynamic as possible so it can be used with other cryptocurrencies than bitcoins.
+Maybe it would even be impossible, because they need to parse dynamic values from the bitcoin network.
 
 Luckily, we can use toBigInt on underlying directly instead of going over toLong.
+So, instead of converting the underlying to Long and and back to BigInt we convert underlying directly to BigInt.
 

### 2019-06-12 doc/chapters/towards_verifying_addition.tex
more explnation

@@
 It does not support upper type bound on the class itself.
-To track this, there is the \href{https://github.com/epfl-lara/stainless/issues/519}{issue \mypound519} on GitHub.
+To track this, \href{https://github.com/epfl-lara/stainless/issues/519}{issue \mypound519} was created on GitHub during this work.
 \begin{lstlisting}[language=scala]
@@
 For the sake of convenience we make this a concrete type by replacing T with Int64.
+Int64,  because Satoshis uses only Int64.
+There are other number typs like UInt16 but for our property we don't need them.
 

### 2019-06-12 doc/chapters/towards_verifying_addition.tex
explanation

@@
 We just replace all occurrence of A with BigInt, since A is not overrode in an implementing class.
-This is not exact the same code, because an implementing class could not override A anymore but that's fine for our verification.
+This is not exact the same code, because an implementing class can not override A anymore but that's fine for our verification.
 
-This was a missing feature of Stainless that was fixed on Mai 28, 2019 with \href{https://github.com/epfl-lara/stainless/pull/470}{pull request \mypound470} on GitHub.
+This was a missing feature in Stainless that was fixed on Mai 28, 2019 with \href{https://github.com/epfl-lara/stainless/pull/470}{pull request \mypound470} on GitHub.
 Now it should work without this change.
@@
 
-Due to the restrictions on BigInt, we can not use the \& function on.
+Due to the restrictions on BigInt, we can not use the \& function either.
 Simplified code:
@@
 
-So, we can replace the \& mask with a bound check whether the result is in range of Long.MinValue and Long.MaxValue.
+So, we can replace the \& mask with a bound check whether the result is in range of Long.MinValue and Long.MaxValue, because Int64 has the same 64-bit range as Long.
 Again the code gets a bit more static and it's not the exact same code anymore.

### 2019-06-12 doc/chapters/towards_verifying_addition.tex
explain more

@@
 
-This is an easy to fix.
+This is easy to fix.
 We just extract the private class out of the object.
-This is not exactly the same code but since the class is still private it can not be extended outside of the file.
+This is not exactly the same code, because other classes in the same file could now access the private class.
+But for our property it does not change anything.
 
@@
 Seems like Stainless does not support the second string parameter of require or at least it throws a warning about it.
-We can safely remove it.
+We can safely remove the string parameters from the requires, since they only serve as error message.
 
@@
 This error is hard to understand, but we can see that there is a missing Long to BigInt conversion.
-So we search for all longs in the code.
+So we search for all Long values in the code.
 \begin{lstlisting}[language=scala]
@@
 Looks like it can not compare a BigInt with a Long value.
-We can easily convert this Long value to a BigInt with string literals parameter.
+We can easily convert this Long value to a BigInt with string literal parameter by passing this numbers as strings to the BigInt constructor.
 
@@
 
-This show, that there is an invalid specification in checkResult and Stainless prints a counter example for it.
+This show, that there is an invalid specification in checkResult and Stainless prints a counterexample for it.
 
-Let's ignore this for a moment and write the verification for our addition with zero.
+Let's ignore this for a moment and write the specification for the \nameref{property_2}.
 

### 2019-06-12 doc/chapters/challenge_stainless.tex
Corrected english

@@
 
-We can either use the sbt plugin or a JAR file to check the code with Stainless.
+We can either use the sbt plugin or a JAR file to check code with Stainless.
 
@@
 
-Having a fixed version configured in the sbt build file, every developer has the same Stainless features available.
+Having a static version configured in the sbt build file, every developer has the same Stainless features available.
 This should prevent incompatibility with new or deprecated features when we use different plugins.
@@
 
-However, there are some drawbacks currently.
+However, currently there are some drawbacks.
 For example the sbt plugin does not always report errors.
-We will read more about that shortly.
+We will see more about that shortly.
 
@@
 
-The bigger problem is, due to the different Scala and sbt versions, the error message, after trying to go in a sbt shell.
+The bigger problem is, due to the different Scala and sbt versions, the following error after trying to go in a sbt shell:
 \begin{verbatim}
@@
 Downgrading Bitcoin-S sbt version to 0.13.17 fixes the error but then it can not load some libraries only compiled for newer versions.
-So this would take too much time to fix and changes the Bitcoin-S code inadvertent.
+So this would take too much time to fix and changes the Bitcoin-S code inadvertently.
 
-The next approach is to use stainless cli instead of sbt.
-Running stainless on all source files does not work, because the dependencies are missing.
+The next approach is to use the stainless cli instead of sbt.
+Running stainless on all source files does not work, because dependencies are missing.
 The parameter \emph{-classpath} can resolve it but the value of this parameter must be the paths to all the dependencies separated by a ':'.
@@
 \emph{.ivy2} is the dependency cache of sbt.
-The \emph{tr} replaces the first char with the second so '\\n' with either ':' or ' '.
+The \emph{tr} replaces the first char with the second so a newline with either ':' or ' '.
 
@@
 Let's go another way, because the errors may take too much time and it might lead to a next error.
-We extract the code needed to verify a transaction mainly the class Transaction and ScriptInterpreter with many other classes their depending on.
+We extract the code needed to verify a transaction mainly the class Transaction and ScriptInterpreter with many other classes they're depending on.
 
-After this extraction Stainless was successfully integrated with both, sbt and JAR.
+After this extraction Stainless was successfully integrated with both sbt and JAR.
 
@@
 But it also ended with no output.
-So we can not able to change the code so Stainless would accept it since we do not know what to change.
+So we are not able to change the code so Stainless would accept it since we do not know what to change.
 

### 2019-06-12 doc/chapters/conclusion.tex
Corrected english

--- 2019-05-28 doc/chapters/conclusion.tex
@@
 
-\section{Future Work}
+As we saw, it is really hard to verify software not written with verification in mind.
+It needs a lot of changes in the software because verification is mathematical and the current software is written mostly in object-oriented style.
+Software written in the functional paradigm would be much easier to reason about.
+
+Thus, either Stainless must find ways to translate more of built-in object-oriented patterns of Scala to their verification tool or developers must invest more in functional programming.
+
+The code we verified in the previous chapter is not the original code from Bitcoin-S.
+So we can not guarantee the correctness of the addition of Satoshis with zero in Bitcoin-S.
+Not all changes we made were as trivial as the replacement of objects with case objects.
+For these non-trivial changes, as seen in section \ref{sec:bound_check}, we cannot say whether they are equivalent to the original implementation or not.

### 2019-06-12 doc/chapters/introduction.tex
Corrected english

@@
 \textit{require} is placed at the beginning of the function body.
-\textit{require} specifies, if the parameter of the function hold a certain condition.
+\textit{require} specifies, if the parameter of the function holds against a certain condition.
 
@@
 \textit{ensuring} is placed at the end of the function after the body.
-\textit{ensuring} specifies, if the return value or the result of the function hold a certain condition.
+\textit{ensuring} specifies, if the return value or the result of the function satisfies a certain condition.
 
@@
 
-The following example demonstrates a simple formal specification for the function calculating factorial.
+The following example demonstrates a simple formal specification for the function calculating a factorial.
 This is a modified example from the section \href{https://epfl-lara.github.io/stainless/intro.html}{Introduction} of the Stainless documentation \cite{Stainless:documentation}, so Stainless reports an error verifying it.
@@
 
-It recursively calculates factorial of an integer number.
+It recursively calculates the factorial of an integer number.
 The input of the function is constrained with \textit{require} to a non-negative value.
@@
 Stainless supports only a subset of Scala.
-They call is Pure Scala.
+They call this Pure Scala.
 You can find the specification of Pure Scala on the Stainless documentation website \cite{Stainless:documentation} in the section \href{https://epfl-lara.github.io/stainless/purescala.html}{Pure Scala}.
 
-In addition, Stainless has its own library with annotations, reimplementaion of some core data types, collections and input-output functions and more.
+In addition, Stainless has its own library with annotations, reimplementation of some core data types, collections and input-output functions and more.
 Some of them are described on the website in the section \href{https://epfl-lara.github.io/stainless/library.html}{Stainless Library}.
-There are more details about library in the source code of Stainless on GitHub \cite{Stainless:github} in the folder \href{https://github.com/epfl-lara/stainless/tree/master/frontends/library/stainless}{\textit{frontends/library/stainless}}.
+There are more details about the library in the source code of Stainless on GitHub \cite{Stainless:github} in the folder \href{https://github.com/epfl-lara/stainless/tree/master/frontends/library/stainless}{\textit{frontends/library/stainless}}.
 
@@
 Thus, we go through the relevant code parts to verify those properties, look at some challenges with Stainless and see why we should write software with verification in mind.
-We will rewrite the code to Pure Scala, because Bitcoin-S is not written in it.
+In order to be able to verify Bitcoin-S, we have to rewrite parts of it in the subset of Scala supported by Stainless.
 
@@
 
-In Bitcoin-S there is a class Satoshis representing an amount of bitcoins.
+In Bitcoin-S there is a class \texttt{Satoshis} representing an amount of bitcoins.
 We look at the verification of the addition of Satoshis with zero Satoshis.

### 2019-06-12 doc/chapters/towards_verifying_addition.tex
changed title
Corrected english

@@
 
-After realizing, that it would consume too much time to rewrite the Bitcoin-S code and even the extracted part with checkTransaction, the smallest unit in Bitcoin-S-Core that is worthwhile to verify was extracted.
+After realizing that it would consume too much time to rewrite the Bitcoin-S code and even the extracted part with checkTransaction, the smallest unit in Bitcoin-S-Core that is worthwhile to verify was extracted.
 This could be the addition of two CurrencyUnits.
@@
 Here, we can just change the objects from object to case object.
-Stainless recommendation is to use objects for modules and case objects as algebraic data type.
+Stainless recommendation is to use objects for modules and case objects as algebraic data types.
 
-This is, due to the internal design of Scala.
+This is due to the internal design of Scala.
 It's possible to reason about case object but not about object.
@@
 Unfortunately not.
-The problem is, CurrencyUnit uses one of its implementing class Satoshis.
+The problem is, CurrencyUnit uses one of its implementing classes: Satoshis.
 
@@
 One such restriction is, that BigInt does not support dynamic BigInt construction.
-Thus, the constructor parameter of BitInt must be a literal arguments.
+Thus, the constructor parameter of BitInt must be a literal argument.
 
@@
 
-This would be really hard to refactor, because Bitcoin-S tries to be as much dynamic as possible so it can be used with other cryptocurrencies than bitcoins.
-Maybe it would even be impossible, because they need to parse dynamic values from the bitcoin network.
+This would be really hard to refactor, because Bitcoin-S tries to be as dynamic as possible so it can be used with cryptocurrencies other than bitcoins.
+Maybe it could be impossible, because they need to parse dynamic values from the bitcoin network.
 
-Luckily, we can use toBigInt on underlying directly instead of going over toLong.
-So, instead of converting the underlying to Long and and back to BigInt we convert underlying directly to BigInt.
+Luckily, we can use toBigInt on the field \texttt{underlying} directly instead of toLong.
+So, instead of converting the underlying to Long and back to BigInt we convert underlying directly to BigInt.
 
@@
 This is a missing feature in Stainless.
-It does not support upper type bound on the class itself.
+It does not support upper type boundaries on the class itself.
 To track this, \href{https://github.com/epfl-lara/stainless/issues/519}{issue \mypound519} was created on GitHub during this work.
@@
 
-For the sake of convenience we make this a concrete type by replacing T with Int64.
+Despite this, in order to be able to continue, we make this a concrete type by replacing T with Int64.
 Int64,  because Satoshis uses only Int64.
-There are other number typs like UInt16 but for our property we don't need them.
+There are other number types like UInt16 but for our property we don't need them.
 
@@
 This is easy.
-We just replace all occurrence of A with BigInt, since A is not overrode in an implementing class.
-This is not exact the same code, because an implementing class can not override A anymore but that's fine for our verification.
+We just replace all occurrence of A with BigInt, since A is not overwritten in an implementing class.
+This is not the exact same code, because an implementing class can not override A anymore but that's fine for our verification.
 
@@
 
-This is a bound check.
-It checks, if the result of the addition is in range of the specified type, which is now the hard coded Int64.
+This is a bounds check.
+It checks if the result of the addition is in range of the specified type, which is now the hard coded Int64.
 
@@
 Seems like Stainless does not support the second string parameter of require or at least it throws a warning about it.
-We can safely remove the string parameters from the requires, since they only serve as error message.
+We can safely remove the string parameters from the requires, since they only serve as error messages.
 
@@
 Looks like it can not compare a BigInt with a Long value.
-We can easily convert this Long value to a BigInt with string literal parameter by passing this numbers as strings to the BigInt constructor.
+We can easily convert this Long value to a BigInt with a string literal parameter by passing these numbers as strings to the BigInt constructor.
 
@@
 
-This show, that there is an invalid specification in checkResult and Stainless prints a counterexample for it.
+This shows that there is an invalid specification in checkResult and Stainless prints a counterexample for it.
 
@@
 
-\section{Writing Specification for Verification}
+\section{Writing Specification for the Property}
 

### 2019-06-12 doc/chapters/trying_verify_checktx.tex
Corrected english

@@
 This chapter describes the part of Bitcoin-S needed to verify the \nameref{property_1} described before.
-We are going to create a transaction and show you the relevant parts of the method checkTransaction, where transactions are checked against some properties.
+We are going to create a transaction and show the relevant parts of the method \texttt{checkTransaction}, where transactions are checked against some properties.
 Then, we will see the bug in Bitcoin-S found during this work and its fix.
@@
 
-Some code in this section is copied or adapted from Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
+Some code in this section is copied or adapted from the Bitcoin-S-Core transaction builder example \cite{BitcoinSCore:txbuilderexample}.
 Bitcoin-S-Core has a bitcoin transaction builder class with the following constructor:
@@
 
-The return type Future does not make sense here, since the implementing constructor calls either Future.successful or Future.fromTry which returns an already resolved Future.
+The return type Future does not make sense here, since the implementation calls either Future.successful or Future.fromTry which returns an already resolved Future.
 This might be for future purposes.
@@
 
-First, we need a some money.
+First, we need some money.
 Thus, we create a fake transaction with one single output.
@@
 Finally, we define the fee rate in satoshis per one byte transaction size as well as some bitcoin network parameters.
-The bitcoin network parameters are not important, so we use some static values for testing purpose.
+The bitcoin network parameters are not important, so we use some static values normally used when testing.
 \begin{lstlisting}[language=scala]
@@
 
-We can pass a transaction and it returns a Boolean whether the transaction is valid or not.
-So for example when we pass the transaction built before the returned value would be true, because it's a valid transaction.
-It might not be accepted by the bitcoin network but for a transaction on its own its valid.
+We can pass a transaction and it returns a Boolean indicating whether the transaction is valid or not.
+So for example when we pass the transaction we built before the returned value would be true, because it's a valid transaction.
+It might not be accepted by the bitcoin network but for a transaction on its own it's valid.
 We can not check context with it, because we can only pass one transaction.
@@
 
-Here the relevant code of checkTransaction again:
+Here is the relevant code of checkTransaction again:
 \begin{lstlisting}[language=scala]
@@
 
-What happens, if we have two TransactionOutPoints (previousOutputs) with a different index but referencing the same Transaction ID (txId)?
+What happens if we have two TransactionOutPoints (previousOutputs) with a different index but referencing the same Transaction ID (txId)?
 
@@
 
-This was fixed in \href{https://github.com/bitcoin-s/bitcoin-s/pull/435}{pull request \mypound435} on GitHub at April 23, 2019, through this work along a unit test to prevent this bug from appearing again in the future.
+This was fixed in \href{https://github.com/bitcoin-s/bitcoin-s/pull/435}{pull request \mypound435} on GitHub at April 23, 2019, through this work along with a unit test to prevent this bug from appearing again in the future.
 
@@
 Trying to integrate Stainless in Bitcoin-S caused a lot of troubles, mainly because of version conflicts.
-You can find more about it in chapter \ref{chap:appendix_arb}.
+For more details see chapter \ref{chap:appendix_arb}.
 
 After integrating Stainless in Bitcoin-S, there were many errors.
-It takes too much time to fix them all so it should be easier to extract the classes needed for checkTransaction.
+It takes too much time to fix them all so it should be easier to extract the classes needed for the \texttt{checkTransaction} function.
 

### 2019-06-12 doc/chapters/conclusion.tex
changed conclusion

@@
 
-As we saw, it is really hard to verify software not written with verification in mind.
-It needs a lot of changes in the software because verification is mathematical and the current software is written mostly in object-oriented style.
+Because of the limitations of the verication tool, we could only verify a rewritten version of the original Bitcoin-S code.
+So we can not guarantee the correctness of the addition of Satoshis with zero in Bitcoin-S.
+Not all changes we made were as trivial as the replacement of objects with case objects.
+For these non-trivial changes, as seen for example the bound check in section \ref{sec:bound_check}, we cannot say whether they are equivalent to the original implementation or not.
+
+So code should be written specically with formal verication in mind, in order to successfully verify it.
+Otherwise, it needs a lot of changes in the software because verification is mathematical and the current software is written mostly in object-oriented style.
 Software written in the functional paradigm would be much easier to reason about.
@@
 
-The code we verified in the previous chapter is not the original code from Bitcoin-S.
-So we can not guarantee the correctness of the addition of Satoshis with zero in Bitcoin-S.
-Not all changes we made were as trivial as the replacement of objects with case objects.
-For these non-trivial changes, as seen in section \ref{sec:bound_check}, we cannot say whether they are equivalent to the original implementation or not.
+Also, we found that trying to verify code reveals bugs as shown in section \ref{sec:bugfix}. Finally, our work led to some feedback to the Stainless developers to improve the tool.

### 2019-06-13 doc/chapters/towards_verifying_addition.tex
added ,
Corrected properties names

@@
-\chapter{Towards Verifying the Addition with Zero Property}
+\chapter{Towards Verifying the Addition-with-Zero Property}
 \label{chap:verify_add}
@@
 This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
-The \href{https://github.com/epfl-lara/stainless/issues/520}{issue \mypound520} on Stainless GitHub gives some thoughts if you want to know more.
+The \href{https://github.com/epfl-lara/stainless/issues/520}{issue \mypound520} on Stainless GitHub gives some thoughts, if you want to know more.
 

### 2019-06-19 doc/chapters/towards_verifying_addition.tex
added todos

@@
 CurrencyUnits is an abstract class in Bitcoin-S, representing currencies like Satoshis.
+
+\section{Extracting the relevant Code}
+
+\todo{ramon: write this section}
+\todo{ramon: keep close to original code}
+\todo{ramon: for every transformation: before and after}
+\todo{ramon,anna: for every transformation: why does it preserve the semantics}
+\todo{ramon,anna: add comments}
+\todo{ramon,anna: explain what the code does and why}
 
@@
 
-When we run Stainless on it, it throws the following errors:
+When we run Stainless on this code (without any properties to prove), it throws the following errors:
 {\footnotesize\begin{verbatim}

### 2019-06-19 paper/paper.tex
initial paper commit

\documentclass[runningheads]{llncs}

\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%

\graphicspath{{./images/}}

\usepackage{hyperref}
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}

%\usepackage{courier}
\usepackage{listings}
\usepackage[dvipsnames]{xcolor}
\colorlet{link}{MidnightBlue}
\hypersetup{
  colorlinks   = true,
  linkbordercolor=link,
  pdfborderstyle={/S/U/W .5},% border style will be underline of width 1pt
     linkcolor  = link,
     citecolor  = link,
     urlcolor   = link
}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{
  language=scala,
  basicstyle=\footnotesize\ttfamily,
  breaklines=true,
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  numbers=left,
  %frame=single, % Border around box
  numbersep=-7pt,
  numberstyle=\color{gray},
  stringstyle=\color{mauve}
}

\newcommand{\todo}[1]{{\par \color{red}#1}}

\begin{document}

\title{Towards Verifying the Bitcoin-S Library}

\author{Ramon Boss \and Kai Brünnler \and Anna Doukmak}

\authorrunning{R. Boss et al.}
\institute{Bern University of Applied Sciences, CH-2501 Biel, Switzerland
\email{\{ramon.boss,kai.bruennler,anna.doukmak\}@bfh.ch}}

\maketitle

\begin{abstract}
  We try to verify properties of the Bitcoin-S library, a Scala
  implementation of parts of the Bitcoin protocol. We use the
  Stainless verifier which supports programs in subset of Scala called
  the \emph{Pure Scala Fragment}. We first try to verify the property
  that regular transactions never create new money. It turns out that
  there is too much code involved that lies outside of the supported
  fragment to make this feasible. However, in the process we uncover
  and fix a bug in Bitcoin-S. We then turn to a much simpler (and less
  interesting) property: that adding zero satoshis to a given amount
  of satoshis yields the given amount of satoshis. Here as well a
  significant part of the relevant code lies outside of the supported
  fragment. However, after a series of equivalent transformations, we
  arrive at code that we successfully verify.

\keywords{Bitcoin  \and Scala \and Bitcoin-S \and Stainless.}
\end{abstract}

\section{Introduction}

\todo{kai: motivation, bitcoin inflation bug, scala better suited for formal verification, bitcoin-s handling millions of dollars, this is a bit of a case study}

\subsection{Formal Verification with Stainless}
\label{sec:stainless}

kai: successor of Leon verifier \cite{DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}

Stainless is developed by "Lab for Automated Reasoning and Analysis"
(LARA) at EPFL's School of Computer and Communication Sciences.  Using
this framework we can verify the correctness of Scala programs.  The
following overview of the framework is taken from the section
\href{https://epfl-lara.github.io/stainless/intro.html}{Introduction}
of the Stainless documentation \cite{Stainless:documentation}.

Stainless statically verifies that a program satisfies a given
specification and that a program will not crash at runtime.  The
framework covers all possible inputs and finds a counter examples for
possible failures in a program which violate the given specification.

The main functions used to write a specification are \textit{require}
and \textit{ensuring}.

With \textit{require} we define a precondition of a function, which we
want to verify.  \textit{require} is placed at the beginning of the
function body.  \textit{require} specifies, if the parameter of the
function holds against a certain condition.

With \textit{ensuring}, we define the postcondition of a function.
\textit{ensuring} is placed at the end of the function after the body.
\textit{ensuring} specifies, if the return value or the result of the
function satisfies a certain condition.

On invoking Stainless, it tries to prove that the postcondition always
holds, assuming the given precondition does hold.

The following example demonstrates a simple formal specification for
the function calculating a factorial.  This is a modified example from
the section
\href{https://epfl-lara.github.io/stainless/intro.html}{Introduction}
of the Stainless documentation \cite{Stainless:documentation}, so
Stainless reports an error verifying it.
\begin{lstlisting}[language=Scala]
  def factorial(n: Int): Int = {
      require(n >= 0)
      if (n == 0) {
        1
      } else {
        n * factorial(n - 1)
      }
  } ensuring(res => res >= 0)
\end{lstlisting}

It recursively calculates the factorial of an integer number.
The input of the function is constrained with \textit{require} to a non-negative value.
The result should also be non-negative.
Thus, Stainless will verify whether the result of the factorial calculation is non-negative for all non-negative inputs.

Stainless can produce 3 outcomes of postcondition verification:
\textit{valid}, \textit{invalid} and \textit{unknown}.  If the
postcondition is \textit{valid} Stainless could prove that for any
inputs constrained in the precondition, the postcondition always
holds.  Reporting the postcondition as \textit{invalid} the framework
could find at least one counterexample which satisfies the
precondition but violates the postcondition.  If Stainless is unable
to prove the postcondition or find a counterexample it reports the
outcome \textit{unknown}.  In this case a timeout or an internal error
occurred.  Furthermore, Stainless checks for calls in the code to the
function with invalid arguments violating the precondition.  For
example if we invoke factorial with a negative number it reports
\textit{invalid}.

Let's return to our incorrect example with factorial.
Stainless reports the following result for the factorial verification from our example:
\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{output1.png}
	\caption{Output of Stainless verification for calculating factorial of Int number}
	\label{fig:output1}
\end{figure}

Stainless disproves the postcondition and gives the number 17 as a
counterexample.  Due to an overflow in the 32 bit Int the result
becomes a negative value.  The program would work correct by changing
the type from Int to BigInt.

Stainless supports only a subset of Scala. \todo{Anna It supports
  blabla, in particular it supports not blablabla. (check class with
  var)}

\todo{why does stainless use its own BigInt etc?}

They call this Pure Scala.  You can find the specification of Pure
Scala on the Stainless documentation website
\cite{Stainless:documentation} in the section
\href{https://epfl-lara.github.io/stainless/purescala.html}{Pure
  Scala}.

In addition, Stainless has its own library with annotations,
reimplementation of some core data types, collections and input-output
functions and more.  Some of them are described on the website in the
section
\href{https://epfl-lara.github.io/stainless/library.html}{Stainless
  Library}.  There are more details about the library in the source
code of Stainless on GitHub \cite{Stainless:github} in the folder
\href{https://github.com/epfl-lara/stainless/tree/master/frontends/library/stainless}{\textit{frontends/library/stainless}}.

Stainless verifies Scala programs using SMT (satisfiability modulo theories) solvers.

\subsection{The Properties to Verify}

Bitcoin-S is a large project found on GitHub implementing many
functionalities.  In this work we look at two of them, namely the
\nameref{property_1} and the \nameref{property_2} described shortly.
Thus, we go through the relevant code parts to verify those
properties, look at some challenges with Stainless and see why we
should write software with verification in mind.  In order to be able
to verify Bitcoin-S, we have to rewrite parts of it in the subset of
Scala supported by Stainless.

\subsubsection{No-Inflation Property}
\label{property_1}

The decision to verify this property was inspired by the bug found in
Bitcoin Core in September 2018 \cite{cve201817144} which allowed to
spend the same unspent transaction output (UTXO) in a transaction
multiple times.  Hence, new coins could be created out of the air.
Thus, we try to verify the property, that a non-coinbase transaction
cannot generate new coins.  Let's name it the \nameref{property_1}.

As we can see later in chapter \ref{chap:verify_check}, we must
rewrite a huge part of the code implementing this property.  This
reimplementation into Pure Scala needs a lot of time.  So, we adjust
the plan and verify another functionality of Bitcoin-S.  Nevertheless,
during the analysis of the \nameref{property_1} we look at a bug in
Bitcoin-S found during this work and see the code changes for the
bugfix in section \ref{sec:bugfix}.

\subsubsection{Addition-with-Zero Property}
\label{property_2}

In Bitcoin-S there is a class \texttt{Satoshis} representing an amount of bitcoins.
We look at the verification of the addition of Satoshis with zero Satoshis.
This operation should result in the same amount of Satoshis.
Let's call it the \nameref{property_2}.

Using Stainless, we see the successful verification of this property.
But the process of the verification with the tool requires many
changes in the code, so that Stainless can accept it.  We look at all
needed modifications in chapter \ref{chap:verify_add}.

\section{Trying to Verify the No-Inflation Property}
\label{chap:verify_check}

This chapter describes the part of Bitcoin-S needed to verify the
\nameref{property_1} described before.  We are going to create a
transaction and show the relevant parts of the method
\texttt{checkTransaction}, where transactions are checked against some
properties.  Then, we will see the bug in Bitcoin-S found during this
work and its fix.  In the end we see why the \nameref{property_1}
needed to be changed to the \nameref{property_2}.

\subsection{Creation of a Transaction}

Some code in this section is copied or adapted from the Bitcoin-S-Core
transaction builder example \cite{BitcoinSCore:txbuilderexample}.
Bitcoin-S-Core has a bitcoin transaction builder class with the
following constructor:
\begin{lstlisting}[language=scala]
  BitcoinTxBuilder(
    destinations: Seq[TransactionOutput], // where to send the money
    utxos: BitcoinTxBuilder.UTXOMap,      // unspent transaction outputs
    feeRate: FeeUnit,                     // fee rate per byte
    changeSPK: ScriptPubKey,              // where to send the change
    network: BitcoinNetwork               // bitcoin network information
  ): Future[BitcoinTxBuilder]
\end{lstlisting}

The return type Future does not make sense here, since the
implementation calls either Future.successful or Future.fromTry which
returns an already resolved Future.  This might be for future
purposes.

Now we create a transaction.

First, we need some money.  Thus, we create a fake transaction with
one single output.  This transaction can be parsed from the bitcoin
network, but we create one manually in order to see this process.
\begin{lstlisting}[language=scala]
  val privKey = ECPrivateKey.freshPrivateKey
  val creditingSPK = P2PKHScriptPubKey(pubKey = privKey.publicKey)

  val amount = Satoshis(Int64(10000))

  val utxo = TransactionOutput(currencyUnit = amount, scriptPubKey = creditingSPK)

  val prevTx = BaseTransaction(
    version = Int32.one,
    inputs = List.empty,
    outputs = List(utxo),
    lockTime = UInt32.zero
  )
\end{lstlisting}

On line one and two we create a new keypair to sign the next
transaction and have a scriptPubKey where the bitcoins are.  This is
our keypair.  So the money is transferred to our public key.  Line
four specifies the amount of satoshis we have in the transaction.
Then we create the actual transaction from line 6 to 13.

Now that we have some bitcoins, we create the new transaction where we
want to spend them.

First, we need some out points.  They point to outputs of previous
transactions.  We use the index zero, because the previous transaction
has only one output that becomes the first index zero.  If there were
two previous outputs, the second output would become the index 1 and
so on.
\begin{lstlisting}[language=scala]
  val outPoint = TransactionOutPoint(prevTx.txId, UInt32.zero)

  val utxoSpendingInfo = BitcoinUTXOSpendingInfo(
    outPoint = outPoint,
    output = utxo,
    signers = List(privKey),
    redeemScriptOpt = None,
    scriptWitnessOpt = None,
    hashType = HashType.sigHashAll
  )

  val utxos = List(utxoSpendingInfo)
\end{lstlisting}

This utxos are the inputs of our transaction.

Second, we need destinations to spend the bitcoins to.
For the sake of convenience we create only one.
\begin{lstlisting}[language=scala]
  val destinationAmount = Satoshis(Int64(5000))

  val destinationSPK = P2PKHScriptPubKey(pubKey = ECPrivateKey.freshPrivateKey.publicKey)

  val destinations = List(
    TransactionOutput(currencyUnit = destinationAmount, scriptPubKey = destinationSPK)
  )
\end{lstlisting}

We spend 5000 satoshis to the newly created random public key.

Finally, we define the fee rate in satoshis per one byte transaction
size as well as some bitcoin network parameters.  The bitcoin network
parameters are not important, so we use some static values normally
used when testing.

\begin{lstlisting}[language=scala]
  val feeRate = SatoshisPerByte(Satoshis.one)

  val networkParams = RegTest // some static values for testing
\end{lstlisting}

Now lets build the transaction with those data.
\begin{lstlisting}[language=scala]
  val txBuilderF: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
    destinations = destinations, // where to send the money
    utxos = utxos,               // unspent transaction outputs
    feeRate = feeRate,           // fee rate per byte
    changeSPK = creditingSPK,    // where to send the change
    network = networkParams      // bitcoin network information
  )

  val signedTxF: Future[Transaction] = txBuilderF
    .flatMap(_.sign)                       // call sign on the transaction builder
    .map {
      (tx: Transaction) => println(tx.hex) // transaction in hex for the bitcoin network
    }
\end{lstlisting}

\todo{ramon: change code to get an actual transaction, completely understand and explain the code}

Line one to seven creates a transaction builder which is then signed
on line ten.  We can now use our transaction object on line twelve.
For example, after calling \emph{hex} on it, we can send the returned
string to the bitcoin network.

\subsection{Validation of a Transaction}

Bitcoin-S offers a function called \emph{checkTransaction} located in the ScriptInterpreter object.
This is its type signature:
\begin{lstlisting}[language=scala]
  checkTransaction(transaction: Transaction): Boolean
\end{lstlisting}

\todo{kai: code zeigen}

We can pass a transaction and it returns a Boolean indicating whether
the transaction is valid or not.  So for example when we pass the
transaction we built before the returned value would be true, because
it's a valid transaction.  It might not be accepted by the bitcoin
network but for a transaction on its own it's valid.  We can not check
context with it, because we can only pass one transaction.

There are several checks in checkTransaction.
For example, it checks if there is either no input or no output.
In this case we get false.

The relevant part for the bug we found:
\begin{lstlisting}[language=scala]
  val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
  val noDuplicateInputs = prevOutputTxIds.distinct.size == prevOutputTxIds.size
\end{lstlisting}

It gathers all transaction ids referenced by the out points.  When we
call \emph{distinct} on the returned list, we get a list with
duplicate removed.  If the size of the new list is the same as the
size of the old, we know that there was no duplicate transaction id,
because, as said, distinct removes the duplicates.

\subsection{Fixing a Bug in Bitcoin-S}
\label{sec:bugfix}

We can see that there is a bug in the checkTransaction function from
before, recognized and fixed through this work.

\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{bitcoin-s-pr-comment}
	\caption{Comment on our PR \#435 on GitHub}
\end{figure}

Here is the relevant code of checkTransaction again:
\begin{lstlisting}[language=scala]
  val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
  val noDuplicateInputs = prevOutputTxIds.distinct.size == prevOutputTxIds.size
\end{lstlisting}

What happens if we have two TransactionOutPoints (previousOutputs) with a different index but referencing the same Transaction ID (txId)?

According to the Bitcoin protocol this is possible.  A transaction can
have multiple outputs that should be referenceable by the next
transaction.  So this is clearly a bug.

What should not be possible is a transaction referencing the same
output twice.  This bug occurred in Bitcoin Core known as
CVE-2018–17144 which was patched on September 18,
2018. \cite{cve201817144}

Here, Bitcoin-S did a bit too much and marked all transaction as
invalid, if they referenced the same transaction twice.  The fix is,
to check on TransactionOutPoint instead of TransactionOutPoint.txId,
because TransactionOutPoint contains the txId as well as the output
index it references.  So in pseudo code, we check on the tuple (tx,
index) instead of (tx).  The fixed code:
\begin{lstlisting}[language=scala]
  val prevOutputs = transaction.inputs.map(_.previousOutput)
  val noDuplicateInputs = prevOutputs.distinct.size == prevOutputs.size
\end{lstlisting}

Since TransactionOutPoint is a case class and Scala has a built in ==
for case classes there is no need to implement TransactionOutPoint.==.
\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{bitcoin-s-pr}
	\caption{Line changes for PR \#435 from GitHub}
\end{figure}

This was fixed in
\href{https://github.com/bitcoin-s/bitcoin-s/pull/435}{pull request
  number 435} on GitHub at April 23, 2019, through this work along
with a unit test to prevent this bug from appearing again in the
future.

\subsection{Adjusting No-Inflation Property}

\todo{anna: what does it mean to integrate}

Trying to integrate Stainless in Bitcoin-S caused a lot of troubles,
mainly because of version conflicts.  For more details see chapter
\ref{chap:appendix_arb}.

After integrating Stainless in Bitcoin-S, there were many errors.  It
takes too much time to fix them all so it should be easier to extract
the classes needed for the \texttt{checkTransaction} function.
\todo{anna: which errors?}

The extracted code has more than 1500 lines.  After running Stainless
on it, it still throws a huge bunch of errors about what Stainless can
not reason about.  After fixing some of those errors, there appear new
ones.  So this would require changing nearly everything of the
extracted code.

Let's adjust the property from the \nameref{property_1} to the
\nameref{property_2}, because there is not enough time to fix all
these errors and write the verification.  This is way less interesting
to verify but gives still many insights in what needs to be done to
verify some parts of Bitcoin-S and other source code.

So let's look at the \nameref{property_2}.

\section{Towards Verifying the Addition-with-Zero Property}
\label{chap:verify_add}

After realizing that it would consume too much time to rewrite the
Bitcoin-S code and even the extracted part with checkTransaction, the
smallest unit in Bitcoin-S-Core that is worthwhile to verify was
extracted.  This could be the addition of two CurrencyUnits.  To make
it even easier, the addition of CurrencyUnits with zero.
CurrencyUnits is an abstract class in Bitcoin-S, representing
currencies like Satoshis.

\subsection{Extracting the relevant Code}

\todo{ramon: write this section}
\todo{ramon: keep close to original code}
\todo{ramon: for every transformation: before and after}
\todo{ramon,anna: for every transformation: why does it preserve the semantics}
\todo{ramon,anna: add comments}
\todo{ramon,anna: explain what the code does and why}

Here we can see the extracted code needed for the addition of CurrencyUnits:
\lstinputlisting[language=scala]{../code/currsniporig/src/main/scala/code/initial/Code.scala}

The additions' signature looks like this:
\begin{lstlisting}[language=scala]
  +(c: CurrencyUnit): CurrencyUnit
\end{lstlisting}

When we run Stainless on this code (without any properties to prove), it throws the following errors:
{\footnotesize
\begin{verbatim}
  [ Error  ] Code.scala:41:1: Objects cannot extend classes or implement traits,
             use a case object instead
           object Int64 extends BaseNumbers[Int64] {
           ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
  [ Error  ] Code.scala:60:3: Stainless doesn't support abstract type members
              type A
              ^^^^^^
  [ Error  ] Code.scala:77:33: Only literal arguments are allowed for BigInt.
              def toBigInt: BigInt = BigInt(toLong)
                                            ^^^^^^
  [ Error  ] Code.scala:84:1: Objects cannot extend classes or implement traits,
             use a case object instead
            object Satoshis extends BaseNumbers[Satoshis] {
            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
  [  Info  ] Shutting down executor service.
\end{verbatim}}

So let's see how we can fix those errors.

\subsection{Turning Object into Case Object}

Stainless output:
{\footnotesize
\begin{verbatim}
  [ Error  ] Code.scala:41:1: Objects cannot extend classes or implement traits,
             use a case object instead
           object Int64 extends BaseNumbers[Int64] {
           ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
  [ Error  ] Code.scala:84:1: Objects cannot extend classes or implement traits,
            use a case object instead
            object Satoshis extends BaseNumbers[Satoshis] {
            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
\end{verbatim}}

Here, we can just change the objects from object to case object.
Stainless recommendation is to use objects for modules and case objects as algebraic data types.

This is due to the internal design of Scala.
It's possible to reason about case object but not about object.
This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
The \href{https://github.com/epfl-lara/stainless/issues/520}{issue number 520} on Stainless GitHub gives some thoughts, if you want to know more.

\subsection{Getting Rid of Abstract Type Member}

Stainless output:
{\footnotesize
\begin{verbatim}
  [ Error  ] Code.scala:60:3: Stainless doesn't support abstract type members
              type A
              ^^^^^^
\end{verbatim}}

This should be easy to rewrite by using generics instead of an abstract type, right?
Unfortunately not.
The problem is, CurrencyUnit uses one of its implementing classes: Satoshis.

Simplified code.
\begin{lstlisting}[language=scala]
  sealed abstract class CurrencyUnit {
    type A

    def +(c: CurrencyUnit): CurrencyUnit =
      Satoshis(satoshis.underlying + c.satoshis.underlying)

    protected def underlying: A
  }

  sealed abstract class Satoshis extends CurrencyUnit {
    override type A = Int64
  }
\end{lstlisting}

What happens, if we typify CurrencyUnit with A, meaning to make it generic with type A?

Satoshis extends CurrencyUnit with type Int64, so it would be of type CurrencyUnit[Int64].
That's too specific, because the return type of the addition is then CurrencyUnit[Int64] not CurrencyUnit[A].
Maybe the Bitcoin-S developers should reimplement this part and not use Satoshis directly.

Since there is no easy way to fix it and the code should stay as much as possible the original, we just remove the abstract type and set it to Int64.
This limits the verification a bit, but as we only want to verify the addition in satoshis, that's OK.

\subsection{Replacing the BigInt Constructor Argument With A String Literal}

Stainless output:
{\footnotesize
\begin{verbatim}
  [ Error  ] Code.scala:77:33: Only literal arguments are allowed for BigInt.
              def toBigInt: BigInt = BigInt(toLong)
                                            ^^^^^^
\end{verbatim}}

As described before, Stainless supports only a subset of Scala.
The BigInt from the Stainless library is a bit restricted.
One such restriction is, that BigInt does not support dynamic BigInt construction.
Thus, the constructor parameter of BitInt must be a literal argument.

Again, a simplified code version.
\begin{lstlisting}[language=scala]
  sealed abstract class Satoshis extends CurrencyUnit { 
    def toBigInt: BigInt = BigInt(toLong)
    def toLong: Long = underlying.toLong
  }
\end{lstlisting}

This would be really hard to refactor, because Bitcoin-S tries to be as dynamic as possible so it can be used with cryptocurrencies other than bitcoins.
Maybe it could be impossible, because they need to parse dynamic values from the bitcoin network.

Luckily, we can use toBigInt on the field \texttt{underlying} directly instead of toLong.
So, instead of converting the underlying to Long and back to BigInt we convert underlying directly to BigInt.

After fixing all Stainless errors, a new error appears.

\subsection{Getting Rid of Special Generics}

Stainless output:
{\footnotesize
\begin{verbatim}
  [ Error  ] Code.scala:7:30: Unknown type parameter type T
             sealed abstract class Number[T <: Number[T]]
                                         ^^^^^^^^^^^^^^
\end{verbatim}}

This is a missing feature in Stainless.
It does not support upper type boundaries on the class itself.
To track this, \href{https://github.com/epfl-lara/stainless/issues/519}{issue 519} was created on GitHub during this work.
\begin{lstlisting}[language=scala]
  sealed abstract class Number[T <: Number[T]] extends BasicArithmetic[T]  
\end{lstlisting}

Despite this, in order to be able to continue, we make this a concrete type by replacing T with Int64.
Int64,  because Satoshis uses only Int64.
There are other number types like UInt16 but for our property we don't need them.

Now, there are two new errors.

\subsection{Getting Rid of Concrete Type}

Stainless output:
{\footnotesize
\begin{verbatim}
  [Warning ] Code.scala:9:3: Could not extract tree in class: type A =
             BigInt (class scala.reflect.internal.Trees$TypeDef)
             type A = BigInt
             ^^^^^^^^^^^^^^^
\end{verbatim}}

This is easy.
We just replace all occurrence of A with BigInt, since A is not overwritten in an implementing class.
This is not the exact same code, because an implementing class can not override A anymore but that's fine for our verification.

This was a missing feature in Stainless that was fixed on Mai 28, 2019 with \href{https://github.com/epfl-lara/stainless/pull/470}{pull request 470} on GitHub.
Now it should work without this change.

\subsection{Replacing the BigInt \&-Function With Bound Check}
\label{sec:bound_check}

Stainless output:
{\footnotesize
\begin{verbatim}
  [ Error  ] Code.scala:22:14: Unknown call to & on result (BigInt) with arguments
             List(Number.this.andMask) of type List(BigInt)
            require((result & andMask) == result,
                    ^^^^^^^^^^^^^^^^
\end{verbatim}}

Due to the restrictions on BigInt, we can not use the \& function either.
Simplified code:
\begin{lstlisting}[language=scala]
  sealed abstract class Number extends BasicArithmetic[Int64] {
    def andMask: BigInt

    override def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))

    private def checkResult(result: BigInt): BigInt = {
      require((result & andMask) == result, "Result was out of bounds, got: " + result)
      result
    }
  }
\end{lstlisting}

This is a bounds check.
It checks if the result of the addition is in range of the specified type, which is now the hard coded Int64.

So, we can replace the \& mask with a bound check whether the result is in range of Long.MinValue and Long.MaxValue, because Int64 has the same 64-bit range as Long.
Again the code gets a bit more static and it's not the exact same code anymore.

Running Stainless produces again new errors.

\subsection{Extracting Private Inner Classes}

Stainless output:
{\footnotesize
\begin{verbatim}
  [Warning ] Code.scala:90:3: Could not extract tree in class: case private
             class SatoshisImpl extends Satoshis with Product with Serializable {
\end{verbatim}}

Stainless can not extract the private class inside the object.
Bitcoin-S uses this a lot, because they separate the class from its implementation.
Simplified:
\begin{lstlisting}[language=scala]
  object Int64 extends BaseNumbers[Int64] {
    private case class Int64Impl(underlying: BigInt) extends Int64 
  }
\end{lstlisting}

This is easy to fix.
We just extract the private class out of the object.
This is not exactly the same code, because other classes in the same file could now access the private class.
But for our property it does not change anything.

Now we get some weird warnings about require.

\subsection{Removing Second Parameter of Require}

Stainless output:
{\footnotesize
\begin{verbatim}
  [Warning ] Code.scala:51:3: Could not extract tree in class: scala.this.Predef
             .require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(
             -9223372036854775808L)), "Number was too small for a int64, got: "
             .+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
            require(underlying >= -9223372036854775808L,
            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
  [Warning ] Code.scala:53:3: Could not extract tree in class: scala.this.Predef
             .require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(
             9223372036854775807L)), "Number was too big for a int64, got: "
             .+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
            require(underlying <= 9223372036854775807L,
            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
  [ Error  ] checkResult$0 depends on missing dependencies: require$1.
\end{verbatim}}

Seems like Stainless does not support the second string parameter of require or at least it throws a warning about it.
We can safely remove the string parameters from the requires, since they only serve as error messages.

A new error appears.

\subsection{Rewriting BigInt Comparison with Long}

Stainless output:
{\footnotesize
\begin{verbatim}
  [ Error  ] inv$4 depends on missing dependencies: long2bigInt$0.
\end{verbatim}}

This error is hard to understand, but we can see that there is a missing Long to BigInt conversion.
So we search for all Long values in the code.
\begin{lstlisting}[language=scala]
  private case class Int64Impl(underlying: BigInt) extends Int64 {
    require(underlying >= -9223372036854775808L)
    require(underlying <= 9223372036854775807L)
  }
\end{lstlisting}

Looks like it can not compare a BigInt with a Long value.
We can easily convert this Long value to a BigInt with a string literal parameter by passing these numbers as strings to the BigInt constructor.

Finally, we get some output from Stainless about the verification in the code.
{\footnotesize
\begin{verbatim}
[...]
[Warning ]  => INVALID
[Warning ] Found counter-example:
[Warning ]   thiss: { x: Object | @unchecked isNumber(x) } -> Int64Impl(0)
[Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(9223372036854775808)
[...]
[  Info  ] +  precond. (call checkResult(thiss, underlying(thiss) + u ...)  invalid  Code.scala:16:45
[...]
[  Info  ] total: 5  valid: 4  (4 from cache) invalid: 1  unknown: 0  time:  1.317
\end{verbatim}}

This shows that there is an invalid specification in checkResult and Stainless prints a counterexample for it.

Let's ignore this for a moment and write the specification for the \nameref{property_2}.

\subsection{Writing Specification for the Property}

As specified, our verification must only support addition with zero.
So we restrict the parameter to be zero in the precondition.
\begin{lstlisting}[language=scala]
  require(c.satoshis == Satoshis.zero)
\end{lstlisting}

We ensure the result is the same value as \emph{this} in the postcondition.
\begin{lstlisting}[language=scala]
  ensuring(res => res.satoshis == this.satoshis)
\end{lstlisting}

We can use equals (==) directly on Satoshis, because it is a case class.
The addition will now look like this:
\begin{lstlisting}[language=scala]
  override def +(c: CurrencyUnit): CurrencyUnit = {
    require(c.satoshis == Satoshis.zero)
    Satoshis(satoshis.underlying + c.satoshis.underlying)
  } ensuring(res => res.satoshis == this.satoshis)
\end{lstlisting}

That's all we need to verify our addition.

Now we will look into the previous error.

\subsection{Propagating Require}

There is another problem with Bitcoin-S.
Bitcoin-S-Core uses require as a fail-fast method whereas Stainless needs it to verify the code.

The Stainless output again:
{\footnotesize
\begin{verbatim}
  [Warning ] Found counter-example:
  [Warning ]   thiss: { x: Object | @unchecked isNumber(x) } -> Int64Impl(0)
  [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(9223372036854775808)
\end{verbatim}}

Corresponding code:

\begin{lstlisting}[language=scala]
  sealed abstract class Number extends BasicArithmetic[Int64] {
    override def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))

    private def checkResult(result: BigInt): BigInt = {
      require(
        result <= BigInt("9223372036854775807")
        && result >= BigInt("-9223372036854775808")
      )
      result
    }
  }
\end{lstlisting}

But how does Stainless find a counter example ignoring the require in checkResult?
Since Stainless is a static verification tool, it tests every possibility.
So it can use a number bigger than the maximum Int64 and pass it to the addition.
The require in checkResult fails.

Thus, we need to add the restriction from checkResult to the addition too.
\begin{lstlisting}[language=scala]
  override def +(num: Int64): Int64 = {
    require(
          num.underlying <= BigInt("9223372036854775807")
      &&  num.underlying >= BigInt("-9223372036854775808")
      && this.underlying <= BigInt("9223372036854775807")
      && this.underlying >= BigInt("-9223372036854775808")
    )
    apply(checkResult(underlying + num.underlying))
  }
\end{lstlisting}

Stainless finds another counter example:
{\footnotesize
\begin{verbatim}
  [Warning ] Found counter-example:
  [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(1)
  [Warning ]   thiss: { x: Object | @unchecked isNumber(x) } ->
                 Int64Impl(9223372036854775807)
\end{verbatim}}

Sure, when adding one to the maximum Int64 the require does not hold anymore.
Since we do only allow zero as a parameter, the easiest way is to restrict it to zero here too.

\subsection{Result}

Finally, everything is green and correctly verified.
\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{final_verify_output}
	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
\end{figure}

The verified code.
\lstinputlisting[language=scala]{../code/currsniporig/src/main/scala/code/end/Code.scala}

But is it really the original code that we verified?
And why did we have to change that much?
This and other questions will be answered in the next chapter.

\section{Practical Challenges with Stainless}
\label{chap:appendix_arb}

In this chapter we look at the practical challenges that could occur
by using Stainless.  Since Stainless is still in development and there
is no 1.x release there might still be breaking changes and
improvements fixing problems described now.

\subsection{sbt vs JAR}

We can either use the sbt plugin or a JAR file to check code with Stainless.

Invoking the JAR on our source code Stainless will verify it.  If we
have a bigger project, this becomes really tricky, because we must
pass all files needed including the dependencies.  This is in contrast
to the sbt plugin, where we can integrate Stainless in our compilation
process.  When we call compile, Stainless verifies the code and stops
the compilation if the verification fails.

Having a static version configured in the sbt build file, every
developer has the same Stainless features available.  This should
prevent incompatibility with new or deprecated features when we use
different plugins.

So the sbt plugin has clear advantages over the JAR file since its
integrated directly.  We do not have to download it manually and find
the right version and if we bump the version we can just edit it in
the build file and every developer is on the same version again.

However, currently there are some drawbacks.  For example the sbt
plugin does not always report errors.  We will see more about that
shortly.

\subsection{Integration into Bitcoin-S}

During this work, Stainless updated the sbt plugin to support sbt
1.2.8 from 0.13.17 and Scala 2.12.8 from 2.11.12.  So this section
might be out of date now.

Stainless requires and Scala recommends Java SE Development Kit 8.
Newer Java versions won't work.

To use the latest version of the sbt tool you have to build it
locally.  You can run \texttt{sbt universal:stage} in the cloned
Stainless git repository.  This generates
\emph{frontends/scalac/target/universal/stage/bin/stainless-scalac}.

Bitcoin-S-Core uses sbt 1.2.8 and Scala 2.12.8, while Stainless sbt
plugin is on sbt 0.13.17 and Scala 2.11.12.

Sbt introduced new features in the 1.x release used by Bitcoin-S.
Most of them can be written the sbt 0.13.17 way.

The bigger problem is, due to the different Scala and sbt versions,
the following error after trying to go in a sbt shell:
\begin{verbatim}
  [warn] There may be incompatibilities among your library dependencies; run 'evicted'
         to see detailed eviction warnings.
  [error] java.lang.NoClassDefFoundError: sbt/SourcePosition
  ...
  Project loading failed: (r)etry, (q)uit, (l)ast, or (i)gnore?
\end{verbatim}

Downgrading Bitcoin-S sbt version to 0.13.17 fixes the error but then
it can not load some libraries only compiled for newer versions.  So
this would take too much time to fix and changes the Bitcoin-S code
inadvertently.

The next approach is to use the stainless cli instead of sbt.  Running
stainless on all source files does not work, because dependencies are
missing.  The parameter \emph{-classpath} can resolve it but the value
of this parameter must be the paths to all the dependencies separated
by a ':'.  Finally, \emph{core} depends on \emph{secp256k1jni},
another package of Bitcoin-S written in Java.  So this needs to be in
the source files to.

The final command looks like this in \emph{core} folder of Bitcoin-S:
\begin{lstlisting}[language=bash]
  $ stainless
    -classpath ".:$(find ~/.ivy2/ -type f -name *.jar | tr '\n' ':')"
    $(find . -type f -name *.scala | tr '\n' ' ')
    $(find ../secp256k1jni -type f -name *.java | tr '\n' ' ')
\end{lstlisting}

\emph{.ivy2} is the dependency cache of sbt.  The \emph{tr} replaces
the first char with the second so a newline with either ':' or ' '.

With this command, Stainless throws the next error:
\begin{verbatim}
  [Internal] Error: object scala.reflect.macros.internal.macroImpl in compiler mirror
             not found.. Trace:
  [Internal] - scala.reflect.internal.MissingRequirementError$.signal
             (MissingRequirementError.scala:17)
  ...
  [Internal] object scala.reflect.macros.internal.macroImpl in compiler mirror not found.
  [Internal] Please inform the authors of Inox about this message
\end{verbatim}

So we can not know how many errors will face us.  Let's go another
way, because the errors may take too much time and it might lead to a
next error.  We extract the code needed to verify a transaction mainly
the class Transaction and ScriptInterpreter with many other classes
they're depending on.

After this extraction Stainless was successfully integrated with both
sbt and JAR.

Running \texttt{sbt compile} in the project with Stainless ended
without error.  But it also ended with no output.  So we are not able
to change the code so Stainless would accept it since we do not know
what to change.

So the sbt plugin does not always complain where the JAR file did.
The open
\href{https://github.com/epfl-lara/stainless/issues/484}{issue 484 on
  GitHub} might describe exactly this error.

Now we can finally run Stainless on our code.  But this leads us to
the next findings.  We must rewrite most of the code, as described in
the previous chapters.

\section{Conclusion}
\label{chap:conclusion}

Because of the limitations of the verication tool, we could only
verify a rewritten version of the original Bitcoin-S code.  So we can
not guarantee the correctness of the addition of Satoshis with zero in
Bitcoin-S.  Not all changes we made were as trivial as the replacement
of objects with case objects.  For these non-trivial changes, as seen
for example the bound check in section \ref{sec:bound_check}, we
cannot say whether they are equivalent to the original implementation
or not.

So code should be written specically with formal verication in mind,
in order to successfully verify it.  Otherwise, it needs a lot of
changes in the software because verification is mathematical and the
current software is written mostly in object-oriented style.  Software
written in the functional paradigm would be much easier to reason
about.

Thus, either Stainless must find ways to translate more of built-in
object-oriented patterns of Scala to their verification tool or
developers must invest more in functional programming.

Also, we found that trying to verify code reveals bugs as shown in
section \ref{sec:bugfix}.  Finally, our work led to some feedback to
the Stainless developers to improve the tool.

% \begin{table}
% \caption{Table captions should be placed above the
% tables.}\label{tab1}
% \begin{tabular}{|l|l|l|}
% \hline
% Heading level &  Example & Font size and style\\
% \hline
% Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
% 1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
% 2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
% 3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
% 4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
% \hline
% \end{tabular}
% \end{table}

% \noindent Displayed equations are centered and set on a separate
% line.
% \begin{equation}
% x + y = z
% \end{equation}
% Please try to avoid rasterized images for line-art diagrams and
% schemas. Whenever possible, use vector graphics instead (see
% Fig.~\ref{fig1}).

% \begin{figure}
% \includegraphics[width=\textwidth]{fig1.eps}
% \caption{A figure caption is always placed below the illustration.
% Please note that short captions are centered, while long ones are
% justified by the macro package automatically.} \label{fig1}
% \end{figure}

% \begin{theorem}
% This is a sample theorem. The run-in heading is set in bold, while
% the following text appears in italics. Definitions, lemmas,
% propositions, and corollaries are styled the same way.
% \end{theorem}
% %
% % the environments 'definition', 'lemma', 'proposition', 'corollary',
% % 'remark', and 'example' are defined in the LLNCS documentclass as well.
% %
% \begin{proof}
% Proofs, examples, and remarks have the initial word in italics,
% while the following text appears in normal font.
% \end{proof}
% For citations of references, we prefer the use of square brackets
% and consecutive numbers. Citations using labels or the author/year
% convention are also acceptable. The following bibliography provides
% a sample reference list with entries for journal
% articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
% book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
% and a homepage~\cite{ref_url1}. Multiple citations are grouped
% \cite{ref_article1,ref_lncs1,ref_book1},
% \cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
%

\bibliographystyle{splncs04}
\bibliography{bibliography}

\end{document}

### 2019-06-20 code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala
copied initial code to reduced folder

@@ -1,0 +1,145 @@
+package addition.reduced.currency
+
+import org.bitcoins.core.consensus.Consensus
+import org.bitcoins.core.number.{BaseNumbers, BasicArithmetic, Int64}
+import org.bitcoins.core.protocol.NetworkElement
+import org.bitcoins.core.serializers.RawSatoshisSerializer
+import org.bitcoins.core.util.Factory
+import scodec.bits.ByteVector
+
+sealed abstract class CurrencyUnit
+  extends NetworkElement
+    with BasicArithmetic[CurrencyUnit] {
+  type A
+
+  def satoshis: Satoshis
+
+  def >=(c: CurrencyUnit): Boolean = {
+    satoshis.underlying >= c.satoshis.underlying
+  }
+
+  def >(c: CurrencyUnit): Boolean = {
+    satoshis.underlying > c.satoshis.underlying
+  }
+
+  def <(c: CurrencyUnit): Boolean = {
+    satoshis.underlying < c.satoshis.underlying
+  }
+
+  def <=(c: CurrencyUnit): Boolean = {
+    satoshis.underlying <= c.satoshis.underlying
+  }
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  override def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
+
+  override def -(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying - c.satoshis.underlying)
+  }
+
+  override def *(factor: BigInt): CurrencyUnit = {
+    Satoshis(satoshis.underlying * factor)
+  }
+
+  override def *(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying * c.satoshis.underlying)
+  }
+
+  def unary_- : CurrencyUnit = {
+    Satoshis(-satoshis.underlying)
+  }
+
+  override def bytes: ByteVector = satoshis.bytes
+
+  def toBigDecimal: BigDecimal
+
+  protected def underlying: A
+}
+
+sealed abstract class Satoshis extends CurrencyUnit {
+  override type A = Int64
+
+  override def toString: String = {
+    val num = toLong
+    val postFix = if (num == 1) "sat" else "sats"
[…]

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/BasicArithmetic.scala

@@ -1,0 +1,47 @@
+package addition.reduced.number
+
+import scala.util.Try
+
+/**
+  * @define mulSafe
+  * Some classes have restrictions on upper bounds
+  * for it's underlying value. This might cause the `*`
+  * operator to throw. This method wraps it in a `Try`
+  * block.
+  */
+trait BasicArithmetic[N] {
+
+  def +(n: N): N
+
+  /**
+    * Some classes have restrictions on upper bounds
+    * for it's underlying value. This might cause the `+`
+    * operator to throw. This method wraps it in a `Try`
+    * block.
+    */
+  def addSafe(n: N): Try[N] = Try { this + n }
+
+  def -(n: N): N
+
+  /**
+    * Some classes have restrictions on lower bounds
+    * for it's underlying value. This might cause the `-`
+    * operator to throw. This method wraps it in a `Try`
+    * block.
+    */
+  def subtractSafe(n: N): Try[N] = Try { this - n }
+
+  def *(factor: BigInt): N
+
+  /**
+    * $mulSafe
+    */
+  def multiplySafe(factor: BigInt): Try[N] = Try { this * factor }
+
+  def *(factor: N): N
+
+  /**
+    * $mulSafe
+    */
+  def multiplySafe(factor: N): Try[N] = Try { this * factor }
+}

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/NumberType.scala

@@ -1,0 +1,408 @@
+package addition.reduced.number
+
+import org.bitcoins.core.protocol.NetworkElement
+import org.bitcoins.core.util.{BitcoinSUtil, Factory, NumberUtil}
+import scodec.bits.ByteVector
+
+import scala.util.{Failure, Success, Try}
+
+/**
+  * Created by chris on 6/4/16.
+  */
+/**
+  * This abstract class is meant to represent a signed and unsigned number in C
+  * This is useful for dealing with codebases/protocols that rely on C's
+  * unsigned integer types
+  */
+sealed abstract class Number[T <: Number[T]]
+  extends NetworkElement
+    with BasicArithmetic[T] {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toInt: Int = toBigInt.bigInteger.intValueExact()
+  def toLong: Long = toBigInt.bigInteger.longValueExact()
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => T
+
+  override def +(num: T): T = apply(checkResult(underlying + num.underlying))
+  override def -(num: T): T = apply(checkResult(underlying - num.underlying))
+  override def *(factor: BigInt): T = apply(checkResult(underlying * factor))
+  override def *(num: T): T = apply(checkResult(underlying * num.underlying))
+
+  def >(num: T): Boolean = underlying > num.underlying
+  def >=(num: T): Boolean = underlying >= num.underlying
+  def <(num: T): Boolean = underlying < num.underlying
+  def <=(num: T): Boolean = underlying <= num.underlying
+
+  def <<(num: Int): T = this.<<(apply(num))
+  def >>(num: Int): T = this.>>(apply(num))
+
+  def <<(num: T): T = {
+    checkIfInt(num).map { _ =>
+      apply((underlying << num.toInt) & andMask)
+    }.get
+  }
+
+  def >>(num: T): T = {
[…]

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/NumberType.scala
removed all numbers expect Int64

@@ -101,87 +101,6 @@ sealed abstract class Number[T <: Number[T]]
 sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
 
 /**
-  * Represents an unsigned number in our number system
-  * Instances of this are [[UInt32]] or [[UInt64]]
-  */
-sealed abstract class UnsignedNumber[T <: Number[T]] extends Number[T]
-
-/** This number type is useful for dealing with [[org.bitcoins.core.util.Bech32]]
-  * related applications. The native encoding for Bech32 is a 5 bit number which
-  * is what this abstraction is meant to  be used for
-  */
-sealed abstract class UInt5 extends UnsignedNumber[UInt5] {
-  override def apply: A => UInt5 = UInt5(_)
-
-  override def andMask: BigInt = 0x1f
-
-  def byte: Byte = toInt.toByte
-
-  def toUInt8: UInt8 = UInt8(toInt)
-
-  override def hex: String = toUInt8.hex
-}
-
-sealed abstract class UInt8 extends UnsignedNumber[UInt8] {
-  override def apply: A => UInt8 = UInt8(_)
-
-  override def hex: String = BitcoinSUtil.encodeHex(toInt.toShort).slice(2, 4)
-
-  override def andMask = 0xff
-
-  def toUInt5: UInt5 = {
-    //this will throw if not in range of a UInt5, come back and look later
-    UInt5(toInt)
-  }
-}
-
-/**
-  * Represents a uint32_t in C
-  */
-sealed abstract class UInt32 extends UnsignedNumber[UInt32] {
-  override def apply: A => UInt32 = UInt32(_)
-  override def hex: String = BitcoinSUtil.encodeHex(toLong).slice(8, 16)
-
-  override def andMask = 0xffffffffL
-}
-
-/**
-  * Represents a uint64_t in C
-  */
-sealed abstract class UInt64 extends UnsignedNumber[UInt64] {
-  override def hex: String = encodeHex(underlying)
-  override def apply: A => UInt64 = UInt64(_)
-  override def andMask = 0xffffffffffffffffL
-
-  /**
-    * Converts a [[BigInt]] to a 8 byte hex representation.
-    * [[BigInt]] will only allocate 1 byte for numbers like 1 which require 1 byte, giving us the hex representation 01
[…]

### 2019-06-20 code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala
removed NetworkElement and Factory from classes

@@ -2,14 +2,9 @@ package addition.reduced.currency
 
 import org.bitcoins.core.consensus.Consensus
 import org.bitcoins.core.number.{BaseNumbers, BasicArithmetic, Int64}
-import org.bitcoins.core.protocol.NetworkElement
-import org.bitcoins.core.serializers.RawSatoshisSerializer
-import org.bitcoins.core.util.Factory
-import scodec.bits.ByteVector
 
 sealed abstract class CurrencyUnit
-  extends NetworkElement
-    with BasicArithmetic[CurrencyUnit] {
+  extends BasicArithmetic[CurrencyUnit] {
   type A
 
   def satoshis: Satoshis
@@ -54,8 +49,6 @@ sealed abstract class CurrencyUnit
     Satoshis(-satoshis.underlying)
   }
 
-  override def bytes: ByteVector = satoshis.bytes
-
   def toBigDecimal: BigDecimal
 
   protected def underlying: A
@@ -70,8 +63,6 @@ sealed abstract class Satoshis extends CurrencyUnit {
     s"$num $postFix"
   }
 
-  override def bytes: ByteVector = RawSatoshisSerializer.write(this)
-
   override def satoshis: Satoshis = this
 
   override def toBigDecimal = BigDecimal(toBigInt)
@@ -83,15 +74,13 @@ sealed abstract class Satoshis extends CurrencyUnit {
   def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
 }
 
-object Satoshis extends Factory[Satoshis] with BaseNumbers[Satoshis] {
+object Satoshis extends BaseNumbers[Satoshis] {
 
   val min = Satoshis(Int64.min)
   val max = Satoshis(Int64.max)
   val zero = Satoshis(Int64.zero)
   val one = Satoshis(Int64.one)
 
-  override def fromBytes(bytes: ByteVector): Satoshis =
-    RawSatoshisSerializer.read(bytes)
   def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
 
   private case class SatoshisImpl(underlying: Int64) extends Satoshis
@@ -103,8 +92,6 @@ sealed abstract class Bitcoins extends CurrencyUnit {
   override def toString: String = s"$toBigDecimal BTC"
 
   override def toBigDecimal: BigDecimal = underlying
-
-  override def hex: String = satoshis.hex
 
   override def satoshis: Satoshis = {
     val sat = underlying * CurrencyUnits.btcToSatoshiScalar

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/NumberType.scala

@@ -1,9 +1,5 @@
 package addition.reduced.number
 
-import org.bitcoins.core.protocol.NetworkElement
-import org.bitcoins.core.util.{BitcoinSUtil, Factory, NumberUtil}
-import scodec.bits.ByteVector
-
 import scala.util.{Failure, Success, Try}
 
 /**
@@ -15,8 +11,7 @@ import scala.util.{Failure, Success, Try}
   * unsigned integer types
   */
 sealed abstract class Number[T <: Number[T]]
-  extends NetworkElement
-    with BasicArithmetic[T] {
+  extends BasicArithmetic[T] {
   type A = BigInt
 
   /** The underlying scala number used to to hold the number */
@@ -90,8 +85,6 @@ sealed abstract class Number[T <: Number[T]]
       Success(())
     }
   }
-
-  override def bytes: ByteVector = BitcoinSUtil.decodeHex(hex)
 }
 
 /**
@@ -106,7 +99,6 @@ sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
 sealed abstract class Int64 extends SignedNumber[Int64] {
   override def apply: A => Int64 = Int64(_)
   override def andMask = 0xffffffffffffffffL
-  override def hex: String = BitcoinSUtil.encodeHex(toLong)
 }
 
 /**
@@ -120,7 +112,7 @@ trait BaseNumbers[T] {
   def max: T
 }
 
-object Int64 extends Factory[Int64] with BaseNumbers[Int64] {
+object Int64 extends BaseNumbers[Int64] {
   private case class Int64Impl(underlying: BigInt) extends Int64 {
     require(underlying >= -9223372036854775808L,
       "Number was too small for a int64, got: " + underlying)
@@ -133,11 +125,6 @@ object Int64 extends Factory[Int64] with BaseNumbers[Int64] {
 
   lazy val min = Int64(-9223372036854775808L)
   lazy val max = Int64(9223372036854775807L)
-
-  override def fromBytes(bytes: ByteVector): Int64 = {
-    require(bytes.size <= 8, "We cannot have an Int64 be larger than 8 bytes")
-    Int64(BigInt(bytes.toArray).toLong)
-  }
 
   def apply(long: Long): Int64 = Int64(BigInt(long))
 

### 2019-06-20 code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala
removed Bitcoins class

@@ -1,6 +1,5 @@
 package addition.reduced.currency
 
-import org.bitcoins.core.consensus.Consensus
 import org.bitcoins.core.number.{BaseNumbers, BasicArithmetic, Int64}
 
 sealed abstract class CurrencyUnit
@@ -86,35 +85,6 @@ object Satoshis extends BaseNumbers[Satoshis] {
   private case class SatoshisImpl(underlying: Int64) extends Satoshis
 }
 
-sealed abstract class Bitcoins extends CurrencyUnit {
-  override type A = BigDecimal
-
-  override def toString: String = s"$toBigDecimal BTC"
-
-  override def toBigDecimal: BigDecimal = underlying
-
-  override def satoshis: Satoshis = {
-    val sat = underlying * CurrencyUnits.btcToSatoshiScalar
-    Satoshis(Int64(sat.toLongExact))
-  }
-}
-
-object Bitcoins extends BaseNumbers[Bitcoins] {
-  val min = Bitcoins((-Consensus.maxMoney).satoshis)
-  val max = Bitcoins(Consensus.maxMoney.satoshis)
-  val zero = Bitcoins(Satoshis.zero)
-  val one = Bitcoins(1)
-
-  def apply(satoshis: Satoshis): Bitcoins = {
-    val b: BigDecimal = satoshis.toLong * CurrencyUnits.satoshisToBTCScalar
-    Bitcoins(b)
-  }
-
-  def apply(underlying: BigDecimal): Bitcoins = BitcoinsImpl(underlying)
-
-  private case class BitcoinsImpl(underlying: BigDecimal) extends Bitcoins
-}
-
 object CurrencyUnits {
 
   /** The number you need to multiply BTC by to get it's satoshis */
@@ -126,7 +96,6 @@ object CurrencyUnits {
   val negativeSatoshi = Satoshis(Int64(-1))
 
   def toSatoshis(unit: CurrencyUnit): Satoshis = unit match {
-    case b: Bitcoins => b.satoshis
     case x: Satoshis => x
   }
 }

### 2019-06-20 code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala
fixed import

@@ -1,6 +1,6 @@
 package addition.reduced.currency
 
-import org.bitcoins.core.number.{BaseNumbers, BasicArithmetic, Int64}
+import addition.reduced.number.{BaseNumbers, BasicArithmetic, Int64}
 
 sealed abstract class CurrencyUnit
   extends BasicArithmetic[CurrencyUnit] {

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/NumberType.scala
fixed comment

@@ -89,7 +89,7 @@ sealed abstract class Number[T <: Number[T]]
 
 /**
   * Represents a signed number in our number system
-  * Instances of this are [[Int32]] or [[Int64]]
+  * Instances of this is [[Int64]]
   */
 sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
 

### 2019-06-20 code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala
removed subtraction and multiplication

@@ -32,22 +32,6 @@ sealed abstract class CurrencyUnit
     Satoshis(satoshis.underlying + c.satoshis.underlying)
   }
 
-  override def -(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying - c.satoshis.underlying)
-  }
-
-  override def *(factor: BigInt): CurrencyUnit = {
-    Satoshis(satoshis.underlying * factor)
-  }
-
-  override def *(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying * c.satoshis.underlying)
-  }
-
-  def unary_- : CurrencyUnit = {
-    Satoshis(-satoshis.underlying)
-  }
-
   def toBigDecimal: BigDecimal
 
   protected def underlying: A

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/BasicArithmetic.scala

@@ -10,38 +10,5 @@ import scala.util.Try
   * block.
   */
 trait BasicArithmetic[N] {
-
   def +(n: N): N
-
-  /**
-    * Some classes have restrictions on upper bounds
-    * for it's underlying value. This might cause the `+`
-    * operator to throw. This method wraps it in a `Try`
-    * block.
-    */
-  def addSafe(n: N): Try[N] = Try { this + n }
-
-  def -(n: N): N
-
-  /**
-    * Some classes have restrictions on lower bounds
-    * for it's underlying value. This might cause the `-`
-    * operator to throw. This method wraps it in a `Try`
-    * block.
-    */
-  def subtractSafe(n: N): Try[N] = Try { this - n }
-
-  def *(factor: BigInt): N
-
-  /**
-    * $mulSafe
-    */
-  def multiplySafe(factor: BigInt): Try[N] = Try { this * factor }
-
-  def *(factor: N): N
-
-  /**
-    * $mulSafe
-    */
-  def multiplySafe(factor: N): Try[N] = Try { this * factor }
 }

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/NumberType.scala

@@ -32,9 +32,6 @@ sealed abstract class Number[T <: Number[T]]
   def apply: A => T
 
   override def +(num: T): T = apply(checkResult(underlying + num.underlying))
-  override def -(num: T): T = apply(checkResult(underlying - num.underlying))
-  override def *(factor: BigInt): T = apply(checkResult(underlying * factor))
-  override def *(num: T): T = apply(checkResult(underlying * num.underlying))
 
   def >(num: T): Boolean = underlying > num.underlying
   def >=(num: T): Boolean = underlying >= num.underlying

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/NumberType.scala
removed binary operations and unary_-

@@ -38,30 +38,6 @@ sealed abstract class Number[T <: Number[T]]
   def <(num: T): Boolean = underlying < num.underlying
   def <=(num: T): Boolean = underlying <= num.underlying
 
-  def <<(num: Int): T = this.<<(apply(num))
-  def >>(num: Int): T = this.>>(apply(num))
-
-  def <<(num: T): T = {
-    checkIfInt(num).map { _ =>
-      apply((underlying << num.toInt) & andMask)
-    }.get
-  }
-
-  def >>(num: T): T = {
-    //this check is for weird behavior with the jvm and shift rights
-    //https://stackoverflow.com/questions/47519140/bitwise-shift-right-with-long-not-equaling-zero/47519728#47519728
-    if (num.toLong > 63) apply(0)
-    else {
-      checkIfInt(num).map { _ =>
-        apply(underlying >> num.toInt)
-      }.get
-    }
-  }
-
-  def |(num: T): T = apply(checkResult(underlying | num.underlying))
-  def &(num: T): T = apply(checkResult(underlying & num.underlying))
-  def unary_- : T = apply(-underlying)
-
   /**
     * Checks if the given result is within the range
     * of this number type
@@ -70,17 +46,6 @@ sealed abstract class Number[T <: Number[T]]
     require((result & andMask) == result,
       "Result was out of bounds, got: " + result)
     result
-  }
-
-  /** Checks if the given nubmer is within range of a Int */
-  private def checkIfInt(num: T): Try[Unit] = {
-    if (num.toBigInt >= Int.MaxValue || num.toBigInt <= Int.MinValue) {
-      Failure(
-        new IllegalArgumentException(
-          "Num was not in range of int, got: " + num))
-    } else {
-      Success(())
-    }
   }
 }
 

### 2019-06-20 code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala
removed comparision operators expect == and !=

@@ -8,22 +8,6 @@ sealed abstract class CurrencyUnit
 
   def satoshis: Satoshis
 
-  def >=(c: CurrencyUnit): Boolean = {
-    satoshis.underlying >= c.satoshis.underlying
-  }
-
-  def >(c: CurrencyUnit): Boolean = {
-    satoshis.underlying > c.satoshis.underlying
-  }
-
-  def <(c: CurrencyUnit): Boolean = {
-    satoshis.underlying < c.satoshis.underlying
-  }
-
-  def <=(c: CurrencyUnit): Boolean = {
-    satoshis.underlying <= c.satoshis.underlying
-  }
-
   def !=(c: CurrencyUnit): Boolean = !(this == c)
 
   def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/NumberType.scala

@@ -33,11 +33,6 @@ sealed abstract class Number[T <: Number[T]]
 
   override def +(num: T): T = apply(checkResult(underlying + num.underlying))
 
-  def >(num: T): Boolean = underlying > num.underlying
-  def >=(num: T): Boolean = underlying >= num.underlying
-  def <(num: T): Boolean = underlying < num.underlying
-  def <=(num: T): Boolean = underlying <= num.underlying
-
   /**
     * Checks if the given result is within the range
     * of this number type

### 2019-06-20 code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala
removed toBigDecimal

@@ -16,8 +16,6 @@ sealed abstract class CurrencyUnit
     Satoshis(satoshis.underlying + c.satoshis.underlying)
   }
 
-  def toBigDecimal: BigDecimal
-
   protected def underlying: A
 }
 
@@ -31,8 +29,6 @@ sealed abstract class Satoshis extends CurrencyUnit {
   }
 
   override def satoshis: Satoshis = this
-
-  override def toBigDecimal = BigDecimal(toBigInt)
 
   def toBigInt: BigInt = BigInt(toLong)
 

### 2019-06-20 code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala
removed toString and toInt

@@ -22,12 +22,6 @@ sealed abstract class CurrencyUnit
 sealed abstract class Satoshis extends CurrencyUnit {
   override type A = Int64
 
-  override def toString: String = {
-    val num = toLong
-    val postFix = if (num == 1) "sat" else "sats"
-    s"$num $postFix"
-  }
-
   override def satoshis: Satoshis = this
 
   def toBigInt: BigInt = BigInt(toLong)
@@ -47,19 +41,4 @@ object Satoshis extends BaseNumbers[Satoshis] {
   def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
 
   private case class SatoshisImpl(underlying: Int64) extends Satoshis
-}
-
-object CurrencyUnits {
-
-  /** The number you need to multiply BTC by to get it's satoshis */
-  val btcToSatoshiScalar: Long = 100000000
-  val satoshisToBTCScalar: BigDecimal = BigDecimal(1.0) / btcToSatoshiScalar
-  val oneBTC: CurrencyUnit = Satoshis(Int64(btcToSatoshiScalar))
-  val oneMBTC: CurrencyUnit = Satoshis(Int64(btcToSatoshiScalar / 1000))
-  val zero: CurrencyUnit = Satoshis.zero
-  val negativeSatoshi = Satoshis(Int64(-1))
-
-  def toSatoshis(unit: CurrencyUnit): Satoshis = unit match {
-    case x: Satoshis => x
-  }
 }

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/NumberType.scala

@@ -17,7 +17,6 @@ sealed abstract class Number[T <: Number[T]]
   /** The underlying scala number used to to hold the number */
   protected def underlying: A
 
-  def toInt: Int = toBigInt.bigInteger.intValueExact()
   def toLong: Long = toBigInt.bigInteger.longValueExact()
   def toBigInt: BigInt = underlying
 

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/BasicArithmetic.scala
removed unused imports

@@ -1,7 +1,5 @@
 package addition.reduced.number
 
-import scala.util.Try
-
 /**
   * @define mulSafe
   * Some classes have restrictions on upper bounds

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/NumberType.scala

@@ -1,7 +1,5 @@
 package addition.reduced.number
 
-import scala.util.{Failure, Success, Try}
-
 /**
   * Created by chris on 6/4/16.
   */

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/NumberType.scala
removed unused comment

@@ -1,9 +1,6 @@
 package addition.reduced.number
 
 /**
-  * Created by chris on 6/4/16.
-  */
-/**
   * This abstract class is meant to represent a signed and unsigned number in C
   * This is useful for dealing with codebases/protocols that rely on C's
   * unsigned integer types

### 2019-06-20 paper/paper.tex
added initial code and todos

@@
   %frame=single, % Border around box
-  numbersep=-7pt,
+  %numbersep=-7pt,
   numberstyle=\color{gray},
@@
 
+\todo{
+kai text
+
+extracted 3 classes (they are in code/addition/src/main/scala/addition/original)
+
+changes from original to reduced (new folder):
+  - package name
+
+  - removed numbers other than Int64
+
+  - removed extending Factory and NetworkElement abstract class. This includes some hex/byte conversion eg fromHex, hex, bytes, fromBytes, ...
+    \href{https://bitcoin-s.org/docs/core/core-intro\#the-basics}{description in section \#the-basics (cannot add link with hashtag)}
+    \href{https://github.com/bitcoin-s/bitcoin-s/blob/master/core/src/main/scala/org/bitcoins/core/protocol/NetworkElement.scala}{NetworkElement class}
+    \href{https://github.com/bitcoin-s/bitcoin-s/blob/master/core/src/main/scala/org/bitcoins/core/util/Factory.scala}{Factory class}
+
+  - removed Bitcoins class
+
+  - removed subtraction and multiplication, binary operations <<, >>, etc, comparision operator <=, >=, etc but not == and !=
+
+  - removed toBigDecimal
+
+  - removed object CurrencyUnits containing some variables to transform satoshis to btc (not used)
+
+  The code we use for the following sections is in reduced folder.
+}
+
 \todo{ramon: write this section}
@@
 Here we can see the extracted code needed for the addition of CurrencyUnits:
-\lstinputlisting[language=scala]{../code/currsniporig/src/main/scala/code/initial/Code.scala}
+\lstinputlisting[language=scala]{../code/addition/src/main/scala/addition/reduced/number/BasicArithmetic.scala}
+\lstinputlisting[language=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}
+\lstinputlisting[language=scala]{../code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala}
 

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/BasicArithmetic.scala
removed unused comment

@@ -1,12 +1,5 @@
 package addition.reduced.number
 
-/**
-  * @define mulSafe
-  * Some classes have restrictions on upper bounds
-  * for it's underlying value. This might cause the `*`
-  * operator to throw. This method wraps it in a `Try`
-  * block.
-  */
 trait BasicArithmetic[N] {
   def +(n: N): N
 }

### 2019-06-20 paper/paper.tex
removed BasicArithmetic

@@
 
-extracted 3 classes (they are in code/addition/src/main/scala/addition/original)
+extracted 2 files (they are in code/addition/src/main/scala/addition/original)
 
@@
 
-  - removed extending Factory and NetworkElement abstract class. This includes some hex/byte conversion eg fromHex, hex, bytes, fromBytes, ...
+  - removed extending Factory, NetworkElement and BasicArithmetic. This includes some hex/byte conversion eg fromHex, hex, bytes, fromBytes, ... Just interfaces never referenced
     \href{https://bitcoin-s.org/docs/core/core-intro\#the-basics}{description in section \#the-basics (cannot add link with hashtag)}
@@
 Here we can see the extracted code needed for the addition of CurrencyUnits:
-\lstinputlisting[language=scala]{../code/addition/src/main/scala/addition/reduced/number/BasicArithmetic.scala}
 \lstinputlisting[language=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}

### 2019-06-20 code/addition/src/main/scala/addition/original/number/BasicArithmetic.scala

@@ -1,47 +1,0 @@
-package org.bitcoins.core.number
-
-import scala.util.Try
-
-/**
-  * @define mulSafe
-  * Some classes have restrictions on upper bounds
-  * for it's underlying value. This might cause the `*`
-  * operator to throw. This method wraps it in a `Try`
-  * block.
-  */
-trait BasicArithmetic[N] {
-
-  def +(n: N): N
-
-  /**
-    * Some classes have restrictions on upper bounds
-    * for it's underlying value. This might cause the `+`
-    * operator to throw. This method wraps it in a `Try`
-    * block.
-    */
-  def addSafe(n: N): Try[N] = Try { this + n }
-
-  def -(n: N): N
-
-  /**
-    * Some classes have restrictions on lower bounds
-    * for it's underlying value. This might cause the `-`
-    * operator to throw. This method wraps it in a `Try`
-    * block.
-    */
-  def subtractSafe(n: N): Try[N] = Try { this - n }
-
-  def *(factor: BigInt): N
-
-  /**
-    * $mulSafe
-    */
-  def multiplySafe(factor: BigInt): Try[N] = Try { this * factor }
-
-  def *(factor: N): N
-
-  /**
-    * $mulSafe
-    */
-  def multiplySafe(factor: N): Try[N] = Try { this * factor }
-}

### 2019-06-20 code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala

@@ -1,9 +1,8 @@
 package addition.reduced.currency
 
-import addition.reduced.number.{BaseNumbers, BasicArithmetic, Int64}
+import addition.reduced.number.{BaseNumbers, Int64}
 
-sealed abstract class CurrencyUnit
-  extends BasicArithmetic[CurrencyUnit] {
+sealed abstract class CurrencyUnit {
   type A
 
   def satoshis: Satoshis
@@ -12,7 +11,7 @@ sealed abstract class CurrencyUnit
 
   def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
 
-  override def +(c: CurrencyUnit): CurrencyUnit = {
+  def +(c: CurrencyUnit): CurrencyUnit = {
     Satoshis(satoshis.underlying + c.satoshis.underlying)
   }
 

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/BasicArithmetic.scala

@@ -1,5 +1,0 @@
-package addition.reduced.number
-
-trait BasicArithmetic[N] {
-  def +(n: N): N
-}

### 2019-06-20 code/addition/src/main/scala/addition/reduced/number/NumberType.scala

@@ -5,8 +5,7 @@ package addition.reduced.number
   * This is useful for dealing with codebases/protocols that rely on C's
   * unsigned integer types
   */
-sealed abstract class Number[T <: Number[T]]
-  extends BasicArithmetic[T] {
+sealed abstract class Number[T <: Number[T]] {
   type A = BigInt
 
   /** The underlying scala number used to to hold the number */
@@ -25,7 +24,7 @@ sealed abstract class Number[T <: Number[T]]
   /** Factory function to create the underlying T, for instance a UInt32 */
   def apply: A => T
 
-  override def +(num: T): T = apply(checkResult(underlying + num.underlying))
+  def +(num: T): T = apply(checkResult(underlying + num.underlying))
 
   /**
     * Checks if the given result is within the range

### 2019-06-20 code/addition/bin/stainless.sh
added binary to execute stainless and created directory with modified code

@@ -1,0 +1,9 @@
+#!/usr/bin/env bash
+
+DIR="$( cd "$( dirname "${BASH_SOURCE[0]}" )" >/dev/null 2>&1 && pwd )"
+CODE_DIR="modified"
+
+JAVA_HOME=/usr/lib/jvm/java-8-openjdk; PATH=$JAVA_HOME/bin:$PATH; java \
+  -jar $DIR/../lib/stainless-0.1.0-93dbd33.jar \
+  $DIR/../src/main/scala/addition/$CODE_DIR/currency/CurrencyUnits.scala \
+  $DIR/../src/main/scala/addition/$CODE_DIR/number/NumberType.scala

### 2019-06-20 code/addition/src/main/scala/addition/modified/currency/CurrencyUnits.scala

@@ -1,0 +1,43 @@
+package addition.modified.currency
+
+import addition.modified.number.{BaseNumbers, Int64}
+
+sealed abstract class CurrencyUnit {
+  type A
+
+  def satoshis: Satoshis
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
+
+  protected def underlying: A
+}
+
+sealed abstract class Satoshis extends CurrencyUnit {
+  override type A = Int64
+
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = BigInt(toLong)
+
+  def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}
+
+object Satoshis extends BaseNumbers[Satoshis] {
+
+  val min = Satoshis(Int64.min)
+  val max = Satoshis(Int64.max)
+  val zero = Satoshis(Int64.zero)
+  val one = Satoshis(Int64.one)
+
+  def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
+
+  private case class SatoshisImpl(underlying: Int64) extends Satoshis
+}

### 2019-06-20 code/addition/src/main/scala/addition/modified/number/NumberType.scala

@@ -1,0 +1,82 @@
+package addition.modified.number
+
+/**
+  * This abstract class is meant to represent a signed and unsigned number in C
+  * This is useful for dealing with codebases/protocols that rely on C's
+  * unsigned integer types
+  */
+sealed abstract class Number[T <: Number[T]] {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toLong: Long = toBigInt.bigInteger.longValueExact()
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => T
+
+  def +(num: T): T = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber[Int64] {
+  override def apply: A => Int64 = Int64(_)
+  override def andMask = 0xffffffffffffffffL
+}
+
+/**
+  * Represents various numbers that should be implemented
+  * inside of any companion object for a number
+  */
+trait BaseNumbers[T] {
+  def zero: T
+  def one: T
+  def min: T
+  def max: T
+}
+
+object Int64 extends BaseNumbers[Int64] {
+  private case class Int64Impl(underlying: BigInt) extends Int64 {
+    require(underlying >= -9223372036854775808L,
+      "Number was too small for a int64, got: " + underlying)
+    require(underlying <= 9223372036854775807L,
+      "Number was too big for a int64, got: " + underlying)
+  }
+
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(-9223372036854775808L)
+  lazy val max = Int64(9223372036854775807L)
[…]

### 2019-06-20 paper/paper.tex
add references

@@
 %\usepackage{courier}
-\usepackage{listings}
-\usepackage[dvipsnames]{xcolor}
-\colorlet{link}{MidnightBlue}
+\usepackage{listings, color}
 \hypersetup{
   colorlinks   = true,
-  linkbordercolor=link,
-  pdfborderstyle={/S/U/W .5},% border style will be underline of width 1pt
-     linkcolor  = link,
-     citecolor  = link,
-     urlcolor   = link
+  linkcolor  = blue,
+  citecolor  = blue,
+  urlcolor   = blue
 }
@@
 
-kai: successor of Leon verifier \cite{DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}
+kai: successor of Leon verifier \cite{DBLP:conf/ecoop/BlancKKS13,DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}
 

### 2019-06-20 paper/paper.tex
work on intro

@@
 }
+
+% nicer bold
+\renewcommand{\bfdefault}{b}%
+
+% make paragraph headings bold instead of italics
+\renewcommand{\paragraph}{\textbf}%
 
@@
 \begin{abstract}
-  We try to verify properties of the Bitcoin-S library, a Scala
+  We try to verify properties of the bitcoin-s library, a Scala
   implementation of parts of the Bitcoin protocol. We use the
@@
   the \emph{Pure Scala Fragment}. We first try to verify the property
-  that regular transactions never create new money. It turns out that
+  that regular transactions do not create new money. It turns out that
   there is too much code involved that lies outside of the supported
   fragment to make this feasible. However, in the process we uncover
-  and fix a bug in Bitcoin-S. We then turn to a much simpler (and less
+  and fix a bug in bitcoin-s. We then turn to a much simpler (and less
   interesting) property: that adding zero satoshis to a given amount
@@
 
-\subsection{Formal Verification with Stainless}
-\label{sec:stainless}
-
-kai: successor of Leon verifier \cite{DBLP:conf/ecoop/BlancKKS13,DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}
-
-Stainless is developed by "Lab for Automated Reasoning and Analysis"
-(LARA) at EPFL's School of Computer and Communication Sciences.  Using
-this framework we can verify the correctness of Scala programs.  The
-following overview of the framework is taken from the section
-\href{https://epfl-lara.github.io/stainless/intro.html}{Introduction}
-of the Stainless documentation \cite{Stainless:documentation}.
-
-Stainless statically verifies that a program satisfies a given
-specification and that a program will not crash at runtime.  The
-framework covers all possible inputs and finds a counter examples for
-possible failures in a program which violate the given specification.
-
-The main functions used to write a specification are \textit{require}
-and \textit{ensuring}.
-
-With \textit{require} we define a precondition of a function, which we
-want to verify.  \textit{require} is placed at the beginning of the
-function body.  \textit{require} specifies, if the parameter of the
-function holds against a certain condition.
-
-With \textit{ensuring}, we define the postcondition of a function.
-\textit{ensuring} is placed at the end of the function after the body.
-\textit{ensuring} specifies, if the return value or the result of the
-function satisfies a certain condition.
-
-On invoking Stainless, it tries to prove that the postcondition always
-holds, assuming the given precondition does hold.
-
-The following example demonstrates a simple formal specification for
-the function calculating a factorial.  This is a modified example from
-the section
-\href{https://epfl-lara.github.io/stainless/intro.html}{Introduction}
-of the Stainless documentation \cite{Stainless:documentation}, so
-Stainless reports an error verifying it.
+\paragraph{The Stainless Verifier.} Stainless is the successor of the
+Leon verifier
+\cite{DBLP:conf/ecoop/BlancKKS13,DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}
+and is developed at EPF Lausanne
+\cite{Stainless:documentation,Stainless:github}. It is intended to be
+used by programmers without training in formal verification and thus
+allows to write specifications in Scala and focusses on counterexample
+finding in addition to proving correctness.
+
+The following example from \cite{Stainless:documentation} demonstrates
+this. Notice how a precondition is specified using the function
+\emph{require} and a postcondition using \emph{ensuring}.
+
 \begin{lstlisting}[language=Scala]
@@
 
-It recursively calculates the factorial of an integer number.
-The input of the function is constrained with \textit{require} to a non-negative value.
-The result should also be non-negative.
-Thus, Stainless will verify whether the result of the factorial calculation is non-negative for all non-negative inputs.
-
-Stainless can produce 3 outcomes of postcondition verification:
-\textit{valid}, \textit{invalid} and \textit{unknown}.  If the
-postcondition is \textit{valid} Stainless could prove that for any
-inputs constrained in the precondition, the postcondition always
-holds.  Reporting the postcondition as \textit{invalid} the framework
-could find at least one counterexample which satisfies the
-precondition but violates the postcondition.  If Stainless is unable
-to prove the postcondition or find a counterexample it reports the
-outcome \textit{unknown}.  In this case a timeout or an internal error
-occurred.  Furthermore, Stainless checks for calls in the code to the
-function with invalid arguments violating the precondition.  For
-example if we invoke factorial with a negative number it reports
-\textit{invalid}.
-
-Let's return to our incorrect example with factorial.
-Stainless reports the following result for the factorial verification from our example:
+Our program happens to not satisfy the specification. An overflow in
+the 32-bit Int leads to a negative result for the input 17, as
+Stainless reports in Figure~\ref{fig:failed}. Changing the type from
+Int to BigInt will result in a successful verification.
+
 \begin{figure}
@@
 	\caption{Output of Stainless verification for calculating factorial of Int number}
-	\label{fig:output1}
+	\label{fig:failed}
 \end{figure}
 
-Stainless disproves the postcondition and gives the number 17 as a
-counterexample.  Due to an overflow in the 32 bit Int the result
-becomes a negative value.  The program would work correct by changing
-the type from Int to BigInt.
-
-Stainless supports only a subset of Scala. \todo{Anna It supports
-  blabla, in particular it supports not blablabla. (check class with
-  var)}
+\paragraph{The Pure Scala Fragment.} Stainless supports only a subset
+of Scala. \todo{Anna It supports blabla, in particular it supports not
+  blablabla. (check class with var)}
 
@@
 \href{https://github.com/epfl-lara/stainless/tree/master/frontends/library/stainless}{\textit{frontends/library/stainless}}.
-
-Stainless verifies Scala programs using SMT (satisfiability modulo theories) solvers.
 

### 2019-06-20 paper/paper.tex
more work on intro

@@
   significant part of the relevant code lies outside of the supported
-  fragment. However, after a series of equivalent transformations, we
+  fragment. However, after a series of equivalent transformations we
   arrive at code that we successfully verify.
@@
 
-\todo{kai: motivation, bitcoin inflation bug, scala better suited for formal verification, bitcoin-s handling millions of dollars, this is a bit of a case study}
+For software handling cryptocurrency, correctness is clearly crucial.
+However, even in very well-tested software such as Bitcoin Core,
+serious bugs occur. The most recent example is the bug found in
+September 2018 \cite{cve201817144} which essentially allowed to
+arbitrarily create new coins. Such software is thus a worthwhile
+target for formal verification.
+
+\paragraph{The bitcoin-s Library.} Here, we try to verify two
+properties of the bitcoin-s library
+\cite{BitcoinS:website,BitcoinS:github}, which implements parts of the
+Bitcoin protocol in Scala. In particular, it allows to serialize,
+deserialize, sign and validate transactions. The library uses
+immutable data structures and algebraic data types but is not written
+with formal verification in mind. According to the website, the
+library is used in production, ``handling millions of dollars each
+day'' \cite{BitcoinS:website}.
 
@@
 and is developed at EPF Lausanne
-\cite{Stainless:documentation,Stainless:github}. It is intended to be
+\cite{Stainless:github}. It is intended to be
 used by programmers without training in formal verification and thus
@@
 
-The following example from \cite{Stainless:documentation} demonstrates
-this. Notice how a precondition is specified using the function
-\emph{require} and a postcondition using \emph{ensuring}.
+The following example from the Stainless documentation
+\cite{Stainless:documentation} demonstrates this. Notice how a
+precondition is specified using the function \emph{require} and a
+postcondition using \emph{ensuring}.
 
@@
 
-\subsection{The Properties to Verify}
-
-Bitcoin-S is a large project found on GitHub implementing many
-functionalities.  In this work we look at two of them, namely the
-\nameref{property_1} and the \nameref{property_2} described shortly.
-Thus, we go through the relevant code parts to verify those
-properties, look at some challenges with Stainless and see why we
-should write software with verification in mind.  In order to be able
-to verify Bitcoin-S, we have to rewrite parts of it in the subset of
-Scala supported by Stainless.
-
-\subsubsection{No-Inflation Property}
-\label{property_1}
-
-The decision to verify this property was inspired by the bug found in
-Bitcoin Core in September 2018 \cite{cve201817144} which allowed to
-spend the same unspent transaction output (UTXO) in a transaction
-multiple times.  Hence, new coins could be created out of the air.
+\section{Trying to Verify the No-Inflation Property}
+
 Thus, we try to verify the property, that a non-coinbase transaction
@@
 bugfix in section \ref{sec:bugfix}.
-
-\subsubsection{Addition-with-Zero Property}
-\label{property_2}
-
-In Bitcoin-S there is a class \texttt{Satoshis} representing an amount of bitcoins.
-We look at the verification of the addition of Satoshis with zero Satoshis.
-This operation should result in the same amount of Satoshis.
-Let's call it the \nameref{property_2}.
-
-Using Stainless, we see the successful verification of this property.
-But the process of the verification with the tool requires many
-changes in the code, so that Stainless can accept it.  We look at all
-needed modifications in chapter \ref{chap:verify_add}.
-
-\section{Trying to Verify the No-Inflation Property}
-\label{chap:verify_check}
 
@@
 \section{Towards Verifying the Addition-with-Zero Property}
+
 \label{chap:verify_add}
+
+In Bitcoin-S there is a class \texttt{Satoshis} representing an amount of bitcoins.
+We look at the verification of the addition of Satoshis with zero Satoshis.
+This operation should result in the same amount of Satoshis.
+Let's call it the \nameref{property_2}.
+
+Using Stainless, we see the successful verification of this property.
+But the process of the verification with the tool requires many
+changes in the code, so that Stainless can accept it.  We look at all
+needed modifications in chapter \ref{chap:verify_add}.
 

### 2019-06-20 paper/paper.tex
more work on intro

@@
 
-\lstset{
+\lstdefinestyle{scala}{
   language=scala,
-  basicstyle=\footnotesize\ttfamily,
+%  basicstyle=\footnotesize\ttfamily,
+  basicstyle=\scriptsize\ttfamily, % kai: temporary 
   breaklines=true,
@@
 }
+
+\lstdefinestyle{stainless}{numbers=none, breaklines=false,  basicstyle=\footnotesize\ttfamily,}
 
@@
 arbitrarily create new coins. Such software is thus a worthwhile
-target for formal verification.
-
-\paragraph{The bitcoin-s Library.} Here, we try to verify two
-properties of the bitcoin-s library
-\cite{BitcoinS:website,BitcoinS:github}, which implements parts of the
-Bitcoin protocol in Scala. In particular, it allows to serialize,
-deserialize, sign and validate transactions. The library uses
-immutable data structures and algebraic data types but is not written
-with formal verification in mind. According to the website, the
-library is used in production, ``handling millions of dollars each
-day'' \cite{BitcoinS:website}.
+target for formal verification. In this work, we set out to verify
+properties of the bitcoin-s library with the Stainless verifier.
+
+\paragraph{The Bitcoin-S Library.} The bitcoin-s library is an
+implementation of parts of the Bitcoin protocol in Scala
+\cite{BitcoinS:website,BitcoinS:github}. In particular, it allows to
+serialize, deserialize, sign and validate transactions. The library
+uses immutable data structures and algebraic data types but is not
+written with formal verification in mind. According to the website,
+the library is used in production, handling significant amounts of
+cryptocurrency each day \cite{BitcoinS:website}.
 
@@
 
-\begin{lstlisting}[language=Scala]
+\begin{lstlisting}[style=scala]
   def factorial(n: Int): Int = {
@@
 
-Our program happens to not satisfy the specification. An overflow in
+Our program happens not to satisfy the specification. An overflow in
 the 32-bit Int leads to a negative result for the input 17, as
@@
 following constructor:
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala, basicstyle=\scriptsize\tt]
   BitcoinTxBuilder(
@@
 network, but we create one manually in order to see this process.
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   val privKey = ECPrivateKey.freshPrivateKey
@@
 so on.
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   val outPoint = TransactionOutPoint(prevTx.txId, UInt32.zero)
@@
 For the sake of convenience we create only one.
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   val destinationAmount = Satoshis(Int64(5000))
@@
 
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   val feeRate = SatoshisPerByte(Satoshis.one)
@@
 Now lets build the transaction with those data.
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   val txBuilderF: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
@@
 This is its type signature:
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   checkTransaction(transaction: Transaction): Boolean
@@
 
-\todo{kai: code zeigen}
+\begin{lstlisting}[style=scala]
+/**
+  * Checks the validity of a transaction in accordance to bitcoin core's CheckTransaction function
+  * https://github.com/bitcoin/bitcoin/blob/f7a21dae5dbf71d5bc00485215e84e6f2b309d0a/src/main.cpp#L939.
+  */
+def checkTransaction(transaction: Transaction): Boolean = {
+  val inputOutputsNotZero =
+    !(transaction.inputs.isEmpty || transaction.outputs.isEmpty)
+  val txNotLargerThanBlock = transaction.bytes.size < Consensus.maxBlockSize
+  val outputsSpendValidAmountsOfMoney = !transaction.outputs.exists(o =>
+    o.value < CurrencyUnits.zero || o.value > Consensus.maxMoney)
+
+  val outputValues = transaction.outputs.map(_.value)
+  val totalSpentByOutputs: CurrencyUnit =
+    outputValues.fold(CurrencyUnits.zero)(_ + _)
+  val allOutputsValidMoneyRange = validMoneyRange(totalSpentByOutputs)
+  val prevOutputs = transaction.inputs.map(_.previousOutput)
+  val noDuplicateInputs = prevOutputs.distinct.size == prevOutputs.size
+
+  val isValidScriptSigForCoinbaseTx = transaction.isCoinbase match {
+    case true =>
+      transaction.inputs.head.scriptSignature.asmBytes.size >= 2 &&
+        transaction.inputs.head.scriptSignature.asmBytes.size <= 100
+    case false =>
+      //since this is not a coinbase tx we cannot have any empty previous outs inside of inputs
+      !transaction.inputs.exists(_.previousOutput == EmptyTransactionOutPoint)
+  }
+  inputOutputsNotZero && txNotLargerThanBlock && outputsSpendValidAmountsOfMoney && noDuplicateInputs &&
+  allOutputsValidMoneyRange && noDuplicateInputs && isValidScriptSigForCoinbaseTx
+}
+\end{lstlisting}
 
@@
 The relevant part for the bug we found:
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
@@
 
-\begin{figure}
-	\centering
-		\includegraphics[width=\textwidth]{bitcoin-s-pr-comment}
-	\caption{Comment on our PR \#435 on GitHub}
-\end{figure}
-
 Here is the relevant code of checkTransaction again:
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
@@
 index) instead of (tx).  The fixed code:
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   val prevOutputs = transaction.inputs.map(_.previousOutput)
@@
 for case classes there is no need to implement TransactionOutPoint.==.
-\begin{figure}
-	\centering
-		\includegraphics[width=\textwidth]{bitcoin-s-pr}
-	\caption{Line changes for PR \#435 from GitHub}
-\end{figure}
 
@@
 Here we can see the extracted code needed for the addition of CurrencyUnits:
-\lstinputlisting[language=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}
-\lstinputlisting[language=scala]{../code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala}
 
 The additions' signature looks like this:
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   +(c: CurrencyUnit): CurrencyUnit
@@
 When we run Stainless on this code (without any properties to prove), it throws the following errors:
-{\footnotesize
-\begin{verbatim}
+
+\begin{lstlisting}[style=stainless]
   [ Error  ] Code.scala:41:1: Objects cannot extend classes or implement traits,
@@
   [  Info  ] Shutting down executor service.
-\end{verbatim}}
+\end{lstlisting}
 
@@
 Simplified code.
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   sealed abstract class CurrencyUnit {
@@
 Again, a simplified code version.
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   sealed abstract class Satoshis extends CurrencyUnit { 
@@
 To track this, \href{https://github.com/epfl-lara/stainless/issues/519}{issue 519} was created on GitHub during this work.
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   sealed abstract class Number[T <: Number[T]] extends BasicArithmetic[T]  
@@
 Simplified code:
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   sealed abstract class Number extends BasicArithmetic[Int64] {
@@
 Simplified:
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   object Int64 extends BaseNumbers[Int64] {
@@
 So we search for all Long values in the code.
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   private case class Int64Impl(underlying: BigInt) extends Int64 {
@@
 So we restrict the parameter to be zero in the precondition.
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   require(c.satoshis == Satoshis.zero)
@@
 We ensure the result is the same value as \emph{this} in the postcondition.
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   ensuring(res => res.satoshis == this.satoshis)
@@
 The addition will now look like this:
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   override def +(c: CurrencyUnit): CurrencyUnit = {
@@
 
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   sealed abstract class Number extends BasicArithmetic[Int64] {
@@
 Thus, we need to add the restriction from checkResult to the addition too.
-\begin{lstlisting}[language=scala]
+\begin{lstlisting}[style=scala]
   override def +(num: Int64): Int64 = {
@@
 The verified code.
-\lstinputlisting[language=scala]{../code/currsniporig/src/main/scala/code/end/Code.scala}
+\lstinputlisting[style=scala]{../code/currsniporig/src/main/scala/code/end/Code.scala}
 

### 2019-06-20 paper/paper.tex
new errors

@@
 \begin{lstlisting}[style=stainless]
-  [ Error  ] Code.scala:41:1: Objects cannot extend classes or implement traits,
-             use a case object instead
+  [ Error  ] currency/CurrencyUnits.scala:6:3: Stainless doesn't support abstract type members
+             type A
+             ^^^^^^
+  [ Error  ] currency/CurrencyUnits.scala:26:33: Only literal arguments are allowed for BigInt.
+              def toBigInt: BigInt = BigInt(toLong)
+                                            ^^^^^^
+  [ Error  ] currency/CurrencyUnits.scala:33:1: Objects cannot extend classes or implement traits, use a case object instead
+            object Satoshis extends BaseNumbers[Satoshis] {
+            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+  [ Error  ] number/NumberType.scala:65:1: Objects cannot extend classes or implement traits, use a case object instead
             object Int64 extends BaseNumbers[Int64] {
             ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-  [ Error  ] Code.scala:60:3: Stainless doesn't support abstract type members
-              type A
-              ^^^^^^
-  [ Error  ] Code.scala:77:33: Only literal arguments are allowed for BigInt.
-              def toBigInt: BigInt = BigInt(toLong)
-                                            ^^^^^^
-  [ Error  ] Code.scala:84:1: Objects cannot extend classes or implement traits,
-             use a case object instead
-            object Satoshis extends BaseNumbers[Satoshis] {
-            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
   [  Info  ] Shutting down executor service.

### 2019-06-20 paper/paper.tex
auto linebreak for style stainless

@@
 
-\lstdefinestyle{stainless}{numbers=none, breaklines=false,  basicstyle=\footnotesize\ttfamily,}
+\lstdefinestyle{stainless}{
+  numbers=none,
+  breaklines=true,
+  breakautoindent=true,
+  breakindent=63pt,
+  basicstyle=\footnotesize\ttfamily
+}
 
@@
 Stainless output:
-{\footnotesize
-\begin{verbatim}
-  [ Error  ] Code.scala:41:1: Objects cannot extend classes or implement traits,
-             use a case object instead
+\begin{lstlisting}[style=stainless]
+  [ Error  ] currency/CurrencyUnits.scala:33:1: Objects cannot extend classes or implement traits, use a case object instead
+            object Satoshis extends BaseNumbers[Satoshis] {
+            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+  [ Error  ] number/NumberType.scala:65:1: Objects cannot extend classes or implement traits, use a case object instead
             object Int64 extends BaseNumbers[Int64] {
             ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-  [ Error  ] Code.scala:84:1: Objects cannot extend classes or implement traits,
-            use a case object instead
-            object Satoshis extends BaseNumbers[Satoshis] {
-            ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-\end{verbatim}}
+\end{lstlisting}
+
+Code before:
+\begin{lstlisting}[style=scala]
+object Int64 extends BaseNumbers[Int64] { ... }
+object Satoshis extends BaseNumbers[Satoshis] { ... }
+\end{lstlisting}
+
+Code after:
+\begin{lstlisting}[style=scala]
+case object Int64 extends BaseNumbers[Int64] { ... }
+case object Satoshis extends BaseNumbers[Satoshis] { ... }
+\end{lstlisting}
 
@@
 Stainless recommendation is to use objects for modules and case objects as algebraic data types.
+
+\todo{Ramon does this change semantics?}
 

### 2019-06-20 paper/paper.tex
verbatim -> lstlisting

@@
 Stainless output:
-{\footnotesize
-\begin{verbatim}
+\begin{lstlisting}[style=stainless]
   [ Error  ] Code.scala:60:3: Stainless doesn't support abstract type members
@@
               ^^^^^^
-\end{verbatim}}
+\end{lstlisting}
 
@@
 Stainless output:
-{\footnotesize
-\begin{verbatim}
+\begin{lstlisting}[style=stainless]
   [ Error  ] Code.scala:77:33: Only literal arguments are allowed for BigInt.
@@
                                             ^^^^^^
-\end{verbatim}}
+\end{lstlisting}
 
@@
 Stainless output:
-{\footnotesize
-\begin{verbatim}
+\begin{lstlisting}[style=stainless]
   [ Error  ] Code.scala:7:30: Unknown type parameter type T
@@
                                          ^^^^^^^^^^^^^^
-\end{verbatim}}
+\end{lstlisting}
 
@@
 Stainless output:
-{\footnotesize
-\begin{verbatim}
+\begin{lstlisting}[style=stainless]
   [Warning ] Code.scala:9:3: Could not extract tree in class: type A =
@@
              ^^^^^^^^^^^^^^^
-\end{verbatim}}
+\end{lstlisting}
 
@@
 Stainless output:
-{\footnotesize
-\begin{verbatim}
+\begin{lstlisting}[style=stainless]
   [ Error  ] Code.scala:22:14: Unknown call to & on result (BigInt) with arguments
@@
                     ^^^^^^^^^^^^^^^^
-\end{verbatim}}
+\end{lstlisting}
 
@@
 Stainless output:
-{\footnotesize
-\begin{verbatim}
+\begin{lstlisting}[style=stainless]
   [Warning ] Code.scala:90:3: Could not extract tree in class: case private
              class SatoshisImpl extends Satoshis with Product with Serializable {
-\end{verbatim}}
+\end{lstlisting}
 
@@
 Stainless output:
-{\footnotesize
-\begin{verbatim}
+\begin{lstlisting}[style=stainless]
   [Warning ] Code.scala:51:3: Could not extract tree in class: scala.this.Predef
@@
   [ Error  ] checkResult$0 depends on missing dependencies: require$1.
-\end{verbatim}}
+\end{lstlisting}
 
@@
 Stainless output:
-{\footnotesize
-\begin{verbatim}
+\begin{lstlisting}[style=stainless]
   [ Error  ] inv$4 depends on missing dependencies: long2bigInt$0.
-\end{verbatim}}
+\end{lstlisting}
 
@@
 Finally, we get some output from Stainless about the verification in the code.
-{\footnotesize
-\begin{verbatim}
+\begin{lstlisting}[style=stainless]
 [...]
@@
 [  Info  ] total: 5  valid: 4  (4 from cache) invalid: 1  unknown: 0  time:  1.317
-\end{verbatim}}
+\end{lstlisting}
 
@@
 The Stainless output again:
-{\footnotesize
-\begin{verbatim}
+\begin{lstlisting}[style=stainless]
   [Warning ] Found counter-example:
@@
   [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(9223372036854775808)
-\end{verbatim}}
+\end{lstlisting}
 
@@
 Stainless finds another counter example:
-{\footnotesize
-\begin{verbatim}
+\begin{lstlisting}[style=stainless]
   [Warning ] Found counter-example:
@@
                  Int64Impl(9223372036854775807)
-\end{verbatim}}
+\end{lstlisting}
 

### 2019-06-20 code/addition/src/main/scala/addition/modified/currency/CurrencyUnits.scala
object to case object

@@ -30,7 +30,7 @@ sealed abstract class Satoshis extends CurrencyUnit {
   def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
 }
 
-object Satoshis extends BaseNumbers[Satoshis] {
+case object Satoshis extends BaseNumbers[Satoshis] {
 
   val min = Satoshis(Int64.min)
   val max = Satoshis(Int64.max)

### 2019-06-20 code/addition/src/main/scala/addition/modified/number/NumberType.scala

@@ -62,7 +62,7 @@ trait BaseNumbers[T] {
   def max: T
 }
 
-object Int64 extends BaseNumbers[Int64] {
+case object Int64 extends BaseNumbers[Int64] {
   private case class Int64Impl(underlying: BigInt) extends Int64 {
     require(underlying >= -9223372036854775808L,
       "Number was too small for a int64, got: " + underlying)

### 2019-06-20 code/addition/src/main/scala/addition/modified/currency/CurrencyUnits.scala
remove abstract type A

@@ -3,8 +3,6 @@ package addition.modified.currency
 import addition.modified.number.{BaseNumbers, Int64}
 
 sealed abstract class CurrencyUnit {
-  type A
-
   def satoshis: Satoshis
 
   def !=(c: CurrencyUnit): Boolean = !(this == c)
@@ -15,12 +13,10 @@ sealed abstract class CurrencyUnit {
     Satoshis(satoshis.underlying + c.satoshis.underlying)
   }
 
-  protected def underlying: A
+  protected def underlying: Int64
 }
 
 sealed abstract class Satoshis extends CurrencyUnit {
-  override type A = Int64
-
   override def satoshis: Satoshis = this
 
   def toBigInt: BigInt = BigInt(toLong)

### 2019-06-20 code/addition/src/main/scala/addition/modified/currency/CurrencyUnits.scala
use toBigInt directly instead of going over toLong

@@ -19,7 +19,7 @@ sealed abstract class CurrencyUnit {
 sealed abstract class Satoshis extends CurrencyUnit {
   override def satoshis: Satoshis = this
 
-  def toBigInt: BigInt = BigInt(toLong)
+  def toBigInt: BigInt = underlying.toBigInt
 
   def toLong: Long = underlying.toLong
 

### 2019-06-20 code/addition/src/main/scala/addition/modified/number/NumberType.scala
removed generic type T from Number class

@@ -5,7 +5,7 @@ package addition.modified.number
   * This is useful for dealing with codebases/protocols that rely on C's
   * unsigned integer types
   */
-sealed abstract class Number[T <: Number[T]] {
+sealed abstract class Number {
   type A = BigInt
 
   /** The underlying scala number used to to hold the number */
@@ -22,9 +22,9 @@ sealed abstract class Number[T <: Number[T]] {
   def andMask: BigInt
 
   /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => T
+  def apply: A => Int64
 
-  def +(num: T): T = apply(checkResult(underlying + num.underlying))
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
 
   /**
     * Checks if the given result is within the range
@@ -41,12 +41,12 @@ sealed abstract class Number[T <: Number[T]] {
   * Represents a signed number in our number system
   * Instances of this is [[Int64]]
   */
-sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
+sealed abstract class SignedNumber extends Number
 
 /**
   * Represents a int64_t in C
   */
-sealed abstract class Int64 extends SignedNumber[Int64] {
+sealed abstract class Int64 extends SignedNumber {
   override def apply: A => Int64 = Int64(_)
   override def andMask = 0xffffffffffffffffL
 }

### 2019-06-20 code/addition/src/main/scala/addition/modified/currency/CurrencyUnits.scala
remove toLong

@@ -21,8 +21,6 @@ sealed abstract class Satoshis extends CurrencyUnit {
 
   def toBigInt: BigInt = underlying.toBigInt
 
-  def toLong: Long = underlying.toLong
-
   def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
 }
 

### 2019-06-20 code/addition/src/main/scala/addition/modified/number/NumberType.scala

@@ -11,7 +11,6 @@ sealed abstract class Number {
   /** The underlying scala number used to to hold the number */
   protected def underlying: A
 
-  def toLong: Long = toBigInt.bigInteger.longValueExact()
   def toBigInt: BigInt = underlying
 
   /**

### 2019-06-20 paper/paper.tex
add some before/after code

@@
 \begin{lstlisting}[style=stainless]
-  [ Error  ] Code.scala:60:3: Stainless doesn't support abstract type members
+  [ Error  ] currency/CurrencyUnits.scala:6:3: Stainless doesn't support abstract type members
              type A
@@
 
-Simplified code.
+Before:
 \begin{lstlisting}[style=scala]
@@
 
-    def +(c: CurrencyUnit): CurrencyUnit =
+  def satoshis: Satoshis
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  def +(c: CurrencyUnit): CurrencyUnit = {
     Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
 
@@
   override type A = Int64
+
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = BigInt(toLong)
+
+  def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
 }
@@
 
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class CurrencyUnit {
+  def satoshis: Satoshis
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
+
+  protected def underlying: Int64
+}
+
+sealed abstract class Satoshis extends CurrencyUnit {
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = BigInt(toLong)
+
+  def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}
+\end{lstlisting}
+
 \subsection{Replacing the BigInt Constructor Argument With A String Literal}
@@
 \begin{lstlisting}[style=stainless]
-  [ Error  ] Code.scala:77:33: Only literal arguments are allowed for BigInt.
+  [ Error  ] currency/CurrencyUnits.scala:26:33: Only literal arguments are allowed for BigInt.
               def toBigInt: BigInt = BigInt(toLong)
@@
 
-Again, a simplified code version.
+Before:
 \begin{lstlisting}[style=scala]
 sealed abstract class Satoshis extends CurrencyUnit {
+  override def satoshis: Satoshis = this
+
   def toBigInt: BigInt = BigInt(toLong)
+
   def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Satoshis extends CurrencyUnit {
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = underlying.toBigInt
+
+  def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
 }
@@
 \begin{lstlisting}[style=stainless]
-  [ Error  ] Code.scala:7:30: Unknown type parameter type T
-             sealed abstract class Number[T <: Number[T]]
+  [ Error  ] number/NumberType.scala:8:30: Unknown type parameter type T
+  sealed abstract class Number[T <: Number[T]] {
                                ^^^^^^^^^^^^^^
+
 \end{lstlisting}
@@
 To track this, \href{https://github.com/epfl-lara/stainless/issues/519}{issue 519} was created on GitHub during this work.
-\begin{lstlisting}[style=scala]
-  sealed abstract class Number[T <: Number[T]] extends BasicArithmetic[T]  
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number[T <: Number[T]] {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toLong: Long = toBigInt.bigInteger.longValueExact()
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => T
+
+  def +(num: T): T = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber[Int64] {
+  override def apply: A => Int64 = Int64(_)
+  override def andMask = 0xffffffffffffffffL
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toLong: Long = toBigInt.bigInteger.longValueExact()
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber extends Number
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: A => Int64 = Int64(_)
+  override def andMask = 0xffffffffffffffffL
+}
 \end{lstlisting}
@@
 Now, there are two new errors.
+
+\subsection{Removing toLong}
+
+Stainless output:
+\begin{lstlisting}[style=stainless]
+  [ Error  ] number/NumberType.scala:14:22: Unknown call to bigInteger on Number.this.toBigInt (BigInt) with arguments List() of type List()
+             def toLong: Long = toBigInt.bigInteger.longValueExact()
+                                ^^^^^^^^^^^^^^^^^^^
+\end{lstlisting}
+
+Now that we do not depend on toLong anymore lets just remove it from out code.
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Satoshis extends CurrencyUnit {
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = underlying.toBigInt
+
+  def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}  
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Satoshis extends CurrencyUnit {
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = underlying.toBigInt
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}  
+\end{lstlisting}
+
+And before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toLong: Long = toBigInt.bigInteger.longValueExact()
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}  
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+\end{lstlisting}
 

### 2019-06-21 paper/paper.tex
more stainless error outputs

@@
 \begin{lstlisting}[style=stainless]
-  [Warning ] Code.scala:9:3: Could not extract tree in class: type A =
-             BigInt (class scala.reflect.internal.Trees$TypeDef)
+  [Warning ] number/NumberType.scala:9:3: Could not extract tree in class: type A = BigInt (class scala.reflect.internal.Trees$TypeDef)
                type A = BigInt
                ^^^^^^^^^^^^^^^
+
 \end{lstlisting}
@@
 \begin{lstlisting}[style=stainless]
-  [ Error  ] Code.scala:22:14: Unknown call to & on result (BigInt) with arguments
-             List(Number.this.andMask) of type List(BigInt)
+  [ Error  ] number/NumberType.scala:33:14: Unknown call to & on result (BigInt) with arguments List(Number.this.andMask) of type List(BigInt)
                require((result & andMask) == result,
                        ^^^^^^^^^^^^^^^^
+
 \end{lstlisting}
@@
 \begin{lstlisting}[style=stainless]
-  [Warning ] Code.scala:90:3: Could not extract tree in class: case private
-             class SatoshisImpl extends Satoshis with Product with Serializable {
+  [Warning ] currency/CurrencyUnits.scala:36:3: Could not extract tree in class: case private class SatoshisImpl extends Satoshis with Product with Serializable {
+  [Warning ]   <caseaccessor> <paramaccessor> private[this] val underlying: addition.modified.number.Int64 = _;
+  [Warning ]   <stable> <caseaccessor> <accessor> <paramaccessor> def underlying: addition.modified.number.Int64 = SatoshisImpl.this.underlying;
+  [Warning ]   def <init>(underlying: addition.modified.number.Int64): addition.modified.currency.Satoshis.SatoshisImpl = {
+  [Warning ]     SatoshisImpl.super.<init>();
+  [Warning ]     ()
+  [Warning ]   };
+  [Warning ]   <synthetic> def copy(underlying: addition.modified.number.Int64 = underlying): addition.modified.currency.Satoshis.SatoshisImpl = new SatoshisImpl(underlying);
+  [Warning ]   <synthetic> def copy$default$1: addition.modified.number.Int64 = SatoshisImpl.this.underlying;
+  [Warning ]   override <synthetic> def productPrefix: String = "SatoshisImpl";
+  [Warning ]   <synthetic> def productArity: Int = 1;
+  [Warning ]   <synthetic> def productElement(x$1: Int): Any = x$1 match {
+  [Warning ]     case 0 => SatoshisImpl.this.underlying
+  [Warning ]     case _ => throw new IndexOutOfBoundsException(x$1.toString())
+  [Warning ]   };
+  [Warning ]   override <synthetic> def productIterator: Iterator[Any] = runtime.this.ScalaRunTime.typedProductIterator[Any](SatoshisImpl.this);
+  [Warning ]   <synthetic> def canEqual(x$1: Any): Boolean = x$1.$isInstanceOf[addition.modified.currency.Satoshis.SatoshisImpl]();
+  [Warning ]   override <synthetic> def hashCode(): Int = ScalaRunTime.this._hashCode(SatoshisImpl.this);
+  [Warning ]   override <synthetic> def toString(): String = ScalaRunTime.this._toString(SatoshisImpl.this);
+  [Warning ]   override <synthetic> def equals(x$1: Any): Boolean = SatoshisImpl.this.eq(x$1.asInstanceOf[Object]).||(x$1 match {
+  [Warning ]   case (_: addition.modified.currency.Satoshis.SatoshisImpl) => true
+  [Warning ]   case _ => false
+  [Warning ] }.&&({
+  [Warning ]     <synthetic> val SatoshisImpl$1: addition.modified.currency.Satoshis.SatoshisImpl = x$1.asInstanceOf[addition.modified.currency.Satoshis.SatoshisImpl];
+  [Warning ]     SatoshisImpl.this.underlying.==(SatoshisImpl$1.underlying).&&(SatoshisImpl$1.canEqual(SatoshisImpl.this))
+  [Warning ]   }))
+  [Warning ] } (class scala.reflect.internal.Trees$ClassDef)
+                private case class SatoshisImpl(underlying: Int64) extends Satoshis
+                ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
 \end{lstlisting}

### 2019-06-21 code/addition/src/main/scala/addition/modified/currency/CurrencyUnits.scala
extract private inner class

@@ -25,13 +25,12 @@ sealed abstract class Satoshis extends CurrencyUnit {
 }
 
 case object Satoshis extends BaseNumbers[Satoshis] {
-
   val min = Satoshis(Int64.min)
   val max = Satoshis(Int64.max)
   val zero = Satoshis(Int64.zero)
   val one = Satoshis(Int64.one)
 
   def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
-
-  private case class SatoshisImpl(underlying: Int64) extends Satoshis
 }
+
+private case class SatoshisImpl(underlying: Int64) extends Satoshis

### 2019-06-21 paper/paper.tex
add before and after for private class

@@
 Bitcoin-S uses this a lot, because they separate the class from its implementation.
-Simplified:
-\begin{lstlisting}[style=scala]
-  object Int64 extends BaseNumbers[Int64] {
-    private case class Int64Impl(underlying: BigInt) extends Int64 
-  }
+
+Before:
+\begin{lstlisting}[style=scala]
+case object Satoshis extends BaseNumbers[Satoshis] {
+  val min = Satoshis(Int64.min)
+  val max = Satoshis(Int64.max)
+  val zero = Satoshis(Int64.zero)
+  val one = Satoshis(Int64.one)
+
+  def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
+
+  private case class SatoshisImpl(underlying: Int64) extends Satoshis
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+case object Satoshis extends BaseNumbers[Satoshis] {
+
+  val min = Satoshis(Int64.min)
+  val max = Satoshis(Int64.max)
+  val zero = Satoshis(Int64.zero)
+  val one = Satoshis(Int64.one)
+
+  def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
+}
+
+private case class SatoshisImpl(underlying: Int64) extends Satoshis
 \end{lstlisting}

### 2019-06-21 code/addition/src/main/scala/addition/modified/number/NumberType.scala
replace BigInt '&' with bound check

@@ -13,13 +13,6 @@ sealed abstract class Number {
 
   def toBigInt: BigInt = underlying
 
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
-  def andMask: BigInt
-
   /** Factory function to create the underlying T, for instance a UInt32 */
   def apply: A => Int64
 
@@ -30,7 +23,9 @@ sealed abstract class Number {
     * of this number type
     */
   private def checkResult(result: BigInt): A = {
-    require((result & andMask) == result,
+    require(
+         result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"),
       "Result was out of bounds, got: " + result)
     result
   }
@@ -47,7 +42,6 @@ sealed abstract class SignedNumber extends Number
   */
 sealed abstract class Int64 extends SignedNumber {
   override def apply: A => Int64 = Int64(_)
-  override def andMask = 0xffffffffffffffffL
 }
 
 /**

### 2019-06-21 paper/paper.tex
add before and after for & function

@@
 Due to the restrictions on BigInt, we can not use the \& function either.
-Simplified code:
-\begin{lstlisting}[style=scala]
-  sealed abstract class Number extends BasicArithmetic[Int64] {
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
   def andMask: BigInt
 
-    override def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-    private def checkResult(result: BigInt): BigInt = {
-      require((result & andMask) == result, "Result was out of bounds, got: " + result)
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
     result
   }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber extends Number
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: A => Int64 = Int64(_)
+  override def andMask = 0xffffffffffffffffL
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require(
+          result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"),
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber extends Number
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: A => Int64 = Int64(_)
 }

### 2019-06-21 paper/paper.tex
add before and after for private inner class

@@
 
+And:
+\begin{lstlisting}[style=stainless]
+  [Warning ] number/NumberType.scala:59:3: Could not extract tree in class: case private class Int64Impl extends Int64 with Product with Serializable {
+    [Warning ]   <caseaccessor> <paramaccessor> private[this] val underlying: BigInt = _;
+    [Warning ]   <stable> <caseaccessor> <accessor> <paramaccessor> def underlying: BigInt = Int64Impl.this.underlying;
+    [Warning ]   def <init>(underlying: BigInt): addition.modified.number.Int64.Int64Impl = {
+    [Warning ]     Int64Impl.super.<init>();
+    [Warning ]     ()
+    [Warning ]   };
+    [Warning ]   scala.this.Predef.require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(-9223372036854775808L)), "Number was too small for a int64, got: ".+(Int64Impl.this.underlying));
+    [Warning ]   scala.this.Predef.require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(9223372036854775807L)), "Number was too big for a int64, got: ".+(Int64Impl.this.underlying));
+    [Warning ]   <synthetic> def copy(underlying: BigInt = underlying): addition.modified.number.Int64.Int64Impl = new Int64Impl(underlying);
+    [Warning ]   <synthetic> def copy$default$1: BigInt = Int64Impl.this.underlying;
+    [Warning ]   override <synthetic> def productPrefix: String = "Int64Impl";
+    [Warning ]   <synthetic> def productArity: Int = 1;
+    [Warning ]   <synthetic> def productElement(x$1: Int): Any = x$1 match {
+    [Warning ]     case 0 => Int64Impl.this.underlying
+    [Warning ]     case _ => throw new IndexOutOfBoundsException(x$1.toString())
+    [Warning ]   };
+    [Warning ]   override <synthetic> def productIterator: Iterator[Any] = runtime.this.ScalaRunTime.typedProductIterator[Any](Int64Impl.this);
+    [Warning ]   <synthetic> def canEqual(x$1: Any): Boolean = x$1.$isInstanceOf[addition.modified.number.Int64.Int64Impl]();
+    [Warning ]   override <synthetic> def hashCode(): Int = ScalaRunTime.this._hashCode(Int64Impl.this);
+    [Warning ]   override <synthetic> def toString(): String = ScalaRunTime.this._toString(Int64Impl.this);
+    [Warning ]   override <synthetic> def equals(x$1: Any): Boolean = Int64Impl.this.eq(x$1.asInstanceOf[Object]).||(x$1 match {
+    [Warning ]   case (_: addition.modified.number.Int64.Int64Impl) => true
+    [Warning ]   case _ => false
+    [Warning ] }.&&({
+    [Warning ]     <synthetic> val Int64Impl$1: addition.modified.number.Int64.Int64Impl = x$1.asInstanceOf[addition.modified.number.Int64.Int64Impl];
+    [Warning ]     Int64Impl.this.underlying.==(Int64Impl$1.underlying).&&(Int64Impl$1.canEqual(Int64Impl.this))
+    [Warning ]   }))
+    [Warning ] } (class scala.reflect.internal.Trees$ClassDef)
+                 private case class Int64Impl(underlying: BigInt) extends Int64 {
+                 ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+\end{lstlisting}
+
 Stainless can not extract the private class inside the object.
@@
 private case class SatoshisImpl(underlying: Int64) extends Satoshis
+\end{lstlisting}
+
+And before:
+\begin{lstlisting}[style=scala]
+case object Int64 extends BaseNumbers[Int64] {
+  private case class Int64Impl(underlying: BigInt) extends Int64 {
+    require(underlying >= -9223372036854775808L,
+      "Number was too small for a int64, got: " + underlying)
+    require(underlying <= 9223372036854775807L,
+      "Number was too big for a int64, got: " + underlying)
+  }
+
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(-9223372036854775808L)
+  lazy val max = Int64(9223372036854775807L)
+
+  def apply(long: Long): Int64 = Int64(BigInt(long))
+
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+case object Int64 extends BaseNumbers[Int64] {
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(-9223372036854775808L)
+  lazy val max = Int64(9223372036854775807L)
+
+  def apply(long: Long): Int64 = Int64(BigInt(long))
+
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= -9223372036854775808L,
+    "Number was too small for a int64, got: " + underlying)
+  require(underlying <= 9223372036854775807L,
+    "Number was too big for a int64, got: " + underlying)
+}
 \end{lstlisting}

### 2019-06-21 code/addition/src/main/scala/addition/modified/number/NumberType.scala
extracted private inner class

@@ -56,13 +56,6 @@ trait BaseNumbers[T] {
 }
 
 case object Int64 extends BaseNumbers[Int64] {
-  private case class Int64Impl(underlying: BigInt) extends Int64 {
-    require(underlying >= -9223372036854775808L,
-      "Number was too small for a int64, got: " + underlying)
-    require(underlying <= 9223372036854775807L,
-      "Number was too big for a int64, got: " + underlying)
-  }
-
   lazy val zero = Int64(0)
   lazy val one = Int64(1)
 
@@ -72,4 +65,11 @@ case object Int64 extends BaseNumbers[Int64] {
   def apply(long: Long): Int64 = Int64(BigInt(long))
 
   def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= -9223372036854775808L,
+    "Number was too small for a int64, got: " + underlying)
+  require(underlying <= 9223372036854775807L,
+    "Number was too big for a int64, got: " + underlying)
 }

### 2019-06-21 code/addition/src/main/scala/addition/modified/number/NumberType.scala
removed second parameter from require (error message)

@@ -25,8 +25,7 @@ sealed abstract class Number {
   private def checkResult(result: BigInt): A = {
     require(
          result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"),
-      "Result was out of bounds, got: " + result)
+      && result >= BigInt("-9223372036854775808"))
     result
   }
 }
@@ -68,8 +67,6 @@ case object Int64 extends BaseNumbers[Int64] {
 }
 
 private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= -9223372036854775808L,
-    "Number was too small for a int64, got: " + underlying)
-  require(underlying <= 9223372036854775807L,
-    "Number was too big for a int64, got: " + underlying)
+  require(underlying >= -9223372036854775808L)
+  require(underlying <= 9223372036854775807L)
 }

### 2019-06-21 paper/paper.tex
add code for removing require

@@
 \begin{lstlisting}[style=stainless]
-  [Warning ] Code.scala:51:3: Could not extract tree in class: scala.this.Predef
-             .require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(
-             -9223372036854775808L)), "Number was too small for a int64, got: "
-             .+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
+  [Warning ] number/NumberType.scala:9:3: Could not extract tree in class: type A = BigInt (class scala.reflect.internal.Trees$TypeDef)
+            type A = BigInt
+            ^^^^^^^^^^^^^^^
+  [Warning ] number/NumberType.scala:71:3: Could not extract tree in class: scala.this.Predef.require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(-9223372036854775808L)), "Number was too small for a int64, got: ".+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
             require(underlying >= -9223372036854775808L,
             ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-  [Warning ] Code.scala:53:3: Could not extract tree in class: scala.this.Predef
-             .require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(
-             9223372036854775807L)), "Number was too big for a int64, got: "
-             .+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
+  [Warning ] number/NumberType.scala:73:3: Could not extract tree in class: scala.this.Predef.require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(9223372036854775807L)), "Number was too big for a int64, got: ".+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
             require(underlying <= 9223372036854775807L,
@@
   [ Error  ] checkResult$0 depends on missing dependencies: require$1.
+
 \end{lstlisting}
@@
 We can safely remove the string parameters from the requires, since they only serve as error messages.
+
+Before:
+\begin{lstlisting}[style=scala]
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= -9223372036854775808L,
+    "Number was too small for a int64, got: " + underlying)
+  require(underlying <= 9223372036854775807L,
+    "Number was too big for a int64, got: " + underlying)
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= -9223372036854775808L)
+  require(underlying <= 9223372036854775807L)
+}
+\end{lstlisting}
+
+And before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require(
+         result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"),
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require(
+          result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"))
+    result
+  }
+}
+\end{lstlisting}
 

### 2019-06-21 code/addition/src/main/scala/addition/modified/number/NumberType.scala
changed from BigInt with Long comparison to BigInt with BigInt comparison

@@ -67,6 +67,6 @@ case object Int64 extends BaseNumbers[Int64] {
 }
 
 private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= -9223372036854775808L)
-  require(underlying <= 9223372036854775807L)
+  require(underlying >= BigInt("-9223372036854775808"))
+  require(underlying <= BigInt("9223372036854775807"))
 }

### 2019-06-21 code/addition/src/main/scala/addition/modified/number/NumberType.scala
replace every A defined as type A = BigInt with BigInt directly

@@ -6,15 +6,13 @@ package addition.modified.number
   * unsigned integer types
   */
 sealed abstract class Number {
-  type A = BigInt
-
   /** The underlying scala number used to to hold the number */
-  protected def underlying: A
+  protected def underlying: BigInt
 
   def toBigInt: BigInt = underlying
 
   /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
+  def apply: BigInt => Int64
 
   def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
 
@@ -22,7 +20,7 @@ sealed abstract class Number {
     * Checks if the given result is within the range
     * of this number type
     */
-  private def checkResult(result: BigInt): A = {
+  private def checkResult(result: BigInt): BigInt = {
     require(
          result <= BigInt("9223372036854775807")
       && result >= BigInt("-9223372036854775808"))
@@ -40,7 +38,7 @@ sealed abstract class SignedNumber extends Number
   * Represents a int64_t in C
   */
 sealed abstract class Int64 extends SignedNumber {
-  override def apply: A => Int64 = Int64(_)
+  override def apply: BigInt => Int64 = Int64(_)
 }
 
 /**

### 2019-06-21 paper/paper.tex
changed from BigInt with Long comparison to BigInt with BigInt comparison
before and after for concrete type

@@
 
+\end{lstlisting}
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require(
+          result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"))
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber extends Number
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: A => Int64 = Int64(_)
+}  
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: BigInt
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: BigInt => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): BigInt = {
+    require(
+         result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"))
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber extends Number
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: BigInt => Int64 = Int64(_)
+}
 \end{lstlisting}
@@
 So we search for all Long values in the code.
+
+Before:
 \begin{lstlisting}[style=scala]
@@
   require(underlying <= 9223372036854775807L)
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= BigInt("-9223372036854775808"))
+  require(underlying <= BigInt("9223372036854775807"))
 }  

### 2019-06-21 code/addition/src/main/scala/addition/modified/number/NumberType.scala
change Long to BigInt

@@ -56,10 +56,8 @@ case object Int64 extends BaseNumbers[Int64] {
   lazy val zero = Int64(0)
   lazy val one = Int64(1)
 
-  lazy val min = Int64(-9223372036854775808L)
-  lazy val max = Int64(9223372036854775807L)
-
-  def apply(long: Long): Int64 = Int64(BigInt(long))
+  lazy val min = Int64(BigInt("-9223372036854775808"))
+  lazy val max = Int64(BigInt("9223372036854775807"))
 
   def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
 }

### 2019-06-21 paper/paper.tex
add more code example to BigInt to Long conversion
add stainless output

@@
 
-\subsection{Rewriting BigInt Comparison with Long}
+\subsection{Rewriting BigInt to Long Conversions}
 
@@
   [ Error  ] inv$4 depends on missing dependencies: long2bigInt$0.
+  [ Error  ] apply$14 depends on missing dependencies: BigInt$0, apply$15.
 \end{lstlisting}
@@
 
+And before:
+\begin{lstlisting}[style=scala]
+case object Int64 extends BaseNumbers[Int64] {
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(-9223372036854775808L)
+  lazy val max = Int64(9223372036854775807L)
+
+  def apply(long: Long): Int64 = Int64(BigInt(long))
+
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+case object Int64 extends BaseNumbers[Int64] {
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(BigInt("-9223372036854775808"))
+  lazy val max = Int64(BigInt("9223372036854775807"))
+
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+\end{lstlisting}
+
 Looks like it can not compare a BigInt with a Long value.
@@
 Finally, we get some output from Stainless about the verification in the code.
-\begin{lstlisting}[style=stainless]
-[...]
-[Warning ]  => INVALID
-[Warning ] Found counter-example:
-[Warning ]   thiss: { x: Object | @unchecked isNumber(x) } -> Int64Impl(0)
-[Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(9223372036854775808)
-[...]
-[  Info  ] +  precond. (call checkResult(thiss, underlying(thiss) + u ...)  invalid  Code.scala:16:45
-[...]
-[  Info  ] total: 5  valid: 4  (4 from cache) invalid: 1  unknown: 0  time:  1.317
-\end{lstlisting}
+
+\todo{kai special chars from stainless output what now?}
+% \begin{lstlisting}[style=stainless][Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
+%   [  Info  ]  - Checking cache: 'adt invariant' VC for apply @33:39...
+%   [  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
+%   [  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
+%   [  Info  ] Cache hit: 'adt invariant' VC for apply @33:39...
+%   [  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
+%   [  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
+%   [Warning ]  => INVALID
+%   [Warning ] Found counter-example:
+%   [Warning ]   bigInt: BigInt -> -9223372036854775809
+%   [  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+%   [  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+%   [  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+%   [  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
+%   [Warning ]  => INVALID
+%   [Warning ] Found counter-example:
+%   [Warning ]   num: Number   -> Int64Impl(9223372036854775807)
+%   [Warning ]   thiss: Number -> Int64Impl(1)
+%   [  Info  ]   ┌───────────────────┐
+%   [  Info  ] ╔═╡ stainless summary ╞══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╗
+%   [  Info  ] ║ └───────────────────┘                                                                                                                      ║
+%   [  Info  ] ║ +      precond. (call checkResult(thiss, underlying(thiss) + u ...)  invalid           U:smt-z3  number/NumberType.scala:17:36       0.110 ║
+%   [  Info  ] ║ apply  adt invariant                                                 valid from cache            currency/CurrencyUnits.scala:33:39  1.136 ║
+%   [  Info  ] ║ apply  adt invariant                                                 invalid           U:smt-z3  number/NumberType.scala:62:38       0.394 ║
+%   [  Info  ] ╟┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄╢
+%   [  Info  ] ║ total: 3    valid: 1    (1 from cache) invalid: 2    unknown: 0    time:   1.640                                                           ║
+%   [  Info  ] ╚════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╝
+%   [  Info  ] Shutting down executor service.  
+% \end{lstlisting}
 

### 2019-06-21 paper/paper.tex
add specification

@@
 We can use equals (==) directly on Satoshis, because it is a case class.
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class CurrencyUnit {
+  def satoshis: Satoshis
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
+
+  protected def underlying: Int64
+}
+\end{lstlisting}
+
 The addition will now look like this:
 \begin{lstlisting}[style=scala]
+sealed abstract class CurrencyUnit {
+  def satoshis: Satoshis
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
   override def +(c: CurrencyUnit): CurrencyUnit = {
@@
   } ensuring(res => res.satoshis == this.satoshis)
+
+  protected def underlying: Int64
+}
 \end{lstlisting}

### 2019-06-21 code/addition/src/main/scala/addition/modified/currency/CurrencyUnits.scala

@@ -10,8 +10,9 @@ sealed abstract class CurrencyUnit {
   def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
 
   def +(c: CurrencyUnit): CurrencyUnit = {
+    require(c.satoshis == Satoshis.zero)
     Satoshis(satoshis.underlying + c.satoshis.underlying)
-  }
+  } ensuring(res => res.satoshis == this.satoshis)
 
   protected def underlying: Int64
 }

### 2019-06-21 paper/paper.tex
more stainless output

@@
 
-The Stainless output again:
-\begin{lstlisting}[style=stainless]
-  [Warning ] Found counter-example:
-  [Warning ]   thiss: { x: Object | @unchecked isNumber(x) } -> Int64Impl(0)
-  [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(9223372036854775808)
-\end{lstlisting}
+Stainless output now: \todo{kai special chars again}
+% \begin{lstlisting}[style=stainless][Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
+%   [  Info  ]  - Checking cache: 'postcondition' VC for + @12:3...
+%   [  Info  ]  - Checking cache: 'adt invariant' VC for apply @34:39...
+%   [  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
+%   [  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
+%   [  Info  ] Cache hit: 'adt invariant' VC for apply @34:39...
+%   [  Info  ] Cache hit: 'postcondition' VC for + @12:3...
+%   [  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
+%   [  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
+%   [Warning ]  => INVALID
+%   [Warning ] Found counter-example:
+%   [Warning ]   bigInt: BigInt -> -9223372036854775809
+%   [  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+%   [  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+%   [  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+%   [  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
+%   [Warning ]  => INVALID
+%   [Warning ] Found counter-example:
+%   [Warning ]   num: Number   -> Int64Impl(9223372036854775807)
+%   [Warning ]   thiss: Number -> Int64Impl(1)
+%   [  Info  ]   ┌───────────────────┐
+%   [  Info  ] ╔═╡ stainless summary ╞═════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╗
+%   [  Info  ] ║ └───────────────────┘                                                                                                                                                                             ║
+%   [  Info  ] ║ +      postcondition                                                 valid from cache            ../../../../../bin/../src/main/scala/addition/modified/currency/CurrencyUnits.scala:12:3   1.266 ║
+%   [  Info  ] ║ +      precond.
+(call checkResult(thiss, underlying(thiss) + u ...)  invalid           U:smt-z3  ../../../../../bin/../src/main/scala/addition/modified/number/NumberType.scala:17:36       0.089 ║
+%   [  Info  ] ║ apply  adt invariant                                                 valid from cache            ../../../../../bin/../src/main/scala/addition/modified/currency/CurrencyUnits.scala:34:39  0.194 ║
+%   [  Info  ] ║ apply  adt invariant                                                 invalid           U:smt-z3  ../../../../../bin/../src/main/scala/addition/modified/number/NumberType.scala:62:38       0.348 ║
+%   [  Info  ] ╟┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄╢
+%   [  Info  ] ║ total: 4    valid: 2    (2 from cache) invalid: 2    unknown: 0    time:   1.897                                                                                                                  ║
+%   [  Info  ] ╚═══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╝
+%   [  Info  ] Shutting down executor service.
+% \end{lstlisting}
 
@@
 \begin{lstlisting}[style=scala]
-  sealed abstract class Number extends BasicArithmetic[Int64] {
-    override def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
+sealed abstract class Number {
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: BigInt
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: BigInt => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
   private def checkResult(result: BigInt): BigInt = {
@@
           result <= BigInt("9223372036854775807")
-        && result >= BigInt("-9223372036854775808")
-      )
+      && result >= BigInt("-9223372036854775808"))
     result

### 2019-06-22 paper/paper.tex
fixed paths

--- 2019-06-21 paper/paper.tex
@@
 
-The Stainless output again:
-\begin{lstlisting}[style=stainless]
-  [Warning ] Found counter-example:
-  [Warning ]   thiss: { x: Object | @unchecked isNumber(x) } -> Int64Impl(0)
-  [Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(9223372036854775808)
-\end{lstlisting}
+Stainless output now: \todo{kai special chars again}
+% \begin{lstlisting}[style=stainless][Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
+%   [  Info  ]  - Checking cache: 'postcondition' VC for + @12:3...
+%   [  Info  ]  - Checking cache: 'adt invariant' VC for apply @34:39...
+%   [  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
+%   [  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
+%   [  Info  ] Cache hit: 'adt invariant' VC for apply @34:39...
+%   [  Info  ] Cache hit: 'postcondition' VC for + @12:3...
+%   [  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
+%   [  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
+%   [Warning ]  => INVALID
+%   [Warning ] Found counter-example:
+%   [Warning ]   bigInt: BigInt -> -9223372036854775809
+%   [  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+%   [  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+%   [  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+%   [  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
+%   [Warning ]  => INVALID
+%   [Warning ] Found counter-example:
+%   [Warning ]   num: Number   -> Int64Impl(9223372036854775807)
+%   [Warning ]   thiss: Number -> Int64Impl(1)
+%   [  Info  ]   ┌───────────────────┐
+%   [  Info  ] ╔═╡ stainless summary ╞══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╗
+%   [  Info  ] ║ └───────────────────┘                                                                                                                      ║
+%   [  Info  ] ║ +      postcondition                                                 valid from cache            currency/CurrencyUnits.scala:12:3   1.266 ║
+%   [  Info  ] ║ +      precond. (call checkResult(thiss, underlying(thiss) + u ...)  invalid           U:smt-z3  number/NumberType.scala:17:36       0.089 ║
+%   [  Info  ] ║ apply  adt invariant                                                 valid from cache            currency/CurrencyUnits.scala:34:39  0.194 ║
+%   [  Info  ] ║ apply  adt invariant                                                 invalid           U:smt-z3  number/NumberType.scala:62:38       0.348 ║
+%   [  Info  ] ╟┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄╢
+%   [  Info  ] ║ total: 4    valid: 2    (2 from cache) invalid: 2    unknown: 0    time:   1.897                                                           ║
+%   [  Info  ] ╚════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╝
+%   [  Info  ] Shutting down executor service.
+% \end{lstlisting}
 

### 2019-06-22 code/addition/src/main/scala/addition/modified/number/NumberType.scala
propagate require

@@ -14,7 +14,13 @@ sealed abstract class Number {
   /** Factory function to create the underlying T, for instance a UInt32 */
   def apply: BigInt => Int64
 
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+  def +(num: Int64): Int64 = {
+    require(
+         num == Int64.zero
+      && this.underlying <= BigInt("9223372036854775807")
+      && this.underlying >= BigInt("-9223372036854775808"))
+    apply(checkResult(underlying + num.underlying))
+  }
 
   /**
     * Checks if the given result is within the range
@@ -38,7 +44,13 @@ sealed abstract class SignedNumber extends Number
   * Represents a int64_t in C
   */
 sealed abstract class Int64 extends SignedNumber {
-  override def apply: BigInt => Int64 = Int64(_)
+  override def apply: BigInt => Int64 = bigInt => {
+    require(
+         bigInt >= BigInt("-9223372036854775808")
+      && bigInt <= BigInt("9223372036854775807"))
+
+    Int64(bigInt)
+  }
 }
 
 /**
@@ -59,7 +71,13 @@ case object Int64 extends BaseNumbers[Int64] {
   lazy val min = Int64(BigInt("-9223372036854775808"))
   lazy val max = Int64(BigInt("9223372036854775807"))
 
-  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+  def apply(bigInt: BigInt): Int64 = {
+    require(
+         bigInt >= BigInt("-9223372036854775808")
+      && bigInt <= BigInt("9223372036854775807"))
+
+    Int64Impl(bigInt)
+  }
 }
 
 private case class Int64Impl(underlying: BigInt) extends Int64 {

### 2019-06-22 paper/paper.tex
add code about propagate require

@@
 Corresponding code:
-
 \begin{lstlisting}[style=scala]
@@
 Since we do only allow zero as a parameter, the easiest way is to restrict it to zero here too.
+
+Code after:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: BigInt
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: BigInt => Int64
+
+  def +(num: Int64): Int64 = {
+    require(
+          num == Int64.zero
+      && this.underlying <= BigInt("9223372036854775807")
+      && this.underlying >= BigInt("-9223372036854775808"))
+    apply(checkResult(underlying + num.underlying))
+  }
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): BigInt = {
+    require(
+          result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"))
+    result
+  }
+}
+\end{lstlisting}
+
+The same problem occurs in the Int64 constructor.
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: BigInt => Int64 = Int64(_)
+}
+
+case object Int64 extends BaseNumbers[Int64] {
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(BigInt("-9223372036854775808"))
+  lazy val max = Int64(BigInt("9223372036854775807"))
+
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= BigInt("-9223372036854775808"))
+  require(underlying <= BigInt("9223372036854775807"))
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: BigInt => Int64 = bigInt => {
+    require(
+          bigInt >= BigInt("-9223372036854775808")
+      && bigInt <= BigInt("9223372036854775807"))
+
+    Int64(bigInt)
+  }
+}
+
+case object Int64 extends BaseNumbers[Int64] {
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(BigInt("-9223372036854775808"))
+  lazy val max = Int64(BigInt("9223372036854775807"))
+
+  def apply(bigInt: BigInt): Int64 = {
+    require(
+         bigInt >= BigInt("-9223372036854775808")
+      && bigInt <= BigInt("9223372036854775807"))
+
+    Int64Impl(bigInt)
+  }
+}
+
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= BigInt("-9223372036854775808"))
+  require(underlying <= BigInt("9223372036854775807"))
+}
+\end{lstlisting}
 

### 2019-06-22 paper/paper.tex
add final stainless output
fix indentation

@@
                                         ^^^^^^^^^^^^^^
-
 \end{lstlisting}
@@
               ^^^^^^^^^^^^^^^
-
 \end{lstlisting}
@@
                       ^^^^^^^^^^^^^^^^
-
 \end{lstlisting}
@@
 [ Error  ] checkResult$0 depends on missing dependencies: require$1.
-
 \end{lstlisting}
@@
 Stainless output now: \todo{kai special chars again}
-% \begin{lstlisting}[style=stainless][Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
+% \begin{lstlisting}[style=stainless]
+% [Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
 % [  Info  ]  - Checking cache: 'postcondition' VC for + @12:3...
@@
 
-Finally, everything is green and correctly verified.
-\begin{figure}
-	\centering
-		\includegraphics[width=\textwidth]{final_verify_output}
-	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
-\end{figure}
+Finally, everything is green and correctly verified. \todo{kai special chars}
+% \begin{lstlisting}[style=scala]
+% [Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
+% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), 1))' VC for one @69:18...
+% [  Info  ] Cache hit: 'precond. (call apply(Int64(), 1))' VC for one @69:18...
+% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), bigInt))' VC for apply @52:5...
+% [  Info  ] Cache miss: 'precond. (call apply(Int64(), bigInt))' VC for apply @52:5...
+% [  Info  ]  - Now solving 'precond. (call apply(Int64(), bigInt))' VC for apply @52:5...
+% [  Info  ]  - Result for 'precond. (call apply(Int64(), bigInt))' VC for apply @52:5:
+% [  Info  ]  => VALID
+% [  Info  ]  - Checking cache: 'precond. (call +(underlying(satoshis(thiss)), underlyin ...)' VC for + @14:14...
+% [  Info  ]  - Checking cache: 'postcondition' VC for + @12:3...
+% [  Info  ] Cache hit: 'precond. (call +(underlying(satoshis(thiss)), underlyin ...)' VC for + @14:14...
+% [  Info  ] Cache miss: 'postcondition' VC for + @12:3...
+% [  Info  ]  - Now solving 'postcondition' VC for + @12:3...
+% [  Info  ]  - Result for 'postcondition' VC for + @12:3:
+% [  Info  ]  => VALID
+% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), -9223372036854775808))' VC for min @71:18...
+% [  Info  ] Cache hit: 'precond. (call apply(Int64(), -9223372036854775808))' VC for min @71:18...
+% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @34:39...
+% [  Info  ] Cache hit: 'adt invariant' VC for apply @34:39...
+% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), 9223372036854775807))' VC for max @72:18...
+% [  Info  ] Cache hit: 'precond. (call apply(Int64(), 9223372036854775807))' VC for max @72:18...
+% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @79:5...
+% [  Info  ] Cache hit: 'adt invariant' VC for apply @79:5...
+% [  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @22:11...
+% [  Info  ] Cache hit: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @22:11...
+% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), 0))' VC for zero @68:19...
+% [  Info  ] Cache hit: 'precond. (call apply(Int64(), 0))' VC for zero @68:19...
+% [  Info  ]   ┌───────────────────┐
+% [  Info  ] ╔═╡ stainless summary ╞══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╗
+% [  Info  ] ║ └───────────────────┘                                                                                                                      ║
+% [  Info  ] ║ +      postcondition                                                 valid             U:smt-z3  currency/CurrencyUnits.scala:12:3   0.643 ║
+% [  Info  ] ║ +      precond. (call +(underlying(satoshis(thiss)), underlyin ...)  valid from cache            currency/CurrencyUnits.scala:14:14  0.097 ║
+% [  Info  ] ║ +      precond. (call checkResult(thiss, underlying(thiss) + u ...)  valid from cache            number/NumberType.scala:22:11       0.015 ║
+% [  Info  ] ║ apply  adt invariant                                                 valid from cache            currency/CurrencyUnits.scala:34:39  0.005 ║
+% [  Info  ] ║ apply  precond. (call apply(Int64(), bigInt))                        valid             U:smt-z3  number/NumberType.scala:52:5        0.187 ║
+% [  Info  ] ║ apply  adt invariant                                                 valid from cache            number/NumberType.scala:79:5        0.017 ║
+% [  Info  ] ║ max    precond. (call apply(Int64(), 9223372036854775807))           valid from cache            number/NumberType.scala:72:18       0.001 ║
+% [  Info  ] ║ min    precond. (call apply(Int64(), -9223372036854775808))          valid from cache            number/NumberType.scala:71:18       0.001 ║
+% [  Info  ] ║ one    precond. (call apply(Int64(), 1))                             valid from cache            number/NumberType.scala:69:18       0.915 ║
+% [  Info  ] ║ zero   precond. (call apply(Int64(), 0))                             valid from cache            number/NumberType.scala:68:19       0.001 ║
+% [  Info  ] ╟┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄╢
+% [  Info  ] ║ total: 10   valid: 10   (8 from cache) invalid: 0    unknown: 0    time:   1.882                                                           ║
+% [  Info  ] ╚════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╝
+% [  Info  ] Shutting down executor service.
+% \end{lstlisting}
+
+% \begin{figure}
+% 	\centering
+% 		\includegraphics[width=\textwidth]{final_verify_output}
+% 	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
+% \end{figure}
 
 The verified code.
-\lstinputlisting[style=scala]{../code/currsniporig/src/main/scala/code/end/Code.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/modified/number/NumberType.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/modified/currency/CurrencyUnits.scala}
 

### 2019-06-22 paper/paper.tex
Added more text to Pure Scala Fragment

@@
 \paragraph{The Pure Scala Fragment.} Stainless supports only a subset
-of Scala. \todo{Anna It supports blabla, in particular it supports not
-  blablabla. (check class with var)}
-
-\todo{why does stainless use its own BigInt etc?}
-
-They call this Pure Scala.  You can find the specification of Pure
+of Scala. They call this Pure Scala.  You can find the specification of Pure
 Scala on the Stainless documentation website
@@
   Scala}.
+\todo{Kai: check the text
+
+It supports Algebraic Data Type definitions in the form of 
+Abstract Class and Case Class/Case Object,
+inheritance of an Abstract Class by a Case Class/Case Object,
+Object for grouping classes and functions,  
+short-circuit interpretation of boolean expressions, 
+Generics with invariant type parameters and simple hierarchy 
+(subclasses define the same type parameters in the same order),
+default values of function parameters,
+Pattern matching, 
+Values,
+anonymous classes and more.
+
+As an extension to Pure Scala Stainless supports some imperative features.
+It supports 
+declaring a variable in a local scope of a function, 
+declaring a variable in trait without assigning a default value,
+while loops,
+case classes with \textit{var} arguments.
+
+In particular Stainless does not support
+Class definition,
+inheritance by Object,
+abstract type members,
+variance annotations, upper and lower type bounds in generics,
+variables in classes and objects,
+private inner classes,
+messages as a second argument in the Require function and more.}
+
+\todo{why does stainless use its own BigInt etc?}
 

### 2019-06-22 paper/paper.tex
Added Reason of own library.
Added text to 2.4 about Stainless integration.

@@
 
-\todo{why does stainless use its own BigInt etc?}
-
 In addition, Stainless has its own library with annotations,
 reimplementation of some core data types, collections and input-output
-functions and more.  Some of them are described on the website in the
-section
+functions and more. \todo{Kai: check the text.
+This allows Stainless to map correctly operations 
+to mathematical function inside of SMT solver.
+Some elements of the library} are described on the website in the section
 \href{https://epfl-lara.github.io/stainless/library.html}{Stainless
@@
 
-\todo{anna: what does it mean to integrate}
-
+\todo{Kai: check the text
+In order to verify a project, Stainless must be integrated into it.
+We can integrate it in an sbt project adding the Stainless Plugin 
+and the required resolver to pugins.sbt. 
+Another option to use Stainless is to import its libraries in a program code 
+and verify a program from command line using the pre-packaged Stainless JAR file 
+or using the Stainless script built from source.}
 Trying to integrate Stainless in Bitcoin-S caused a lot of troubles,

### 2019-06-23 paper/paper.tex
changed transaction code to see actual transaction

--- 2019-06-22 paper/paper.tex
@@
 
-  val signedTxF: Future[Transaction] = txBuilderF
-    .flatMap(_.sign)                       // call sign on the transaction builder
-    .map {
-      (tx: Transaction) => println(tx.hex) // transaction in hex for the bitcoin network
-    }
-\end{lstlisting}
-
-\todo{ramon: change code to get an actual transaction, completely understand and explain the code}
+  val txF: Future[Transaction] = txBuilderF.flatMap(_.sign)
+
+  val tx: Transaction = Await.result(signedTxF, 1 second)
+\end{lstlisting}
+
+\todo{ramon: completely understand and explain the code}
 

### 2019-06-23 paper/paper.tex
Added text about errors to the section Adjusting no-inflation property

@@
 functions and more. \todo{Kai: check the text.
+
 This allows Stainless to map correctly operations 
@@
 \todo{Kai: check the text
+
 In order to verify a project, Stainless must be integrated into it.
@@
 
-After integrating Stainless in Bitcoin-S, there were many errors.  It
-takes too much time to fix them all so it should be easier to extract
+\todo{Kai: about errors, is it enough to reference to the chapter 4.2 where they are described??}
+\todo{After integrating the Stainless plugin in the Bitcoin-S sbt project, 
+there were many errors because of the different sbt versions.  
+Some errors are described in the section 4.2.}
+It takes too much time to fix them all so it should be easier to extract
 the classes needed for the \texttt{checkTransaction} function.
-\todo{anna: which errors?}
 

### 2019-06-24 paper/paper.tex
add appendix
move transformations to appendix

@@
 
-\subsection{Fixing a Bug in Bitcoin-S}
-\label{sec:bugfix}
+\paragraph{Stainless integration problems.}
+
+\todo{Kai: check the text}
+
+In order to verify a project, Stainless must be integrated into it.
+We can integrate it in an sbt project adding the Stainless Plugin 
+and the required resolver to pugins.sbt. 
+Another option to use Stainless is to import its libraries in a program code 
+and verify a program from command line using the pre-packaged Stainless JAR file 
+or using the Stainless script built from source.
+Trying to integrate Stainless in Bitcoin-S caused a lot of troubles,
+mainly because of version conflicts.  For more details see chapter
+\ref{chap:appendix_arb}.
+
+\todo{Kai: about errors, is it enough to reference to the chapter 4.2 where they are described??}
+\todo{After integrating the Stainless plugin in the Bitcoin-S sbt project, 
+there were many errors because of the different sbt versions.  
+Some errors are described in the section 4.2.}
+It takes too much time to fix them all so it should be easier to extract
+the classes needed for the \texttt{checkTransaction} function.
+
+\paragraph{Extracting the relevant code.}
+The extracted code has more than 1500 lines.  After running Stainless
+on it, it still throws a huge bunch of errors about what Stainless can
+not reason about. After fixing some of those errors, there appear new
+ones.  So this would require changing nearly everything of the
+extracted code.
+
+Let's adjust the property from the \nameref{property_1} to the
+\nameref{property_2}, because there is not enough time to fix all
+these errors and write the verification.  This is way less interesting
+to verify but gives still many insights in what needs to be done to
+verify some parts of Bitcoin-S and other source code.
+
+\paragraph{Fixing a Bug in Bitcoin-S}
 
@@
 future.
-
-\subsection{Adjusting No-Inflation Property}
-
-\todo{Kai: check the text
-
-In order to verify a project, Stainless must be integrated into it.
-We can integrate it in an sbt project adding the Stainless Plugin 
-and the required resolver to pugins.sbt. 
-Another option to use Stainless is to import its libraries in a program code 
-and verify a program from command line using the pre-packaged Stainless JAR file 
-or using the Stainless script built from source.}
-Trying to integrate Stainless in Bitcoin-S caused a lot of troubles,
-mainly because of version conflicts.  For more details see chapter
-\ref{chap:appendix_arb}.
-
-\todo{Kai: about errors, is it enough to reference to the chapter 4.2 where they are described??}
-\todo{After integrating the Stainless plugin in the Bitcoin-S sbt project, 
-there were many errors because of the different sbt versions.  
-Some errors are described in the section 4.2.}
-It takes too much time to fix them all so it should be easier to extract
-the classes needed for the \texttt{checkTransaction} function.
-
-The extracted code has more than 1500 lines.  After running Stainless
-on it, it still throws a huge bunch of errors about what Stainless can
-not reason about. After fixing some of those errors, there appear new
-ones.  So this would require changing nearly everything of the
-extracted code.
-
-Let's adjust the property from the \nameref{property_1} to the
-\nameref{property_2}, because there is not enough time to fix all
-these errors and write the verification.  This is way less interesting
-to verify but gives still many insights in what needs to be done to
-verify some parts of Bitcoin-S and other source code.
-
-So let's look at the \nameref{property_2}.
 
@@
 
-When we run Stainless on this code (without any properties to prove), it throws the following errors:
-
-\begin{lstlisting}[style=stainless]
-[ Error  ] currency/CurrencyUnits.scala:6:3: Stainless doesn't support abstract type members
-            type A
-            ^^^^^^
-[ Error  ] currency/CurrencyUnits.scala:26:33: Only literal arguments are allowed for BigInt.
-            def toBigInt: BigInt = BigInt(toLong)
-                                          ^^^^^^
-[ Error  ] currency/CurrencyUnits.scala:33:1: Objects cannot extend classes or implement traits, use a case object instead
-          object Satoshis extends BaseNumbers[Satoshis] {
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-[ Error  ] number/NumberType.scala:65:1: Objects cannot extend classes or implement traits, use a case object instead
-          object Int64 extends BaseNumbers[Int64] {
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-[  Info  ] Shutting down executor service.
-\end{lstlisting}
-
-So let's see how we can fix those errors.
-
-\subsection{Turning Object into Case Object}
-
-Stainless output:
-\begin{lstlisting}[style=stainless]
-[ Error  ] currency/CurrencyUnits.scala:33:1: Objects cannot extend classes or implement traits, use a case object instead
-          object Satoshis extends BaseNumbers[Satoshis] {
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-[ Error  ] number/NumberType.scala:65:1: Objects cannot extend classes or implement traits, use a case object instead
-          object Int64 extends BaseNumbers[Int64] {
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-\end{lstlisting}
-
-Code before:
-\begin{lstlisting}[style=scala]
-object Int64 extends BaseNumbers[Int64] { ... }
-object Satoshis extends BaseNumbers[Satoshis] { ... }
-\end{lstlisting}
-
-Code after:
-\begin{lstlisting}[style=scala]
-case object Int64 extends BaseNumbers[Int64] { ... }
-case object Satoshis extends BaseNumbers[Satoshis] { ... }
-\end{lstlisting}
-
-Here, we can just change the objects from object to case object.
-Stainless recommendation is to use objects for modules and case objects as algebraic data types.
-
-\todo{Ramon does this change semantics?}
-
-This is due to the internal design of Scala.
-It's possible to reason about case object but not about object.
-This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
-The \href{https://github.com/epfl-lara/stainless/issues/520}{issue number 520} on Stainless GitHub gives some thoughts, if you want to know more.
-
-\subsection{Getting Rid of Abstract Type Member}
-
-Stainless output:
-\begin{lstlisting}[style=stainless]
-[ Error  ] currency/CurrencyUnits.scala:6:3: Stainless doesn't support abstract type members
-            type A
-            ^^^^^^
-\end{lstlisting}
-
-This should be easy to rewrite by using generics instead of an abstract type, right?
-Unfortunately not.
-The problem is, CurrencyUnit uses one of its implementing classes: Satoshis.
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class CurrencyUnit {
-  type A
-
-  def satoshis: Satoshis
-
-  def !=(c: CurrencyUnit): Boolean = !(this == c)
-
-  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
-  def +(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying + c.satoshis.underlying)
-  }
-
-  protected def underlying: A
-}
-
-sealed abstract class Satoshis extends CurrencyUnit {
-  override type A = Int64
-
-  override def satoshis: Satoshis = this
-
-  def toBigInt: BigInt = BigInt(toLong)
-
-  def toLong: Long = underlying.toLong
-
-  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
-}
-\end{lstlisting}
-
-What happens, if we typify CurrencyUnit with A, meaning to make it generic with type A?
-
-Satoshis extends CurrencyUnit with type Int64, so it would be of type CurrencyUnit[Int64].
-That's too specific, because the return type of the addition is then CurrencyUnit[Int64] not CurrencyUnit[A].
-Maybe the Bitcoin-S developers should reimplement this part and not use Satoshis directly.
-
-Since there is no easy way to fix it and the code should stay as much as possible the original, we just remove the abstract type and set it to Int64.
-This limits the verification a bit, but as we only want to verify the addition in satoshis, that's OK.
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class CurrencyUnit {
-  def satoshis: Satoshis
-
-  def !=(c: CurrencyUnit): Boolean = !(this == c)
-
-  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
-  def +(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying + c.satoshis.underlying)
-  }
-
-  protected def underlying: Int64
-}
-
-sealed abstract class Satoshis extends CurrencyUnit {
-  override def satoshis: Satoshis = this
-
-  def toBigInt: BigInt = BigInt(toLong)
-
-  def toLong: Long = underlying.toLong
-
-  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
-}
-\end{lstlisting}
-
-\subsection{Replacing the BigInt Constructor Argument With A String Literal}
-
-Stainless output:
-\begin{lstlisting}[style=stainless]
-[ Error  ] currency/CurrencyUnits.scala:26:33: Only literal arguments are allowed for BigInt.
-            def toBigInt: BigInt = BigInt(toLong)
-                                          ^^^^^^
-\end{lstlisting}
-
-As described before, Stainless supports only a subset of Scala.
-The BigInt from the Stainless library is a bit restricted.
-One such restriction is, that BigInt does not support dynamic BigInt construction.
-Thus, the constructor parameter of BitInt must be a literal argument.
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Satoshis extends CurrencyUnit {
-  override def satoshis: Satoshis = this
-
-  def toBigInt: BigInt = BigInt(toLong)
-
-  def toLong: Long = underlying.toLong
-
-  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Satoshis extends CurrencyUnit {
-  override def satoshis: Satoshis = this
-
-  def toBigInt: BigInt = underlying.toBigInt
-
-  def toLong: Long = underlying.toLong
-
-  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
-}
-\end{lstlisting}
-
-This would be really hard to refactor, because Bitcoin-S tries to be as dynamic as possible so it can be used with cryptocurrencies other than bitcoins.
-Maybe it could be impossible, because they need to parse dynamic values from the bitcoin network.
-
-Luckily, we can use toBigInt on the field \texttt{underlying} directly instead of toLong.
-So, instead of converting the underlying to Long and back to BigInt we convert underlying directly to BigInt.
-
-After fixing all Stainless errors, a new error appears.
-
-\subsection{Getting Rid of Special Generics}
-
-Stainless output:
-\begin{lstlisting}[style=stainless]
-[ Error  ] number/NumberType.scala:8:30: Unknown type parameter type T
-          sealed abstract class Number[T <: Number[T]] {
-                                        ^^^^^^^^^^^^^^
-\end{lstlisting}
-
-This is a missing feature in Stainless.
-It does not support upper type boundaries on the class itself.
-To track this, \href{https://github.com/epfl-lara/stainless/issues/519}{issue 519} was created on GitHub during this work.
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number[T <: Number[T]] {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toLong: Long = toBigInt.bigInteger.longValueExact()
-  def toBigInt: BigInt = underlying
-
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
-  def andMask: BigInt
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => T
-
-  def +(num: T): T = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require((result & andMask) == result,
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}
-
-/**
-  * Represents a signed number in our number system
-  * Instances of this is [[Int64]]
-  */
-sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
-
-/**
-  * Represents a int64_t in C
-  */
-sealed abstract class Int64 extends SignedNumber[Int64] {
-  override def apply: A => Int64 = Int64(_)
-  override def andMask = 0xffffffffffffffffL
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toLong: Long = toBigInt.bigInteger.longValueExact()
-  def toBigInt: BigInt = underlying
-
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
-  def andMask: BigInt
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require((result & andMask) == result,
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}
-
-/**
-  * Represents a signed number in our number system
-  * Instances of this is [[Int64]]
-  */
-sealed abstract class SignedNumber extends Number
-
-/**
-  * Represents a int64_t in C
-  */
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: A => Int64 = Int64(_)
-  override def andMask = 0xffffffffffffffffL
-}
-\end{lstlisting}
-
-Despite this, in order to be able to continue, we make this a concrete type by replacing T with Int64.
-Int64,  because Satoshis uses only Int64.
-There are other number types like UInt16 but for our property we don't need them.
-
-Now, there are two new errors.
-
-\subsection{Removing toLong}
-
-Stainless output:
-\begin{lstlisting}[style=stainless]
-[ Error  ] number/NumberType.scala:14:22: Unknown call to bigInteger on Number.this.toBigInt (BigInt) with arguments List() of type List()
-            def toLong: Long = toBigInt.bigInteger.longValueExact()
-                              ^^^^^^^^^^^^^^^^^^^
-\end{lstlisting}
-
-Now that we do not depend on toLong anymore lets just remove it from out code.
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Satoshis extends CurrencyUnit {
-  override def satoshis: Satoshis = this
-
-  def toBigInt: BigInt = underlying.toBigInt
-
-  def toLong: Long = underlying.toLong
-
-  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
-}  
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Satoshis extends CurrencyUnit {
-  override def satoshis: Satoshis = this
-
-  def toBigInt: BigInt = underlying.toBigInt
-
-  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
-}  
-\end{lstlisting}
-
-And before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toLong: Long = toBigInt.bigInteger.longValueExact()
-  def toBigInt: BigInt = underlying
-
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
-  def andMask: BigInt
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require((result & andMask) == result,
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}  
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toBigInt: BigInt = underlying
-
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
-  def andMask: BigInt
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require((result & andMask) == result,
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}
-\end{lstlisting}
-
-\subsection{Getting Rid of Concrete Type}
-
-Stainless output:
-\begin{lstlisting}[style=stainless]
-[Warning ] number/NumberType.scala:9:3: Could not extract tree in class: type A = BigInt (class scala.reflect.internal.Trees$TypeDef)
-              type A = BigInt
-              ^^^^^^^^^^^^^^^
-\end{lstlisting}
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require(
-          result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"))
-    result
-  }
-}
-
-/**
-  * Represents a signed number in our number system
-  * Instances of this is [[Int64]]
-  */
-sealed abstract class SignedNumber extends Number
-
-/**
-  * Represents a int64_t in C
-  */
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: A => Int64 = Int64(_)
-}  
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: BigInt
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: BigInt => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): BigInt = {
-    require(
-         result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"))
-    result
-  }
-}
-
-/**
-  * Represents a signed number in our number system
-  * Instances of this is [[Int64]]
-  */
-sealed abstract class SignedNumber extends Number
-
-/**
-  * Represents a int64_t in C
-  */
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: BigInt => Int64 = Int64(_)
-}
-\end{lstlisting}
-
-This is easy.
-We just replace all occurrence of A with BigInt, since A is not overwritten in an implementing class.
-This is not the exact same code, because an implementing class can not override A anymore but that's fine for our verification.
-
-This was a missing feature in Stainless that was fixed on Mai 28, 2019 with \href{https://github.com/epfl-lara/stainless/pull/470}{pull request 470} on GitHub.
-Now it should work without this change.
-
-\subsection{Replacing the BigInt \&-Function With Bound Check}
-\label{sec:bound_check}
-
-Stainless output:
-\begin{lstlisting}[style=stainless]
-[ Error  ] number/NumberType.scala:33:14: Unknown call to & on result (BigInt) with arguments List(Number.this.andMask) of type List(BigInt)
-              require((result & andMask) == result,
-                      ^^^^^^^^^^^^^^^^
-\end{lstlisting}
-
-Due to the restrictions on BigInt, we can not use the \& function either.
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toBigInt: BigInt = underlying
-
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
-  def andMask: BigInt
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require((result & andMask) == result,
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}
-
-/**
-  * Represents a signed number in our number system
-  * Instances of this is [[Int64]]
-  */
-sealed abstract class SignedNumber extends Number
-
-/**
-  * Represents a int64_t in C
-  */
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: A => Int64 = Int64(_)
-  override def andMask = 0xffffffffffffffffL
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require(
-          result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"),
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}
-
-/**
-  * Represents a signed number in our number system
-  * Instances of this is [[Int64]]
-  */
-sealed abstract class SignedNumber extends Number
-
-/**
-  * Represents a int64_t in C
-  */
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: A => Int64 = Int64(_)
-}
-\end{lstlisting}
-
-This is a bounds check.
-It checks if the result of the addition is in range of the specified type, which is now the hard coded Int64.
-
-So, we can replace the \& mask with a bound check whether the result is in range of Long.MinValue and Long.MaxValue, because Int64 has the same 64-bit range as Long.
-Again the code gets a bit more static and it's not the exact same code anymore.
-
-Running Stainless produces again new errors.
-
-\subsection{Extracting Private Inner Classes}
-
-Stainless output:
-\begin{lstlisting}[style=stainless]
-[Warning ] currency/CurrencyUnits.scala:36:3: Could not extract tree in class: case private class SatoshisImpl extends Satoshis with Product with Serializable {
-[Warning ]   <caseaccessor> <paramaccessor> private[this] val underlying: addition.modified.number.Int64 = _;
-[Warning ]   <stable> <caseaccessor> <accessor> <paramaccessor> def underlying: addition.modified.number.Int64 = SatoshisImpl.this.underlying;
-[Warning ]   def <init>(underlying: addition.modified.number.Int64): addition.modified.currency.Satoshis.SatoshisImpl = {
-[Warning ]     SatoshisImpl.super.<init>();
-[Warning ]     ()
-[Warning ]   };
-[Warning ]   <synthetic> def copy(underlying: addition.modified.number.Int64 = underlying): addition.modified.currency.Satoshis.SatoshisImpl = new SatoshisImpl(underlying);
-[Warning ]   <synthetic> def copy$default$1: addition.modified.number.Int64 = SatoshisImpl.this.underlying;
-[Warning ]   override <synthetic> def productPrefix: String = "SatoshisImpl";
-[Warning ]   <synthetic> def productArity: Int = 1;
-[Warning ]   <synthetic> def productElement(x$1: Int): Any = x$1 match {
-[Warning ]     case 0 => SatoshisImpl.this.underlying
-[Warning ]     case _ => throw new IndexOutOfBoundsException(x$1.toString())
-[Warning ]   };
-[Warning ]   override <synthetic> def productIterator: Iterator[Any] = runtime.this.ScalaRunTime.typedProductIterator[Any](SatoshisImpl.this);
-[Warning ]   <synthetic> def canEqual(x$1: Any): Boolean = x$1.$isInstanceOf[addition.modified.currency.Satoshis.SatoshisImpl]();
-[Warning ]   override <synthetic> def hashCode(): Int = ScalaRunTime.this._hashCode(SatoshisImpl.this);
-[Warning ]   override <synthetic> def toString(): String = ScalaRunTime.this._toString(SatoshisImpl.this);
-[Warning ]   override <synthetic> def equals(x$1: Any): Boolean = SatoshisImpl.this.eq(x$1.asInstanceOf[Object]).||(x$1 match {
-[Warning ]   case (_: addition.modified.currency.Satoshis.SatoshisImpl) => true
-[Warning ]   case _ => false
-[Warning ] }.&&({
-[Warning ]     <synthetic> val SatoshisImpl$1: addition.modified.currency.Satoshis.SatoshisImpl = x$1.asInstanceOf[addition.modified.currency.Satoshis.SatoshisImpl];
-[Warning ]     SatoshisImpl.this.underlying.==(SatoshisImpl$1.underlying).&&(SatoshisImpl$1.canEqual(SatoshisImpl.this))
-[Warning ]   }))
-[Warning ] } (class scala.reflect.internal.Trees$ClassDef)
-              private case class SatoshisImpl(underlying: Int64) extends Satoshis
-              ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
-\end{lstlisting}
-
-And:
-\begin{lstlisting}[style=stainless]
-[Warning ] number/NumberType.scala:59:3: Could not extract tree in class: case private class Int64Impl extends Int64 with Product with Serializable {
-[Warning ]   <caseaccessor> <paramaccessor> private[this] val underlying: BigInt = _;
-[Warning ]   <stable> <caseaccessor> <accessor> <paramaccessor> def underlying: BigInt = Int64Impl.this.underlying;
-[Warning ]   def <init>(underlying: BigInt): addition.modified.number.Int64.Int64Impl = {
-[Warning ]     Int64Impl.super.<init>();
-[Warning ]     ()
-[Warning ]   };
-[Warning ]   scala.this.Predef.require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(-9223372036854775808L)), "Number was too small for a int64, got: ".+(Int64Impl.this.underlying));
-[Warning ]   scala.this.Predef.require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(9223372036854775807L)), "Number was too big for a int64, got: ".+(Int64Impl.this.underlying));
-[Warning ]   <synthetic> def copy(underlying: BigInt = underlying): addition.modified.number.Int64.Int64Impl = new Int64Impl(underlying);
-[Warning ]   <synthetic> def copy$default$1: BigInt = Int64Impl.this.underlying;
-[Warning ]   override <synthetic> def productPrefix: String = "Int64Impl";
-[Warning ]   <synthetic> def productArity: Int = 1;
-[Warning ]   <synthetic> def productElement(x$1: Int): Any = x$1 match {
-[Warning ]     case 0 => Int64Impl.this.underlying
-[Warning ]     case _ => throw new IndexOutOfBoundsException(x$1.toString())
-[Warning ]   };
-[Warning ]   override <synthetic> def productIterator: Iterator[Any] = runtime.this.ScalaRunTime.typedProductIterator[Any](Int64Impl.this);
-[Warning ]   <synthetic> def canEqual(x$1: Any): Boolean = x$1.$isInstanceOf[addition.modified.number.Int64.Int64Impl]();
-[Warning ]   override <synthetic> def hashCode(): Int = ScalaRunTime.this._hashCode(Int64Impl.this);
-[Warning ]   override <synthetic> def toString(): String = ScalaRunTime.this._toString(Int64Impl.this);
-[Warning ]   override <synthetic> def equals(x$1: Any): Boolean = Int64Impl.this.eq(x$1.asInstanceOf[Object]).||(x$1 match {
-[Warning ]   case (_: addition.modified.number.Int64.Int64Impl) => true
-[Warning ]   case _ => false
-[Warning ] }.&&({
-[Warning ]     <synthetic> val Int64Impl$1: addition.modified.number.Int64.Int64Impl = x$1.asInstanceOf[addition.modified.number.Int64.Int64Impl];
-[Warning ]     Int64Impl.this.underlying.==(Int64Impl$1.underlying).&&(Int64Impl$1.canEqual(Int64Impl.this))
-[Warning ]   }))
-[Warning ] } (class scala.reflect.internal.Trees$ClassDef)
-              private case class Int64Impl(underlying: BigInt) extends Int64 {
-              ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-\end{lstlisting}
-
-Stainless can not extract the private class inside the object.
-Bitcoin-S uses this a lot, because they separate the class from its implementation.
-
-Before:
-\begin{lstlisting}[style=scala]
-case object Satoshis extends BaseNumbers[Satoshis] {
-  val min = Satoshis(Int64.min)
-  val max = Satoshis(Int64.max)
-  val zero = Satoshis(Int64.zero)
-  val one = Satoshis(Int64.one)
-
-  def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
-
-  private case class SatoshisImpl(underlying: Int64) extends Satoshis
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-case object Satoshis extends BaseNumbers[Satoshis] {
-
-  val min = Satoshis(Int64.min)
-  val max = Satoshis(Int64.max)
-  val zero = Satoshis(Int64.zero)
-  val one = Satoshis(Int64.one)
-
-  def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
-}
-
-private case class SatoshisImpl(underlying: Int64) extends Satoshis
-\end{lstlisting}
-
-And before:
-\begin{lstlisting}[style=scala]
-case object Int64 extends BaseNumbers[Int64] {
-  private case class Int64Impl(underlying: BigInt) extends Int64 {
-    require(underlying >= -9223372036854775808L,
-      "Number was too small for a int64, got: " + underlying)
-    require(underlying <= 9223372036854775807L,
-      "Number was too big for a int64, got: " + underlying)
-  }
-
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
-
-  lazy val min = Int64(-9223372036854775808L)
-  lazy val max = Int64(9223372036854775807L)
-
-  def apply(long: Long): Int64 = Int64(BigInt(long))
-
-  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-case object Int64 extends BaseNumbers[Int64] {
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
-
-  lazy val min = Int64(-9223372036854775808L)
-  lazy val max = Int64(9223372036854775807L)
-
-  def apply(long: Long): Int64 = Int64(BigInt(long))
-
-  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
-}
-
-private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= -9223372036854775808L,
-    "Number was too small for a int64, got: " + underlying)
-  require(underlying <= 9223372036854775807L,
-    "Number was too big for a int64, got: " + underlying)
-}
-\end{lstlisting}
-
-This is easy to fix.
-We just extract the private class out of the object.
-This is not exactly the same code, because other classes in the same file could now access the private class.
-But for our property it does not change anything.
-
-Now we get some weird warnings about require.
-
-\subsection{Removing Second Parameter of Require}
-
-Stainless output:
-\begin{lstlisting}[style=stainless]
-[Warning ] number/NumberType.scala:9:3: Could not extract tree in class: type A = BigInt (class scala.reflect.internal.Trees$TypeDef)
-          type A = BigInt
-          ^^^^^^^^^^^^^^^
-[Warning ] number/NumberType.scala:71:3: Could not extract tree in class: scala.this.Predef.require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(-9223372036854775808L)), "Number was too small for a int64, got: ".+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
-          require(underlying >= -9223372036854775808L,
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-[Warning ] number/NumberType.scala:73:3: Could not extract tree in class: scala.this.Predef.require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(9223372036854775807L)), "Number was too big for a int64, got: ".+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
-          require(underlying <= 9223372036854775807L,
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-[ Error  ] checkResult$0 depends on missing dependencies: require$1.
-\end{lstlisting}
-
-Seems like Stainless does not support the second string parameter of require or at least it throws a warning about it.
-We can safely remove the string parameters from the requires, since they only serve as error messages.
-
-Before:
-\begin{lstlisting}[style=scala]
-private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= -9223372036854775808L,
-    "Number was too small for a int64, got: " + underlying)
-  require(underlying <= 9223372036854775807L,
-    "Number was too big for a int64, got: " + underlying)
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= -9223372036854775808L)
-  require(underlying <= 9223372036854775807L)
-}
-\end{lstlisting}
-
-And before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require(
-         result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"),
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require(
-          result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"))
-    result
-  }
-}
-\end{lstlisting}
-
-A new error appears.
-
-\subsection{Rewriting BigInt to Long Conversions}
-
-Stainless output:
-\begin{lstlisting}[style=stainless]
-[ Error  ] inv$4 depends on missing dependencies: long2bigInt$0.
-[ Error  ] apply$14 depends on missing dependencies: BigInt$0, apply$15.
-\end{lstlisting}
-
-This error is hard to understand, but we can see that there is a missing Long to BigInt conversion.
-So we search for all Long values in the code.
-
-Before:
-\begin{lstlisting}[style=scala]
-private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= -9223372036854775808L)
-  require(underlying <= 9223372036854775807L)
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= BigInt("-9223372036854775808"))
-  require(underlying <= BigInt("9223372036854775807"))
-}  
-\end{lstlisting}
-
-And before:
-\begin{lstlisting}[style=scala]
-case object Int64 extends BaseNumbers[Int64] {
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
-
-  lazy val min = Int64(-9223372036854775808L)
-  lazy val max = Int64(9223372036854775807L)
-
-  def apply(long: Long): Int64 = Int64(BigInt(long))
-
-  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-case object Int64 extends BaseNumbers[Int64] {
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
-
-  lazy val min = Int64(BigInt("-9223372036854775808"))
-  lazy val max = Int64(BigInt("9223372036854775807"))
-
-  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
-}
-\end{lstlisting}
-
-Looks like it can not compare a BigInt with a Long value.
-We can easily convert this Long value to a BigInt with a string literal parameter by passing these numbers as strings to the BigInt constructor.
-
-Finally, we get some output from Stainless about the verification in the code.
-
-\todo{kai special chars from stainless output what now?}
-% \begin{lstlisting}[style=stainless][Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
-% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @33:39...
-% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
-% [  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
-% [  Info  ] Cache hit: 'adt invariant' VC for apply @33:39...
-% [  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
-% [  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
-% [Warning ]  => INVALID
-% [Warning ] Found counter-example:
-% [Warning ]   bigInt: BigInt -> -9223372036854775809
-% [  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-% [  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-% [  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-% [  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
-% [Warning ]  => INVALID
-% [Warning ] Found counter-example:
-% [Warning ]   num: Number   -> Int64Impl(9223372036854775807)
-% [Warning ]   thiss: Number -> Int64Impl(1)
-% [  Info  ]   ┌───────────────────┐
-% [  Info  ] ╔═╡ stainless summary ╞══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╗
-% [  Info  ] ║ └───────────────────┘                                                                                                                      ║
-% [  Info  ] ║ +      precond. (call checkResult(thiss, underlying(thiss) + u ...)  invalid           U:smt-z3  number/NumberType.scala:17:36       0.110 ║
-% [  Info  ] ║ apply  adt invariant                                                 valid from cache            currency/CurrencyUnits.scala:33:39  1.136 ║
-% [  Info  ] ║ apply  adt invariant                                                 invalid           U:smt-z3  number/NumberType.scala:62:38       0.394 ║
-% [  Info  ] ╟┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄╢
-% [  Info  ] ║ total: 3    valid: 1    (1 from cache) invalid: 2    unknown: 0    time:   1.640                                                           ║
-% [  Info  ] ╚════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╝
-% [  Info  ] Shutting down executor service.  
-% \end{lstlisting}
-
-This shows that there is an invalid specification in checkResult and Stainless prints a counterexample for it.
-
-Let's ignore this for a moment and write the specification for the \nameref{property_2}.
-
-\subsection{Writing Specification for the Property}
-
-As specified, our verification must only support addition with zero.
-So we restrict the parameter to be zero in the precondition.
-\begin{lstlisting}[style=scala]
-  require(c.satoshis == Satoshis.zero)
-\end{lstlisting}
-
-We ensure the result is the same value as \emph{this} in the postcondition.
-\begin{lstlisting}[style=scala]
-  ensuring(res => res.satoshis == this.satoshis)
-\end{lstlisting}
-
-We can use equals (==) directly on Satoshis, because it is a case class.
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class CurrencyUnit {
-  def satoshis: Satoshis
-
-  def !=(c: CurrencyUnit): Boolean = !(this == c)
-
-  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
-  def +(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying + c.satoshis.underlying)
-  }
-
-  protected def underlying: Int64
-}
-\end{lstlisting}
-
-The addition will now look like this:
-\begin{lstlisting}[style=scala]
-sealed abstract class CurrencyUnit {
-  def satoshis: Satoshis
-
-  def !=(c: CurrencyUnit): Boolean = !(this == c)
-
-  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
-  override def +(c: CurrencyUnit): CurrencyUnit = {
-    require(c.satoshis == Satoshis.zero)
-    Satoshis(satoshis.underlying + c.satoshis.underlying)
-  } ensuring(res => res.satoshis == this.satoshis)
-
-  protected def underlying: Int64
-}
-\end{lstlisting}
-
-That's all we need to verify our addition.
-
-Now we will look into the previous error.
-
-\subsection{Propagating Require}
-
-There is another problem with Bitcoin-S.
-Bitcoin-S-Core uses require as a fail-fast method whereas Stainless needs it to verify the code.
-
-Stainless output now: \todo{kai special chars again}
-% \begin{lstlisting}[style=stainless]
-% [Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
-% [  Info  ]  - Checking cache: 'postcondition' VC for + @12:3...
-% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @34:39...
-% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
-% [  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
-% [  Info  ] Cache hit: 'adt invariant' VC for apply @34:39...
-% [  Info  ] Cache hit: 'postcondition' VC for + @12:3...
-% [  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
-% [  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
-% [Warning ]  => INVALID
-% [Warning ] Found counter-example:
-% [Warning ]   bigInt: BigInt -> -9223372036854775809
-% [  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-% [  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-% [  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-% [  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
-% [Warning ]  => INVALID
-% [Warning ] Found counter-example:
-% [Warning ]   num: Number   -> Int64Impl(9223372036854775807)
-% [Warning ]   thiss: Number -> Int64Impl(1)
-% [  Info  ]   ┌───────────────────┐
-% [  Info  ] ╔═╡ stainless summary ╞══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╗
-% [  Info  ] ║ └───────────────────┘                                                                                                                      ║
-% [  Info  ] ║ +      postcondition                                                 valid from cache            currency/CurrencyUnits.scala:12:3   1.266 ║
-% [  Info  ] ║ +      precond. (call checkResult(thiss, underlying(thiss) + u ...)  invalid           U:smt-z3  number/NumberType.scala:17:36       0.089 ║
-% [  Info  ] ║ apply  adt invariant                                                 valid from cache            currency/CurrencyUnits.scala:34:39  0.194 ║
-% [  Info  ] ║ apply  adt invariant                                                 invalid           U:smt-z3  number/NumberType.scala:62:38       0.348 ║
-% [  Info  ] ╟┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄╢
-% [  Info  ] ║ total: 4    valid: 2    (2 from cache) invalid: 2    unknown: 0    time:   1.897                                                           ║
-% [  Info  ] ╚════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╝
-% [  Info  ] Shutting down executor service.
-% \end{lstlisting}
-
-Corresponding code:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: BigInt
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: BigInt => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): BigInt = {
-    require(
-         result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"))
-    result
-  }
-}
-\end{lstlisting}
-
-But how does Stainless find a counter example ignoring the require in checkResult?
-Since Stainless is a static verification tool, it tests every possibility.
-So it can use a number bigger than the maximum Int64 and pass it to the addition.
-The require in checkResult fails.
-
-Thus, we need to add the restriction from checkResult to the addition too.
-\begin{lstlisting}[style=scala]
-override def +(num: Int64): Int64 = {
-  require(
-       num.underlying <= BigInt("9223372036854775807")
-    && num.underlying >= BigInt("-9223372036854775808")
-    && this.underlying <= BigInt("9223372036854775807")
-    && this.underlying >= BigInt("-9223372036854775808")
-  )
-  apply(checkResult(underlying + num.underlying))
-}
-\end{lstlisting}
-
-Stainless finds another counter example:
-\begin{lstlisting}[style=stainless]
-[Warning ] Found counter-example:
-[Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(1)
-[Warning ]   thiss: { x: Object | @unchecked isNumber(x) } ->
-                Int64Impl(9223372036854775807)
-\end{lstlisting}
-
-Sure, when adding one to the maximum Int64 the require does not hold anymore.
-Since we do only allow zero as a parameter, the easiest way is to restrict it to zero here too.
-
-Code after:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: BigInt
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: BigInt => Int64
-
-  def +(num: Int64): Int64 = {
-    require(
-          num == Int64.zero
-      && this.underlying <= BigInt("9223372036854775807")
-      && this.underlying >= BigInt("-9223372036854775808"))
-    apply(checkResult(underlying + num.underlying))
-  }
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): BigInt = {
-    require(
-          result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"))
-    result
-  }
-}
-\end{lstlisting}
-
-The same problem occurs in the Int64 constructor.
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: BigInt => Int64 = Int64(_)
-}
-
-case object Int64 extends BaseNumbers[Int64] {
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
-
-  lazy val min = Int64(BigInt("-9223372036854775808"))
-  lazy val max = Int64(BigInt("9223372036854775807"))
-
-  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
-}
-
-private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= BigInt("-9223372036854775808"))
-  require(underlying <= BigInt("9223372036854775807"))
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: BigInt => Int64 = bigInt => {
-    require(
-          bigInt >= BigInt("-9223372036854775808")
-      && bigInt <= BigInt("9223372036854775807"))
-
-    Int64(bigInt)
-  }
-}
-
-case object Int64 extends BaseNumbers[Int64] {
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
-
-  lazy val min = Int64(BigInt("-9223372036854775808"))
-  lazy val max = Int64(BigInt("9223372036854775807"))
-
-  def apply(bigInt: BigInt): Int64 = {
-    require(
-         bigInt >= BigInt("-9223372036854775808")
-      && bigInt <= BigInt("9223372036854775807"))
-
-    Int64Impl(bigInt)
-  }
-}
-
-private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= BigInt("-9223372036854775808"))
-  require(underlying <= BigInt("9223372036854775807"))
-}
-\end{lstlisting}
+When we run Stainless on this code (without any properties to prove), it throws the following errors: \todo{describe errors: no support for abstract types, unsupported arguments for the BigInt constructor, unsupported inheritance for objects.}
 
@@
 
+\bibliographystyle{splncs04}
+\bibliography{bibliography}
+
+\appendix
+\section{Code Transformations}
+
+\paragraph{Project Setup}
+\begin{itemize}
+\item clone our repo
+\item contains a stainless jar
+\item call the script blahblah
+\end{itemize}
+
+When we run Stainless on this code (without any properties to prove), it throws the following errors:
+
+\begin{lstlisting}[style=stainless]
+[ Error  ] currency/CurrencyUnits.scala:6:3: Stainless doesn't support abstract type members
+            type A
+            ^^^^^^
+[ Error  ] currency/CurrencyUnits.scala:26:33: Only literal arguments are allowed for BigInt.
+            def toBigInt: BigInt = BigInt(toLong)
+                                          ^^^^^^
+[ Error  ] currency/CurrencyUnits.scala:33:1: Objects cannot extend classes or implement traits, use a case object instead
+          object Satoshis extends BaseNumbers[Satoshis] {
+          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+[ Error  ] number/NumberType.scala:65:1: Objects cannot extend classes or implement traits, use a case object instead
+          object Int64 extends BaseNumbers[Int64] {
+          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+[  Info  ] Shutting down executor service.
+\end{lstlisting}
+
+So let's see how we can fix those errors.
+
+\subsection{Turning Object into Case Object}
+
+Stainless output:
+\begin{lstlisting}[style=stainless]
+[ Error  ] currency/CurrencyUnits.scala:33:1: Objects cannot extend classes or implement traits, use a case object instead
+          object Satoshis extends BaseNumbers[Satoshis] {
+          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+[ Error  ] number/NumberType.scala:65:1: Objects cannot extend classes or implement traits, use a case object instead
+          object Int64 extends BaseNumbers[Int64] {
+          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+\end{lstlisting}
+
+Code before:
+\begin{lstlisting}[style=scala]
+object Int64 extends BaseNumbers[Int64] { ... }
+object Satoshis extends BaseNumbers[Satoshis] { ... }
+\end{lstlisting}
+
+Code after:
+\begin{lstlisting}[style=scala]
+case object Int64 extends BaseNumbers[Int64] { ... }
+case object Satoshis extends BaseNumbers[Satoshis] { ... }
+\end{lstlisting}
+
+Here, we can just change the objects from object to case object.
+Stainless recommendation is to use objects for modules and case objects as algebraic data types.
+
+\todo{Ramon does this change semantics?}
+
+This is due to the internal design of Scala.
+It's possible to reason about case object but not about object.
+This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
+The \href{https://github.com/epfl-lara/stainless/issues/520}{issue number 520} on Stainless GitHub gives some thoughts, if you want to know more.
+
+\subsection{Getting Rid of Abstract Type Member}
+
+Stainless output:
+\begin{lstlisting}[style=stainless]
+[ Error  ] currency/CurrencyUnits.scala:6:3: Stainless doesn't support abstract type members
+            type A
+            ^^^^^^
+\end{lstlisting}
+
+This should be easy to rewrite by using generics instead of an abstract type, right?
+Unfortunately not.
+The problem is, CurrencyUnit uses one of its implementing classes: Satoshis.
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class CurrencyUnit {
+  type A
+
+  def satoshis: Satoshis
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
+
+  protected def underlying: A
+}
+
+sealed abstract class Satoshis extends CurrencyUnit {
+  override type A = Int64
+
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = BigInt(toLong)
+
+  def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}
+\end{lstlisting}
+
+What happens, if we typify CurrencyUnit with A, meaning to make it generic with type A?
+
+Satoshis extends CurrencyUnit with type Int64, so it would be of type CurrencyUnit[Int64].
+That's too specific, because the return type of the addition is then CurrencyUnit[Int64] not CurrencyUnit[A].
+Maybe the Bitcoin-S developers should reimplement this part and not use Satoshis directly.
+
+Since there is no easy way to fix it and the code should stay as much as possible the original, we just remove the abstract type and set it to Int64.
+This limits the verification a bit, but as we only want to verify the addition in satoshis, that's OK.
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class CurrencyUnit {
+  def satoshis: Satoshis
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
+
+  protected def underlying: Int64
+}
+
+sealed abstract class Satoshis extends CurrencyUnit {
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = BigInt(toLong)
+
+  def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}
+\end{lstlisting}
+
+\subsection{Replacing the BigInt Constructor Argument With A String Literal}
+
+Stainless output:
+\begin{lstlisting}[style=stainless]
+[ Error  ] currency/CurrencyUnits.scala:26:33: Only literal arguments are allowed for BigInt.
+            def toBigInt: BigInt = BigInt(toLong)
+                                          ^^^^^^
+\end{lstlisting}
+
+As described before, Stainless supports only a subset of Scala.
+The BigInt from the Stainless library is a bit restricted.
+One such restriction is, that BigInt does not support dynamic BigInt construction.
+Thus, the constructor parameter of BitInt must be a literal argument.
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Satoshis extends CurrencyUnit {
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = BigInt(toLong)
+
+  def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Satoshis extends CurrencyUnit {
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = underlying.toBigInt
+
+  def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}
+\end{lstlisting}
+
+This would be really hard to refactor, because Bitcoin-S tries to be as dynamic as possible so it can be used with cryptocurrencies other than bitcoins.
+Maybe it could be impossible, because they need to parse dynamic values from the bitcoin network.
+
+Luckily, we can use toBigInt on the field \texttt{underlying} directly instead of toLong.
+So, instead of converting the underlying to Long and back to BigInt we convert underlying directly to BigInt.
+
+After fixing all Stainless errors, a new error appears.
+
+\subsection{Getting Rid of Special Generics}
+
+Stainless output:
+\begin{lstlisting}[style=stainless]
+[ Error  ] number/NumberType.scala:8:30: Unknown type parameter type T
+          sealed abstract class Number[T <: Number[T]] {
+                                        ^^^^^^^^^^^^^^
+\end{lstlisting}
+
+This is a missing feature in Stainless.
+It does not support upper type boundaries on the class itself.
+To track this, \href{https://github.com/epfl-lara/stainless/issues/519}{issue 519} was created on GitHub during this work.
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number[T <: Number[T]] {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toLong: Long = toBigInt.bigInteger.longValueExact()
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => T
+
+  def +(num: T): T = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber[Int64] {
+  override def apply: A => Int64 = Int64(_)
+  override def andMask = 0xffffffffffffffffL
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toLong: Long = toBigInt.bigInteger.longValueExact()
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber extends Number
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: A => Int64 = Int64(_)
+  override def andMask = 0xffffffffffffffffL
+}
+\end{lstlisting}
+
+Despite this, in order to be able to continue, we make this a concrete type by replacing T with Int64.
+Int64,  because Satoshis uses only Int64.
+There are other number types like UInt16 but for our property we don't need them.
+
+Now, there are two new errors.
+
+\subsection{Removing toLong}
+
+Stainless output:
+\begin{lstlisting}[style=stainless]
+[ Error  ] number/NumberType.scala:14:22: Unknown call to bigInteger on Number.this.toBigInt (BigInt) with arguments List() of type List()
+            def toLong: Long = toBigInt.bigInteger.longValueExact()
+                              ^^^^^^^^^^^^^^^^^^^
+\end{lstlisting}
+
+Now that we do not depend on toLong anymore lets just remove it from out code.
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Satoshis extends CurrencyUnit {
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = underlying.toBigInt
+
+  def toLong: Long = underlying.toLong
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}  
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Satoshis extends CurrencyUnit {
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = underlying.toBigInt
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}  
+\end{lstlisting}
+
+And before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toLong: Long = toBigInt.bigInteger.longValueExact()
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}  
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+\end{lstlisting}
+
+\subsection{Getting Rid of Concrete Type}
+
+Stainless output:
+\begin{lstlisting}[style=stainless]
+[Warning ] number/NumberType.scala:9:3: Could not extract tree in class: type A = BigInt (class scala.reflect.internal.Trees$TypeDef)
+              type A = BigInt
+              ^^^^^^^^^^^^^^^
+\end{lstlisting}
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require(
+          result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"))
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber extends Number
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: A => Int64 = Int64(_)
+}  
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: BigInt
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: BigInt => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): BigInt = {
+    require(
+         result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"))
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber extends Number
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: BigInt => Int64 = Int64(_)
+}
+\end{lstlisting}
+
+This is easy.
+We just replace all occurrence of A with BigInt, since A is not overwritten in an implementing class.
+This is not the exact same code, because an implementing class can not override A anymore but that's fine for our verification.
+
+This was a missing feature in Stainless that was fixed on Mai 28, 2019 with \href{https://github.com/epfl-lara/stainless/pull/470}{pull request 470} on GitHub.
+Now it should work without this change.
+
+\subsection{Replacing the BigInt \&-Function With Bound Check}
+\label{sec:bound_check}
+
+Stainless output:
+\begin{lstlisting}[style=stainless]
+[ Error  ] number/NumberType.scala:33:14: Unknown call to & on result (BigInt) with arguments List(Number.this.andMask) of type List(BigInt)
+              require((result & andMask) == result,
+                      ^^^^^^^^^^^^^^^^
+\end{lstlisting}
+
+Due to the restrictions on BigInt, we can not use the \& function either.
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toBigInt: BigInt = underlying
+
+  /**
+    * This is used to determine the valid amount of bytes in a number
+    * for instance a UInt8 has an andMask of 0xff
+    * a UInt32 has an andMask of 0xffffffff
+    */
+  def andMask: BigInt
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber extends Number
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: A => Int64 = Int64(_)
+  override def andMask = 0xffffffffffffffffL
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require(
+          result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"),
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber extends Number
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: A => Int64 = Int64(_)
+}
+\end{lstlisting}
+
+This is a bounds check.
+It checks if the result of the addition is in range of the specified type, which is now the hard coded Int64.
+
+So, we can replace the \& mask with a bound check whether the result is in range of Long.MinValue and Long.MaxValue, because Int64 has the same 64-bit range as Long.
+Again the code gets a bit more static and it's not the exact same code anymore.
+
+Running Stainless produces again new errors.
+
+\subsection{Extracting Private Inner Classes}
+
+Stainless output:
+\begin{lstlisting}[style=stainless]
+[Warning ] currency/CurrencyUnits.scala:36:3: Could not extract tree in class: case private class SatoshisImpl extends Satoshis with Product with Serializable {
+[Warning ]   <caseaccessor> <paramaccessor> private[this] val underlying: addition.modified.number.Int64 = _;
+[Warning ]   <stable> <caseaccessor> <accessor> <paramaccessor> def underlying: addition.modified.number.Int64 = SatoshisImpl.this.underlying;
+[Warning ]   def <init>(underlying: addition.modified.number.Int64): addition.modified.currency.Satoshis.SatoshisImpl = {
+[Warning ]     SatoshisImpl.super.<init>();
+[Warning ]     ()
+[Warning ]   };
+[Warning ]   <synthetic> def copy(underlying: addition.modified.number.Int64 = underlying): addition.modified.currency.Satoshis.SatoshisImpl = new SatoshisImpl(underlying);
+[Warning ]   <synthetic> def copy$default$1: addition.modified.number.Int64 = SatoshisImpl.this.underlying;
+[Warning ]   override <synthetic> def productPrefix: String = "SatoshisImpl";
+[Warning ]   <synthetic> def productArity: Int = 1;
+[Warning ]   <synthetic> def productElement(x$1: Int): Any = x$1 match {
+[Warning ]     case 0 => SatoshisImpl.this.underlying
+[Warning ]     case _ => throw new IndexOutOfBoundsException(x$1.toString())
+[Warning ]   };
+[Warning ]   override <synthetic> def productIterator: Iterator[Any] = runtime.this.ScalaRunTime.typedProductIterator[Any](SatoshisImpl.this);
+[Warning ]   <synthetic> def canEqual(x$1: Any): Boolean = x$1.$isInstanceOf[addition.modified.currency.Satoshis.SatoshisImpl]();
+[Warning ]   override <synthetic> def hashCode(): Int = ScalaRunTime.this._hashCode(SatoshisImpl.this);
+[Warning ]   override <synthetic> def toString(): String = ScalaRunTime.this._toString(SatoshisImpl.this);
+[Warning ]   override <synthetic> def equals(x$1: Any): Boolean = SatoshisImpl.this.eq(x$1.asInstanceOf[Object]).||(x$1 match {
+[Warning ]   case (_: addition.modified.currency.Satoshis.SatoshisImpl) => true
+[Warning ]   case _ => false
+[Warning ] }.&&({
+[Warning ]     <synthetic> val SatoshisImpl$1: addition.modified.currency.Satoshis.SatoshisImpl = x$1.asInstanceOf[addition.modified.currency.Satoshis.SatoshisImpl];
+[Warning ]     SatoshisImpl.this.underlying.==(SatoshisImpl$1.underlying).&&(SatoshisImpl$1.canEqual(SatoshisImpl.this))
+[Warning ]   }))
+[Warning ] } (class scala.reflect.internal.Trees$ClassDef)
+              private case class SatoshisImpl(underlying: Int64) extends Satoshis
+              ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
+\end{lstlisting}
+
+And:
+\begin{lstlisting}[style=stainless]
+[Warning ] number/NumberType.scala:59:3: Could not extract tree in class: case private class Int64Impl extends Int64 with Product with Serializable {
+[Warning ]   <caseaccessor> <paramaccessor> private[this] val underlying: BigInt = _;
+[Warning ]   <stable> <caseaccessor> <accessor> <paramaccessor> def underlying: BigInt = Int64Impl.this.underlying;
+[Warning ]   def <init>(underlying: BigInt): addition.modified.number.Int64.Int64Impl = {
+[Warning ]     Int64Impl.super.<init>();
+[Warning ]     ()
+[Warning ]   };
+[Warning ]   scala.this.Predef.require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(-9223372036854775808L)), "Number was too small for a int64, got: ".+(Int64Impl.this.underlying));
+[Warning ]   scala.this.Predef.require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(9223372036854775807L)), "Number was too big for a int64, got: ".+(Int64Impl.this.underlying));
+[Warning ]   <synthetic> def copy(underlying: BigInt = underlying): addition.modified.number.Int64.Int64Impl = new Int64Impl(underlying);
+[Warning ]   <synthetic> def copy$default$1: BigInt = Int64Impl.this.underlying;
+[Warning ]   override <synthetic> def productPrefix: String = "Int64Impl";
+[Warning ]   <synthetic> def productArity: Int = 1;
+[Warning ]   <synthetic> def productElement(x$1: Int): Any = x$1 match {
+[Warning ]     case 0 => Int64Impl.this.underlying
+[Warning ]     case _ => throw new IndexOutOfBoundsException(x$1.toString())
+[Warning ]   };
+[Warning ]   override <synthetic> def productIterator: Iterator[Any] = runtime.this.ScalaRunTime.typedProductIterator[Any](Int64Impl.this);
+[Warning ]   <synthetic> def canEqual(x$1: Any): Boolean = x$1.$isInstanceOf[addition.modified.number.Int64.Int64Impl]();
+[Warning ]   override <synthetic> def hashCode(): Int = ScalaRunTime.this._hashCode(Int64Impl.this);
+[Warning ]   override <synthetic> def toString(): String = ScalaRunTime.this._toString(Int64Impl.this);
+[Warning ]   override <synthetic> def equals(x$1: Any): Boolean = Int64Impl.this.eq(x$1.asInstanceOf[Object]).||(x$1 match {
+[Warning ]   case (_: addition.modified.number.Int64.Int64Impl) => true
+[Warning ]   case _ => false
+[Warning ] }.&&({
+[Warning ]     <synthetic> val Int64Impl$1: addition.modified.number.Int64.Int64Impl = x$1.asInstanceOf[addition.modified.number.Int64.Int64Impl];
+[Warning ]     Int64Impl.this.underlying.==(Int64Impl$1.underlying).&&(Int64Impl$1.canEqual(Int64Impl.this))
+[Warning ]   }))
+[Warning ] } (class scala.reflect.internal.Trees$ClassDef)
+              private case class Int64Impl(underlying: BigInt) extends Int64 {
+              ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+\end{lstlisting}
+
+Stainless can not extract the private class inside the object.
+Bitcoin-S uses this a lot, because they separate the class from its implementation.
+
+Before:
+\begin{lstlisting}[style=scala]
+case object Satoshis extends BaseNumbers[Satoshis] {
+  val min = Satoshis(Int64.min)
+  val max = Satoshis(Int64.max)
+  val zero = Satoshis(Int64.zero)
+  val one = Satoshis(Int64.one)
+
+  def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
+
+  private case class SatoshisImpl(underlying: Int64) extends Satoshis
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+case object Satoshis extends BaseNumbers[Satoshis] {
+
+  val min = Satoshis(Int64.min)
+  val max = Satoshis(Int64.max)
+  val zero = Satoshis(Int64.zero)
+  val one = Satoshis(Int64.one)
+
+  def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
+}
+
+private case class SatoshisImpl(underlying: Int64) extends Satoshis
+\end{lstlisting}
+
+And before:
+\begin{lstlisting}[style=scala]
+case object Int64 extends BaseNumbers[Int64] {
+  private case class Int64Impl(underlying: BigInt) extends Int64 {
+    require(underlying >= -9223372036854775808L,
+      "Number was too small for a int64, got: " + underlying)
+    require(underlying <= 9223372036854775807L,
+      "Number was too big for a int64, got: " + underlying)
+  }
+
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(-9223372036854775808L)
+  lazy val max = Int64(9223372036854775807L)
+
+  def apply(long: Long): Int64 = Int64(BigInt(long))
+
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+case object Int64 extends BaseNumbers[Int64] {
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(-9223372036854775808L)
+  lazy val max = Int64(9223372036854775807L)
+
+  def apply(long: Long): Int64 = Int64(BigInt(long))
+
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= -9223372036854775808L,
+    "Number was too small for a int64, got: " + underlying)
+  require(underlying <= 9223372036854775807L,
+    "Number was too big for a int64, got: " + underlying)
+}
+\end{lstlisting}
+
+This is easy to fix.
+We just extract the private class out of the object.
+This is not exactly the same code, because other classes in the same file could now access the private class.
+But for our property it does not change anything.
+
+Now we get some weird warnings about require.
+
+\subsection{Removing Second Parameter of Require}
+
+Stainless output:
+\begin{lstlisting}[style=stainless]
+[Warning ] number/NumberType.scala:9:3: Could not extract tree in class: type A = BigInt (class scala.reflect.internal.Trees$TypeDef)
+          type A = BigInt
+          ^^^^^^^^^^^^^^^
+[Warning ] number/NumberType.scala:71:3: Could not extract tree in class: scala.this.Predef.require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(-9223372036854775808L)), "Number was too small for a int64, got: ".+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
+          require(underlying >= -9223372036854775808L,
+          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+[Warning ] number/NumberType.scala:73:3: Could not extract tree in class: scala.this.Predef.require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(9223372036854775807L)), "Number was too big for a int64, got: ".+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
+          require(underlying <= 9223372036854775807L,
+          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+[ Error  ] checkResult$0 depends on missing dependencies: require$1.
+\end{lstlisting}
+
+Seems like Stainless does not support the second string parameter of require or at least it throws a warning about it.
+We can safely remove the string parameters from the requires, since they only serve as error messages.
+
+Before:
+\begin{lstlisting}[style=scala]
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= -9223372036854775808L,
+    "Number was too small for a int64, got: " + underlying)
+  require(underlying <= 9223372036854775807L,
+    "Number was too big for a int64, got: " + underlying)
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= -9223372036854775808L)
+  require(underlying <= 9223372036854775807L)
+}
+\end{lstlisting}
+
+And before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require(
+         result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"),
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  type A = BigInt
+
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: A
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: A => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): A = {
+    require(
+          result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"))
+    result
+  }
+}
+\end{lstlisting}
+
+A new error appears.
+
+\subsection{Rewriting BigInt to Long Conversions}
+
+Stainless output:
+\begin{lstlisting}[style=stainless]
+[ Error  ] inv$4 depends on missing dependencies: long2bigInt$0.
+[ Error  ] apply$14 depends on missing dependencies: BigInt$0, apply$15.
+\end{lstlisting}
+
+This error is hard to understand, but we can see that there is a missing Long to BigInt conversion.
+So we search for all Long values in the code.
+
+Before:
+\begin{lstlisting}[style=scala]
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= -9223372036854775808L)
+  require(underlying <= 9223372036854775807L)
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= BigInt("-9223372036854775808"))
+  require(underlying <= BigInt("9223372036854775807"))
+}  
+\end{lstlisting}
+
+And before:
+\begin{lstlisting}[style=scala]
+case object Int64 extends BaseNumbers[Int64] {
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(-9223372036854775808L)
+  lazy val max = Int64(9223372036854775807L)
+
+  def apply(long: Long): Int64 = Int64(BigInt(long))
+
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+case object Int64 extends BaseNumbers[Int64] {
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(BigInt("-9223372036854775808"))
+  lazy val max = Int64(BigInt("9223372036854775807"))
+
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+\end{lstlisting}
+
+Looks like it can not compare a BigInt with a Long value.
+We can easily convert this Long value to a BigInt with a string literal parameter by passing these numbers as strings to the BigInt constructor.
+
+Finally, we get some output from Stainless about the verification in the code.
+
+\todo{kai special chars from stainless output what now?}
+% \begin{lstlisting}[style=stainless][Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
+% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @33:39...
+% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
+% [  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
+% [  Info  ] Cache hit: 'adt invariant' VC for apply @33:39...
+% [  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
+% [  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
+% [Warning ]  => INVALID
+% [Warning ] Found counter-example:
+% [Warning ]   bigInt: BigInt -> -9223372036854775809
+% [  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+% [  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+% [  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+% [  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
+% [Warning ]  => INVALID
+% [Warning ] Found counter-example:
+% [Warning ]   num: Number   -> Int64Impl(9223372036854775807)
+% [Warning ]   thiss: Number -> Int64Impl(1)
+% [  Info  ]   ┌───────────────────┐
+% [  Info  ] ╔═╡ stainless summary ╞══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╗
+% [  Info  ] ║ └───────────────────┘                                                                                                                      ║
+% [  Info  ] ║ +      precond. (call checkResult(thiss, underlying(thiss) + u ...)  invalid           U:smt-z3  number/NumberType.scala:17:36       0.110 ║
+% [  Info  ] ║ apply  adt invariant                                                 valid from cache            currency/CurrencyUnits.scala:33:39  1.136 ║
+% [  Info  ] ║ apply  adt invariant                                                 invalid           U:smt-z3  number/NumberType.scala:62:38       0.394 ║
+% [  Info  ] ╟┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄╢
+% [  Info  ] ║ total: 3    valid: 1    (1 from cache) invalid: 2    unknown: 0    time:   1.640                                                           ║
+% [  Info  ] ╚════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╝
+% [  Info  ] Shutting down executor service.  
+% \end{lstlisting}
+
+This shows that there is an invalid specification in checkResult and Stainless prints a counterexample for it.
+
+Let's ignore this for a moment and write the specification for the \nameref{property_2}.
+
+\subsection{Writing Specification for the Property}
+
+As specified, our verification must only support addition with zero.
+So we restrict the parameter to be zero in the precondition.
+\begin{lstlisting}[style=scala]
+  require(c.satoshis == Satoshis.zero)
+\end{lstlisting}
+
+We ensure the result is the same value as \emph{this} in the postcondition.
+\begin{lstlisting}[style=scala]
+  ensuring(res => res.satoshis == this.satoshis)
+\end{lstlisting}
+
+We can use equals (==) directly on Satoshis, because it is a case class.
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class CurrencyUnit {
+  def satoshis: Satoshis
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
+
+  protected def underlying: Int64
+}
+\end{lstlisting}
+
+The addition will now look like this:
+\begin{lstlisting}[style=scala]
+sealed abstract class CurrencyUnit {
+  def satoshis: Satoshis
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  override def +(c: CurrencyUnit): CurrencyUnit = {
+    require(c.satoshis == Satoshis.zero)
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  } ensuring(res => res.satoshis == this.satoshis)
+
+  protected def underlying: Int64
+}
+\end{lstlisting}
+
+That's all we need to verify our addition.
+
+Now we will look into the previous error.
+
+\subsection{Propagating Require}
+
+There is another problem with Bitcoin-S.
+Bitcoin-S-Core uses require as a fail-fast method whereas Stainless needs it to verify the code.
+
+Stainless output now: \todo{kai special chars again}
+% \begin{lstlisting}[style=stainless]
+% [Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
+% [  Info  ]  - Checking cache: 'postcondition' VC for + @12:3...
+% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @34:39...
+% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
+% [  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
+% [  Info  ] Cache hit: 'adt invariant' VC for apply @34:39...
+% [  Info  ] Cache hit: 'postcondition' VC for + @12:3...
+% [  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
+% [  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
+% [Warning ]  => INVALID
+% [Warning ] Found counter-example:
+% [Warning ]   bigInt: BigInt -> -9223372036854775809
+% [  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+% [  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+% [  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+% [  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
+% [Warning ]  => INVALID
+% [Warning ] Found counter-example:
+% [Warning ]   num: Number   -> Int64Impl(9223372036854775807)
+% [Warning ]   thiss: Number -> Int64Impl(1)
+% [  Info  ]   ┌───────────────────┐
+% [  Info  ] ╔═╡ stainless summary ╞══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╗
+% [  Info  ] ║ └───────────────────┘                                                                                                                      ║
+% [  Info  ] ║ +      postcondition                                                 valid from cache            currency/CurrencyUnits.scala:12:3   1.266 ║
+% [  Info  ] ║ +      precond. (call checkResult(thiss, underlying(thiss) + u ...)  invalid           U:smt-z3  number/NumberType.scala:17:36       0.089 ║
+% [  Info  ] ║ apply  adt invariant                                                 valid from cache            currency/CurrencyUnits.scala:34:39  0.194 ║
+% [  Info  ] ║ apply  adt invariant                                                 invalid           U:smt-z3  number/NumberType.scala:62:38       0.348 ║
+% [  Info  ] ╟┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄╢
+% [  Info  ] ║ total: 4    valid: 2    (2 from cache) invalid: 2    unknown: 0    time:   1.897                                                           ║
+% [  Info  ] ╚════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╝
+% [  Info  ] Shutting down executor service.
+% \end{lstlisting}
+
+Corresponding code:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: BigInt
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: BigInt => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): BigInt = {
+    require(
+         result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"))
+    result
+  }
+}
+\end{lstlisting}
+
+But how does Stainless find a counter example ignoring the require in checkResult?
+Since Stainless is a static verification tool, it tests every possibility.
+So it can use a number bigger than the maximum Int64 and pass it to the addition.
+The require in checkResult fails.
+
+Thus, we need to add the restriction from checkResult to the addition too.
+\begin{lstlisting}[style=scala]
+override def +(num: Int64): Int64 = {
+  require(
+       num.underlying <= BigInt("9223372036854775807")
+    && num.underlying >= BigInt("-9223372036854775808")
+    && this.underlying <= BigInt("9223372036854775807")
+    && this.underlying >= BigInt("-9223372036854775808")
+  )
+  apply(checkResult(underlying + num.underlying))
+}
+\end{lstlisting}
+
+Stainless finds another counter example:
+\begin{lstlisting}[style=stainless]
+[Warning ] Found counter-example:
+[Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(1)
+[Warning ]   thiss: { x: Object | @unchecked isNumber(x) } ->
+                Int64Impl(9223372036854775807)
+\end{lstlisting}
+
+Sure, when adding one to the maximum Int64 the require does not hold anymore.
+Since we do only allow zero as a parameter, the easiest way is to restrict it to zero here too.
+
+Code after:
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: BigInt
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: BigInt => Int64
+
+  def +(num: Int64): Int64 = {
+    require(
+          num == Int64.zero
+      && this.underlying <= BigInt("9223372036854775807")
+      && this.underlying >= BigInt("-9223372036854775808"))
+    apply(checkResult(underlying + num.underlying))
+  }
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): BigInt = {
+    require(
+          result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"))
+    result
+  }
+}
+\end{lstlisting}
+
+The same problem occurs in the Int64 constructor.
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: BigInt => Int64 = Int64(_)
+}
+
+case object Int64 extends BaseNumbers[Int64] {
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(BigInt("-9223372036854775808"))
+  lazy val max = Int64(BigInt("9223372036854775807"))
+
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= BigInt("-9223372036854775808"))
+  require(underlying <= BigInt("9223372036854775807"))
+}
+\end{lstlisting}
+
+After:
+\begin{lstlisting}[style=scala]
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: BigInt => Int64 = bigInt => {
+    require(
+          bigInt >= BigInt("-9223372036854775808")
+      && bigInt <= BigInt("9223372036854775807"))
+
+    Int64(bigInt)
+  }
+}
+
+case object Int64 extends BaseNumbers[Int64] {
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(BigInt("-9223372036854775808"))
+  lazy val max = Int64(BigInt("9223372036854775807"))
+
+  def apply(bigInt: BigInt): Int64 = {
+    require(
+         bigInt >= BigInt("-9223372036854775808")
+      && bigInt <= BigInt("9223372036854775807"))
+
+    Int64Impl(bigInt)
+  }
+}
+
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= BigInt("-9223372036854775808"))
+  require(underlying <= BigInt("9223372036854775807"))
+}
+\end{lstlisting}
+
 % \begin{table}
@@
 
-\bibliographystyle{splncs04}
-\bibliography{bibliography}
-
 \end{document}

### 2019-06-24 paper/paper.tex
more changes

@@
 
-\subsection{Creation of a Transaction}
+\paragraph{Creating a Transaction}
 
@@
 
-\subsection{Validation of a Transaction}
+\paragraph{Validating a Transaction}
 
@@
 
-\paragraph{Stainless integration problems.}
-
-\todo{Kai: check the text}
+\paragraph{Trying Stainless on the entire project.}
+
+\begin{itemize}
+\item result sbt: no output
+\item result jar: 
+\end{itemize}
 
@@
 
-\paragraph{Extracting the relevant code.}
-The extracted code has more than 1500 lines.  After running Stainless
-on it, it still throws a huge bunch of errors about what Stainless can
-not reason about. After fixing some of those errors, there appear new
-ones.  So this would require changing nearly everything of the
-extracted code.
-
-Let's adjust the property from the \nameref{property_1} to the
-\nameref{property_2}, because there is not enough time to fix all
-these errors and write the verification.  This is way less interesting
-to verify but gives still many insights in what needs to be done to
-verify some parts of Bitcoin-S and other source code.
+\paragraph{Putting aside the No-Inflation Property.} Naively trying
+Stainless on the entire bitcoin-s codebase results in either no output
+(with sbt) or many errors (with jar). So we extract the relevant code
+to only verify that. However, the extracted code has more than 1500
+lines and liberally uses Scala features outside of the supported
+fragment.  We tried to transform the code into the supported fragment,
+but realize that a better approach is to first verify a simpler
+property with less code involved and then turn back to the
+no-inflation property with more experience. So we turn to the
+addition-with-zero property in the next section. \todo{explain}
 
@@
 
-\subsection{Extracting the relevant Code}
+\paragraph{Extracting the relevant Code}
 
@@
 
-\subsection{Result}
-
-Finally, everything is green and correctly verified. \todo{kai special chars}
-% \begin{lstlisting}[style=scala]
-% [Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
-% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), 1))' VC for one @69:18...
-% [  Info  ] Cache hit: 'precond. (call apply(Int64(), 1))' VC for one @69:18...
-% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), bigInt))' VC for apply @52:5...
-% [  Info  ] Cache miss: 'precond. (call apply(Int64(), bigInt))' VC for apply @52:5...
-% [  Info  ]  - Now solving 'precond. (call apply(Int64(), bigInt))' VC for apply @52:5...
-% [  Info  ]  - Result for 'precond. (call apply(Int64(), bigInt))' VC for apply @52:5:
-% [  Info  ]  => VALID
-% [  Info  ]  - Checking cache: 'precond. (call +(underlying(satoshis(thiss)), underlyin ...)' VC for + @14:14...
-% [  Info  ]  - Checking cache: 'postcondition' VC for + @12:3...
-% [  Info  ] Cache hit: 'precond. (call +(underlying(satoshis(thiss)), underlyin ...)' VC for + @14:14...
-% [  Info  ] Cache miss: 'postcondition' VC for + @12:3...
-% [  Info  ]  - Now solving 'postcondition' VC for + @12:3...
-% [  Info  ]  - Result for 'postcondition' VC for + @12:3:
-% [  Info  ]  => VALID
-% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), -9223372036854775808))' VC for min @71:18...
-% [  Info  ] Cache hit: 'precond. (call apply(Int64(), -9223372036854775808))' VC for min @71:18...
-% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @34:39...
-% [  Info  ] Cache hit: 'adt invariant' VC for apply @34:39...
-% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), 9223372036854775807))' VC for max @72:18...
-% [  Info  ] Cache hit: 'precond. (call apply(Int64(), 9223372036854775807))' VC for max @72:18...
-% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @79:5...
-% [  Info  ] Cache hit: 'adt invariant' VC for apply @79:5...
-% [  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @22:11...
-% [  Info  ] Cache hit: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @22:11...
-% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), 0))' VC for zero @68:19...
-% [  Info  ] Cache hit: 'precond. (call apply(Int64(), 0))' VC for zero @68:19...
-% [  Info  ]   ┌───────────────────┐
-% [  Info  ] ╔═╡ stainless summary ╞══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╗
-% [  Info  ] ║ └───────────────────┘                                                                                                                      ║
-% [  Info  ] ║ +      postcondition                                                 valid             U:smt-z3  currency/CurrencyUnits.scala:12:3   0.643 ║
-% [  Info  ] ║ +      precond. (call +(underlying(satoshis(thiss)), underlyin ...)  valid from cache            currency/CurrencyUnits.scala:14:14  0.097 ║
-% [  Info  ] ║ +      precond. (call checkResult(thiss, underlying(thiss) + u ...)  valid from cache            number/NumberType.scala:22:11       0.015 ║
-% [  Info  ] ║ apply  adt invariant                                                 valid from cache            currency/CurrencyUnits.scala:34:39  0.005 ║
-% [  Info  ] ║ apply  precond. (call apply(Int64(), bigInt))                        valid             U:smt-z3  number/NumberType.scala:52:5        0.187 ║
-% [  Info  ] ║ apply  adt invariant                                                 valid from cache            number/NumberType.scala:79:5        0.017 ║
-% [  Info  ] ║ max    precond. (call apply(Int64(), 9223372036854775807))           valid from cache            number/NumberType.scala:72:18       0.001 ║
-% [  Info  ] ║ min    precond. (call apply(Int64(), -9223372036854775808))          valid from cache            number/NumberType.scala:71:18       0.001 ║
-% [  Info  ] ║ one    precond. (call apply(Int64(), 1))                             valid from cache            number/NumberType.scala:69:18       0.915 ║
-% [  Info  ] ║ zero   precond. (call apply(Int64(), 0))                             valid from cache            number/NumberType.scala:68:19       0.001 ║
-% [  Info  ] ╟┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄╢
-% [  Info  ] ║ total: 10   valid: 10   (8 from cache) invalid: 0    unknown: 0    time:   1.882                                                           ║
-% [  Info  ] ╚════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╝
-% [  Info  ] Shutting down executor service.
-% \end{lstlisting}
-
-% \begin{figure}
-% 	\centering
-% 		\includegraphics[width=\textwidth]{final_verify_output}
-% 	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
-% \end{figure}
+\paragraph{Transforming the Code}
+
+\todo{kai todo}
+
+\paragraph{Result}
+
+Finally, everything is green and correctly verified. 
+\begin{figure}
+	\centering
+		\includegraphics[width=\textwidth]{final_verify_output}
+	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
+\end{figure}
 
@@
 
-But is it really the original code that we verified?
-And why did we have to change that much?
-This and other questions will be answered in the next chapter.
-
-\section{Practical Challenges with Stainless}
-\label{chap:appendix_arb}
+\section{Conclusion}
+\label{chap:conclusion}
+
+Because of the limitations of the verication tool, we could only
+verify a rewritten version of the original Bitcoin-S code.  So we can
+not guarantee the correctness of the addition of Satoshis with zero in
+Bitcoin-S.  Not all changes we made were as trivial as the replacement
+of objects with case objects.  For these non-trivial changes, as seen
+for example the bound check in section \ref{sec:bound_check}, we
+cannot say whether they are equivalent to the original implementation
+or not.
+
+So code should be written specically with formal verication in mind,
+in order to successfully verify it.  Otherwise, it needs a lot of
+changes in the software because verification is mathematical and the
+current software is written mostly in object-oriented style.  Software
+written in the functional paradigm would be much easier to reason
+about.
+
+Thus, either Stainless must find ways to translate more of built-in
+object-oriented patterns of Scala to their verification tool or
+developers must invest more in functional programming.
+
+Also, we found that trying to verify code reveals bugs as shown in
+section \ref{sec:bugfix}.  Finally, our work led to some feedback to
+the Stainless developers to improve the tool.
+
+\bibliographystyle{splncs04}
+\bibliography{bibliography}
+
+\appendix
+\section{Project Setup}
 
@@
 
-\subsection{sbt vs JAR}
+\todo{nur über stainless-sbt-plugin reden damit man weiss wir haben es versucht}
+
+\subsection{stainless-sbt-plugin vs JAR}
 
@@
 However, currently there are some drawbacks.  For example the sbt
-plugin does not always report errors.  We will see more about that
-shortly.
+plugin does not always report errors.
+
+We use the jar for everything.
 
@@
 
-\section{Conclusion}
-\label{chap:conclusion}
-
-Because of the limitations of the verication tool, we could only
-verify a rewritten version of the original Bitcoin-S code.  So we can
-not guarantee the correctness of the addition of Satoshis with zero in
-Bitcoin-S.  Not all changes we made were as trivial as the replacement
-of objects with case objects.  For these non-trivial changes, as seen
-for example the bound check in section \ref{sec:bound_check}, we
-cannot say whether they are equivalent to the original implementation
-or not.
-
-So code should be written specically with formal verication in mind,
-in order to successfully verify it.  Otherwise, it needs a lot of
-changes in the software because verification is mathematical and the
-current software is written mostly in object-oriented style.  Software
-written in the functional paradigm would be much easier to reason
-about.
-
-Thus, either Stainless must find ways to translate more of built-in
-object-oriented patterns of Scala to their verification tool or
-developers must invest more in functional programming.
-
-Also, we found that trying to verify code reveals bugs as shown in
-section \ref{sec:bugfix}.  Finally, our work led to some feedback to
-the Stainless developers to improve the tool.
-
-\bibliographystyle{splncs04}
-\bibliography{bibliography}
-
-\appendix
-\section{Code Transformations}
-
-\paragraph{Project Setup}
 \begin{itemize}
 \item clone our repo
+\item jar vs sbt, you can use both
 \item contains a stainless jar
@@
 \end{itemize}
+
+\section{Code Transformations}
+
+These are the code transformations needed to Verify the Addition-with-Zero Property.
 

### 2019-06-24 paper/paper.tex
more changes
added comments about semantic

@@
 
-\todo{ramon: completely understand and explain the code}
+\todo{kai: do you need more explanation?}
 
@@
 	\centering
-		\includegraphics[width=\textwidth]{final_verify_output}
+		\includegraphics[width=\textwidth]{result_output}
 	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
@@
 
-\todo{Ramon does this change semantics?}
+\todo{Kai this does not change the semantic}
 
@@
 
+\todo{kai this does not change the semantic}
+
 Before:
@@
 
+\todo{kai this does not change semantic}
+
 Before:
@@
 
+\todo{kai this does not change semantic}
+
 This is a missing feature in Stainless.
@@
 
+\todo{kai This does not change semantic}
+
 Now that we do not depend on toLong anymore lets just remove it from out code.
@@
 \end{lstlisting}
+
+\todo{kai here we do not know if it changes semantic}
 
@@
 So we search for all Long values in the code.
+
+\todo{kai this does not change semantic}
 

### 2019-06-24 paper/paper.tex
update email
more intro

@@
 
+%\usepackage{newtxtext}
+%\usepackage{newtxmath}
+\usepackage[zerostyle=b,scaled=.9]{newtxtt}
+
 % nicer bold
@@
   language=scala,
-%  basicstyle=\footnotesize\ttfamily,
-  basicstyle=\scriptsize\ttfamily, % kai: temporary 
+  basicstyle=\footnotesize\ttfamily,
+%  basicstyle=\scriptsize\ttfamily, % kai: temporary 
   breaklines=true,
@@
 \institute{Bern University of Applied Sciences, CH-2501 Biel, Switzerland
-\email{\{ramon.boss,kai.bruennler,anna.doukmak\}@bfh.ch}}
+\email{ramon.boss@outlook.com, kai.bruennler@bfh.ch,\\anna.doukmak@gmail.com}}
 
@@
 
-The following example from the Stainless documentation
+The example in Figure~\ref{fig:factorial} from the Stainless documentation
 \cite{Stainless:documentation} demonstrates this. Notice how a
@@
 
+\begin{figure}
 \begin{lstlisting}[style=scala]
@@
 \end{lstlisting}
-
+	\caption{Factorial program with specification}
+	\label{fig:factorial}
+\end{figure}
 Our program happens not to satisfy the specification. An overflow in
@@
 
-\paragraph{The Pure Scala Fragment.} Stainless supports only a subset
-of Scala. They call this Pure Scala.  You can find the specification of Pure
-Scala on the Stainless documentation website
+\paragraph{The Pure Scala Fragment.} The Scala fragment supported by
+Stainless is described in the Stainless documentation
 \cite{Stainless:documentation} in the section
@@
   Scala}.
-\todo{Kai: check the text
-
-It supports Algebraic Data Type definitions in the form of 
-Abstract Class and Case Class/Case Object,
-inheritance of an Abstract Class by a Case Class/Case Object,
-Object for grouping classes and functions,  
-short-circuit interpretation of boolean expressions, 
-Generics with invariant type parameters and simple hierarchy 
-(subclasses define the same type parameters in the same order),
-default values of function parameters,
-Pattern matching, 
-Values,
-anonymous classes and more.
-
-As an extension to Pure Scala Stainless supports some imperative features.
-It supports 
-declaring a variable in a local scope of a function, 
-declaring a variable in trait without assigning a default value,
-while loops,
-case classes with \textit{var} arguments.
-
-In particular Stainless does not support
-Class definition,
-inheritance by Object,
-abstract type members,
-variance annotations, upper and lower type bounds in generics,
-variables in classes and objects,
-private inner classes,
-messages as a second argument in the Require function and more.}
-
-In addition, Stainless has its own library with annotations,
-reimplementation of some core data types, collections and input-output
-functions and more. \todo{Kai: check the text.
-
-This allows Stainless to map correctly operations 
-to mathematical function inside of SMT solver.
-Some elements of the library} are described on the website in the section
-\href{https://epfl-lara.github.io/stainless/library.html}{Stainless
-  Library}.  There are more details about the library in the source
-code of Stainless on GitHub \cite{Stainless:github} in the folder
-\href{https://github.com/epfl-lara/stainless/tree/master/frontends/library/stainless}{\textit{frontends/library/stainless}}.
+
+It comprises algebraic data types in the form of abstract classes,
+case classes and case objects, objects for grouping classes and
+functions, boolean expressions with short-circuit interpretation,
+generics with invariant type parameters, default values of function
+parameters, pattern matching, local and anonymous classes and more.
+
+In addition to Pure Scala Stainless also supports some imperative
+features, such as using a (mutable) variable in a local scope of a
+function and while loops. They turn out not to be relevant for the
+current work.
+
+What will turn out to be more relevant are the following Scala
+features which Stainless does not support, such as: (concrete) class
+definitions, inheritance by objects, abstract type members, variance
+annotations, upper and lower type bounds in generics and private inner
+classes.
+
+In addition, Stainless has its own library of some core data types and
+functions which need to be mapped correctly to functions inside of the
+SMT solver that Stainless ultimately relies on. Those data types in
+general do not have all the methods of the Scala data types. We will see an example of a method 
+of the BigInt type that Scala provides but Stainless does not.
+\todo{question viktor: where is BigInt??}
 
@@
 following constructor:
-\begin{lstlisting}[style=scala, basicstyle=\scriptsize\tt]
+\begin{lstlisting}[style=scala]
   BitcoinTxBuilder(
-    destinations: Seq[TransactionOutput], // where to send the money
-    utxos: BitcoinTxBuilder.UTXOMap,      // unspent transaction outputs
-    feeRate: FeeUnit,                     // fee rate per byte
-    changeSPK: ScriptPubKey,              // where to send the change
-    network: BitcoinNetwork               // bitcoin network information
+    destinations: Seq[TransactionOutput], 
+    utxos: BitcoinTxBuilder.UTXOMap,      
+    feeRate: FeeUnit,                     
+    changeSPK: ScriptPubKey,              
+    network: BitcoinNetwork               
   ): Future[BitcoinTxBuilder]

### 2019-06-25 paper/paper.tex
more intro
remove irrelevant stainless messages

@@
 
+\todo{anna: stainless unterstützt anscheinend type bounds prinzipiell
+  (nur nicht selbstreferenzierende)? siehe B.4, stainless issue 519}
+
 In addition, Stainless has its own library of some core data types and
@@
 SMT solver that Stainless ultimately relies on. Those data types in
-general do not have all the methods of the Scala data types. We will see an example of a method 
-of the BigInt type that Scala provides but Stainless does not.
-\todo{question viktor: where is BigInt??}
+general do not have all the methods of the Scala data types. We will
+see an example of a method of the BigInt type that Scala provides but
+Stainless does not.  \todo{question viktor: where is BigInt??}
 
@@
 \end{lstlisting}
-
-\todo{kai: do you need more explanation?}
 
@@
 [Warning ] currency/CurrencyUnits.scala:36:3: Could not extract tree in class: case private class SatoshisImpl extends Satoshis with Product with Serializable {
-[Warning ]   <caseaccessor> <paramaccessor> private[this] val underlying: addition.modified.number.Int64 = _;
-[Warning ]   <stable> <caseaccessor> <accessor> <paramaccessor> def underlying: addition.modified.number.Int64 = SatoshisImpl.this.underlying;
-[Warning ]   def <init>(underlying: addition.modified.number.Int64): addition.modified.currency.Satoshis.SatoshisImpl = {
-[Warning ]     SatoshisImpl.super.<init>();
-[Warning ]     ()
-[Warning ]   };
-[Warning ]   <synthetic> def copy(underlying: addition.modified.number.Int64 = underlying): addition.modified.currency.Satoshis.SatoshisImpl = new SatoshisImpl(underlying);
-[Warning ]   <synthetic> def copy$default$1: addition.modified.number.Int64 = SatoshisImpl.this.underlying;
-[Warning ]   override <synthetic> def productPrefix: String = "SatoshisImpl";
-[Warning ]   <synthetic> def productArity: Int = 1;
-[Warning ]   <synthetic> def productElement(x$1: Int): Any = x$1 match {
-[Warning ]     case 0 => SatoshisImpl.this.underlying
-[Warning ]     case _ => throw new IndexOutOfBoundsException(x$1.toString())
-[Warning ]   };
-[Warning ]   override <synthetic> def productIterator: Iterator[Any] = runtime.this.ScalaRunTime.typedProductIterator[Any](SatoshisImpl.this);
-[Warning ]   <synthetic> def canEqual(x$1: Any): Boolean = x$1.$isInstanceOf[addition.modified.currency.Satoshis.SatoshisImpl]();
-[Warning ]   override <synthetic> def hashCode(): Int = ScalaRunTime.this._hashCode(SatoshisImpl.this);
-[Warning ]   override <synthetic> def toString(): String = ScalaRunTime.this._toString(SatoshisImpl.this);
-[Warning ]   override <synthetic> def equals(x$1: Any): Boolean = SatoshisImpl.this.eq(x$1.asInstanceOf[Object]).||(x$1 match {
-[Warning ]   case (_: addition.modified.currency.Satoshis.SatoshisImpl) => true
-[Warning ]   case _ => false
-[Warning ] }.&&({
-[Warning ]     <synthetic> val SatoshisImpl$1: addition.modified.currency.Satoshis.SatoshisImpl = x$1.asInstanceOf[addition.modified.currency.Satoshis.SatoshisImpl];
-[Warning ]     SatoshisImpl.this.underlying.==(SatoshisImpl$1.underlying).&&(SatoshisImpl$1.canEqual(SatoshisImpl.this))
-[Warning ]   }))
+...
 [Warning ] } (class scala.reflect.internal.Trees$ClassDef)
@@
 [Warning ] number/NumberType.scala:59:3: Could not extract tree in class: case private class Int64Impl extends Int64 with Product with Serializable {
-[Warning ]   <caseaccessor> <paramaccessor> private[this] val underlying: BigInt = _;
-[Warning ]   <stable> <caseaccessor> <accessor> <paramaccessor> def underlying: BigInt = Int64Impl.this.underlying;
-[Warning ]   def <init>(underlying: BigInt): addition.modified.number.Int64.Int64Impl = {
-[Warning ]     Int64Impl.super.<init>();
-[Warning ]     ()
-[Warning ]   };
-[Warning ]   scala.this.Predef.require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(-9223372036854775808L)), "Number was too small for a int64, got: ".+(Int64Impl.this.underlying));
-[Warning ]   scala.this.Predef.require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(9223372036854775807L)), "Number was too big for a int64, got: ".+(Int64Impl.this.underlying));
-[Warning ]   <synthetic> def copy(underlying: BigInt = underlying): addition.modified.number.Int64.Int64Impl = new Int64Impl(underlying);
-[Warning ]   <synthetic> def copy$default$1: BigInt = Int64Impl.this.underlying;
-[Warning ]   override <synthetic> def productPrefix: String = "Int64Impl";
-[Warning ]   <synthetic> def productArity: Int = 1;
-[Warning ]   <synthetic> def productElement(x$1: Int): Any = x$1 match {
-[Warning ]     case 0 => Int64Impl.this.underlying
-[Warning ]     case _ => throw new IndexOutOfBoundsException(x$1.toString())
-[Warning ]   };
-[Warning ]   override <synthetic> def productIterator: Iterator[Any] = runtime.this.ScalaRunTime.typedProductIterator[Any](Int64Impl.this);
-[Warning ]   <synthetic> def canEqual(x$1: Any): Boolean = x$1.$isInstanceOf[addition.modified.number.Int64.Int64Impl]();
-[Warning ]   override <synthetic> def hashCode(): Int = ScalaRunTime.this._hashCode(Int64Impl.this);
-[Warning ]   override <synthetic> def toString(): String = ScalaRunTime.this._toString(Int64Impl.this);
-[Warning ]   override <synthetic> def equals(x$1: Any): Boolean = Int64Impl.this.eq(x$1.asInstanceOf[Object]).||(x$1 match {
-[Warning ]   case (_: addition.modified.number.Int64.Int64Impl) => true
-[Warning ]   case _ => false
-[Warning ] }.&&({
-[Warning ]     <synthetic> val Int64Impl$1: addition.modified.number.Int64.Int64Impl = x$1.asInstanceOf[addition.modified.number.Int64.Int64Impl];
-[Warning ]     Int64Impl.this.underlying.==(Int64Impl$1.underlying).&&(Int64Impl$1.canEqual(Int64Impl.this))
-[Warning ]   }))
+...
 [Warning ] } (class scala.reflect.internal.Trees$ClassDef)
@@
 
-\todo{kai special chars from stainless output what now?}
-% \begin{lstlisting}[style=stainless][Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
-% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @33:39...
-% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
-% [  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
-% [  Info  ] Cache hit: 'adt invariant' VC for apply @33:39...
-% [  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
-% [  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
-% [Warning ]  => INVALID
-% [Warning ] Found counter-example:
-% [Warning ]   bigInt: BigInt -> -9223372036854775809
-% [  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-% [  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-% [  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-% [  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
-% [Warning ]  => INVALID
-% [Warning ] Found counter-example:
-% [Warning ]   num: Number   -> Int64Impl(9223372036854775807)
-% [Warning ]   thiss: Number -> Int64Impl(1)
-% [  Info  ]   ┌───────────────────┐
-% [  Info  ] ╔═╡ stainless summary ╞══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╗
-% [  Info  ] ║ └───────────────────┘                                                                                                                      ║
-% [  Info  ] ║ +      precond. (call checkResult(thiss, underlying(thiss) + u ...)  invalid           U:smt-z3  number/NumberType.scala:17:36       0.110 ║
-% [  Info  ] ║ apply  adt invariant                                                 valid from cache            currency/CurrencyUnits.scala:33:39  1.136 ║
-% [  Info  ] ║ apply  adt invariant                                                 invalid           U:smt-z3  number/NumberType.scala:62:38       0.394 ║
-% [  Info  ] ╟┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄╢
-% [  Info  ] ║ total: 3    valid: 1    (1 from cache) invalid: 2    unknown: 0    time:   1.640                                                           ║
-% [  Info  ] ╚════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╝
-% [  Info  ] Shutting down executor service.  
-% \end{lstlisting}
+\begin{lstlisting}[style=stainless]
+[  Info  ]  - Checking cache: 'adt invariant' VC for apply @33:39...
+[  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
+[  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
+[  Info  ] Cache hit: 'adt invariant' VC for apply @33:39...
+[  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
+[  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
+[Warning ]  => INVALID
+[Warning ] Found counter-example:
+[Warning ]   bigInt: BigInt -> -9223372036854775809
+[  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+[  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+[  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+[  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
+[Warning ]  => INVALID
+[Warning ] Found counter-example:
+[Warning ]   num: Number   -> Int64Impl(9223372036854775807)
+[Warning ]   thiss: Number -> Int64Impl(1)
+\end{lstlisting}
 
@@
 
-Stainless output now: \todo{kai special chars again}
-% \begin{lstlisting}[style=stainless]
-% [Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
-% [  Info  ]  - Checking cache: 'postcondition' VC for + @12:3...
-% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @34:39...
-% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
-% [  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
-% [  Info  ] Cache hit: 'adt invariant' VC for apply @34:39...
-% [  Info  ] Cache hit: 'postcondition' VC for + @12:3...
-% [  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
-% [  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
-% [Warning ]  => INVALID
-% [Warning ] Found counter-example:
-% [Warning ]   bigInt: BigInt -> -9223372036854775809
-% [  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-% [  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-% [  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-% [  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
-% [Warning ]  => INVALID
-% [Warning ] Found counter-example:
-% [Warning ]   num: Number   -> Int64Impl(9223372036854775807)
-% [Warning ]   thiss: Number -> Int64Impl(1)
-% [  Info  ]   ┌───────────────────┐
-% [  Info  ] ╔═╡ stainless summary ╞══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╗
-% [  Info  ] ║ └───────────────────┘                                                                                                                      ║
-% [  Info  ] ║ +      postcondition                                                 valid from cache            currency/CurrencyUnits.scala:12:3   1.266 ║
-% [  Info  ] ║ +      precond. (call checkResult(thiss, underlying(thiss) + u ...)  invalid           U:smt-z3  number/NumberType.scala:17:36       0.089 ║
-% [  Info  ] ║ apply  adt invariant                                                 valid from cache            currency/CurrencyUnits.scala:34:39  0.194 ║
-% [  Info  ] ║ apply  adt invariant                                                 invalid           U:smt-z3  number/NumberType.scala:62:38       0.348 ║
-% [  Info  ] ╟┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄╢
-% [  Info  ] ║ total: 4    valid: 2    (2 from cache) invalid: 2    unknown: 0    time:   1.897                                                           ║
-% [  Info  ] ╚════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╝
-% [  Info  ] Shutting down executor service.
-% \end{lstlisting}
+Stainless output now:
+\begin{lstlisting}[style=stainless]
+[  Info  ]  - Checking cache: 'postcondition' VC for + @12:3...
+[  Info  ]  - Checking cache: 'adt invariant' VC for apply @34:39...
+[  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
+[  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
+[  Info  ] Cache hit: 'adt invariant' VC for apply @34:39...
+[  Info  ] Cache hit: 'postcondition' VC for + @12:3...
+[  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
+[  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
+[Warning ]  => INVALID
+[Warning ] Found counter-example:
+[Warning ]   bigInt: BigInt -> -9223372036854775809
+[  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+[  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+[  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
+[  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
+[Warning ]  => INVALID
+[Warning ] Found counter-example:
+[Warning ]   num: Number   -> Int64Impl(9223372036854775807)
+[Warning ]   thiss: Number -> Int64Impl(1)
+\end{lstlisting}
 
@@
 
-% \begin{table}
-% \caption{Table captions should be placed above the
-% tables.}\label{tab1}
-% \begin{tabular}{|l|l|l|}
-% \hline
-% Heading level &  Example & Font size and style\\
-% \hline
-% Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-% 1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-% 2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-% 3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-% 4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-% \hline
-% \end{tabular}
-% \end{table}
-
-% \noindent Displayed equations are centered and set on a separate
-% line.
-% \begin{equation}
-% x + y = z
-% \end{equation}
-% Please try to avoid rasterized images for line-art diagrams and
-% schemas. Whenever possible, use vector graphics instead (see
-% Fig.~\ref{fig1}).
-
-% \begin{figure}
-% \includegraphics[width=\textwidth]{fig1.eps}
-% \caption{A figure caption is always placed below the illustration.
-% Please note that short captions are centered, while long ones are
-% justified by the macro package automatically.} \label{fig1}
-% \end{figure}
-
-% \begin{theorem}
-% This is a sample theorem. The run-in heading is set in bold, while
-% the following text appears in italics. Definitions, lemmas,
-% propositions, and corollaries are styled the same way.
-% \end{theorem}
-% %
-% % the environments 'definition', 'lemma', 'proposition', 'corollary',
-% % 'remark', and 'example' are defined in the LLNCS documentclass as well.
-% %
-% \begin{proof}
-% Proofs, examples, and remarks have the initial word in italics,
-% while the following text appears in normal font.
-% \end{proof}
-% For citations of references, we prefer the use of square brackets
-% and consecutive numbers. Citations using labels or the author/year
-% convention are also acceptable. The following bibliography provides
-% a sample reference list with entries for journal
-% articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
-% book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-% and a homepage~\cite{ref_url1}. Multiple citations are grouped
-% \cite{ref_article1,ref_lncs1,ref_book1},
-% \cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-%
-
 \end{document}

### 2019-06-26 paper/paper.tex
change titles

@@
 
-\subsection{Turning Object into Case Object}
+\subsection{An Inheriting Object}
 
@@
 
-\subsection{Getting Rid of Abstract Type Member}
+\subsection{Abstract Type Member}
 
@@
 
-\subsection{Replacing the BigInt Constructor Argument With A String Literal}
+\subsection{Non-Literal BigInt Constructor Argument}
 
@@
 
-\subsection{Getting Rid of Special Generics}
+\subsection{Self-Reference in Type Parameter Bound}
 
@@
 
-\subsection{Removing toLong}
-
+\subsection{Missing Member \texttt{bigInteger} in BigInt}
+\todo{Ramon is .bigInteger missing in Stainless?}
 Stainless output:
@@
 
-\subsection{Getting Rid of Concrete Type}
+\subsection{Type Member}
 
@@
 
-\subsection{Replacing the BigInt \&-Function With Bound Check}
+\subsection{Missing Bitwise-And Method on BigInt}
 \label{sec:bound_check}
@@
 
-\subsection{Extracting Private Inner Classes}
+\subsection{Inner Class in Case Object}
 
@@
 
-\subsection{Removing Second Parameter of Require}
+\subsection{Message Parameter in Require}
 
@@
 
-\subsection{Rewriting BigInt to Long Conversions}
+\subsection{Missing Implicit Long to BigInt Conversion}
 
@@
 Let's ignore this for a moment and write the specification for the \nameref{property_2}.
+
+\subsection{Missing BigInt Constructor with Long Argument}
+\todo{Split code from above}
 

### 2019-06-26 code/addition/src/main/scala/addition/modified/currency/CurrencyUnits.scala
Updated postcondition

@@ -1,6 +1,7 @@
 package addition.modified.currency
 
 import addition.modified.number.{BaseNumbers, Int64}
+import stainless.lang._
 
 sealed abstract class CurrencyUnit {
   def satoshis: Satoshis
@@ -10,12 +11,12 @@ sealed abstract class CurrencyUnit {
   def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
 
   def +(c: CurrencyUnit): CurrencyUnit = {
-    require(c.satoshis == Satoshis.zero)
     Satoshis(satoshis.underlying + c.satoshis.underlying)
-  } ensuring(res => res.satoshis == this.satoshis)
+  } ensuring(res =>
+    (c.satoshis == Satoshis.zero) ==>
+      (res.satoshis == this.satoshis))
 
-  protected def underlying: Int64
-}
+  protected def underlying: Int64}
 
 sealed abstract class Satoshis extends CurrencyUnit {
   override def satoshis: Satoshis = this

### 2019-06-26 code/addition/src/main/scala/addition/modified/number/NumberType.scala

@@ -14,13 +14,7 @@ sealed abstract class Number {
   /** Factory function to create the underlying T, for instance a UInt32 */
   def apply: BigInt => Int64
 
-  def +(num: Int64): Int64 = {
-    require(
-         num == Int64.zero
-      && this.underlying <= BigInt("9223372036854775807")
-      && this.underlying >= BigInt("-9223372036854775808"))
-    apply(checkResult(underlying + num.underlying))
-  }
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
 
   /**
     * Checks if the given result is within the range
@@ -44,13 +38,7 @@ sealed abstract class SignedNumber extends Number
   * Represents a int64_t in C
   */
 sealed abstract class Int64 extends SignedNumber {
-  override def apply: BigInt => Int64 = bigInt => {
-    require(
-         bigInt >= BigInt("-9223372036854775808")
-      && bigInt <= BigInt("9223372036854775807"))
-
-    Int64(bigInt)
-  }
+  override def apply: BigInt => Int64 = Int64(_)
 }
 
 /**
@@ -71,13 +59,7 @@ case object Int64 extends BaseNumbers[Int64] {
   lazy val min = Int64(BigInt("-9223372036854775808"))
   lazy val max = Int64(BigInt("9223372036854775807"))
 
-  def apply(bigInt: BigInt): Int64 = {
-    require(
-         bigInt >= BigInt("-9223372036854775808")
-      && bigInt <= BigInt("9223372036854775807"))
-
-    Int64Impl(bigInt)
-  }
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
 }
 
 private case class Int64Impl(underlying: BigInt) extends Int64 {

### 2019-06-26 paper/paper.tex
more intro

@@
 \documentclass[runningheads]{llncs}
+
+% to have bookmarks in pdf
+% remove for final version
+\usepackage[open,openlevel=5]{bookmark}
+\setcounter{tocdepth}{5}
 
@@
 definitions, inheritance by objects, abstract type members, variance
-annotations, upper and lower type bounds in generics and private inner
-classes.
-
-\todo{anna: stainless unterstützt anscheinend type bounds prinzipiell
-  (nur nicht selbstreferenzierende)? siehe B.4, stainless issue 519}
+annotations and private inner classes.
 
@@
 SMT solver that Stainless ultimately relies on. Those data types in
-general do not have all the methods of the Scala data types. We will
-see an example of a method of the BigInt type that Scala provides but
-Stainless does not.  \todo{question viktor: where is BigInt??}
-
-\section{Trying to Verify the No-Inflation Property}
-
-Thus, we try to verify the property, that a non-coinbase transaction
-cannot generate new coins.  Let's name it the \nameref{property_1}.
-
-As we can see later in chapter \ref{chap:verify_check}, we must
-rewrite a huge part of the code implementing this property.  This
-reimplementation into Pure Scala needs a lot of time.  So, we adjust
-the plan and verify another functionality of Bitcoin-S.  Nevertheless,
-during the analysis of the \nameref{property_1} we look at a bug in
-Bitcoin-S found during this work and see the code changes for the
-bugfix in section \ref{sec:bugfix}.
-
-This chapter describes the part of Bitcoin-S needed to verify the
-\nameref{property_1} described before.  We are going to create a
-transaction and show the relevant parts of the method
-\texttt{checkTransaction}, where transactions are checked against some
-properties.  Then, we will see the bug in Bitcoin-S found during this
-work and its fix.  In the end we see why the \nameref{property_1}
-needed to be changed to the \nameref{property_2}.
-
-\paragraph{Creating a Transaction}
-
-Some code in this section is copied or adapted from the Bitcoin-S-Core
+general do not have all the methods of the Scala data types. For
+example, the BigInt type in Scala has a methods for bitwise operations
+while the BigInt type in Stainless does not.
+
+\paragraph{Outline.} In the next section we try to verify the property
+that a regular (non-coinbase) transaction cannot generate new
+coins. We call it the \emph{no-inflation property}. Trying to verify
+it, we uncover and fix a bug in the bitcoin-s library. We find that
+there is too much code involved that lies outside of the supported
+fragment to currently make this verification feasible. Instead, in
+Section~3 we turn to a simpler property to verify. The simplest
+possible property we could think of was the fact that adding zero
+satoshis to a given amount of satoshis yields the given amount of
+satoshis. We call it the \emph{addition-with-zero property}. Here as
+well we see that a significant part of the code lies outside of the
+supported fragment. However, after a series of equivalent
+transformations we arrive at code that we successfully
+verify. \todo{conclusions: what now?}
+
+\section{The No-Inflation Property}
+
+\paragraph{Creating a Transaction} Some code in this section is copied or adapted from the Bitcoin-S-Core
 transaction builder example \cite{BitcoinSCore:txbuilderexample}.
@@
 For the sake of convenience we create only one.
+
 \begin{lstlisting}[style=scala]
@@
 
-\section{Towards Verifying the Addition-with-Zero Property}
+\section{The Addition-with-Zero Property}
 

### 2019-06-26 paper/paper.tex
work on section 2

@@
 %\usepackage{newtxmath}
-\usepackage[zerostyle=b,scaled=.9]{newtxtt}
+\usepackage[zerostyle=b,scaled=.88]{newtxtt}
 
@@
 
-\paragraph{Outline.} In the next section we try to verify the property
-that a regular (non-coinbase) transaction cannot generate new
-coins. We call it the \emph{no-inflation property}. Trying to verify
-it, we uncover and fix a bug in the bitcoin-s library. We find that
-there is too much code involved that lies outside of the supported
-fragment to currently make this verification feasible. Instead, in
-Section~3 we turn to a simpler property to verify. The simplest
-possible property we could think of was the fact that adding zero
-satoshis to a given amount of satoshis yields the given amount of
-satoshis. We call it the \emph{addition-with-zero property}. Here as
-well we see that a significant part of the code lies outside of the
-supported fragment. However, after a series of equivalent
-transformations we arrive at code that we successfully
-verify. \todo{conclusions: what now?}
+\paragraph{Outline and Properties to Verify.} In the next section we
+try to verify the property that a regular (non-coinbase) transaction
+cannot generate new coins. We call it the \emph{no-inflation
+  property}. Trying to verify it, we uncover and fix a bug in the
+bitcoin-s library. We then find that there is too much code involved
+that lies outside of the supported fragment to currently make this
+verification feasible. Instead, we turn to a simpler property to
+verify. The simplest possible property we can think of is the fact
+that adding zero satoshis to a given amount of satoshis yields the
+given amount of satoshis. We call it the \emph{addition-with-zero
+  property} and we try to verify it in Section~3. Here as well we see
+that a significant part of the code lies outside of the supported
+fragment. However, after a series of equivalent transformations we
+arrive at code that we successfully verify.
 
@@
 
-\paragraph{Creating a Transaction} Some code in this section is copied or adapted from the Bitcoin-S-Core
-transaction builder example \cite{BitcoinSCore:txbuilderexample}.
-Bitcoin-S-Core has a bitcoin transaction builder class with the
-following constructor:
-\begin{lstlisting}[style=scala]
-  BitcoinTxBuilder(
-    destinations: Seq[TransactionOutput], 
-    utxos: BitcoinTxBuilder.UTXOMap,      
-    feeRate: FeeUnit,                     
-    changeSPK: ScriptPubKey,              
-    network: BitcoinNetwork               
-  ): Future[BitcoinTxBuilder]
-\end{lstlisting}
-
-The return type Future does not make sense here, since the
-implementation calls either Future.successful or Future.fromTry which
-returns an already resolved Future.  This might be for future
-purposes.
-
-Now we create a transaction.
-
-First, we need some money.  Thus, we create a fake transaction with
-one single output.  This transaction can be parsed from the bitcoin
-network, but we create one manually in order to see this process.
+A crucial function for the verification of the no-inflation property
+is the \texttt{checkTransaction} function, shown in
+Figure~\ref{fig:checktrans}. To better understand it, we first see how
+to create a transaction.
+
+\paragraph{Creating a Transaction}
+
+% The bitcoin-s library has a bitcoin
+% transaction builder class with the following constructor:
+
+% \begin{lstlisting}[style=scala]
+% def apply(
+%       destinations: Seq[TransactionOutput],
+%       utxos: BitcoinTxBuilder.UTXOMap,
+%       feeRate: FeeUnit,
+%       changeSPK: ScriptPubKey,
+%       network: BitcoinNetwork): Future[BitcoinTxBuilder] = {...}
+% \end{lstlisting}
+
+% The purpose of the return type Future is not clear to us here, since
+% the constructor always returns a completed future. This might be for
+% future purposes.
+
+To create a transaction, we first need some coins -- an unspent
+transaction output. We could load an actual unspent transaction output
+from the bitcoin network, but we create one manually in order to see
+this process. So we first create an (invalid) transaction with one
+output in Figure~\ref{fig:prevtx}.
+
+\begin{figure}
 \begin{lstlisting}[style=scala]
@@
 
-On line one and two we create a new keypair to sign the next
-transaction and have a scriptPubKey where the bitcoins are.  This is
-our keypair.  So the money is transferred to our public key.  Line
-four specifies the amount of satoshis we have in the transaction.
-Then we create the actual transaction from line 6 to 13.
-
-Now that we have some bitcoins, we create the new transaction where we
-want to spend them.
+  \caption{Creating a transaction output to spend}
+  \label{fig:prevtx}
+\end{figure}
+
+We first create a keypair, then a lock script with its public key,
+then the amount of satoshis, then a transaction output (utxo) for that
+amount and locked with that script. Finally we create the actual
+transaction with that output and no inputs. Of course, that is not a
+valid transaction, because it creates coins out of nothing. In
+particular, checkTransaction(prevTx) is false.
+
+Now that we have a transaction output, we create a new transaction to
+spend it.
 
@@
 so on.
+\begin{figure}
 \begin{lstlisting}[style=scala]
@@
   val utxos = List(utxoSpendingInfo)
-\end{lstlisting}
-
-This utxos are the inputs of our transaction.
-
-Second, we need destinations to spend the bitcoins to.
-For the sake of convenience we create only one.
-
-\begin{lstlisting}[style=scala]
+
   val destinationAmount = Satoshis(Int64(5000))
@@
   )
-\end{lstlisting}
-
-We spend 5000 satoshis to the newly created random public key.
-
-Finally, we define the fee rate in satoshis per one byte transaction
-size as well as some bitcoin network parameters.  The bitcoin network
-parameters are not important, so we use some static values normally
-used when testing.
-
-\begin{lstlisting}[style=scala]
+
   val feeRate = SatoshisPerByte(Satoshis.one)
@@
   val networkParams = RegTest // some static values for testing
-\end{lstlisting}
-
-Now lets build the transaction with those data.
-\begin{lstlisting}[style=scala]
+
   val txBuilderF: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
@@
 \end{lstlisting}
-
+  \caption{Creating a transaction}
+  \label{fig:tx}
+\end{figure}
+
+This utxos are the inputs of our transaction.
+Second, we need destinations to spend the bitcoins to.
+For the sake of convenience we create only one.
+We spend 5000 satoshis to the newly created random public key.
+Finally, we define the fee rate in satoshis per one byte transaction
+size as well as some bitcoin network parameters.  The bitcoin network
+parameters are not important, so we use some static values normally
+used when testing.
+
+Now lets build the transaction with those data.
 Line one to seven creates a transaction builder which is then signed
@@
 Bitcoin-S offers a function called \emph{checkTransaction} located in the ScriptInterpreter object.
-This is its type signature:
-\begin{lstlisting}[style=scala]
-  checkTransaction(transaction: Transaction): Boolean
-\end{lstlisting}
-
+
+\begin{figure}
 \begin{lstlisting}[style=scala]
@@
   val allOutputsValidMoneyRange = validMoneyRange(totalSpentByOutputs)
-  val prevOutputs = transaction.inputs.map(_.previousOutput)
-  val noDuplicateInputs = prevOutputs.distinct.size == prevOutputs.size
+  val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
+  val noDuplicateInputs = prevOutputTxIds.distinct.size == prevOutputTxIds.size
 
@@
 \end{lstlisting}
+
+  \caption{The \texttt{checkTransaction} function in the \texttt{ScriptInterpreter} object}
+  \label{fig:checktrans}
+\end{figure}
 
@@
 
-\todo{Kai: about errors, is it enough to reference to the chapter 4.2 where they are described??}
-\todo{After integrating the Stainless plugin in the Bitcoin-S sbt project, 
+After integrating the Stainless plugin in the Bitcoin-S sbt project, 
 there were many errors because of the different sbt versions.  
-Some errors are described in the section 4.2.}
+Some errors are described in the section 4.2.
 It takes too much time to fix them all so it should be easier to extract
@@
 
-\todo{
-kai text
-
-extracted 2 files (they are in code/addition/src/main/scala/addition/original)
-
-changes from original to reduced (new folder):
-  - package name
-
-  - removed numbers other than Int64
+The relevant code is in two files: \texttt{CurrencyUnits.scala} and \texttt{NumberType.scala}.
+
+reference: code/addition/src/main/scala/addition/original
+
+From those files we removed all code that is not needed for our
+verification. For example, we removed all number types except for
+Int64 (so Int32, UInt64, etc.) because we do not use them. We also
+removed the superclasses Factory and NetworkElement of CurrencyUnit
+and Number, respectively because the inherited members are not used by
+the relevant code. Also, we removed all binary operations on Number that are not used.
 
@@
 }
-
-\todo{ramon: write this section}
-\todo{ramon: keep close to original code}
-\todo{ramon: for every transformation: before and after}
-\todo{ramon,anna: for every transformation: why does it preserve the semantics}
-\todo{ramon,anna: add comments}
-\todo{ramon,anna: explain what the code does and why}
 
@@
 the Stainless developers to improve the tool.
+
+\todo{conclusions: what's future work? how to
+  change bitcoin-s? how to extend stainless?}
 

### 2019-06-26 paper/paper.tex
remove needless transformation

@@
   The code we use for the following sections is in reduced folder.
-}
 
@@
 
-When we run Stainless on this code (without any properties to prove), it throws the following errors: \todo{describe errors: no support for abstract types, unsupported arguments for the BigInt constructor, unsupported inheritance for objects.}
-
-\paragraph{Transforming the Code}
+When we run Stainless on this code (without any properties to prove),
+it throws the following errors: \todo{describe errors: no support for
+  abstract types, unsupported arguments for the BigInt constructor,
+  unsupported inheritance for objects.}
+
+\paragraph{Transforming the Code.}
 
 \todo{kai todo}
+
+\paragraph{The Specification.}
+
+As specified, our verification must only support addition with zero.
+So we restrict the parameter to be zero in the precondition.
+\begin{lstlisting}[style=scala]
+  require(c.satoshis == Satoshis.zero)
+\end{lstlisting}
+
+We ensure the result is the same value as \emph{this} in the postcondition.
+\begin{lstlisting}[style=scala]
+  ensuring(res => res.satoshis == this.satoshis)
+\end{lstlisting}
+
+We can use equals (==) directly on Satoshis, because it is a case class.
+
+Before:
+\begin{lstlisting}[style=scala]
+sealed abstract class CurrencyUnit {
+  def satoshis: Satoshis
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
+
+  protected def underlying: Int64
+}
+\end{lstlisting}
+
+The addition will now look like this:
+\begin{lstlisting}[style=scala]
+sealed abstract class CurrencyUnit {
+  def satoshis: Satoshis
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  override def +(c: CurrencyUnit): CurrencyUnit = {
+    require(c.satoshis == Satoshis.zero)
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  } ensuring(res => res.satoshis == this.satoshis)
+
+  protected def underlying: Int64
+}
+\end{lstlisting}
+
+That's all we need to verify our addition.
 
@@
 
-\subsection{Writing Specification for the Property}
-
-As specified, our verification must only support addition with zero.
-So we restrict the parameter to be zero in the precondition.
-\begin{lstlisting}[style=scala]
-  require(c.satoshis == Satoshis.zero)
-\end{lstlisting}
-
-We ensure the result is the same value as \emph{this} in the postcondition.
-\begin{lstlisting}[style=scala]
-  ensuring(res => res.satoshis == this.satoshis)
-\end{lstlisting}
-
-We can use equals (==) directly on Satoshis, because it is a case class.
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class CurrencyUnit {
-  def satoshis: Satoshis
-
-  def !=(c: CurrencyUnit): Boolean = !(this == c)
-
-  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
-  def +(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying + c.satoshis.underlying)
-  }
-
-  protected def underlying: Int64
-}
-\end{lstlisting}
-
-The addition will now look like this:
-\begin{lstlisting}[style=scala]
-sealed abstract class CurrencyUnit {
-  def satoshis: Satoshis
-
-  def !=(c: CurrencyUnit): Boolean = !(this == c)
-
-  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
-  override def +(c: CurrencyUnit): CurrencyUnit = {
-    require(c.satoshis == Satoshis.zero)
-    Satoshis(satoshis.underlying + c.satoshis.underlying)
-  } ensuring(res => res.satoshis == this.satoshis)
-
-  protected def underlying: Int64
-}
-\end{lstlisting}
-
-That's all we need to verify our addition.
-
-Now we will look into the previous error.
-
-\subsection{Propagating Require}
-
-There is another problem with Bitcoin-S.
-Bitcoin-S-Core uses require as a fail-fast method whereas Stainless needs it to verify the code.
-
-Stainless output now:
-\begin{lstlisting}[style=stainless]
-[  Info  ]  - Checking cache: 'postcondition' VC for + @12:3...
-[  Info  ]  - Checking cache: 'adt invariant' VC for apply @34:39...
-[  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
-[  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
-[  Info  ] Cache hit: 'adt invariant' VC for apply @34:39...
-[  Info  ] Cache hit: 'postcondition' VC for + @12:3...
-[  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
-[  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
-[Warning ]  => INVALID
-[Warning ] Found counter-example:
-[Warning ]   bigInt: BigInt -> -9223372036854775809
-[  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-[  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-[  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-[  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
-[Warning ]  => INVALID
-[Warning ] Found counter-example:
-[Warning ]   num: Number   -> Int64Impl(9223372036854775807)
-[Warning ]   thiss: Number -> Int64Impl(1)
-\end{lstlisting}
-
-Corresponding code:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: BigInt
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: BigInt => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): BigInt = {
-    require(
-         result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"))
-    result
-  }
-}
-\end{lstlisting}
-
-But how does Stainless find a counter example ignoring the require in checkResult?
-Since Stainless is a static verification tool, it tests every possibility.
-So it can use a number bigger than the maximum Int64 and pass it to the addition.
-The require in checkResult fails.
-
-Thus, we need to add the restriction from checkResult to the addition too.
-\begin{lstlisting}[style=scala]
-override def +(num: Int64): Int64 = {
-  require(
-       num.underlying <= BigInt("9223372036854775807")
-    && num.underlying >= BigInt("-9223372036854775808")
-    && this.underlying <= BigInt("9223372036854775807")
-    && this.underlying >= BigInt("-9223372036854775808")
-  )
-  apply(checkResult(underlying + num.underlying))
-}
-\end{lstlisting}
-
-Stainless finds another counter example:
-\begin{lstlisting}[style=stainless]
-[Warning ] Found counter-example:
-[Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(1)
-[Warning ]   thiss: { x: Object | @unchecked isNumber(x) } ->
-                Int64Impl(9223372036854775807)
-\end{lstlisting}
-
-Sure, when adding one to the maximum Int64 the require does not hold anymore.
-Since we do only allow zero as a parameter, the easiest way is to restrict it to zero here too.
-
-Code after:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: BigInt
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: BigInt => Int64
-
-  def +(num: Int64): Int64 = {
-    require(
-          num == Int64.zero
-      && this.underlying <= BigInt("9223372036854775807")
-      && this.underlying >= BigInt("-9223372036854775808"))
-    apply(checkResult(underlying + num.underlying))
-  }
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): BigInt = {
-    require(
-          result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"))
-    result
-  }
-}
-\end{lstlisting}
-
-The same problem occurs in the Int64 constructor.
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: BigInt => Int64 = Int64(_)
-}
-
-case object Int64 extends BaseNumbers[Int64] {
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
-
-  lazy val min = Int64(BigInt("-9223372036854775808"))
-  lazy val max = Int64(BigInt("9223372036854775807"))
-
-  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
-}
-
-private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= BigInt("-9223372036854775808"))
-  require(underlying <= BigInt("9223372036854775807"))
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: BigInt => Int64 = bigInt => {
-    require(
-          bigInt >= BigInt("-9223372036854775808")
-      && bigInt <= BigInt("9223372036854775807"))
-
-    Int64(bigInt)
-  }
-}
-
-case object Int64 extends BaseNumbers[Int64] {
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
-
-  lazy val min = Int64(BigInt("-9223372036854775808"))
-  lazy val max = Int64(BigInt("9223372036854775807"))
-
-  def apply(bigInt: BigInt): Int64 = {
-    require(
-         bigInt >= BigInt("-9223372036854775808")
-      && bigInt <= BigInt("9223372036854775807"))
-
-    Int64Impl(bigInt)
-  }
-}
-
-private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= BigInt("-9223372036854775808"))
-  require(underlying <= BigInt("9223372036854775807"))
-}
-\end{lstlisting}
-
 \end{document}

### 2019-06-26 paper/paper.tex
start working on transformations

@@
 
-These are the code transformations needed to Verify the Addition-with-Zero Property.
-
-When we run Stainless on this code (without any properties to prove), it throws the following errors:
-
-\begin{lstlisting}[style=stainless]
-[ Error  ] currency/CurrencyUnits.scala:6:3: Stainless doesn't support abstract type members
-            type A
-            ^^^^^^
-[ Error  ] currency/CurrencyUnits.scala:26:33: Only literal arguments are allowed for BigInt.
-            def toBigInt: BigInt = BigInt(toLong)
-                                          ^^^^^^
-[ Error  ] currency/CurrencyUnits.scala:33:1: Objects cannot extend classes or implement traits, use a case object instead
-          object Satoshis extends BaseNumbers[Satoshis] {
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-[ Error  ] number/NumberType.scala:65:1: Objects cannot extend classes or implement traits, use a case object instead
-          object Int64 extends BaseNumbers[Int64] {
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-[  Info  ] Shutting down executor service.
-\end{lstlisting}
-
-So let's see how we can fix those errors.
-
-\subsection{An Inheriting Object}
-
-Stainless output:
+Here we see in detail how to transform the bitcoin-s code into the
+Scala fragment supported by Stainless. All subsections start with the
+code on which Stainless is called, followed by the Stainless error
+message(s) and finally a description of the changes we make to the
+code.
+
+We claim that all transformations are equivalent in the sense that if
+the addition-with-zero property holds for the transformed code, then
+it also holds for the code before the transformation.
+
+\subsection{Inheriting Objects}
+
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}
+
 \begin{lstlisting}[style=stainless]
@@
 
-Code before:
-\begin{lstlisting}[style=scala]
-object Int64 extends BaseNumbers[Int64] { ... }
-object Satoshis extends BaseNumbers[Satoshis] { ... }
-\end{lstlisting}
-
-Code after:
-\begin{lstlisting}[style=scala]
-case object Int64 extends BaseNumbers[Int64] { ... }
-case object Satoshis extends BaseNumbers[Satoshis] { ... }
-\end{lstlisting}
-
-Here, we can just change the objects from object to case object.
-Stainless recommendation is to use objects for modules and case objects as algebraic data types.
-
-\todo{Kai this does not change the semantic}
-
-This is due to the internal design of Scala.
-It's possible to reason about case object but not about object.
-This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
-The \href{https://github.com/epfl-lara/stainless/issues/520}{issue number 520} on Stainless GitHub gives some thoughts, if you want to know more.
-
-\subsection{Abstract Type Member}
-
-Stainless output:
+Here, we can just turn the objects into case objects by literally just
+changing the word \texttt{object} into \texttt{case object} on lines
+33 and 65 in the two respective files.
+
+That transformation is equivalent. Case objects have some additional
+properties (in particular, being serializable) and they inherit from
+\texttt{Product} instead of \texttt{AnyRef}, but none of our code
+depends on any of that.
+
+\subsection{Abstract Type Members}
+
 \begin{lstlisting}[style=stainless]
@@
 
-\todo{kai this does not change the semantic}
-
 Before:
@@
 That's too specific, because the return type of the addition is then CurrencyUnit[Int64] not CurrencyUnit[A].
+
 Maybe the Bitcoin-S developers should reimplement this part and not use Satoshis directly.
-
-Since there is no easy way to fix it and the code should stay as much as possible the original, we just remove the abstract type and set it to Int64.
-This limits the verification a bit, but as we only want to verify the addition in satoshis, that's OK.
+\todo{ramon: explain. what would be the correct code?}
+
+Since we only want to verify the addition of satoshis, and the Satoshis class overrides A with Int64 anyway,  we just remove the abstract type and set it to Int64.
 

### 2019-06-27 paper/paper.tex
more work on transformations

@@
 
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}
 \lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}
 
 \begin{lstlisting}[style=stainless]
+[ Error  ] number/NumberType.scala:65:1: Objects cannot extend classes or implement traits, use a case object instead
+          object Int64 extends BaseNumbers[Int64] {
+          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
 [ Error  ] currency/CurrencyUnits.scala:33:1: Objects cannot extend classes or implement traits, use a case object instead
@@
           ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-[ Error  ] number/NumberType.scala:65:1: Objects cannot extend classes or implement traits, use a case object instead
-          object Int64 extends BaseNumbers[Int64] {
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
 \end{lstlisting}
@@
 changing the word \texttt{object} into \texttt{case object} on lines
-33 and 65 in the two respective files.
+65 and 33 in the two respective files.
 
@@
 
-This should be easy to rewrite by using generics instead of an abstract type, right?
-Unfortunately not.
-The problem is, CurrencyUnit uses one of its implementing classes: Satoshis.
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class CurrencyUnit {
-  type A
-
-  def satoshis: Satoshis
-
-  def !=(c: CurrencyUnit): Boolean = !(this == c)
-
-  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
-  def +(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying + c.satoshis.underlying)
-  }
-
-  protected def underlying: A
-}
-
-sealed abstract class Satoshis extends CurrencyUnit {
-  override type A = Int64
-
-  override def satoshis: Satoshis = this
-
-  def toBigInt: BigInt = BigInt(toLong)
-
-  def toLong: Long = underlying.toLong
-
-  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
-}
-\end{lstlisting}
-
-What happens, if we typify CurrencyUnit with A, meaning to make it generic with type A?
-
-Satoshis extends CurrencyUnit with type Int64, so it would be of type CurrencyUnit[Int64].
-That's too specific, because the return type of the addition is then CurrencyUnit[Int64] not CurrencyUnit[A].
-
-Maybe the Bitcoin-S developers should reimplement this part and not use Satoshis directly.
-\todo{ramon: explain. what would be the correct code?}
-
-Since we only want to verify the addition of satoshis, and the Satoshis class overrides A with Int64 anyway,  we just remove the abstract type and set it to Int64.
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class CurrencyUnit {
-  def satoshis: Satoshis
-
-  def !=(c: CurrencyUnit): Boolean = !(this == c)
-
-  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
-  def +(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying + c.satoshis.underlying)
-  }
-
-  protected def underlying: Int64
-}
-
-sealed abstract class Satoshis extends CurrencyUnit {
-  override def satoshis: Satoshis = this
-
-  def toBigInt: BigInt = BigInt(toLong)
-
-  def toLong: Long = underlying.toLong
-
-  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
-}
-\end{lstlisting}
+Note that we can not simply replace the unsupported abstract type
+member by a (supported) type parameter. The problem is that the
+CurrencyUnit class uses one of its implementing classes: Satoshis.
+
+Satoshis would have to instantiate a potential type parameter with
+type Int64, so it would extend CurrencyUnit[Int64]. But that is too
+specific, because the return type of the \texttt{+}-method would then then be
+CurrencyUnit[Int64] not CurrencyUnit[A].
+
+Since we only want to verify the addition of satoshis, and the
+Satoshis class overrides A with Int64 anyway, we just remove the
+abstract type and set it to Int64.
+
+We remove line 6 and line 22 from CurrencyUnits.scala (to maintain
+line numbers we actually replace them with empty lines for now) and in
+line 18 we replace A by Int64.
 
@@
 
-Stainless output:
 \begin{lstlisting}[style=stainless]
@@
 
-As described before, Stainless supports only a subset of Scala.
-The BigInt from the Stainless library is a bit restricted.
-One such restriction is, that BigInt does not support dynamic BigInt construction.
-Thus, the constructor parameter of BitInt must be a literal argument.
-
-\todo{kai this does not change semantic}
-
-Before:
+As described before, the types in the Stainless library are more
+restricted than their Scala library counterparts. In particular, the
+Stainless BigInt constructor is restricted to literal arguments.
+
+After:
 \begin{lstlisting}[style=scala]
@@
 
-  def toBigInt: BigInt = BigInt(toLong)
+  def toBigInt: BigInt = underlying.toBigInt
 
@@
 
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Satoshis extends CurrencyUnit {
-  override def satoshis: Satoshis = this
-
-  def toBigInt: BigInt = underlying.toBigInt
-
-  def toLong: Long = underlying.toLong
-
-  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
-}
-\end{lstlisting}
-
-This would be really hard to refactor, because Bitcoin-S tries to be as dynamic as possible so it can be used with cryptocurrencies other than bitcoins.
-Maybe it could be impossible, because they need to parse dynamic values from the bitcoin network.
-
-Luckily, we can use toBigInt on the field \texttt{underlying} directly instead of toLong.
-So, instead of converting the underlying to Long and back to BigInt we convert underlying directly to BigInt.
-
-After fixing all Stainless errors, a new error appears.
+Here we can simply use toBigInt on the field \texttt{underlying}
+directly.  So, instead of converting the underlying to Long and back
+to BigInt we convert underlying directly to BigInt.
+
+We replace line 26 by
+\begin{lstlisting}[style=scala]
+def toBigInt: BigInt = underlying.toBigint
+\end{lstlisting}
+
+\todo{explain why equivalence or why not equivalence}
 
@@
 
-\todo{kai this does not change semantic}
-
-This is a missing feature in Stainless.
-It does not support upper type boundaries on the class itself.
-To track this, \href{https://github.com/epfl-lara/stainless/issues/519}{issue 519} was created on GitHub during this work.
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number[T <: Number[T]] {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toLong: Long = toBigInt.bigInteger.longValueExact()
-  def toBigInt: BigInt = underlying
-
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
-  def andMask: BigInt
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => T
-
-  def +(num: T): T = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require((result & andMask) == result,
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}
-
-/**
-  * Represents a signed number in our number system
-  * Instances of this is [[Int64]]
-  */
-sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
-
-/**
-  * Represents a int64_t in C
-  */
-sealed abstract class Int64 extends SignedNumber[Int64] {
-  override def apply: A => Int64 = Int64(_)
-  override def andMask = 0xffffffffffffffffL
-}
-\end{lstlisting}
-
-After:
+Stainless does not currently support a class with a type parameter
+with a type boundary that contains the class itself. We opened an
+issue \cite{Stainless:issue519} on the Stainless repository and the
+Stainless developers have targeted Stainless version 0.4 to support
+such self-referential type boundaries.
+
+For now, since our code only uses Number with type parameter T
+instantiated to Int64, we just remove the type parameter and replace
+it by Int64. We respectively replace lines 8, 44 and 49 by
 \begin{lstlisting}[style=scala]
 sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toLong: Long = toBigInt.bigInteger.longValueExact()
-  def toBigInt: BigInt = underlying
-
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
-  def andMask: BigInt
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require((result & andMask) == result,
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}
-
-/**
-  * Represents a signed number in our number system
-  * Instances of this is [[Int64]]
-  */
+\end{lstlisting}
+
+\begin{lstlisting}[style=scala]
 sealed abstract class SignedNumber extends Number
-
-/**
-  * Represents a int64_t in C
-  */
+\end{lstlisting}
+
+\begin{lstlisting}[style=scala]
 sealed abstract class Int64 extends SignedNumber {
-  override def apply: A => Int64 = Int64(_)
-  override def andMask = 0xffffffffffffffffL
-}
-\end{lstlisting}
-
-Despite this, in order to be able to continue, we make this a concrete type by replacing T with Int64.
-Int64,  because Satoshis uses only Int64.
-There are other number types like UInt16 but for our property we don't need them.
-
-Now, there are two new errors.
+\end{lstlisting}
+
+and replace T by Int64 in lines 25-27.
 
 \subsection{Missing Member \texttt{bigInteger} in BigInt}
+
 \todo{Ramon is .bigInteger missing in Stainless?}
+
 Stainless output:

### 2019-06-27 paper/paper.tex
more work on transformations

@@
 
-\todo{explain why equivalence or why not equivalence}
+This is an equivalent transformation: the only thing that might go
+wrong in the detour via \texttt{Long} is that the BigInt underlying
+Int64 in turn underlying Satoshis does not fit into a Long. However,
+the only constructor of Int64 ensures exactly that.
 
@@
 \subsection{Missing Member \texttt{bigInteger} in BigInt}
+
+KAI CONTINUE HERE
 

### 2019-06-28 paper/paper.tex
fnished work on transformations

@@
 
+\section{NumberType.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}
+
+\section{CurrencyUnits.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala}
+
 \section{Code Transformations}
@@
 Scala fragment supported by Stainless. All subsections start with the
-code on which Stainless is called, followed by the Stainless error
-message(s) and finally a description of the changes we make to the
-code.
+Stainless error message(s) and finally a description of the changes we
+make to the code.
 
@@
 \subsection{Inheriting Objects}
-
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala}
 
@@
 
-KAI CONTINUE HERE
-
-\todo{Ramon is .bigInteger missing in Stainless?}
-
-Stainless output:
 \begin{lstlisting}[style=stainless]
@@
 \end{lstlisting}
-
-\todo{kai This does not change semantic}
-
-Now that we do not depend on toLong anymore lets just remove it from out code.
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Satoshis extends CurrencyUnit {
-  override def satoshis: Satoshis = this
-
-  def toBigInt: BigInt = underlying.toBigInt
-
-  def toLong: Long = underlying.toLong
-
-  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
-}  
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Satoshis extends CurrencyUnit {
-  override def satoshis: Satoshis = this
-
-  def toBigInt: BigInt = underlying.toBigInt
-
-  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
-}  
-\end{lstlisting}
-
-And before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toLong: Long = toBigInt.bigInteger.longValueExact()
-  def toBigInt: BigInt = underlying
-
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
-  def andMask: BigInt
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require((result & andMask) == result,
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}  
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toBigInt: BigInt = underlying
-
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
-  def andMask: BigInt
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require((result & andMask) == result,
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}
-\end{lstlisting}
+\todo{stainless error emssage missing}
+
+The Scala class BigInt is essentially a wrapper around
+java.math.BigInteger. BigInt has a member bigInteger which is the
+underlying instance of the Java class. The Java class has a method
+longValueExact which returns a long only if the BigInteger fits into a
+long, otherwise throws exception. Stainless does not support Java
+classes and in particular its BigInt has no member bigInteger.
+
+However, our code never calls toLong anymore, so we just remove it. We
+replace line 14 in NumberType.scala and line 28 in CurrencyUnits.scala
+by an empty line.
 
@@
 
-\todo{kai this does not change semantic}
-
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require(
-          result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"))
-    result
-  }
-}
-
-/**
-  * Represents a signed number in our number system
-  * Instances of this is [[Int64]]
-  */
-sealed abstract class SignedNumber extends Number
-
-/**
-  * Represents a int64_t in C
-  */
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: A => Int64 = Int64(_)
-}  
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: BigInt
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: BigInt => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): BigInt = {
-    require(
-         result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"))
-    result
-  }
-}
-
-/**
-  * Represents a signed number in our number system
-  * Instances of this is [[Int64]]
-  */
-sealed abstract class SignedNumber extends Number
-
-/**
-  * Represents a int64_t in C
-  */
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: BigInt => Int64 = Int64(_)
-}
-\end{lstlisting}
-
-This is easy.
-We just replace all occurrence of A with BigInt, since A is not overwritten in an implementing class.
-This is not the exact same code, because an implementing class can not override A anymore but that's fine for our verification.
-
-This was a missing feature in Stainless that was fixed on Mai 28, 2019 with \href{https://github.com/epfl-lara/stainless/pull/470}{pull request 470} on GitHub.
-Now it should work without this change.
+Our version of Stainless does not support type members. We just
+replace all occurrence of A with BigInt, since A is never overwritten
+in an implementing class.
+
+We remove line 9 in NumberType.scala and replace A by BigInt in lines
+12, 25, 33 and 50.
+
+In the mean time Stainless has implemented support for type member
+\cite{Stainless:pull470}.  Since version 0.2 verification should
+succeed without this change.
 
 \subsection{Missing Bitwise-And Method on BigInt}
-\label{sec:bound_check}
-
-Stainless output:
+
 \begin{lstlisting}[style=stainless]
-[ Error  ] number/NumberType.scala:33:14: Unknown call to & on result (BigInt) with arguments List(Number.this.andMask) of type List(BigInt)
+[ Error  ] number/NumberType.scala:34:14: Unknown call to & on result (BigInt) with arguments List(Number.this.andMask) of type List(BigInt)
               require((result & andMask) == result,
@@
 
-\todo{kai here we do not know if it changes semantic}
-
-Due to the restrictions on BigInt, we can not use the \& function either.
-Before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toBigInt: BigInt = underlying
-
-  /**
-    * This is used to determine the valid amount of bytes in a number
-    * for instance a UInt8 has an andMask of 0xff
-    * a UInt32 has an andMask of 0xffffffff
-    */
-  def andMask: BigInt
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require((result & andMask) == result,
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}
-
-/**
-  * Represents a signed number in our number system
-  * Instances of this is [[Int64]]
-  */
-sealed abstract class SignedNumber extends Number
-
-/**
-  * Represents a int64_t in C
-  */
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: A => Int64 = Int64(_)
-  override def andMask = 0xffffffffffffffffL
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require(
-          result <= BigInt("9223372036854775807")
+Contrary to Scala BigInt, the Stainless BigInt class does not support
+bitwise operations, in particular not the \&-method for bitwise and.
+
+The bitwise and with the andMask on line 34 is a bounds check.  It
+checks if the result parameter is in range of the specified type,
+which in our case is the hard coded Int64.
+
+So, we replace the \& mask with a check whether the result
+is in range of Long.MinValue and Long.MaxValue, because Int64 has the
+same 64-bit range as Long.
+
+We replace lines 34-35 by the following, squeezed into two lines to
+preserve line numbers. Note that the BigInt constructor requires a
+literal:
+
+\begin{lstlisting}[style=scala]
+require(result <= BigInt("9223372036854775807")  
   && result >= BigInt("-9223372036854775808"),
       "Result was out of bounds, got: " + result)
-    result
-  }
-}
-
-/**
-  * Represents a signed number in our number system
-  * Instances of this is [[Int64]]
-  */
-sealed abstract class SignedNumber extends Number
-
-/**
-  * Represents a int64_t in C
-  */
-sealed abstract class Int64 extends SignedNumber {
-  override def apply: A => Int64 = Int64(_)
-}
-\end{lstlisting}
-
-This is a bounds check.
-It checks if the result of the addition is in range of the specified type, which is now the hard coded Int64.
-
-So, we can replace the \& mask with a bound check whether the result is in range of Long.MinValue and Long.MaxValue, because Int64 has the same 64-bit range as Long.
-Again the code gets a bit more static and it's not the exact same code anymore.
-
-Running Stainless produces again new errors.
+\end{lstlisting}
+
+\todo{kai here we do not know if it changes semantic. ramon: oh, we don't?}
 
@@
 
-Stainless output:
 \begin{lstlisting}[style=stainless]
-[Warning ] currency/CurrencyUnits.scala:36:3: Could not extract tree in class: case private class SatoshisImpl extends Satoshis with Product with Serializable {
+[Warning ] number/NumberType.scala:66:3: Could not extract tree in class: case private class Int64Impl extends Int64 with Product with Serializable {
+...
+[Warning ] } (class scala.reflect.internal.Trees$ClassDef)
+              private case class Int64Impl(underlying: BigInt) extends Int64 {
+              ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
+...
+[Warning ] currency/CurrencyUnits.scala:42:3: Could not extract tree in class: case private class SatoshisImpl extends Satoshis with Product with Serializable {
 ...
@@
 
-And:
+Stainless does not support inner classes in a case object. Bitcoin-s
+uses this a lot to separate the class \todo{ramon: i don't understand} from its implementation.
+
+This is easy to fix. We just move the inner classes out of the case objects. They do
+not interfere with any other code.
+
+We remove lines 66-71 in NumberType.scala and insert them at the end of the file.
+We remove line 42 in CurrencyUnits.scala and insert it at the end of the file.
+
+\subsection{Message Parameter in Require}
+
 \begin{lstlisting}[style=stainless]
-[Warning ] number/NumberType.scala:59:3: Could not extract tree in class: case private class Int64Impl extends Int64 with Product with Serializable {
-...
-[Warning ] } (class scala.reflect.internal.Trees$ClassDef)
-              private case class Int64Impl(underlying: BigInt) extends Int64 {
-              ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-\end{lstlisting}
-
-Stainless can not extract the private class inside the object.
-Bitcoin-S uses this a lot, because they separate the class from its implementation.
-
-\todo{kai this does not change semantic}
-
-Before:
-\begin{lstlisting}[style=scala]
-case object Satoshis extends BaseNumbers[Satoshis] {
-  val min = Satoshis(Int64.min)
-  val max = Satoshis(Int64.max)
-  val zero = Satoshis(Int64.zero)
-  val one = Satoshis(Int64.one)
-
-  def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
-
-  private case class SatoshisImpl(underlying: Int64) extends Satoshis
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-case object Satoshis extends BaseNumbers[Satoshis] {
-
-  val min = Satoshis(Int64.min)
-  val max = Satoshis(Int64.max)
-  val zero = Satoshis(Int64.zero)
-  val one = Satoshis(Int64.one)
-
-  def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
-}
-
-private case class SatoshisImpl(underlying: Int64) extends Satoshis
-\end{lstlisting}
-
-And before:
-\begin{lstlisting}[style=scala]
-case object Int64 extends BaseNumbers[Int64] {
-  private case class Int64Impl(underlying: BigInt) extends Int64 {
-    require(underlying >= -9223372036854775808L,
-      "Number was too small for a int64, got: " + underlying)
-    require(underlying <= 9223372036854775807L,
-      "Number was too big for a int64, got: " + underlying)
-  }
-
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
-
-  lazy val min = Int64(-9223372036854775808L)
-  lazy val max = Int64(9223372036854775807L)
-
-  def apply(long: Long): Int64 = Int64(BigInt(long))
-
-  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-case object Int64 extends BaseNumbers[Int64] {
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
-
-  lazy val min = Int64(-9223372036854775808L)
-  lazy val max = Int64(9223372036854775807L)
-
-  def apply(long: Long): Int64 = Int64(BigInt(long))
-
-  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
-}
-
-private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= -9223372036854775808L,
-    "Number was too small for a int64, got: " + underlying)
-  require(underlying <= 9223372036854775807L,
-    "Number was too big for a int64, got: " + underlying)
-}
-\end{lstlisting}
-
-This is easy to fix.
-We just extract the private class out of the object.
-This is not exactly the same code, because other classes in the same file could now access the private class.
-But for our property it does not change anything.
-
-Now we get some weird warnings about require.
-
-\subsection{Message Parameter in Require}
-
-Stainless output:
-\begin{lstlisting}[style=stainless]
-[Warning ] number/NumberType.scala:9:3: Could not extract tree in class: type A = BigInt (class scala.reflect.internal.Trees$TypeDef)
-          type A = BigInt
-          ^^^^^^^^^^^^^^^
-[Warning ] number/NumberType.scala:71:3: Could not extract tree in class: scala.this.Predef.require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(-9223372036854775808L)), "Number was too small for a int64, got: ".+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
+[Warning ] number/NumberType.scala:67:3: Could not extract tree in class: scala.this.Predef.require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(-9223372036854775808L)), "Number was too small for a int64, got: ".+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
           require(underlying >= -9223372036854775808L,
           ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-[Warning ] number/NumberType.scala:73:3: Could not extract tree in class: scala.this.Predef.require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(9223372036854775807L)), "Number was too big for a int64, got: ".+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
+[Warning ] number/NumberType.scala:69:3: Could not extract tree in class: scala.this.Predef.require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(9223372036854775807L)), "Number was too big for a int64, got: ".+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
           require(underlying <= 9223372036854775807L,
@@
 
-Seems like Stainless does not support the second string parameter of require or at least it throws a warning about it.
-We can safely remove the string parameters from the requires, since they only serve as error messages.
-
-\todo{kai this does not change semantic}
-
-Before:
-\begin{lstlisting}[style=scala]
-private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= -9223372036854775808L,
-    "Number was too small for a int64, got: " + underlying)
-  require(underlying <= 9223372036854775807L,
-    "Number was too big for a int64, got: " + underlying)
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-private case class Int64Impl(underlying: BigInt) extends Int64 {
+Here the require(condition, message) is used as an assertion: if the
+condition is false the fail with message.  Stainless does not support
+the message parameter of require.  For the verification we can simply
+remove that parameter.
+
+\subsection{Missing Implicit Long to BigInt Conversion}
+
+\begin{lstlisting}[style=stainless]
+[ Error  ] inv$4 depends on missing dependencies: long2bigInt$0.
+\end{lstlisting}
+
+This error message does not specify a line number and it is not clear
+what ``inv'' is. However, the Scala BigInt has implict conversions
+from Long and they are missing in the Stainless BigInt.
+
+We replace the two require clauses at lines 67 and following in NumberType.scala
+
+\begin{lstlisting}[style=scala]
   require(underlying >= -9223372036854775808L)
   require(underlying <= 9223372036854775807L)
-}
-\end{lstlisting}
-
-And before:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require(
-         result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"),
-      "Result was out of bounds, got: " + result)
-    result
-  }
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-  type A = BigInt
-
-  /** The underlying scala number used to to hold the number */
-  protected def underlying: A
-
-  def toBigInt: BigInt = underlying
-
-  /** Factory function to create the underlying T, for instance a UInt32 */
-  def apply: A => Int64
-
-  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
-
-  /**
-    * Checks if the given result is within the range
-    * of this number type
-    */
-  private def checkResult(result: BigInt): A = {
-    require(
-          result <= BigInt("9223372036854775807")
-      && result >= BigInt("-9223372036854775808"))
-    result
-  }
-}
-\end{lstlisting}
-
-A new error appears.
-
-\subsection{Missing Implicit Long to BigInt Conversion}
-
-Stainless output:
+\end{lstlisting}
+
+by:
+\begin{lstlisting}[style=scala]
+  require(underlying >= BigInt(-9223372036854775808L))
+  require(underlying <= BigInt(9223372036854775807L))
+\end{lstlisting}
+
+We also replace lines 76 and 77
+\begin{lstlisting}[style=scala]
+  lazy val min = Int64(-9223372036854775808L)
+  lazy val max = Int64(9223372036854775807L)
+\end{lstlisting}
+
+by:
+\begin{lstlisting}[style=scala]
+  lazy val min = Int64(BigInt(-9223372036854775808L))
+  lazy val max = Int64(BigInt(9223372036854775807L))
+}
+\end{lstlisting}
+
+\subsection{Missing BigInt Constructor with Long Argument}
+
 \begin{lstlisting}[style=stainless]
-[ Error  ] inv$4 depends on missing dependencies: long2bigInt$0.
 [ Error  ] apply$14 depends on missing dependencies: BigInt$0, apply$15.
@@
 
-This error is hard to understand, but we can see that there is a missing Long to BigInt conversion.
-So we search for all Long values in the code.
-
-\todo{kai this does not change semantic}
-
-Before:
-\begin{lstlisting}[style=scala]
-private case class Int64Impl(underlying: BigInt) extends Int64 {
-  require(underlying >= -9223372036854775808L)
-  require(underlying <= 9223372036854775807L)
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-private case class Int64Impl(underlying: BigInt) extends Int64 {
+Here again, the Scala BigInt has a constructor with a Long argument
+which is missing in the Stainless BigInt.
+
+We simply replace the Long values in a BigInt constructor call with a string
+literal.
+
+The lines from the previous transformation respectively change into
+the following:
+
+\begin{lstlisting}[style=scala]
   require(underlying >= BigInt("-9223372036854775808"))
   require(underlying <= BigInt("9223372036854775807"))
-}  
-\end{lstlisting}
-
-And before:
-\begin{lstlisting}[style=scala]
-case object Int64 extends BaseNumbers[Int64] {
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
-
-  lazy val min = Int64(-9223372036854775808L)
-  lazy val max = Int64(9223372036854775807L)
-
-  def apply(long: Long): Int64 = Int64(BigInt(long))
-
-  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
-}
-\end{lstlisting}
-
-After:
-\begin{lstlisting}[style=scala]
-case object Int64 extends BaseNumbers[Int64] {
-  lazy val zero = Int64(0)
-  lazy val one = Int64(1)
-
+\end{lstlisting}
+
+and
+
+\begin{lstlisting}[style=scala]
   lazy val min = Int64(BigInt("-9223372036854775808"))
   lazy val max = Int64(BigInt("9223372036854775807"))
-
-  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
-}
-\end{lstlisting}
-
-Looks like it can not compare a BigInt with a Long value.
-We can easily convert this Long value to a BigInt with a string literal parameter by passing these numbers as strings to the BigInt constructor.
-
-Finally, we get some output from Stainless about the verification in the code.
-
-\begin{lstlisting}[style=stainless]
-[  Info  ]  - Checking cache: 'adt invariant' VC for apply @33:39...
-[  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
-[  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
-[  Info  ] Cache hit: 'adt invariant' VC for apply @33:39...
-[  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
-[  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
-[Warning ]  => INVALID
-[Warning ] Found counter-example:
-[Warning ]   bigInt: BigInt -> -9223372036854775809
-[  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-[  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-[  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
-[  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
-[Warning ]  => INVALID
-[Warning ] Found counter-example:
-[Warning ]   num: Number   -> Int64Impl(9223372036854775807)
-[Warning ]   thiss: Number -> Int64Impl(1)
-\end{lstlisting}
-
-This shows that there is an invalid specification in checkResult and Stainless prints a counterexample for it.
-
-Let's ignore this for a moment and write the specification for the \nameref{property_2}.
-
-\subsection{Missing BigInt Constructor with Long Argument}
-\todo{Split code from above}
+}
+\end{lstlisting}
 

### 2019-06-28 code/addition/bin/stainless.sh
updated script

@@ -1,9 +1,13 @@
 #!/usr/bin/env bash
 
+if [ "$1" == "" ]; then
+  echo "Missing path parameter. Usage: stainless.sh <path to code>"
+  echo "Taking default value '.'"
+fi
+
 DIR="$( cd "$( dirname "${BASH_SOURCE[0]}" )" >/dev/null 2>&1 && pwd )"
 CODE_DIR="modified"
 
 JAVA_HOME=/usr/lib/jvm/java-8-openjdk; PATH=$JAVA_HOME/bin:$PATH; java \
   -jar $DIR/../lib/stainless-0.1.0-93dbd33.jar \
-  $DIR/../src/main/scala/addition/$CODE_DIR/currency/CurrencyUnits.scala \
-  $DIR/../src/main/scala/addition/$CODE_DIR/number/NumberType.scala
+  $(find $1 -type f -name *.scala | tr '\n' ' ')

### 2019-06-28 paper/paper.tex
fix
add information about sbt/JAR

@@
 
+\todo{
+Kai:
+
+sbt-plugin vs JAR:
+Can use either sbt-plugin or JAR
+
+More about sbt on https://www.scala-sbt.org/1.x/docs/index.html (features are also listed there)
+Installation see Stainless Doc 'Installing Stainless' section
+
+Use JAR:
+Download all library dependencies including dependencies of dependencies, pass libraries + source to JAR
+
+Use sbt-plugin: integrate in sbt build tool with addSbtPlugin() method
+
+Why JAR:
+sbt-plugin problems (v 0.1 uses older versions (v0.2 not), no output to console see github \#484)
+otherwise use sbt-plugin because of library management and replicability (everyone uses the same Stainless version configured in sbt)
+
+Reproduce results:
+prereq: Java 8 in /usr/lib/jvm/java-8-openjdk
+1. clone repo
+2. run ./code/bin/stainless.sh <path to code>
+   eg. ./code/bin/stainless.sh code/addition/src/main/scala/addition/reduced/
+}
+
 \subsection{stainless-sbt-plugin vs JAR}
@@
 Stainless does not currently support a class with a type parameter
-with a type boundary that contains the class itself. We opened an
+with a type boundary that contains the type parameter itself. We opened an
 issue \cite{Stainless:issue519} on the Stainless repository and the

### 2019-06-28 paper/paper.tex
work on project setup

@@
 
-In this chapter we look at the practical challenges that could occur
-by using Stainless.  Since Stainless is still in development and there
-is no 1.x release there might still be breaking changes and
-improvements fixing problems described now.
-
-\todo{nur über stainless-sbt-plugin reden damit man weiss wir haben es versucht}
-
-\todo{
-Kai:
-
-sbt-plugin vs JAR:
-Can use either sbt-plugin or JAR
+\paragraph{Using the Stainless-sbt-plugin vs Using the Stanless-JAR}
+
+We can either use the sbt plugin or a JAR file to check code with Stainless.
 
@@
 Installation see Stainless Doc 'Installing Stainless' section
-
-Use JAR:
-Download all library dependencies including dependencies of dependencies, pass libraries + source to JAR
-
-Use sbt-plugin: integrate in sbt build tool with addSbtPlugin() method
-
-Why JAR:
-sbt-plugin problems (v 0.1 uses older versions (v0.2 not), no output to console see github \#484)
-otherwise use sbt-plugin because of library management and replicability (everyone uses the same Stainless version configured in sbt)
-
-Reproduce results:
-prereq: Java 8 in /usr/lib/jvm/java-8-openjdk
-1. clone repo
-2. run ./code/bin/stainless.sh <path to code>
-   eg. ./code/bin/stainless.sh code/addition/src/main/scala/addition/reduced/
-}
-
-\subsection{stainless-sbt-plugin vs JAR}
-
-We can either use the sbt plugin or a JAR file to check code with Stainless.
 
@@
 
+Use JAR:
+Download all library dependencies including dependencies of dependencies, pass libraries + source to JAR
+
+Use sbt-plugin: integrate in sbt build tool with addSbtPlugin() method
+
+Why JAR:
+sbt-plugin problems (v 0.1 uses older versions (v0.2 not), no output to console see github \#484)
+otherwise use sbt-plugin because of library management and replicability (everyone uses the same Stainless version configured in sbt)
+
+\paragraph{Reproducing our Results.}
+
+Prerequisites:
+\begin{enumerate}
+\item Java 8 in /usr/lib/jvm/java-8-openjdk (Stainless requires and Scala recommends Java SE Development Kit 8. Newer Java versions won't work.)
+\end{enumerate}
+
+\begin{lstlisting}[language=bash]
+ clone repo
+ run ./code/bin/stainless.sh <path to code>
+   eg. ./code/bin/stainless.sh code/addition/src/main/scala/addition/reduced/
+\end{lstlisting}
+
 \subsection{Integration into Bitcoin-S}
@@
 might be out of date now.
-
-Stainless requires and Scala recommends Java SE Development Kit 8.
-Newer Java versions won't work.
 
@@
   GitHub} might describe exactly this error.
-
-Now we can finally run Stainless on our code.  But this leads us to
-the next findings.  We must rewrite most of the code, as described in
-the previous chapters.
-
-\begin{itemize}
-\item clone our repo
-\item jar vs sbt, you can use both
-\item contains a stainless jar
-\item call the script blahblah
-\end{itemize}
 

### 2019-06-28 code/bin/stainless.sh
remove static JAVA_HOME

@@ -8,6 +8,6 @@ fi
 DIR="$( cd "$( dirname "${BASH_SOURCE[0]}" )" >/dev/null 2>&1 && pwd )"
 CODE_DIR="modified"
 
-JAVA_HOME=/usr/lib/jvm/java-8-openjdk; PATH=$JAVA_HOME/bin:$PATH; java \
+PATH=$JAVA_HOME/bin:$PATH; java \
   -jar $DIR/../lib/stainless-0.1.0-93dbd33.jar \
   $(find $1 -type f -name *.scala | tr '\n' ' ')

### 2019-06-28 code/addition/src/main/scala/addition/verified/currency/CurrencyUnits.scala
restore previous modified folder and rename to verified

@@ -1,0 +1,38 @@
+package addition.modified.currency
+
+import addition.modified.number.{BaseNumbers, Int64}
+import stainless.lang._
+
+sealed abstract class CurrencyUnit {
+  def satoshis: Satoshis
+
+  def !=(c: CurrencyUnit): Boolean = !(this == c)
+
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+
+  def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  } ensuring(res =>
+    (c.satoshis == Satoshis.zero) ==>
+      (res.satoshis == this.satoshis))
+
+  protected def underlying: Int64}
+
+sealed abstract class Satoshis extends CurrencyUnit {
+  override def satoshis: Satoshis = this
+
+  def toBigInt: BigInt = underlying.toBigInt
+
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}
+
+case object Satoshis extends BaseNumbers[Satoshis] {
+  val min = Satoshis(Int64.min)
+  val max = Satoshis(Int64.max)
+  val zero = Satoshis(Int64.zero)
+  val one = Satoshis(Int64.one)
+
+  def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
+}
+
+private case class SatoshisImpl(underlying: Int64) extends Satoshis

### 2019-06-28 code/addition/src/main/scala/addition/verified/number/NumberType.scala

@@ -1,0 +1,68 @@
+package addition.modified.number
+
+/**
+  * This abstract class is meant to represent a signed and unsigned number in C
+  * This is useful for dealing with codebases/protocols that rely on C's
+  * unsigned integer types
+  */
+sealed abstract class Number {
+  /** The underlying scala number used to to hold the number */
+  protected def underlying: BigInt
+
+  def toBigInt: BigInt = underlying
+
+  /** Factory function to create the underlying T, for instance a UInt32 */
+  def apply: BigInt => Int64
+
+  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))
+
+  /**
+    * Checks if the given result is within the range
+    * of this number type
+    */
+  private def checkResult(result: BigInt): BigInt = {
+    require(
+         result <= BigInt("9223372036854775807")
+      && result >= BigInt("-9223372036854775808"))
+    result
+  }
+}
+
+/**
+  * Represents a signed number in our number system
+  * Instances of this is [[Int64]]
+  */
+sealed abstract class SignedNumber extends Number
+
+/**
+  * Represents a int64_t in C
+  */
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: BigInt => Int64 = Int64(_)
+}
+
+/**
+  * Represents various numbers that should be implemented
+  * inside of any companion object for a number
+  */
+trait BaseNumbers[T] {
+  def zero: T
+  def one: T
+  def min: T
+  def max: T
+}
+
+case object Int64 extends BaseNumbers[Int64] {
+  lazy val zero = Int64(0)
+  lazy val one = Int64(1)
+
+  lazy val min = Int64(BigInt("-9223372036854775808"))
+  lazy val max = Int64(BigInt("9223372036854775807"))
+
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= BigInt("-9223372036854775808"))
+  require(underlying <= BigInt("9223372036854775807"))
+}

### 2019-06-28 code/addition/src/main/scala/allgreen/currency/CurrencyUnits.scala
removed package addition

@@ -1,6 +1,6 @@
-package addition.modified.currency
+package allgreen.currency
 
-import addition.modified.number.{BaseNumbers, Int64}
+import allgreen.number.{BaseNumbers, Int64}
 import stainless.lang._
 
 sealed abstract class CurrencyUnit {

### 2019-06-28 code/addition/src/main/scala/allgreen/number/NumberType.scala

@@ -1,4 +1,4 @@
-package addition.modified.number
+package allgreen.number
 
 /**
   * This abstract class is meant to represent a signed and unsigned number in C

### 2019-06-28 code/addition/src/main/scala/reduced/currency/CurrencyUnits.scala

@@ -1,6 +1,6 @@
-package addition.reduced.currency
+package reduced.currency
 
-import addition.reduced.number.{BaseNumbers, Int64}
+import reduced.number.{BaseNumbers, Int64}
 
 sealed abstract class CurrencyUnit {
   type A

### 2019-06-28 code/addition/src/main/scala/reduced/number/NumberType.scala

@@ -1,4 +1,4 @@
-package addition.reduced.number
+package reduced.number
 
 /**
   * This abstract class is meant to represent a signed and unsigned number in C

### 2019-06-28 code/addition/src/main/scala/verified/currency/CurrencyUnits.scala

@@ -1,6 +1,6 @@
-package addition.modified.currency
+package verified.currency
 
-import addition.modified.number.{BaseNumbers, Int64}
+import verified.number.{BaseNumbers, Int64}
 import stainless.lang._
 
 sealed abstract class CurrencyUnit {

### 2019-06-28 code/addition/src/main/scala/verified/number/NumberType.scala

@@ -1,4 +1,4 @@
-package addition.modified.number
+package verified.number
 
 /**
   * This abstract class is meant to represent a signed and unsigned number in C

### 2019-06-29 paper/paper.tex
more work

@@
   implementation of parts of the Bitcoin protocol. We use the
-  Stainless verifier which supports programs in subset of Scala called
-  the \emph{Pure Scala Fragment}. We first try to verify the property
-  that regular transactions do not create new money. It turns out that
+  Stainless verifier which supports programs in a subset of Scala
+  called \emph{Pure Scala}. We first try to verify the property that
+  regular transactions do not create new money. It turns out that
   there is too much code involved that lies outside of the supported
@@
 \cite{DBLP:conf/ecoop/BlancKKS13,DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}
-and is developed at EPF Lausanne
-\cite{Stainless:github}. It is intended to be
-used by programmers without training in formal verification and thus
-allows to write specifications in Scala and focusses on counterexample
-finding in addition to proving correctness.
+and is developed at EPF Lausanne \cite{Stainless:github}. It is
+intended to be used by programmers without training in formal
+verification. To facilitate that, it accepts specifications written in
+the programming language itself (Scala). Also, it focusses on
+counterexample finding in addition to proving
+correctness. Counterexamples are useful to programmers while
+correctness proofs are not -- correctness is obvious or does not hold,
+and often both at the same time.
 
-The example in Figure~\ref{fig:factorial} from the Stainless documentation
-\cite{Stainless:documentation} demonstrates this. Notice how a
-precondition is specified using the function \emph{require} and a
-postcondition using \emph{ensuring}.
+The example in Figure~\ref{fig:factorial} adapted from the Stainless
+documentation \cite{Stainless:documentation} shows how the verifier is
+used. Notice how a precondition is specified using \emph{require} and
+a postcondition using \emph{ensuring}.
 
@@
 \end{figure}
-Our program happens not to satisfy the specification. An overflow in
-the 32-bit Int leads to a negative result for the input 17, as
-Stainless reports in Figure~\ref{fig:failed}. Changing the type from
-Int to BigInt will result in a successful verification.
+Our program does not satisfy the specification. An overflow in the
+32-bit Int leads to a negative result for the input 17, as Stainless
+reports in Figure~\ref{fig:failed}. Changing the type from Int to
+BigInt will result in a successful verification.
 
@@
 		\includegraphics[width=\textwidth]{output1.png}
-	\caption{Output of Stainless verification for calculating factorial of Int number}
+	\caption{Stainless output for the factorial program}
 	\label{fig:failed}
@@
 
-What will turn out to be more relevant are the following Scala
-features which Stainless does not support, such as: (concrete) class
-definitions, inheritance by objects, abstract type members, variance
-annotations and private inner classes.
+What will turn out to be more relevant are the Scala features which
+Stainless does not support, such as: (concrete) class definitions,
+inheritance by objects, abstract type members, and inner classes in
+case objects.
 
-In addition, Stainless has its own library of some core data types and
+Also, Stainless has its own library of some core data types and
 functions which need to be mapped correctly to functions inside of the
@@
 try to verify the property that a regular (non-coinbase) transaction
-cannot generate new coins. We call it the \emph{no-inflation
+can not generate new coins. We call it the \emph{no-inflation
   property}. Trying to verify it, we uncover and fix a bug in the
@@
 that lies outside of the supported fragment to currently make this
-verification feasible. Instead, we turn to a simpler property to
-verify. The simplest possible property we can think of is the fact
-that adding zero satoshis to a given amount of satoshis yields the
-given amount of satoshis. We call it the \emph{addition-with-zero
-  property} and we try to verify it in Section~3. Here as well we see
-that a significant part of the code lies outside of the supported
-fragment. However, after a series of equivalent transformations we
-arrive at code that we successfully verify.
+verification feasible. So we turn to a simpler property to verify. The
+simplest possible property we can think of is the fact that adding
+zero satoshis to a given amount of satoshis yields the given amount of
+satoshis. We call it the \emph{addition-with-zero property} and we try
+to verify it in Section~3. Here as well we see that a significant part
+of the code lies outside of the supported fragment. However, after a
+series of equivalent transformations we arrive at code that we
+successfully verify.
 
 \section{The No-Inflation Property}
-
-A crucial function for the verification of the no-inflation property
-is the \texttt{checkTransaction} function, shown in
-Figure~\ref{fig:checktrans}. To better understand it, we first see how
-to create a transaction.
-
-\paragraph{Creating a Transaction}
-
-% The bitcoin-s library has a bitcoin
-% transaction builder class with the following constructor:
-
-% \begin{lstlisting}[style=scala]
-% def apply(
-%       destinations: Seq[TransactionOutput],
-%       utxos: BitcoinTxBuilder.UTXOMap,
-%       feeRate: FeeUnit,
-%       changeSPK: ScriptPubKey,
-%       network: BitcoinNetwork): Future[BitcoinTxBuilder] = {...}
-% \end{lstlisting}
-
-% The purpose of the return type Future is not clear to us here, since
-% the constructor always returns a completed future. This might be for
-% future purposes.
-
-To create a transaction, we first need some coins -- an unspent
-transaction output. We could load an actual unspent transaction output
-from the bitcoin network, but we create one manually in order to see
-this process. So we first create an (invalid) transaction with one
-output in Figure~\ref{fig:prevtx}.
-
-\begin{figure}
-\begin{lstlisting}[style=scala]
-  val privKey = ECPrivateKey.freshPrivateKey
-  val creditingSPK = P2PKHScriptPubKey(pubKey = privKey.publicKey)
-
-  val amount = Satoshis(Int64(10000))
-
-  val utxo = TransactionOutput(currencyUnit = amount, scriptPubKey = creditingSPK)
-
-  val prevTx = BaseTransaction(
-    version = Int32.one,
-    inputs = List.empty,
-    outputs = List(utxo),
-    lockTime = UInt32.zero
-  )
-\end{lstlisting}
-
-  \caption{Creating a transaction output to spend}
-  \label{fig:prevtx}
-\end{figure}
-
-We first create a keypair, then a lock script with its public key,
-then the amount of satoshis, then a transaction output (utxo) for that
-amount and locked with that script. Finally we create the actual
-transaction with that output and no inputs. Of course, that is not a
-valid transaction, because it creates coins out of nothing. In
-particular, checkTransaction(prevTx) is false.
-
-Now that we have a transaction output, we create a new transaction to
-spend it.
-
-First, we need some out points.  They point to outputs of previous
-transactions.  We use the index zero, because the previous transaction
-has only one output that becomes the first index zero.  If there were
-two previous outputs, the second output would become the index 1 and
-so on.
-\begin{figure}
-\begin{lstlisting}[style=scala]
-  val outPoint = TransactionOutPoint(prevTx.txId, UInt32.zero)
-
-  val utxoSpendingInfo = BitcoinUTXOSpendingInfo(
-    outPoint = outPoint,
-    output = utxo,
-    signers = List(privKey),
-    redeemScriptOpt = None,
-    scriptWitnessOpt = None,
-    hashType = HashType.sigHashAll
-  )
-
-  val utxos = List(utxoSpendingInfo)
-
-  val destinationAmount = Satoshis(Int64(5000))
-
-  val destinationSPK = P2PKHScriptPubKey(pubKey = ECPrivateKey.freshPrivateKey.publicKey)
-
-  val destinations = List(
-    TransactionOutput(currencyUnit = destinationAmount, scriptPubKey = destinationSPK)
-  )
-
-  val feeRate = SatoshisPerByte(Satoshis.one)
-
-  val networkParams = RegTest // some static values for testing
-
-  val txBuilderF: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
-    destinations = destinations, // where to send the money
-    utxos = utxos,               // unspent transaction outputs
-    feeRate = feeRate,           // fee rate per byte
-    changeSPK = creditingSPK,    // where to send the change
-    network = networkParams      // bitcoin network information
-  )
-
-  val txF: Future[Transaction] = txBuilderF.flatMap(_.sign)
-
-  val tx: Transaction = Await.result(signedTxF, 1 second)
-\end{lstlisting}
-  \caption{Creating a transaction}
-  \label{fig:tx}
-\end{figure}
-
-This utxos are the inputs of our transaction.
-Second, we need destinations to spend the bitcoins to.
-For the sake of convenience we create only one.
-We spend 5000 satoshis to the newly created random public key.
-Finally, we define the fee rate in satoshis per one byte transaction
-size as well as some bitcoin network parameters.  The bitcoin network
-parameters are not important, so we use some static values normally
-used when testing.
-
-Now lets build the transaction with those data.
-Line one to seven creates a transaction builder which is then signed
-on line ten.  We can now use our transaction object on line twelve.
-For example, after calling \emph{hex} on it, we can send the returned
-string to the bitcoin network.
-
-\paragraph{Validating a Transaction}
-
-Bitcoin-S offers a function called \emph{checkTransaction} located in the ScriptInterpreter object.
 
@@
 
-  \caption{The \texttt{checkTransaction} function in the \texttt{ScriptInterpreter} object}
+  \caption{The \texttt{checkTransaction} function}
   \label{fig:checktrans}
@@
 
-We can pass a transaction and it returns a Boolean indicating whether
-the transaction is valid or not.  So for example when we pass the
-transaction we built before the returned value would be true, because
-it's a valid transaction.  It might not be accepted by the bitcoin
-network but for a transaction on its own it's valid.  We can not check
-context with it, because we can only pass one transaction.
+A crucial function for the verification of the no-inflation property
+is the function which validates transactions: the
+\texttt{checkTransaction} function in the ScriptInterpreter object,
+shown in Figure~\ref{fig:checktrans}. It takes a transaction and if
+some basic checks succeed returns true, otherwise false. For example,
+one of those checks is that both the list of inputs and list of
+outputs need to be non-empty.
 
-There are several checks in checkTransaction.
-For example, it checks if there is either no input or no output.
-In this case we get false.
+To better understand the validation of a transaction in bitcoin-s, it
+is useful to review how transactions are represented and created.
 
-The relevant part for the bug we found:
+\paragraph{Creating a Transaction.} The code in this subsection is
+adapted from the bitcoin-s documentation. To create a transaction, we
+first need some coins -- an unspent transaction output. We could load
+an actual unspent transaction output from the bitcoin network, but we
+create one manually in order to see this process. So we first create
+an (invalid) transaction with one output in Figure~\ref{fig:prevtx}.
+
+\begin{figure}
 \begin{lstlisting}[style=scala]
-  val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
-  val noDuplicateInputs = prevOutputTxIds.distinct.size == prevOutputTxIds.size
+  val privKey = ECPrivateKey.freshPrivateKey
+  val creditingSPK = P2PKHScriptPubKey(pubKey = privKey.publicKey)
+
+  val amount = Satoshis(Int64(10000))
+
+  val utxo = TransactionOutput(currencyUnit = amount, scriptPubKey = creditingSPK)
+
+  val prevTx = BaseTransaction(
+    version = Int32.one,
+    inputs = List.empty,
+    outputs = List(utxo),
+    lockTime = UInt32.zero
+  )
 \end{lstlisting}
 
-It gathers all transaction ids referenced by the out points.  When we
-call \emph{distinct} on the returned list, we get a list with
-duplicate removed.  If the size of the new list is the same as the
-size of the old, we know that there was no duplicate transaction id,
-because, as said, distinct removes the duplicates.
+  \caption{Creating a transaction output to spend}
+  \label{fig:prevtx}
+\end{figure}
 
-\paragraph{Trying Stainless on the entire project.}
+We first create a keypair, then a lock script with its public key,
+then the amount of satoshis, then a transaction output (utxo) for that
+amount and locked with that script. Finally we create the actual
+transaction with that output and no inputs. Of course, that is not a
+valid transaction, because it creates coins out of nothing. In
+particular, checkTransaction(prevTx) returns false, simply because the
+list of inputs is empty.
 
-\begin{itemize}
-\item result sbt: no output
-\item result jar: 
-\end{itemize}
+\begin{figure}
+\begin{lstlisting}[style=scala]
+  val outPoint = TransactionOutPoint(prevTx.txId, UInt32.zero)
 
-In order to verify a project, Stainless must be integrated into it.
-We can integrate it in an sbt project adding the Stainless Plugin 
-and the required resolver to pugins.sbt. 
-Another option to use Stainless is to import its libraries in a program code 
-and verify a program from command line using the pre-packaged Stainless JAR file 
-or using the Stainless script built from source.
-Trying to integrate Stainless in Bitcoin-S caused a lot of troubles,
-mainly because of version conflicts.  For more details see chapter
-\ref{chap:appendix_arb}.
+  val utxoSpendingInfo = BitcoinUTXOSpendingInfo(
+    outPoint = outPoint,
+    output = utxo,
+    signers = List(privKey),
+    redeemScriptOpt = None,
+    scriptWitnessOpt = None,
+    hashType = HashType.sigHashAll
+  )
 
-After integrating the Stainless plugin in the Bitcoin-S sbt project, 
-there were many errors because of the different sbt versions.  
-Some errors are described in the section 4.2.
-It takes too much time to fix them all so it should be easier to extract
-the classes needed for the \texttt{checkTransaction} function.
+  val utxos = List(utxoSpendingInfo)
 
-\paragraph{Putting aside the No-Inflation Property.} Naively trying
-Stainless on the entire bitcoin-s codebase results in either no output
-(with sbt) or many errors (with jar). So we extract the relevant code
-to only verify that. However, the extracted code has more than 1500
-lines and liberally uses Scala features outside of the supported
-fragment.  We tried to transform the code into the supported fragment,
-but realize that a better approach is to first verify a simpler
-property with less code involved and then turn back to the
-no-inflation property with more experience. So we turn to the
-addition-with-zero property in the next section. \todo{explain}
+  val destinationAmount = Satoshis(Int64(5000))
 
-\paragraph{Fixing a Bug in Bitcoin-S}
+  val destinationSPK = P2PKHScriptPubKey(pubKey = ECPrivateKey.freshPrivateKey.publicKey)
 
-We can see that there is a bug in the checkTransaction function from
-before, recognized and fixed through this work.
+  val destinations = List(
+    TransactionOutput(currencyUnit = destinationAmount, scriptPubKey = destinationSPK)
+  )
 
-Here is the relevant code of checkTransaction again:
-\begin{lstlisting}[style=scala]
-  val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
-  val noDuplicateInputs = prevOutputTxIds.distinct.size == prevOutputTxIds.size
+  val feeRate = SatoshisPerByte(Satoshis.one)
+
+  val networkParams = RegTest // some static values for testing
+
+  val txBuilderF: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
+    destinations = destinations, 
+    utxos = utxos,               
+    feeRate = feeRate,           
+    changeSPK = creditingSPK,    // where to send the change
+    network = networkParams      
+  )
+
+  val txF: Future[Transaction] = txBuilderF.flatMap(_.sign)
+
+  val tx: Transaction = Await.result(txF, 1 second)
 \end{lstlisting}
+  \caption{Creating a transaction}
+  \label{fig:tx}
+\end{figure}
 
-What happens if we have two TransactionOutPoints (previousOutputs) with a different index but referencing the same Transaction ID (txId)?
+Now that we have a transaction output, we create a transaction to
+spend it.
 
-According to the Bitcoin protocol this is possible.  A transaction can
-have multiple outputs that should be referenceable by the next
-transaction.  So this is clearly a bug.
+First, we need a so-called \emph{out point}, which points to an output
+of a previous transaction. The second parameter is the index of the
+referenced output. Here we use zero because the previous transaction
+has only one output (with index zero).
 
-What should not be possible is a transaction referencing the same
-output twice.  This bug occurred in Bitcoin Core known as
-CVE-2018–17144 which was patched on September 18,
-2018. \cite{cve201817144}
+Second, we need some information on how to spend the utxo from the
+previous transaction, in particular, how to sign the transaction to
+allow this. That is the information needed to create a transaction
+input, so essentially a transaction input.
 
-Here, Bitcoin-S did a bit too much and marked all transaction as
-invalid, if they referenced the same transaction twice.  The fix is,
-to check on TransactionOutPoint instead of TransactionOutPoint.txId,
-because TransactionOutPoint contains the txId as well as the output
-index it references.  So in pseudo code, we check on the tuple (tx,
-index) instead of (tx).  The fixed code:
-\begin{lstlisting}[style=scala]
+Third, we assemble the list of transaction inputs, in our case just one.
+
+Fourth, we set the amount of satoshis that we want to spend. The Int64
+class emulates C data types in Scala, and we will look at it more in
+the next section.
+
+Fifth, we create a lock script to send the coins to. To be supremely
+secure, we use a public key derived from a private key we immediately
+forget.
+
+Sixth, we create our list of transaction outputs.
+
+Seventh and eighth, we define the fee rate and some bitcoin network parameters.
+
+Nineth, we create a transaction builder with those data.
+
+Tenth, we tell the transaction builder to start signing the transaction.
+
+Finally, we get the actual signed transaction. We could serialize it
+and send it to the Bitcoin network. We can also pass it to the
+checkTransaction function, which will return true.
+
+\paragraph{A Bug in the checkTransaction function.} Note lines 16 and
+17 of the checkTransaction function in Figure~\ref{fig:checktrans}.
+
+The value prevOutputTxIds gathers all transaction IDs referenced by
+the inputs of the current transaction.  When we call \emph{distinct}
+on the returned list, we get the list with duplicates removed.  If the
+size of the new list is the same as the size of the old, we know that
+there is no duplicate transaction ID.
+
+However, this code prevents a transaction with two inputs spending two
+different outputs of a previous transaction. The check is too strict:
+checkTransaction returns false for valid transactions.
+
+The fix is simple: we perform the duplicate check on the
+TransactionOutPoints instead of on their transaction IDs.  A
+TransactionOutPoint contains the txId as well as the output index it
+references.
+
+Specifically, we replace lines 16-17 as follows:
+\begin{lstlisting}[style=scala, firstnumber=16]
   val prevOutputs = transaction.inputs.map(_.previousOutput)
@@
 
-Since TransactionOutPoint is a case class and Scala has a built in ==
-for case classes there is no need to implement TransactionOutPoint.==.
+Note that TransactionOutPoint is a case class and thus has a built in
+== method.
 
-This was fixed in
-\href{https://github.com/bitcoin-s/bitcoin-s/pull/435}{pull request
-  number 435} on GitHub at April 23, 2019, through this work along
-with a unit test to prevent this bug from appearing again in the
-future.
+We submitted the fix together with a unit test to prevent this bug
+from appearing again in the future in a pull request
+\cite{BitcoinS:pull435}, which has already been merged.
+
+\paragraph{An Attempt at Verification.} Naively trying Stainless on
+the entire bitcoin-s codebase results in many errors -- as was to be
+expected. We tried to extract only the relevant code and to verify
+that. However, even the extracted code has more than 1500 lines and
+liberally uses Scala features outside of the supported fragment.  We
+tried to transform the code into the supported fragment, but quickly
+realized that that a better approach is to first verify a simpler
+property with less code involved and later come back to the
+no-inflation property with more experience. So we now turn to the
+addition-with-zero property.
 
@@
 
-\label{chap:verify_add}
-
 In Bitcoin-S there is a class \texttt{Satoshis} representing an amount of bitcoins.
-We look at the verification of the addition of Satoshis with zero Satoshis.
-This operation should result in the same amount of Satoshis.
-Let's call it the \nameref{property_2}.
-
-Using Stainless, we see the successful verification of this property.
-But the process of the verification with the tool requires many
-changes in the code, so that Stainless can accept it.  We look at all
-needed modifications in chapter \ref{chap:verify_add}.
-
-After realizing that it would consume too much time to rewrite the
-Bitcoin-S code and even the extracted part with checkTransaction, the
-smallest unit in Bitcoin-S-Core that is worthwhile to verify was
-extracted.  This could be the addition of two CurrencyUnits.  To make
-it even easier, the addition of CurrencyUnits with zero.
-CurrencyUnits is an abstract class in Bitcoin-S, representing
-currencies like Satoshis.
-
-\paragraph{Extracting the relevant Code}
-
-The relevant code is in two files: \texttt{CurrencyUnits.scala} and \texttt{NumberType.scala}.
-
-reference: code/addition/src/main/scala/addition/original
-
-From those files we removed all code that is not needed for our
-verification. For example, we removed all number types except for
-Int64 (so Int32, UInt64, etc.) because we do not use them. We also
-removed the superclasses Factory and NetworkElement of CurrencyUnit
-and Number, respectively because the inherited members are not used by
-the relevant code. Also, we removed all binary operations on Number that are not used.
-
-  - removed extending Factory, NetworkElement and BasicArithmetic. This includes some hex/byte conversion eg fromHex, hex, bytes, fromBytes, ... Just interfaces never referenced
-    \href{https://bitcoin-s.org/docs/core/core-intro\#the-basics}{description in section \#the-basics (cannot add link with hashtag)}
-    \href{https://github.com/bitcoin-s/bitcoin-s/blob/master/core/src/main/scala/org/bitcoins/core/protocol/NetworkElement.scala}{NetworkElement class}
-    \href{https://github.com/bitcoin-s/bitcoin-s/blob/master/core/src/main/scala/org/bitcoins/core/util/Factory.scala}{Factory class}
-
-  - removed Bitcoins class
-
-  - removed subtraction and multiplication, binary operations <<, >>, etc, comparision operator <=, >=, etc but not == and !=
-
-  - removed toBigDecimal
-
-  - removed object CurrencyUnits containing some variables to transform satoshis to btc (not used)
-
-  The code we use for the following sections is in reduced folder.
-
-Here we can see the extracted code needed for the addition of CurrencyUnits:
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala}
 
@@
 
+\paragraph{Extracting the relevant Code} The relevant code is in two
+files: \texttt{CurrencyUnits.scala} and \texttt{NumberType.scala}.
+
+From those files we removed all code that is not needed for our
+verification. For example, we removed all number types except for
+Int64 (so Int32, UInt64, etc.) because we do not use them. We also
+removed the superclasses Factory and NetworkElement of CurrencyUnit
+and Number, respectively, because the inherited members are not used by
+the relevant code. Also, we removed all binary operations on Number
+that are not used, like subtraction and multiplication.
+
+\paragraph{Transforming the Code.}
+
 When we run Stainless on this code (without any properties to prove),
-it throws the following errors: \todo{describe errors: no support for
+it throws the following errors: describe errors: no support for
   abstract types, unsupported arguments for the BigInt constructor,
@@
 
-\paragraph{Transforming the Code.}
-
-\todo{kai todo}
+\todo{...}
 
 \paragraph{The Specification.}
-
-As specified, our verification must only support addition with zero.
-So we restrict the parameter to be zero in the precondition.
-\begin{lstlisting}[style=scala]
-  require(c.satoshis == Satoshis.zero)
-\end{lstlisting}
-
-We ensure the result is the same value as \emph{this} in the postcondition.
-\begin{lstlisting}[style=scala]
-  ensuring(res => res.satoshis == this.satoshis)
-\end{lstlisting}
 
@@
 
-That's all we need to verify our addition.
-
 \paragraph{Result}
@@
 \end{figure}
-
-The verified code.
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/modified/number/NumberType.scala}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/modified/currency/CurrencyUnits.scala}
 
@@
 
+\clearpage
 \appendix
+
 \section{Project Setup}
 
-\paragraph{Using the Stainless-sbt-plugin vs Using the Stanless-JAR}
+\paragraph{Using the Stainless-sbt-plugin vs Using the Stainless-JAR}
 
@@
 otherwise use sbt-plugin because of library management and replicability (everyone uses the same Stainless version configured in sbt)
+no sbt because
+
+-- version conflicts
+-- no output maybe \href{https://github.com/epfl-lara/stainless/issues/484}{issue 484 on
+  GitHub}
 
@@
 
-\subsection{Integration into Bitcoin-S}
+\section{Extracted Code}
+\label{sec:extracted}
 
-During this work, Stainless updated the sbt plugin to support sbt
-1.2.8 from 0.13.17 and Scala 2.12.8 from 2.11.12.  So this section
-might be out of date now.
+\subsection{NumberType.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/number/NumberType.scala}
 
-To use the latest version of the sbt tool you have to build it
-locally.  You can run \texttt{sbt universal:stage} in the cloned
-Stainless git repository.  This generates
-\emph{frontends/scalac/target/universal/stage/bin/stainless-scalac}.
-
-Bitcoin-S-Core uses sbt 1.2.8 and Scala 2.12.8, while Stainless sbt
-plugin is on sbt 0.13.17 and Scala 2.11.12.
-
-Sbt introduced new features in the 1.x release used by Bitcoin-S.
-Most of them can be written the sbt 0.13.17 way.
-
-The bigger problem is, due to the different Scala and sbt versions,
-the following error after trying to go in a sbt shell:
-\begin{verbatim}
-  [warn] There may be incompatibilities among your library dependencies; run 'evicted'
-         to see detailed eviction warnings.
-  [error] java.lang.NoClassDefFoundError: sbt/SourcePosition
-  ...
-  Project loading failed: (r)etry, (q)uit, (l)ast, or (i)gnore?
-\end{verbatim}
-
-Downgrading Bitcoin-S sbt version to 0.13.17 fixes the error but then
-it can not load some libraries only compiled for newer versions.  So
-this would take too much time to fix and changes the Bitcoin-S code
-inadvertently.
-
-The next approach is to use the stainless cli instead of sbt.  Running
-stainless on all source files does not work, because dependencies are
-missing.  The parameter \emph{-classpath} can resolve it but the value
-of this parameter must be the paths to all the dependencies separated
-by a ':'.  Finally, \emph{core} depends on \emph{secp256k1jni},
-another package of Bitcoin-S written in Java.  So this needs to be in
-the source files to.
-
-The final command looks like this in \emph{core} folder of Bitcoin-S:
-\begin{lstlisting}[language=bash]
-  $ stainless
-    -classpath ".:$(find ~/.ivy2/ -type f -name *.jar | tr '\n' ':')"
-    $(find . -type f -name *.scala | tr '\n' ' ')
-    $(find ../secp256k1jni -type f -name *.java | tr '\n' ' ')
-\end{lstlisting}
-
-\emph{.ivy2} is the dependency cache of sbt.  The \emph{tr} replaces
-the first char with the second so a newline with either ':' or ' '.
-
-With this command, Stainless throws the next error:
-\begin{verbatim}
-  [Internal] Error: object scala.reflect.macros.internal.macroImpl in compiler mirror
-             not found.. Trace:
-  [Internal] - scala.reflect.internal.MissingRequirementError$.signal
-             (MissingRequirementError.scala:17)
-  ...
-  [Internal] object scala.reflect.macros.internal.macroImpl in compiler mirror not found.
-  [Internal] Please inform the authors of Inox about this message
-\end{verbatim}
-
-So we can not know how many errors will face us.  Let's go another
-way, because the errors may take too much time and it might lead to a
-next error.  We extract the code needed to verify a transaction mainly
-the class Transaction and ScriptInterpreter with many other classes
-they're depending on.
-
-After this extraction Stainless was successfully integrated with both
-sbt and JAR.
-
-Running \texttt{sbt compile} in the project with Stainless ended
-without error.  But it also ended with no output.  So we are not able
-to change the code so Stainless would accept it since we do not know
-what to change.
-
-So the sbt plugin does not always complain where the JAR file did.
-The open
-\href{https://github.com/epfl-lara/stainless/issues/484}{issue 484 on
-  GitHub} might describe exactly this error.
-
-\section{NumberType.scala}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}
-
-\section{CurrencyUnits.scala}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala}
+\subsection{CurrencyUnits.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/currency/CurrencyUnits.scala}
 
 \section{Code Transformations} 
+\label{sec:transform}
 
@@
 Stainless BigInt constructor is restricted to literal arguments.
-
-After:
-\begin{lstlisting}[style=scala]
-sealed abstract class Satoshis extends CurrencyUnit {
-  override def satoshis: Satoshis = this
-
-  def toBigInt: BigInt = underlying.toBigInt
-
-  def toLong: Long = underlying.toLong
-
-  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
-}
-\end{lstlisting}
 
@@
 
+\section{Verified Code}
+\label{sec:verified}
+
+\subsection{NumberType.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/verified/number/NumberType.scala}
+
+\subsection{CurrencyUnits.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/verified/currency/CurrencyUnits.scala}
+
 \end{document}

### 2019-06-29 paper/paper.tex
more work

@@
 
-In Bitcoin-S there is a class \texttt{Satoshis} representing an amount of bitcoins.
+It is of course a crucial property we are verifying here: if zero
+satoshis were credited to your account, you would not want your
+balance to change! It is also the simplest meaningful property to
+verify that we can think of. However, the code involved in performing
+an addition of two satoshi amounts in bitcoin-s is non-trivial. The
+reason for that is a pecularity of consensus code: agreement with the
+majority is more important than correctness, whatever correctness
+might mean. The most widely used bitcoin implementation by far is the
+reference implementation Bitcoin Core, written in C++. For consensus
+code, bitcoin-s has little choice but to be in strict agreement with
+the reference implementation. To achieve that, it implements C-like
+data types in Scala and then implements functionality using those
+C-like data types. For example, the Satoshis class, which is used to
+represent an amount of satoshis, is implemented using the class Int64
+which represents the C-type \texttt{int64\_t}.
 
-The additions' signature looks like this:
-\begin{lstlisting}[style=scala]
-  +(c: CurrencyUnit): CurrencyUnit
-\end{lstlisting}
+\paragraph{Extracting the relevant Code} The relevant code for the
+addition of satoshis is in two files: \texttt{CurrencyUnits.scala} and
+\texttt{NumberType.scala}.
 
-\paragraph{Extracting the relevant Code} The relevant code is in two
-files: \texttt{CurrencyUnits.scala} and \texttt{NumberType.scala}.
+\begin{figure}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced-nocomments/number/NumberType.scala}
+  \caption{Extracted Code from NumberType.scala}
+  \label{fig:numbertype}
+\end{figure}
 
-From those files we removed all code that is not needed for our
-verification. For example, we removed all number types except for
-Int64 (so Int32, UInt64, etc.) because we do not use them. We also
-removed the superclasses Factory and NetworkElement of CurrencyUnit
-and Number, respectively, because the inherited members are not used by
-the relevant code. Also, we removed all binary operations on Number
-that are not used, like subtraction and multiplication.
+\begin{figure}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced-nocomments/currency/CurrencyUnits.scala}
+  \caption{Extracted Code from CurrencyUnits.scala}
+  \label{fig:currencyunits}
+\end{figure}
+
+From those files we removed all code that is not needed for the
+verification of our property.  For example, we removed all number
+types except for Int64 (so Int32, UInt64, etc.) because they are not
+used. We also removed the superclasses Factory and NetworkElement of
+CurrencyUnit and Number, respectively, because the inherited members
+are not used. Also, we removed all binary operations on Number that
+are not used, like subtraction and multiplication. The extracted code
+is shown in Figure~\ref{fig:numbertype} and
+Figure~\ref{fig:currencyunits}.
 
@@
   abstract types, unsupported arguments for the BigInt constructor,
-  unsupported inheritance for objects.}
+  unsupported inheritance for objects.
 
@@
 
-Before:
+The addition will now look like this:
 \begin{lstlisting}[style=scala]
-sealed abstract class CurrencyUnit {
-  def satoshis: Satoshis
-
-  def !=(c: CurrencyUnit): Boolean = !(this == c)
-
-  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
   def +(c: CurrencyUnit): CurrencyUnit = {
     Satoshis(satoshis.underlying + c.satoshis.underlying)
-  }
-
-  protected def underlying: Int64
-}
-\end{lstlisting}
-
-The addition will now look like this:
-\begin{lstlisting}[style=scala]
-sealed abstract class CurrencyUnit {
-  def satoshis: Satoshis
-
-  def !=(c: CurrencyUnit): Boolean = !(this == c)
-
-  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
-  override def +(c: CurrencyUnit): CurrencyUnit = {
-    require(c.satoshis == Satoshis.zero)
-    Satoshis(satoshis.underlying + c.satoshis.underlying)
-  } ensuring(res => res.satoshis == this.satoshis)
-
-  protected def underlying: Int64
-}
+  } ensuring(res =>
+    (c.satoshis == Satoshis.zero) ==>
+      (res.satoshis == this.satoshis))
 \end{lstlisting}
@@
 plugin does not always report errors.
+use sbt-plugin because of library management and replicability (everyone uses the same Stainless version configured in sbt)
 
@@
 Why JAR:
-sbt-plugin problems (v 0.1 uses older versions (v0.2 not), no output to console see github \#484)
-otherwise use sbt-plugin because of library management and replicability (everyone uses the same Stainless version configured in sbt)
+
 no sbt because
 
--- version conflicts
+-- version conflicts, stainless 0.1 uses older sbt version than bitcoin-s (v0.2 not)
 -- no output maybe \href{https://github.com/epfl-lara/stainless/issues/484}{issue 484 on

### 2019-06-29 code/addition/src/main/scala/reduced/currency/CurrencyUnits.scala
remove comments

@@ -4,34 +4,24 @@ import reduced.number.{BaseNumbers, Int64}
 
 sealed abstract class CurrencyUnit {
   type A
-
   def satoshis: Satoshis
-
   def !=(c: CurrencyUnit): Boolean = !(this == c)
-
   def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
   def +(c: CurrencyUnit): CurrencyUnit = {
     Satoshis(satoshis.underlying + c.satoshis.underlying)
   }
-
   protected def underlying: A
 }
 
 sealed abstract class Satoshis extends CurrencyUnit {
   override type A = Int64
-
   override def satoshis: Satoshis = this
-
   def toBigInt: BigInt = BigInt(toLong)
-
   def toLong: Long = underlying.toLong
-
   def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
 }
 
 object Satoshis extends BaseNumbers[Satoshis] {
-
   val min = Satoshis(Int64.min)
   val max = Satoshis(Int64.max)
   val zero = Satoshis(Int64.zero)

### 2019-07-02 paper/paper.tex
more work on transformations

@@
 \begin{figure}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced-nocomments/number/NumberType.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/number/NumberType.scala}
   \caption{Extracted Code from NumberType.scala}
@@
 \begin{figure}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced-nocomments/currency/CurrencyUnits.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/currency/CurrencyUnits.scala}
   \caption{Extracted Code from CurrencyUnits.scala}
@@
   lazy val max = Int64(BigInt(9223372036854775807L))
-}
 \end{lstlisting}

### 2019-07-03 paper/paper.tex
changes to paper, factor out project setup

@@
 
+% latin modern, no more bitmap fonts in pdf
+\usepackage{lmodern}
 %\usepackage{newtxtext}
@@
   there is too much code involved that lies outside of the supported
-  fragment to make this feasible. However, in the process we uncover
-  and fix a bug in bitcoin-s. We then turn to a much simpler (and less
-  interesting) property: that adding zero satoshis to a given amount
-  of satoshis yields the given amount of satoshis. Here as well a
-  significant part of the relevant code lies outside of the supported
-  fragment. However, after a series of equivalent transformations we
-  arrive at code that we successfully verify.
+  fragment to make this feasible. However, in the process we find and
+  fix a bug in bitcoin-s. We then turn to a much simpler property:
+  that adding zero satoshis to a given amount of satoshis yields the
+  given amount of satoshis. Here as well a significant part of the
+  relevant code lies outside of the supported fragment. We perform a
+  series of equivalent transformations on it until we arrive at code that we
+  successfully verify. In that process we find and fix a second bug in
+  bitcoin-s.
 
@@
 \cite{BitcoinS:website,BitcoinS:github}. In particular, it allows to
-serialize, deserialize, sign and validate transactions. The library
+serialize, deserialize, sign and validate Bitcoin transactions. The library
 uses immutable data structures and algebraic data types but is not
@@
 \end{figure}
-Our program does not satisfy the specification. An overflow in the
+Our function does not satisfy the specification. An overflow in the
 32-bit Int leads to a negative result for the input 17, as Stainless
@@
 to verify it in Section~3. Here as well we see that a significant part
-of the code lies outside of the supported fragment. However, after a
-series of equivalent transformations we arrive at code that we
-successfully verify.
+of the code lies outside of the supported fragment. We perform a
+  series of equivalent transformations on it until we arrive at code that we
+  successfully verify. In that process we find and fix a second bug in
+  bitcoin-s.
 
@@
 
-\paragraph{A Bug in the checkTransaction function.} Note lines 16 and
+\paragraph{A Bug in the checkTransaction Function.} Note lines 16 and
 17 of the checkTransaction function in Figure~\ref{fig:checktrans}.
@@
 
-\paragraph{Extracting the relevant Code} The relevant code for the
+\paragraph{Extracting the Relevant Code} The relevant code for the
 addition of satoshis is in two files: \texttt{CurrencyUnits.scala} and
@@
 Figure~\ref{fig:currencyunits}.
+
+\paragraph{A Bug in the checkTransaction Function.}
 
@@
 
-\section{Project Setup}
+% \section{Extracted Code}
+% \label{sec:extracted}
 
-\paragraph{Using the Stainless-sbt-plugin vs Using the Stainless-JAR}
+% \subsection{NumberType.scala}
+% \lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/number/NumberType.scala}
 
-We can either use the sbt plugin or a JAR file to check code with Stainless.
-
-More about sbt on https://www.scala-sbt.org/1.x/docs/index.html (features are also listed there)
-Installation see Stainless Doc 'Installing Stainless' section
-
-Invoking the JAR on our source code Stainless will verify it.  If we
-have a bigger project, this becomes really tricky, because we must
-pass all files needed including the dependencies.  This is in contrast
-to the sbt plugin, where we can integrate Stainless in our compilation
-process.  When we call compile, Stainless verifies the code and stops
-the compilation if the verification fails.
-
-Having a static version configured in the sbt build file, every
-developer has the same Stainless features available.  This should
-prevent incompatibility with new or deprecated features when we use
-different plugins.
-
-So the sbt plugin has clear advantages over the JAR file since its
-integrated directly.  We do not have to download it manually and find
-the right version and if we bump the version we can just edit it in
-the build file and every developer is on the same version again.
-
-However, currently there are some drawbacks.  For example the sbt
-plugin does not always report errors.
-use sbt-plugin because of library management and replicability (everyone uses the same Stainless version configured in sbt)
-
-We use the jar for everything.
-
-Use JAR:
-Download all library dependencies including dependencies of dependencies, pass libraries + source to JAR
-
-Use sbt-plugin: integrate in sbt build tool with addSbtPlugin() method
-
-Why JAR:
-
-no sbt because
-
--- version conflicts, stainless 0.1 uses older sbt version than bitcoin-s (v0.2 not)
--- no output maybe \href{https://github.com/epfl-lara/stainless/issues/484}{issue 484 on
-  GitHub}
-
-\paragraph{Reproducing our Results.}
-
-Prerequisites:
-\begin{enumerate}
-\item Java 8 in /usr/lib/jvm/java-8-openjdk (Stainless requires and Scala recommends Java SE Development Kit 8. Newer Java versions won't work.)
-\end{enumerate}
-
-\begin{lstlisting}[language=bash]
- clone repo
- run ./code/bin/stainless.sh <path to code>
-   eg. ./code/bin/stainless.sh code/addition/src/main/scala/addition/reduced/
-\end{lstlisting}
-
-\section{Extracted Code}
-\label{sec:extracted}
-
-\subsection{NumberType.scala}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/number/NumberType.scala}
-
-\subsection{CurrencyUnits.scala}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/currency/CurrencyUnits.scala}
+% \subsection{CurrencyUnits.scala}
+% \lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/currency/CurrencyUnits.scala}
 
@@
 
-\section{Verified Code}
-\label{sec:verified}
+% \section{Verified Code}
+% \label{sec:verified}
 
-\subsection{NumberType.scala}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/verified/number/NumberType.scala}
+% \subsection{NumberType.scala}
+% \lstinputlisting[style=scala]{../code/addition/src/main/scala/verified/number/NumberType.scala}
 
-\subsection{CurrencyUnits.scala}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/verified/currency/CurrencyUnits.scala}
+% \subsection{CurrencyUnits.scala}
+% \lstinputlisting[style=scala]{../code/addition/src/main/scala/verified/currency/CurrencyUnits.scala}
 

### 2019-07-03 paper/paper.tex
changes to paper

@@
 
-What will turn out to be more relevant are the Scala features which
-Stainless does not support, such as: (concrete) class definitions,
-inheritance by objects, abstract type members, and inner classes in
-case objects.
+What will turn out to be more relevant for us are the Scala features
+which Stainless does not support, such as: (concrete) class
+definitions, inheritance by objects, abstract type members, and inner
+classes in case objects.
 
@@
 general do not have all the methods of the Scala data types. For
-example, the BigInt type in Scala has a methods for bitwise operations
+example, the BigInt type in Scala has methods for bitwise operations
 while the BigInt type in Stainless does not.
@@
 
+why fix the preconditions:
+\begin{itemize}
+\item all green: new errors stand out more
+\item at runtime, errors are caught earlier, which is preferrable
+\end{itemize}
+
 \paragraph{The Specification.}
@@
 The addition will now look like this:
-\begin{lstlisting}[style=scala]
-  def +(c: CurrencyUnit): CurrencyUnit = {
-    Satoshis(satoshis.underlying + c.satoshis.underlying)
-  } ensuring(res =>
-    (c.satoshis == Satoshis.zero) ==>
-      (res.satoshis == this.satoshis))
-\end{lstlisting}
+\lstinputlisting[style=scala, firstline=15, lastline=18, firstnumber=15]{../code/addition/src/main/scala/verified/currency/CurrencyUnits.scala}
 

### 2019-07-03 code/addition/src/main/scala/reduced/currency/CurrencyUnits.scala
minimize code

@@ -5,7 +5,6 @@ import reduced.number.{BaseNumbers, Int64}
 sealed abstract class CurrencyUnit {
   type A
   def satoshis: Satoshis
-  def !=(c: CurrencyUnit): Boolean = !(this == c)
   def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
   def +(c: CurrencyUnit): CurrencyUnit = {
     Satoshis(satoshis.underlying + c.satoshis.underlying)
@@ -22,10 +21,7 @@ sealed abstract class Satoshis extends CurrencyUnit {
 }
 
 object Satoshis extends BaseNumbers[Satoshis] {
-  val min = Satoshis(Int64.min)
-  val max = Satoshis(Int64.max)
   val zero = Satoshis(Int64.zero)
-  val one = Satoshis(Int64.one)
 
   def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
 

### 2019-07-04 code/addition/src/main/scala/extracted/currency/CurrencyUnits.scala
further reduce code

@@ -1,6 +1,6 @@
-package reduced.currency
+package extracted.currency
 
-import reduced.number.{BaseNumbers, Int64}
+import extracted.number.{BaseNumbers, Int64}
 
 sealed abstract class CurrencyUnit {
   type A
@@ -22,8 +22,6 @@ sealed abstract class Satoshis extends CurrencyUnit {
 
 object Satoshis extends BaseNumbers[Satoshis] {
   val zero = Satoshis(Int64.zero)
-
   def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
-
   private case class SatoshisImpl(underlying: Int64) extends Satoshis
 }

### 2019-07-04 paper/paper.tex
further reduce code

@@
 \end{lstlisting}
-	\caption{Factorial program with specification}
+	\caption{Factorial function with specification}
 	\label{fig:factorial}
@@
 		\includegraphics[width=\textwidth]{output1.png}
-	\caption{Stainless output for the factorial program}
+	\caption{Stainless output for the factorial function}
 	\label{fig:failed}
@@
 
-\paragraph{Extracting the Relevant Code} The relevant code for the
-addition of satoshis is in two files: \texttt{CurrencyUnits.scala} and
-\texttt{NumberType.scala}.
-
 \begin{figure}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/number/NumberType.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/extracted/number/NumberType.scala}
   \caption{Extracted Code from NumberType.scala}
@@
 \begin{figure}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/currency/CurrencyUnits.scala}
+\lstinputlisting[style=scala]{../code/addition/src/main/scala/extracted/currency/CurrencyUnits.scala}
   \caption{Extracted Code from CurrencyUnits.scala}
@@
 
-From those files we removed all code that is not needed for the
-verification of our property.  For example, we removed all number
-types except for Int64 (so Int32, UInt64, etc.) because they are not
-used. We also removed the superclasses Factory and NetworkElement of
-CurrencyUnit and Number, respectively, because the inherited members
-are not used. Also, we removed all binary operations on Number that
-are not used, like subtraction and multiplication. The extracted code
-is shown in Figure~\ref{fig:numbertype} and
-Figure~\ref{fig:currencyunits}.
+\paragraph{Extracting the Relevant Code} The relevant code for the
+addition of satoshis is in two files: \texttt{CurrencyUnits.scala} and
+\texttt{NumberType.scala}. From those files we removed the majority of
+the code because it is not needed for the verification of our
+property.  For example, we removed all number types except for Int64
+(so Int32, UInt64, etc.) because they are not used. We also removed
+the superclasses Factory and NetworkElement of CurrencyUnit and
+Number, respectively, because the inherited members are not
+used. Also, we removed all binary operations on Number that are not
+used, like subtraction and multiplication. The extracted code is shown
+in Figure~\ref{fig:numbertype} and Figure~\ref{fig:currencyunits}.
 
@@
 
-\paragraph{Transforming the Code.}
+\paragraph{Transforming the Code.} Here we see how to transform the
+bitcoin-s code into the Scala fragment supported by Stainless.  All
+transformations are equivalent in the sense that if the
+addition-with-zero property holds for the transformed code, then it
+also holds for the code before the transformation. Here is the list of
+features used by the extracted code but unsupported by Stainless, and
+how to transform the code:
 
-When we run Stainless on this code (without any properties to prove),
-it throws the following errors: describe errors: no support for
-  abstract types, unsupported arguments for the BigInt constructor,
-  unsupported inheritance for objects.
+\begin{enumerate}
+\item Inheriting Objects. Here, we can just turn the objects into case
+  objects by literally just changing the word \texttt{object} into
+  \texttt{case object} on lines 65 and 33 in the two respective files.
 
-\todo{...}
+That transformation is equivalent. Case objects have some additional
+properties (in particular, being serializable) and they inherit from
+\texttt{Product} instead of \texttt{AnyRef}, but none of our code
+depends on any of that.
 
-why fix the preconditions:
-\begin{itemize}
-\item all green: new errors stand out more
-\item at runtime, errors are caught earlier, which is preferrable
-\end{itemize}
+\item Abstract Type Members.
+
+Note that we can not simply replace the unsupported abstract type
+member by a (supported) type parameter. The problem is that the
+CurrencyUnit class uses one of its implementing classes: Satoshis.
+
+Satoshis would have to instantiate a potential type parameter with
+type Int64, so it would extend CurrencyUnit[Int64]. But that is too
+specific, because the return type of the \texttt{+}-method would then then be
+CurrencyUnit[Int64] not CurrencyUnit[A].
+
+Since we only want to verify the addition of satoshis, and the
+Satoshis class overrides A with Int64 anyway, we just remove the
+abstract type and set it to Int64.
+
+We remove line 6 and line 22 from CurrencyUnits.scala (to maintain
+line numbers we actually replace them with empty lines for now) and in
+line 18 we replace A by Int64.
+
+\item{Non-Literal BigInt Constructor Argument}
+
+As described before, the types in the Stainless library are more
+restricted than their Scala library counterparts. In particular, the
+Stainless BigInt constructor is restricted to literal arguments.
+
+Here we can simply use toBigInt on the field \texttt{underlying}
+directly.  So, instead of converting the underlying to Long and back
+to BigInt we convert underlying directly to BigInt.
+
+We replace line 26 by
+\begin{lstlisting}[style=scala]
+def toBigInt: BigInt = underlying.toBigint
+\end{lstlisting}
+
+This is an equivalent transformation: the only thing that might go
+wrong in the detour via \texttt{Long} is that the BigInt underlying
+Int64 in turn underlying Satoshis does not fit into a Long. However,
+the only constructor of Int64 ensures exactly that.
+
+\item{Self-Reference in Type Parameter Bound}
+
+Stainless does not currently support a class with a type parameter
+with a type boundary that contains the type parameter itself. We opened an
+issue \cite{Stainless:issue519} on the Stainless repository and the
+Stainless developers have targeted Stainless version 0.4 to support
+such self-referential type boundaries.
+
+For now, since our code only uses Number with type parameter T
+instantiated to Int64, we just remove the type parameter and replace
+it by Int64. We respectively replace lines 8, 44 and 49 by
+\begin{lstlisting}[style=scala]
+sealed abstract class Number {
+\end{lstlisting}
+
+\begin{lstlisting}[style=scala]
+sealed abstract class SignedNumber extends Number
+\end{lstlisting}
+
+\begin{lstlisting}[style=scala]
+sealed abstract class Int64 extends SignedNumber {
+\end{lstlisting}
+
+and replace T by Int64 in lines 25-27.
+
+\item{Missing Member \texttt{bigInteger} in BigInt}
+
+The Scala class BigInt is essentially a wrapper around
+java.math.BigInteger. BigInt has a member bigInteger which is the
+underlying instance of the Java class. The Java class has a method
+longValueExact which returns a long only if the BigInteger fits into a
+long, otherwise throws exception. Stainless does not support Java
+classes and in particular its BigInt has no member bigInteger.
+
+However, our code never calls toLong anymore, so we just remove it. We
+replace line 14 in NumberType.scala and line 28 in CurrencyUnits.scala
+by an empty line.
+
+\item{Type Member}
+
+Our version of Stainless does not support type members. We just
+replace all occurrence of A with BigInt, since A is never overwritten
+in an implementing class.
+
+We remove line 9 in NumberType.scala and replace A by BigInt in lines
+12, 25, 33 and 50.
+
+In the mean time Stainless has implemented support for type member
+\cite{Stainless:pull470}.  Since version 0.2 verification should
+succeed without this change.
+
+\item{Missing Bitwise-And Method on BigInt}
+
+Contrary to Scala BigInt, the Stainless BigInt class does not support
+bitwise operations, in particular not the \&-method for bitwise and.
+
+The bitwise and with the andMask on line 34 is a bounds check.  It
+checks if the result parameter is in range of the specified type,
+which in our case is the hard coded Int64.
+
+So, we replace the \& mask with a check whether the result
+is in range of Long.MinValue and Long.MaxValue, because Int64 has the
+same 64-bit range as Long.
+
+We replace lines 34-35 by the following, squeezed into two lines to
+preserve line numbers. Note that the BigInt constructor requires a
+literal:
+
+\begin{lstlisting}[style=scala]
+require(result <= BigInt("9223372036854775807")  
+  && result >= BigInt("-9223372036854775808"),
+      "Result was out of bounds, got: " + result)
+\end{lstlisting}
+
+\item{Inner Class in Case Object}
+
+Stainless does not support inner classes in a case object. Bitcoin-s
+uses this a lot to separate the class from its implementation.
+
+This is easy to fix. We just move the inner classes out of the case objects. They do
+not interfere with any other code.
+
+We remove lines 66-71 in NumberType.scala and insert them at the end of the file.
+We remove line 42 in CurrencyUnits.scala and insert it at the end of the file.
+
+\item{Message Parameter in Require}
+
+Here the require(condition, message) is used as an assertion: if the
+condition is false the fail with message.  Stainless does not support
+the message parameter of require.  For the verification we can simply
+remove that parameter.
+
+\item{Missing Implicit Long to BigInt Conversion}
+
+This error message does not specify a line number and it is not clear
+what ``inv'' is. However, the Scala BigInt has implict conversions
+from Long and they are missing in the Stainless BigInt.
+
+We replace the two require clauses at lines 67 and following in NumberType.scala
+
+\begin{lstlisting}[style=scala]
+  require(underlying >= -9223372036854775808L)
+  require(underlying <= 9223372036854775807L)
+\end{lstlisting}
+
+by:
+\begin{lstlisting}[style=scala]
+  require(underlying >= BigInt(-9223372036854775808L))
+  require(underlying <= BigInt(9223372036854775807L))
+\end{lstlisting}
+
+We also replace lines 76 and 77
+\begin{lstlisting}[style=scala]
+  lazy val min = Int64(-9223372036854775808L)
+  lazy val max = Int64(9223372036854775807L)
+\end{lstlisting}
+
+by:
+\begin{lstlisting}[style=scala]
+  lazy val min = Int64(BigInt(-9223372036854775808L))
+  lazy val max = Int64(BigInt(9223372036854775807L))
+\end{lstlisting}
+
+\item{Missing BigInt Constructor with Long Argument}
+
+Here again, the Scala BigInt has a constructor with a Long argument
+which is missing in the Stainless BigInt.
+
+We simply replace the Long values in a BigInt constructor call with a string
+literal.
+
+The lines from the previous transformation respectively change into
+the following:
+
+\begin{lstlisting}[style=scala]
+  require(underlying >= BigInt("-9223372036854775808"))
+  require(underlying <= BigInt("9223372036854775807"))
+\end{lstlisting}
+
+and
+
+\begin{lstlisting}[style=scala]
+  lazy val min = Int64(BigInt("-9223372036854775808"))
+  lazy val max = Int64(BigInt("9223372036854775807"))
+}
+\end{lstlisting}
+\end{enumerate}
 
@@
 		\includegraphics[width=\textwidth]{result_output}
-	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
+	\caption{Stainless output for the transformed code}
 \end{figure}
@@
 \appendix
-
-% \section{Extracted Code}
-% \label{sec:extracted}
-
-% \subsection{NumberType.scala}
-% \lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/number/NumberType.scala}
-
-% \subsection{CurrencyUnits.scala}
-% \lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/currency/CurrencyUnits.scala}
 
@@
 \end{lstlisting}
-\todo{stainless error emssage missing}
 
@@
 
-\todo{kai here we do not know if it changes semantic. ramon: oh, we don't?}
-
 \subsection{Inner Class in Case Object}
@@
 Stainless does not support inner classes in a case object. Bitcoin-s
-uses this a lot to separate the class \todo{ramon: i don't understand} from its implementation.
+uses this a lot to separate the class from its implementation.
 
@@
 
-% \section{Verified Code}
-% \label{sec:verified}
-
-% \subsection{NumberType.scala}
-% \lstinputlisting[style=scala]{../code/addition/src/main/scala/verified/number/NumberType.scala}
-
-% \subsection{CurrencyUnits.scala}
-% \lstinputlisting[style=scala]{../code/addition/src/main/scala/verified/currency/CurrencyUnits.scala}
-
 \end{document}

### 2019-07-04 code/addition/src/main/scala/extracted/number/NumberType.scala

@@ -1,4 +1,4 @@
-package reduced.number
+package extracted.number
 
 sealed abstract class Number[T <: Number[T]] {
   type A = BigInt
@@ -36,7 +36,6 @@ object Int64 extends BaseNumbers[Int64] {
   }
 
   lazy val zero = Int64(0)
-
   def apply(long: Long): Int64 = Int64(BigInt(long))
   def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
 }

### 2019-07-04 code/addition/src/main/scala/verified/currency/CurrencyUnits.scala
more work

@@ -5,28 +5,22 @@ import stainless.lang._
 
 sealed abstract class CurrencyUnit {
   def satoshis: Satoshis
-
   def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
-
   def +(c: CurrencyUnit): CurrencyUnit = {
     Satoshis(satoshis.underlying + c.satoshis.underlying)
   } ensuring (res =>
     (c == Satoshis.zero) ==> (res == this))
-
   protected def underlying: Int64
 }
 
 sealed abstract class Satoshis extends CurrencyUnit {
   override def satoshis: Satoshis = this
-
   def toBigInt: BigInt = underlying.toBigInt
-
   def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
 }
 
 case object Satoshis extends BaseNumbers[Satoshis] {
   val zero = Satoshis(Int64.zero)
-
   def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
 }
 

### 2019-07-04 paper/paper.tex
more work

@@
 \begin{abstract}
-  We try to verify properties of the bitcoin-s library, a Scala
+  We try to verify some properties of the bitcoin-s library, a Scala
   implementation of parts of the Bitcoin protocol. We use the
   Stainless verifier which supports programs in a subset of Scala
-  called \emph{Pure Scala}. We first try to verify the property that
-  regular transactions do not create new money. It turns out that
-  there is too much code involved that lies outside of the supported
-  fragment to make this feasible. However, in the process we find and
-  fix a bug in bitcoin-s. We then turn to a much simpler property:
-  that adding zero satoshis to a given amount of satoshis yields the
-  given amount of satoshis. Here as well a significant part of the
-  relevant code lies outside of the supported fragment. We perform a
-  series of equivalent transformations on it until we arrive at code that we
-  successfully verify. In that process we find and fix a second bug in
+  called \emph{Pure Scala}.  Since bitcoin-s is not written in this
+  fragment, we extract the relevant code from it and perform a series
+  of equivalent transformations until we arrive at code that we
+  successfully verify. In that process we find and fix two bugs in
   bitcoin-s.
 
-\keywords{Bitcoin  \and Scala \and Bitcoin-S \and Stainless.}
+\keywords{Bitcoin  \and Scala \and bitcoin-s \and Stainless.}
 \end{abstract}
@@
 parameters, pattern matching, local and anonymous classes and more.
-
 In addition to Pure Scala Stainless also supports some imperative
@@
 Also, Stainless has its own library of some core data types and
-functions which need to be mapped correctly to functions inside of the
-SMT solver that Stainless ultimately relies on. Those data types in
-general do not have all the methods of the Scala data types. For
-example, the BigInt type in Scala has methods for bitwise operations
-while the BigInt type in Stainless does not.
+functions which are mapped to corresponding data types and functions
+inside of the SMT solver that Stainless ultimately relies on. Those
+data types in general do not have all the methods of the Scala data
+types. For example, the BigInt type in Scala has methods for bitwise
+operations while the BigInt type in Stainless does not.
 
@@
 \begin{lstlisting}[style=scala]
-/**
-  * Checks the validity of a transaction in accordance to bitcoin core's CheckTransaction function
-  * https://github.com/bitcoin/bitcoin/blob/f7a21dae5dbf71d5bc00485215e84e6f2b309d0a/src/main.cpp#L939.
-  */
 def checkTransaction(transaction: Transaction): Boolean = {
@@
     !(transaction.inputs.isEmpty || transaction.outputs.isEmpty)
-  val txNotLargerThanBlock = transaction.bytes.size < Consensus.maxBlockSize
-  val outputsSpendValidAmountsOfMoney = !transaction.outputs.exists(o =>
+  val txNotLargerThanBlock = 
+    transaction.bytes.size < Consensus.maxBlockSize
+  val outputsSpendValidAmountsOfMoney = 
+    !transaction.outputs.exists(o =>
       o.value < CurrencyUnits.zero || o.value > Consensus.maxMoney)
@@
     outputValues.fold(CurrencyUnits.zero)(_ + _)
-  val allOutputsValidMoneyRange = validMoneyRange(totalSpentByOutputs)
+  val allOutputsValidMoneyRange = 
+    validMoneyRange(totalSpentByOutputs)
   val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
-  val noDuplicateInputs = prevOutputTxIds.distinct.size == prevOutputTxIds.size
+  val noDuplicateInputs = 
+    prevOutputTxIds.distinct.size == prevOutputTxIds.size
 
@@
     case false =>
-      //since this is not a coinbase tx we cannot have any empty previous outs inside of inputs
-      !transaction.inputs.exists(_.previousOutput == EmptyTransactionOutPoint)
+      !transaction.inputs.exists(
+        _.previousOutput == EmptyTransactionOutPoint)
   }
-  inputOutputsNotZero && txNotLargerThanBlock && outputsSpendValidAmountsOfMoney && noDuplicateInputs &&
-  allOutputsValidMoneyRange && noDuplicateInputs && isValidScriptSigForCoinbaseTx
+  inputOutputsNotZero && txNotLargerThanBlock && 
+  outputsSpendValidAmountsOfMoney && noDuplicateInputs &&
+  allOutputsValidMoneyRange && noDuplicateInputs && 
+  isValidScriptSigForCoinbaseTx
 }
@@
 A crucial function for the verification of the no-inflation property
-is the function which validates transactions: the
-\texttt{checkTransaction} function in the ScriptInterpreter object,
-shown in Figure~\ref{fig:checktrans}. It takes a transaction and if
-some basic checks succeed returns true, otherwise false. For example,
-one of those checks is that both the list of inputs and list of
-outputs need to be non-empty.
+is the \texttt{checkTransaction} function shown in
+Figure~\ref{fig:checktrans}. Given a transaction it returns true if
+some basic checks succeed, otherwise false. For example, one of those
+checks is that both the list of inputs and list of outputs need to be
+non-empty.
 
@@
 then the amount of satoshis, then a transaction output (utxo) for that
-amount and locked with that script. Finally we create the actual
-transaction with that output and no inputs. Of course, that is not a
-valid transaction, because it creates coins out of nothing. In
-particular, checkTransaction(prevTx) returns false, simply because the
-list of inputs is empty.
+amount and locked with that script. Finally we create a transaction
+with that output and no inputs. Of course, that is not a valid
+transaction, because it creates coins out of nothing. In particular,
+\texttt{checkTransaction(prevTx)} returns false, simply because the list of
+inputs is empty.
 
@@
 Now that we have a transaction output, we create a transaction to
-spend it.
+spend it in Figure~\ref{fig:tx}. First, we need a reference to an
+output of a previous transaction, here called
+\texttt{outPoint}. Second, we add some information on how to spend
+that output, in particular, how to sign the transaction to allow
+this. Now we assemble the list of unspent transaction outputs (utxos),
+in our case just one.
 
-First, we need a so-called \emph{out point}, which points to an output
-of a previous transaction. The second parameter is the index of the
-referenced output. Here we use zero because the previous transaction
-has only one output (with index zero).
+We then set the amount of satoshis that we want to spend. The Int64
+class aims to emulate a C data type in Scala, and we will look at it
+more closely in the next section.
 
-Second, we need some information on how to spend the utxo from the
-previous transaction, in particular, how to sign the transaction to
-allow this. That is the information needed to create a transaction
-input, so essentially a transaction input.
+We then create a lock script (\texttt{destinationSPK}) to receive the
+coins, create our list of transaction outputs (\texttt{destinations}),
+define the fee rate and set some bitcoin network parameters.
 
-Third, we assemble the list of transaction inputs, in our case just one.
-
-Fourth, we set the amount of satoshis that we want to spend. The Int64
-class emulates C data types in Scala, and we will look at it more in
-the next section.
-
-Fifth, we create a lock script to send the coins to. To be supremely
-secure, we use a public key derived from a private key we immediately
-forget.
-
-Sixth, we create our list of transaction outputs.
-
-Seventh and eighth, we define the fee rate and some bitcoin network parameters.
-
-Nineth, we create a transaction builder with those data.
-
-Tenth, we tell the transaction builder to start signing the transaction.
+Now we create a transaction builder with those data and we tell it to
+start signing the transaction in line 34.
 
@@
 
-\paragraph{A Bug in the checkTransaction Function.} Note lines 16 and
-17 of the checkTransaction function in Figure~\ref{fig:checktrans}.
-
-The value prevOutputTxIds gathers all transaction IDs referenced by
-the inputs of the current transaction.  When we call \emph{distinct}
-on the returned list, we get the list with duplicates removed.  If the
-size of the new list is the same as the size of the old, we know that
-there is no duplicate transaction ID.
-
-However, this code prevents a transaction with two inputs spending two
-different outputs of a previous transaction. The check is too strict:
-checkTransaction returns false for valid transactions.
+\paragraph{A Bug in the checkTransaction Function.} Note lines 15-17
+of the checkTransaction function in Figure~\ref{fig:checktrans}. Here,
+value prevOutputTxIds gathers a list of all transaction indentifiers
+referenced by the inputs of the current transaction. If the size of
+this list is the same as the size of this list with duplicates
+removed, we know that no transaction has been referenced twice. This
+prevents a transaction from spending two different outputs of the same
+previous transaction. The check is too strict: checkTransaction
+returns false for valid transactions.
 
 The fix is simple: we perform the duplicate check on the
-TransactionOutPoints instead of on their transaction IDs.  A
+TransactionOutPoints instead of on their transaction identifiers.  A
 TransactionOutPoint contains the txId as well as the output index it
-references.
+references. Note that TransactionOutPoint is a case class and thus has
+a built in == method.
 
-Specifically, we replace lines 16-17 as follows:
-\begin{lstlisting}[style=scala, firstnumber=16]
+Specifically, we replace lines 15-17 as follows:
+\begin{lstlisting}[style=scala, firstnumber=15]
   val prevOutputs = transaction.inputs.map(_.previousOutput)
-  val noDuplicateInputs = prevOutputs.distinct.size == prevOutputs.size
+  val noDuplicateInputs = 
+    prevOutputs.distinct.size == prevOutputs.size
 \end{lstlisting}
 
-Note that TransactionOutPoint is a case class and thus has a built in
-== method.
-
-We submitted the fix together with a unit test to prevent this bug
-from appearing again in the future in a pull request
-\cite{BitcoinS:pull435}, which has already been merged.
+We submitted this fix together with a corresponding unit test to the
+bitcoin-s project in a pull request, which has been merged
+\cite{BitcoinS:pull435}.
 
@@
 the entire bitcoin-s codebase results in many errors -- as was to be
-expected. We tried to extract only the relevant code and to verify
-that. However, even the extracted code has more than 1500 lines and
-liberally uses Scala features outside of the supported fragment.  We
-tried to transform the code into the supported fragment, but quickly
-realized that that a better approach is to first verify a simpler
-property with less code involved and later come back to the
-no-inflation property with more experience. So we now turn to the
-addition-with-zero property.
+expected. We tried to extract only the code relevant to the
+no-inflation-property and to verify that. However, even the extracted
+code has more than 1500 lines and liberally uses Scala features
+outside of the supported fragment.  We tried to transform the code
+into the supported fragment, but quickly realized that a better
+approach is to first verify a simpler property with less code involved
+and later come back to the no-inflation property with more
+experience. So we now turn to the addition-with-zero property.
 
@@
 verify that we can think of. However, the code involved in performing
-an addition of two satoshi amounts in bitcoin-s is non-trivial. The
+the addition of two satoshi amounts in bitcoin-s is non-trivial. The
 reason for that is a pecularity of consensus code: agreement with the
@@
 data types in Scala and then implements functionality using those
-C-like data types. For example, the Satoshis class, which is used to
-represent an amount of satoshis, is implemented using the class Int64
-which represents the C-type \texttt{int64\_t}.
+C-like data types. For example, the Satoshis class, which represents
+an amount of satoshis, is implemented using the class Int64 which
+aims to represent the C-type \texttt{int64\_t}.
 
@@
 \paragraph{Extracting the Relevant Code} The relevant code for the
-addition of satoshis is in two files: \texttt{CurrencyUnits.scala} and
-\texttt{NumberType.scala}. From those files we removed the majority of
-the code because it is not needed for the verification of our
-property.  For example, we removed all number types except for Int64
-(so Int32, UInt64, etc.) because they are not used. We also removed
-the superclasses Factory and NetworkElement of CurrencyUnit and
-Number, respectively, because the inherited members are not
-used. Also, we removed all binary operations on Number that are not
-used, like subtraction and multiplication. The extracted code is shown
-in Figure~\ref{fig:numbertype} and Figure~\ref{fig:currencyunits}.
+addition of satoshis is in two files: CurrencyUnits.scala and
+NumberType.scala. From those files we removed the majority of the code
+because it is not needed for the verification of our property.  For
+example, we removed all number types except for Int64 (so Int32,
+UInt64, etc.) because they are not used. We also removed the
+superclasses Factory and NetworkElement of CurrencyUnit and Number,
+respectively, because the inherited members are not used. Also, we
+removed all binary operations on Number that are not used, like
+subtraction and multiplication. The extracted code is shown in
+Figure~\ref{fig:numbertype} and Figure~\ref{fig:currencyunits}.
 
-\paragraph{A Bug in the checkTransaction Function.}
+\paragraph{A Bug in the checkResult Function.} Note the checkResult
+function on line 12 and the value andMask on line 23 of
+NumberType.scala. The function is intended to catch overflows by
+performing a bitwise conjunction of its argument with andMask and
+comparing the result with the argument. However, because of the way
+Java BigIntegers are represented \cite{wikipedia:twocomp} and because
+bitwise operations implicitly perform a sign extension
+\cite{java:bigint} on the shorter operand, the function does not
+actually catch overflows.
 
-\paragraph{Transforming the Code.} Here we see how to transform the
-bitcoin-s code into the Scala fragment supported by Stainless.  All
-transformations are equivalent in the sense that if the
+While this is a potentially serious bug, it turns out that checkResult
+is only ever called inside a constructor call for a number type (like
+Int64) which contains the intended range check, see lines 32-35. The
+checkResult function thus can, and should, be removed entirely. The
+bitcoin-s developers have acknowledged the bug and we submitted a pull
+request to fix it \cite{BitcoinS:pull565}.
+
+\paragraph{Transforming the Code.} We now turn to the list of Scala
+features used by the extracted code which are not supported by
+Stainless and how to transform the code into the supported fragment.
+All transformations are equivalent in the sense that if the
 addition-with-zero property holds for the transformed code, then it
-also holds for the code before the transformation. Here is the list of
-features used by the extracted code but unsupported by Stainless, and
-how to transform the code:
+also holds for the code before the transformation.
 
-\begin{enumerate}
-\item Inheriting Objects. Here, we can just turn the objects into case
-  objects by literally just changing the word \texttt{object} into
-  \texttt{case object} on lines 65 and 33 in the two respective files.
+\paragraph{Inheriting Objects.} In both files we have objects
+extending the BaseNumbers trait, on lines 30 and 23 respectively,
+which Stainless does not support. We simply turn those onjects into
+case objects. That transformation is equivalent. Case objects have
+various additional properties (for example, being serializable) but
+none of our code depends on the absence of any of that.
 
-That transformation is equivalent. Case objects have some additional
-properties (in particular, being serializable) and they inherit from
-\texttt{Product} instead of \texttt{AnyRef}, but none of our code
-depends on any of that.
+\paragraph{Abstract Type Members.} In CurrencyUnits.scala on line 6
+there is an abstract type that is not supported. Note that we can not
+simply replace it by a (supported) type parameter since the
+CurrencyUnit class uses one of its implementing classes:
+Satoshis. Since the Satoshis class overrides A with Int64 anyway, we
+just remove the abstract type declaration and replace A by Int64
+everywhere.
 
-\item Abstract Type Members.
+\paragraph{Non-Literal BigInt Constructor Argument.} In
+CurrencyUnits.scala on line 18 the BigInt constructor is called with a
+non-literal argument. As described before, the types in the Stainless
+library are more restricted than their Scala library counterparts. In
+particular, the Stainless BigInt constructor is restricted to literal
+arguments. So we simply replace \texttt{toLong} by
+\texttt{underlying.toBigInt}: instead of converting the underlying
+Int64 (which in turn has an underlying BigInt) to Long and then back
+to BigInt we simply directly return the BigInt. This is an equivalent
+transformation: the only thing that might go wrong in the detour via
+\texttt{Long} is that the underlying BigInt does not fit into a
+Long. However, the only constructor of Int64Impl ensures exactly that
+and all functions producing Int64 do so via this constructor.
 
-Note that we can not simply replace the unsupported abstract type
-member by a (supported) type parameter. The problem is that the
-CurrencyUnit class uses one of its implementing classes: Satoshis.
+\paragraph{Self-Reference in Type Parameter Bound.} In
+NumberTypes.scala on lines 3 and 19 are classes a type parameter and a
+type boundary that contains that typarameter itself. Stainless does
+not currently support such self-referential type boundaries. We opened
+an issue \cite{Stainless:issue519} on the Stainless repository and the
+developers have targeted version 0.4 to support self-referential type
+boundaries. Since our code only uses Number with type parameter T
+instantiated to Int64, we just remove the type parameter declaration
+and replace all its occurrences it by Int64.
 
-Satoshis would have to instantiate a potential type parameter with
-type Int64, so it would extend CurrencyUnit[Int64]. But that is too
-specific, because the return type of the \texttt{+}-method would then then be
-CurrencyUnit[Int64] not CurrencyUnit[A].
-
-Since we only want to verify the addition of satoshis, and the
-Satoshis class overrides A with Int64 anyway, we just remove the
-abstract type and set it to Int64.
-
-We remove line 6 and line 22 from CurrencyUnits.scala (to maintain
-line numbers we actually replace them with empty lines for now) and in
-line 18 we replace A by Int64.
-
-\item{Non-Literal BigInt Constructor Argument}
-
-As described before, the types in the Stainless library are more
-restricted than their Scala library counterparts. In particular, the
-Stainless BigInt constructor is restricted to literal arguments.
-
-Here we can simply use toBigInt on the field \texttt{underlying}
-directly.  So, instead of converting the underlying to Long and back
-to BigInt we convert underlying directly to BigInt.
-
-We replace line 26 by
-\begin{lstlisting}[style=scala]
-def toBigInt: BigInt = underlying.toBigint
-\end{lstlisting}
-
-This is an equivalent transformation: the only thing that might go
-wrong in the detour via \texttt{Long} is that the BigInt underlying
-Int64 in turn underlying Satoshis does not fit into a Long. However,
-the only constructor of Int64 ensures exactly that.
-
-\item{Self-Reference in Type Parameter Bound}
-
-Stainless does not currently support a class with a type parameter
-with a type boundary that contains the type parameter itself. We opened an
-issue \cite{Stainless:issue519} on the Stainless repository and the
-Stainless developers have targeted Stainless version 0.4 to support
-such self-referential type boundaries.
-
-For now, since our code only uses Number with type parameter T
-instantiated to Int64, we just remove the type parameter and replace
-it by Int64. We respectively replace lines 8, 44 and 49 by
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-\end{lstlisting}
-
-\begin{lstlisting}[style=scala]
-sealed abstract class SignedNumber extends Number
-\end{lstlisting}
-
-\begin{lstlisting}[style=scala]
-sealed abstract class Int64 extends SignedNumber {
-\end{lstlisting}
-
-and replace T by Int64 in lines 25-27.
-
-\item{Missing Member \texttt{bigInteger} in BigInt}
-
-The Scala class BigInt is essentially a wrapper around
+\paragraph{Missing Member \texttt{bigInteger} in BigInt.} In NumberType
+on line 6 there is a reference to \texttt{bigInteger}. The Scala
+BigInt class is essentially a wrapper around
 java.math.BigInteger. BigInt has a member bigInteger which is the
@@
 long, otherwise throws exception. Stainless does not support Java
-classes and in particular its BigInt has no member bigInteger.
+classes and in particular its BigInt has no member
+bigInteger. However, our code never calls toLong anymore, so we just
+remove it.
 
-However, our code never calls toLong anymore, so we just remove it. We
-replace line 14 in NumberType.scala and line 28 in CurrencyUnits.scala
-by an empty line.
+\paragraph{Type Members.} In NumberType.scala there is a type member
+on line 4. Our version of Stainless (0.1) does not support type
+members. We just remove the declaration and replace all occurrences of
+A with BigInt, since A is never overwritten in an implementing class.
+Note that in the mean time Stainless has implemented support for type
+members \cite{Stainless:pull470}.  Since version 0.2 verification
+should succeed without this change.
 
-\item{Type Member}
+\paragraph{Missing Bitwise-And Method on BigInt.} Contrary to Scala
+BigInt, the Stainless BigInt class does not support bitwise
+operations, in particular not the \&-method used in NumberType.scala
+on line 13. However, as described above, the checkResult function is
+both broken and redundant, so we remove it and all calls to it.
 
-Our version of Stainless does not support type members. We just
-replace all occurrence of A with BigInt, since A is never overwritten
-in an implementing class.
+\paragraph{Inner Class in Case Object.} We have inner classes in
+NumberType.scala on line 31 and in CurrencyUnits.scala on line
+26. Stainless does not support inner classes in a case object. We just
+move the inner classes out of the case objects. They do not interfere
+with any other code.
 
-We remove line 9 in NumberType.scala and replace A by BigInt in lines
-12, 25, 33 and 50.
+\paragraph{Message Parameter in Require.} The calls of the require
+function on lines 32 and 34 in CurrencyUnits.scala have a second
+parameter: the error message. Stainless does not support the message
+parameter. We simply remove it.
 
-In the mean time Stainless has implemented support for type member
-\cite{Stainless:pull470}.  Since version 0.2 verification should
-succeed without this change.
+\paragraph{Missing Implicit Long to BigInt Conversion.} The Scala
+BigInt class has implict conversions from Long which NumberType.scala
+uses on lines 32 and 34 and they are missing in the Stainless
+BigInt. Since a BigInt constructor with a Long argument is also
+missing, we replace the Long literals by an explicit call to the
+BigInt constructor with a literal string argument,
+e.g. \texttt{BigInt("-9223...5808")}.
 
-\item{Missing Bitwise-And Method on BigInt}
+\paragraph{The Specification.} Now that all our code has been
+transformed into the supported fragment, we can finally write our
+specification, shown in Figure~\ref{fig:spec}, and verify it with
+Stainless, as the output in Figure~\ref{fig:result} shows.
 
-Contrary to Scala BigInt, the Stainless BigInt class does not support
-bitwise operations, in particular not the \&-method for bitwise and.
+\begin{figure}
+\lstinputlisting[style=scala, firstline=9, lastline=12, firstnumber=9]{../code/addition/src/main/scala/verified/currency/CurrencyUnits.scala}
+\caption{Addition function with specification}
+\label{fig:spec}
+\end{figure}
 
-The bitwise and with the andMask on line 34 is a bounds check.  It
-checks if the result parameter is in range of the specified type,
-which in our case is the hard coded Int64.
-
-So, we replace the \& mask with a check whether the result
-is in range of Long.MinValue and Long.MaxValue, because Int64 has the
-same 64-bit range as Long.
-
-We replace lines 34-35 by the following, squeezed into two lines to
-preserve line numbers. Note that the BigInt constructor requires a
-literal:
-
-\begin{lstlisting}[style=scala]
-require(result <= BigInt("9223372036854775807")  
-  && result >= BigInt("-9223372036854775808"),
-      "Result was out of bounds, got: " + result)
-\end{lstlisting}
-
-\item{Inner Class in Case Object}
-
-Stainless does not support inner classes in a case object. Bitcoin-s
-uses this a lot to separate the class from its implementation.
-
-This is easy to fix. We just move the inner classes out of the case objects. They do
-not interfere with any other code.
-
-We remove lines 66-71 in NumberType.scala and insert them at the end of the file.
-We remove line 42 in CurrencyUnits.scala and insert it at the end of the file.
-
-\item{Message Parameter in Require}
-
-Here the require(condition, message) is used as an assertion: if the
-condition is false the fail with message.  Stainless does not support
-the message parameter of require.  For the verification we can simply
-remove that parameter.
-
-\item{Missing Implicit Long to BigInt Conversion}
-
-This error message does not specify a line number and it is not clear
-what ``inv'' is. However, the Scala BigInt has implict conversions
-from Long and they are missing in the Stainless BigInt.
-
-We replace the two require clauses at lines 67 and following in NumberType.scala
-
-\begin{lstlisting}[style=scala]
-  require(underlying >= -9223372036854775808L)
-  require(underlying <= 9223372036854775807L)
-\end{lstlisting}
-
-by:
-\begin{lstlisting}[style=scala]
-  require(underlying >= BigInt(-9223372036854775808L))
-  require(underlying <= BigInt(9223372036854775807L))
-\end{lstlisting}
-
-We also replace lines 76 and 77
-\begin{lstlisting}[style=scala]
-  lazy val min = Int64(-9223372036854775808L)
-  lazy val max = Int64(9223372036854775807L)
-\end{lstlisting}
-
-by:
-\begin{lstlisting}[style=scala]
-  lazy val min = Int64(BigInt(-9223372036854775808L))
-  lazy val max = Int64(BigInt(9223372036854775807L))
-\end{lstlisting}
-
-\item{Missing BigInt Constructor with Long Argument}
-
-Here again, the Scala BigInt has a constructor with a Long argument
-which is missing in the Stainless BigInt.
-
-We simply replace the Long values in a BigInt constructor call with a string
-literal.
-
-The lines from the previous transformation respectively change into
-the following:
-
-\begin{lstlisting}[style=scala]
-  require(underlying >= BigInt("-9223372036854775808"))
-  require(underlying <= BigInt("9223372036854775807"))
-\end{lstlisting}
-
-and
-
-\begin{lstlisting}[style=scala]
-  lazy val min = Int64(BigInt("-9223372036854775808"))
-  lazy val max = Int64(BigInt("9223372036854775807"))
-}
-\end{lstlisting}
-\end{enumerate}
-
-\paragraph{The Specification.}
-
-We can use equals (==) directly on Satoshis, because it is a case class.
-
-The addition will now look like this:
-\lstinputlisting[style=scala, firstline=15, lastline=18, firstnumber=15]{../code/addition/src/main/scala/verified/currency/CurrencyUnits.scala}
-
-\paragraph{Result}
-
-Finally, everything is green and correctly verified. 
 \begin{figure}
@@
 	\caption{Stainless output for the transformed code}
+  \label{fig:result}
 \end{figure}
 
-\section{Conclusion}
-\label{chap:conclusion}
+\section{Conclusion and Future Work}
+
+We have show
+-- bugs (already fixed)
+-- suggestions for bitcoin-s
 
 Because of the limitations of the verication tool, we could only
-verify a rewritten version of the original Bitcoin-S code.  So we can
-not guarantee the correctness of the addition of Satoshis with zero in
-Bitcoin-S.  Not all changes we made were as trivial as the replacement
-of objects with case objects.  For these non-trivial changes, as seen
-for example the bound check in section \ref{sec:bound_check}, we
-cannot say whether they are equivalent to the original implementation
-or not.
+verify a rewritten version of the original Bitcoin-S code.
 
 So code should be written specically with formal verication in mind,
-in order to successfully verify it.  Otherwise, it needs a lot of
-changes in the software because verification is mathematical and the
-current software is written mostly in object-oriented style.  Software
-written in the functional paradigm would be much easier to reason
-about.
-
-Thus, either Stainless must find ways to translate more of built-in
-object-oriented patterns of Scala to their verification tool or
-developers must invest more in functional programming.
 
@@
 
-\todo{conclusions: what's future work? how to
-  change bitcoin-s? how to extend stainless?}
+\bibliographystyle{splncs04}
 
-\bibliographystyle{splncs04}
 \bibliography{bibliography}
@@
 
-\section{Code Transformations} 
-\label{sec:transform}
-
-Here we see in detail how to transform the bitcoin-s code into the
-Scala fragment supported by Stainless. All subsections start with the
-Stainless error message(s) and finally a description of the changes we
-make to the code.
-
-We claim that all transformations are equivalent in the sense that if
-the addition-with-zero property holds for the transformed code, then
-it also holds for the code before the transformation.
-
-\subsection{Inheriting Objects}
-
-\begin{lstlisting}[style=stainless]
-[ Error  ] number/NumberType.scala:65:1: Objects cannot extend classes or implement traits, use a case object instead
-          object Int64 extends BaseNumbers[Int64] {
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-[ Error  ] currency/CurrencyUnits.scala:33:1: Objects cannot extend classes or implement traits, use a case object instead
-          object Satoshis extends BaseNumbers[Satoshis] {
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-\end{lstlisting}
-
-Here, we can just turn the objects into case objects by literally just
-changing the word \texttt{object} into \texttt{case object} on lines
-65 and 33 in the two respective files.
-
-That transformation is equivalent. Case objects have some additional
-properties (in particular, being serializable) and they inherit from
-\texttt{Product} instead of \texttt{AnyRef}, but none of our code
-depends on any of that.
-
-\subsection{Abstract Type Members}
-
-\begin{lstlisting}[style=stainless]
-[ Error  ] currency/CurrencyUnits.scala:6:3: Stainless doesn't support abstract type members
-            type A
-            ^^^^^^
-\end{lstlisting}
-
-Note that we can not simply replace the unsupported abstract type
-member by a (supported) type parameter. The problem is that the
-CurrencyUnit class uses one of its implementing classes: Satoshis.
-
-Satoshis would have to instantiate a potential type parameter with
-type Int64, so it would extend CurrencyUnit[Int64]. But that is too
-specific, because the return type of the \texttt{+}-method would then then be
-CurrencyUnit[Int64] not CurrencyUnit[A].
-
-Since we only want to verify the addition of satoshis, and the
-Satoshis class overrides A with Int64 anyway, we just remove the
-abstract type and set it to Int64.
-
-We remove line 6 and line 22 from CurrencyUnits.scala (to maintain
-line numbers we actually replace them with empty lines for now) and in
-line 18 we replace A by Int64.
-
-\subsection{Non-Literal BigInt Constructor Argument}
-
-\begin{lstlisting}[style=stainless]
-[ Error  ] currency/CurrencyUnits.scala:26:33: Only literal arguments are allowed for BigInt.
-            def toBigInt: BigInt = BigInt(toLong)
-                                          ^^^^^^
-\end{lstlisting}
-
-As described before, the types in the Stainless library are more
-restricted than their Scala library counterparts. In particular, the
-Stainless BigInt constructor is restricted to literal arguments.
-
-Here we can simply use toBigInt on the field \texttt{underlying}
-directly.  So, instead of converting the underlying to Long and back
-to BigInt we convert underlying directly to BigInt.
-
-We replace line 26 by
-\begin{lstlisting}[style=scala]
-def toBigInt: BigInt = underlying.toBigint
-\end{lstlisting}
-
-This is an equivalent transformation: the only thing that might go
-wrong in the detour via \texttt{Long} is that the BigInt underlying
-Int64 in turn underlying Satoshis does not fit into a Long. However,
-the only constructor of Int64 ensures exactly that.
-
-\subsection{Self-Reference in Type Parameter Bound}
-
-\begin{lstlisting}[style=stainless]
-[ Error  ] number/NumberType.scala:8:30: Unknown type parameter type T
-          sealed abstract class Number[T <: Number[T]] {
-                                        ^^^^^^^^^^^^^^
-\end{lstlisting}
-
-Stainless does not currently support a class with a type parameter
-with a type boundary that contains the type parameter itself. We opened an
-issue \cite{Stainless:issue519} on the Stainless repository and the
-Stainless developers have targeted Stainless version 0.4 to support
-such self-referential type boundaries.
-
-For now, since our code only uses Number with type parameter T
-instantiated to Int64, we just remove the type parameter and replace
-it by Int64. We respectively replace lines 8, 44 and 49 by
-\begin{lstlisting}[style=scala]
-sealed abstract class Number {
-\end{lstlisting}
-
-\begin{lstlisting}[style=scala]
-sealed abstract class SignedNumber extends Number
-\end{lstlisting}
-
-\begin{lstlisting}[style=scala]
-sealed abstract class Int64 extends SignedNumber {
-\end{lstlisting}
-
-and replace T by Int64 in lines 25-27.
-
-\subsection{Missing Member \texttt{bigInteger} in BigInt}
-
-\begin{lstlisting}[style=stainless]
-[ Error  ] number/NumberType.scala:14:22: Unknown call to bigInteger on Number.this.toBigInt (BigInt) with arguments List() of type List()
-            def toLong: Long = toBigInt.bigInteger.longValueExact()
-                              ^^^^^^^^^^^^^^^^^^^
-\end{lstlisting}
-
-The Scala class BigInt is essentially a wrapper around
-java.math.BigInteger. BigInt has a member bigInteger which is the
-underlying instance of the Java class. The Java class has a method
-longValueExact which returns a long only if the BigInteger fits into a
-long, otherwise throws exception. Stainless does not support Java
-classes and in particular its BigInt has no member bigInteger.
-
-However, our code never calls toLong anymore, so we just remove it. We
-replace line 14 in NumberType.scala and line 28 in CurrencyUnits.scala
-by an empty line.
-
-\subsection{Type Member}
-
-\begin{lstlisting}[style=stainless]
-[Warning ] number/NumberType.scala:9:3: Could not extract tree in class: type A = BigInt (class scala.reflect.internal.Trees$TypeDef)
-              type A = BigInt
-              ^^^^^^^^^^^^^^^
-\end{lstlisting}
-
-Our version of Stainless does not support type members. We just
-replace all occurrence of A with BigInt, since A is never overwritten
-in an implementing class.
-
-We remove line 9 in NumberType.scala and replace A by BigInt in lines
-12, 25, 33 and 50.
-
-In the mean time Stainless has implemented support for type member
-\cite{Stainless:pull470}.  Since version 0.2 verification should
-succeed without this change.
-
-\subsection{Missing Bitwise-And Method on BigInt}
-
-\begin{lstlisting}[style=stainless]
-[ Error  ] number/NumberType.scala:34:14: Unknown call to & on result (BigInt) with arguments List(Number.this.andMask) of type List(BigInt)
-              require((result & andMask) == result,
-                      ^^^^^^^^^^^^^^^^
-\end{lstlisting}
-
-Contrary to Scala BigInt, the Stainless BigInt class does not support
-bitwise operations, in particular not the \&-method for bitwise and.
-
-The bitwise and with the andMask on line 34 is a bounds check.  It
-checks if the result parameter is in range of the specified type,
-which in our case is the hard coded Int64.
-
-So, we replace the \& mask with a check whether the result
-is in range of Long.MinValue and Long.MaxValue, because Int64 has the
-same 64-bit range as Long.
-
-We replace lines 34-35 by the following, squeezed into two lines to
-preserve line numbers. Note that the BigInt constructor requires a
-literal:
-
-\begin{lstlisting}[style=scala]
-require(result <= BigInt("9223372036854775807")  
-  && result >= BigInt("-9223372036854775808"),
-      "Result was out of bounds, got: " + result)
-\end{lstlisting}
-
-\subsection{Inner Class in Case Object}
-
-\begin{lstlisting}[style=stainless]
-[Warning ] number/NumberType.scala:66:3: Could not extract tree in class: case private class Int64Impl extends Int64 with Product with Serializable {
-...
-[Warning ] } (class scala.reflect.internal.Trees$ClassDef)
-              private case class Int64Impl(underlying: BigInt) extends Int64 {
-              ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-...
-[Warning ] currency/CurrencyUnits.scala:42:3: Could not extract tree in class: case private class SatoshisImpl extends Satoshis with Product with Serializable {
-...
-[Warning ] } (class scala.reflect.internal.Trees$ClassDef)
-              private case class SatoshisImpl(underlying: Int64) extends Satoshis
-              ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
-\end{lstlisting}
-
-Stainless does not support inner classes in a case object. Bitcoin-s
-uses this a lot to separate the class from its implementation.
-
-This is easy to fix. We just move the inner classes out of the case objects. They do
-not interfere with any other code.
-
-We remove lines 66-71 in NumberType.scala and insert them at the end of the file.
-We remove line 42 in CurrencyUnits.scala and insert it at the end of the file.
-
-\subsection{Message Parameter in Require}
-
-\begin{lstlisting}[style=stainless]
-[Warning ] number/NumberType.scala:67:3: Could not extract tree in class: scala.this.Predef.require(Int64Impl.this.underlying.>=(math.this.BigInt.long2bigInt(-9223372036854775808L)), "Number was too small for a int64, got: ".+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
-          require(underlying >= -9223372036854775808L,
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-[Warning ] number/NumberType.scala:69:3: Could not extract tree in class: scala.this.Predef.require(Int64Impl.this.underlying.<=(math.this.BigInt.long2bigInt(9223372036854775807L)), "Number was too big for a int64, got: ".+(Int64Impl.this.underlying)) (class scala.reflect.internal.Trees$Apply)
-          require(underlying <= 9223372036854775807L,
-          ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^...
-[ Error  ] checkResult$0 depends on missing dependencies: require$1.
-\end{lstlisting}
-
-Here the require(condition, message) is used as an assertion: if the
-condition is false the fail with message.  Stainless does not support
-the message parameter of require.  For the verification we can simply
-remove that parameter.
-
-\subsection{Missing Implicit Long to BigInt Conversion}
-
-\begin{lstlisting}[style=stainless]
-[ Error  ] inv$4 depends on missing dependencies: long2bigInt$0.
-\end{lstlisting}
-
-This error message does not specify a line number and it is not clear
-what ``inv'' is. However, the Scala BigInt has implict conversions
-from Long and they are missing in the Stainless BigInt.
-
-We replace the two require clauses at lines 67 and following in NumberType.scala
-
-\begin{lstlisting}[style=scala]
-  require(underlying >= -9223372036854775808L)
-  require(underlying <= 9223372036854775807L)
-\end{lstlisting}
-
-by:
-\begin{lstlisting}[style=scala]
-  require(underlying >= BigInt(-9223372036854775808L))
-  require(underlying <= BigInt(9223372036854775807L))
-\end{lstlisting}
-
-We also replace lines 76 and 77
-\begin{lstlisting}[style=scala]
-  lazy val min = Int64(-9223372036854775808L)
-  lazy val max = Int64(9223372036854775807L)
-\end{lstlisting}
-
-by:
-\begin{lstlisting}[style=scala]
-  lazy val min = Int64(BigInt(-9223372036854775808L))
-  lazy val max = Int64(BigInt(9223372036854775807L))
-\end{lstlisting}
-
-\subsection{Missing BigInt Constructor with Long Argument}
-
-\begin{lstlisting}[style=stainless]
-[ Error  ] apply$14 depends on missing dependencies: BigInt$0, apply$15.
-\end{lstlisting}
-
-Here again, the Scala BigInt has a constructor with a Long argument
-which is missing in the Stainless BigInt.
-
-We simply replace the Long values in a BigInt constructor call with a string
-literal.
-
-The lines from the previous transformation respectively change into
-the following:
-
-\begin{lstlisting}[style=scala]
-  require(underlying >= BigInt("-9223372036854775808"))
-  require(underlying <= BigInt("9223372036854775807"))
-\end{lstlisting}
-
-and
-
-\begin{lstlisting}[style=scala]
-  lazy val min = Int64(BigInt("-9223372036854775808"))
-  lazy val max = Int64(BigInt("9223372036854775807"))
-}
-\end{lstlisting}
-
 \end{document}

### 2019-07-04 paper/paper.tex
more work

@@
   implementation of parts of the Bitcoin protocol. We use the
-  Stainless verifier which supports programs in a subset of Scala
+  Stainless verifier which supports programs in a fragment of Scala
   called \emph{Pure Scala}.  Since bitcoin-s is not written in this
@@
 superclasses Factory and NetworkElement of CurrencyUnit and Number,
-respectively, because the inherited members are not used. Also, we
+respectively, because the inherited members are not used. We further
 removed all binary operations on Number that are not used, like
@@
 
-We have show
--- bugs (already fixed)
--- suggestions for bitcoin-s
+We are happy to see some friendly green verifier output. However,
+apart from the bugs we found, the main conclusion from this work is
+that we had to non-trivially transform even a very small portion of
+the code in order to verify it. At the moment, it is unrealistic to
+routinely formally verify properties as part of the bitcoin-s
+development process. However, Stainless development has already
+progressed (e.g.\ type members are supported in recent versions) and
+continues to do so (e.g. self-referential type bounds are on the
+roadmap). Some missing features that we identified are presumably very
+easy to support, like the message parameter in the require
+function. Some other features presumably require more substantial
+work, like bitwise operations on integer types.
 
-Because of the limitations of the verication tool, we could only
-verify a rewritten version of the original Bitcoin-S code.
+On the other hand, bitcoin-s uses features that might not be supported
+even by future Stainless versions, such as (concrete, non-case)
+classes \todo{is that even true? where?} or calls to Java code. Here
+the bitcoin-s code can hopefully be adapted to accomodate formal
+verification.
 
-So code should be written specically with formal verication in mind,
-
-Also, we found that trying to verify code reveals bugs as shown in
-section \ref{sec:bugfix}.  Finally, our work led to some feedback to
-the Stainless developers to improve the tool.
+From our results we conclude that formal verification of bitcoin
+libraries in general and bitcoin-s in particular is a worthwile
+endeavour. We are looking forward to verifying more substantial parts
+of the code in future work.
 
@@
 
+\todo{reference to verified code}
+
 \clearpage

### 2019-07-05 paper/paper.tex
typos etc

@@
 32-bit Int leads to a negative result for the input 17, as Stainless
-reports in Figure~\ref{fig:failed}. Changing the type from Int to
-BigInt will result in a successful verification.
+reports in Figure~\ref{fig:failed}. Changing the type Int to BigInt
+will result in a successful verification.
 
@@
 What will turn out to be more relevant for us are the Scala features
-which Stainless does not support, such as: (concrete) class
+which Stainless does not support, such as: arbitrary class
 definitions, inheritance by objects, abstract type members, and inner
@@
 \texttt{outPoint}. Second, we add some information on how to spend
-that output, in particular, how to sign the transaction to allow
-this. Now we assemble the list of unspent transaction outputs (utxos),
-in our case just one.
+that output, in particular, how to sign the transaction. Now we
+assemble the list of unspent transaction outputs (utxos), in our case
+just one.
 
@@
 
-We then create a lock script (\texttt{destinationSPK}) to receive the
+We then create a lock script (\textit{destinationSPK}) to receive the
 coins, create our list of transaction outputs (\texttt{destinations}),
@@
 of the checkTransaction function in Figure~\ref{fig:checktrans}. Here,
-value prevOutputTxIds gathers a list of all transaction indentifiers
+the value prevOutputTxIds gathers a list of all transaction identifiers
 referenced by the inputs of the current transaction. If the size of
@@
 The fix is simple: we perform the duplicate check on the
-TransactionOutPoints instead of on their transaction identifiers.  A
-TransactionOutPoint contains the txId as well as the output index it
-references. Note that TransactionOutPoint is a case class and thus has
-a built in == method.
+TransactionOutPoint instances instead of on their transaction
+identifiers. Note that TransactionOutPoint is a case class and thus
+its notion of equality is just what we need: equality of of both the
+txId and the output index.
 
@@
 the addition of two satoshi amounts in bitcoin-s is non-trivial. The
-reason for that is a pecularity of consensus code: agreement with the
+reason for that is a peculiarity of consensus code: agreement with the
 majority is more important than correctness, whatever correctness
@@
 extending the BaseNumbers trait, on lines 30 and 23 respectively,
-which Stainless does not support. We simply turn those onjects into
+which Stainless does not support. We simply turn those objects into
 case objects. That transformation is equivalent. Case objects have
 various additional properties (for example, being serializable) but
-none of our code depends on the absence of any of that.
+none of our code depends on the absence of those.
 
@@
 \paragraph{Self-Reference in Type Parameter Bound.} In
-NumberTypes.scala on lines 3 and 19 are classes a type parameter and a
-type boundary that contains that typarameter itself. Stainless does
-not currently support such self-referential type boundaries. We opened
-an issue \cite{Stainless:issue519} on the Stainless repository and the
-developers have targeted version 0.4 to support self-referential type
-boundaries. Since our code only uses Number with type parameter T
-instantiated to Int64, we just remove the type parameter declaration
-and replace all its occurrences it by Int64.
+NumberTypes.scala both on lines 3 and 19 is a class with a type
+parameter and a type boundary that contains that type parameter
+itself. Stainless does not currently support such self-referential
+type boundaries. We opened an issue \cite{Stainless:issue519} on the
+Stainless repository and the developers have targeted version 0.4 to
+support self-referential type boundaries. Since our code only uses
+Number with type parameter T instantiated to Int64, we just remove the
+type parameter declaration and replace all its occurrences by Int64.
 
-\paragraph{Missing Member \texttt{bigInteger} in BigInt.} In NumberType
-on line 6 there is a reference to \texttt{bigInteger}. The Scala
-BigInt class is essentially a wrapper around
+\paragraph{Missing Member \texttt{bigInteger} in BigInt.} In
+NumberType on line 6 there is a reference to \texttt{bigInteger}. The
+Scala BigInt class is essentially a wrapper around
 java.math.BigInteger. BigInt has a member bigInteger which is the
@@
 classes and in particular its BigInt has no member
-bigInteger. However, our code never calls toLong anymore, so we just
+bigInteger. However, our code does not call toLong anymore, so we just
 remove it.
@@
 A with BigInt, since A is never overwritten in an implementing class.
-Note that in the mean time Stainless has implemented support for type
+Note that in the meantime Stainless has implemented support for type
 members \cite{Stainless:pull470}.  Since version 0.2 verification
@@
 even by future Stainless versions, such as (concrete, non-case)
-classes \todo{is that even true? where?} or calls to Java code. Here
-the bitcoin-s code can hopefully be adapted to accomodate formal
+classes \todo{NOT TRUE, no classes found} or calls to Java code. Here
+the bitcoin-s code can hopefully be adapted to accommodate formal
 verification.
@@
 From our results we conclude that formal verification of bitcoin
-libraries in general and bitcoin-s in particular is a worthwile
+libraries in general and bitcoin-s in particular is a worthwhile
 endeavour. We are looking forward to verifying more substantial parts
@@
 
-\todo{reference to verified code}
+\todo{reference to verified code} \todo{say which version of stainless
+  we used, which version of bitcoin-s:
+  91633375a37a170ba260ff4b443582aca4aa0c77 }
+
+\todo{code in text: italic or courier?}
 

### 2019-07-06 paper/paper.tex
remover some double spaces
update todos

@@
 \cite{BitcoinS:website,BitcoinS:github}. In particular, it allows to
-serialize, deserialize, sign and validate Bitcoin transactions. The library
-uses immutable data structures and algebraic data types but is not
-written with formal verification in mind. According to the website,
-the library is used in production, handling significant amounts of
-cryptocurrency each day \cite{BitcoinS:website}.
+serialize, deserialize, sign and validate Bitcoin transactions. The
+library uses immutable data structures and algebraic data types but is
+not specifically written with formal verification in mind. According
+to the website, the library is used in production, handling
+significant amounts of cryptocurrency each day
+\cite{BitcoinS:website}.
 
@@
 On the other hand, bitcoin-s uses features that might not be supported
-even by future Stainless versions, such as (concrete, non-case)
-classes \todo{NOT TRUE, no classes found} or calls to Java code. Here
+even by future Stainless versions, such as calls to Java code. Here
 the bitcoin-s code can hopefully be adapted to accommodate formal
@@
 
-\todo{reference to verified code} \todo{say which version of stainless
+\todo{reference to verified code}
+
+\todo{say which version of stainless
   we used, which version of bitcoin-s:

### 2019-07-06 paper/paper.tex
final touches

@@
 \begin{abstract}
-  We try to verify some properties of the bitcoin-s library, a Scala
+  We try to verify properties of the bitcoin-s library, a Scala
   implementation of parts of the Bitcoin protocol. We use the
@@
 \paragraph{The Stainless Verifier.} Stainless is the successor of the
-Leon verifier
-\cite{DBLP:conf/ecoop/BlancKKS13,DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}
-and is developed at EPF Lausanne \cite{Stainless:github}. It is
-intended to be used by programmers without training in formal
+Leon verifier and is developed at EPF Lausanne
+\cite{DBLP:conf/ecoop/BlancKKS13,DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}. It
+is intended to be used by programmers without training in formal
 verification. To facilitate that, it accepts specifications written in
@@
 documentation \cite{Stainless:documentation} shows how the verifier is
-used. Notice how a precondition is specified using \emph{require} and
-a postcondition using \emph{ensuring}.
+used. Note how a precondition is specified using \texttt{require} and
+a postcondition using \texttt{ensuring}. Our function does not satisfy
+the specification. An overflow in the 32-bit integer type leads to a
+negative result for the input 17, as Stainless reports in
+Figure~\ref{fig:failed}. Changing the type \texttt{Int} to
+\texttt{BigInt} will result in a successful verification.
 
@@
 \end{figure}
-Our function does not satisfy the specification. An overflow in the
-32-bit Int leads to a negative result for the input 17, as Stainless
-reports in Figure~\ref{fig:failed}. Changing the type Int to BigInt
-will result in a successful verification.
 
@@
 \paragraph{The Pure Scala Fragment.} The Scala fragment supported by
-Stainless is described in the Stainless documentation
-\cite{Stainless:documentation} in the section
-\href{https://epfl-lara.github.io/stainless/purescala.html}{Pure
-  Scala}.
-
-It comprises algebraic data types in the form of abstract classes,
-case classes and case objects, objects for grouping classes and
-functions, boolean expressions with short-circuit interpretation,
-generics with invariant type parameters, default values of function
-parameters, pattern matching, local and anonymous classes and more.
-In addition to Pure Scala Stainless also supports some imperative
-features, such as using a (mutable) variable in a local scope of a
-function and while loops. They turn out not to be relevant for the
-current work.
+Stainless comprises algebraic data types in the form of abstract
+classes, case classes and case objects, objects for grouping classes
+and functions, boolean expressions with short-circuit interpretation,
+generics with invariant type parameters, pattern matching, local and
+anonymous classes and more.  In addition to Pure Scala Stainless also
+supports some imperative features, such as using a (mutable) variable
+in a local scope of a function and while loops. They turn out not to
+be relevant for our current work.
 
 What will turn out to be more relevant for us are the Scala features
-which Stainless does not support, such as: arbitrary class
-definitions, inheritance by objects, abstract type members, and inner
-classes in case objects.
-
-Also, Stainless has its own library of some core data types and
-functions which are mapped to corresponding data types and functions
-inside of the SMT solver that Stainless ultimately relies on. Those
-data types in general do not have all the methods of the Scala data
-types. For example, the BigInt type in Scala has methods for bitwise
-operations while the BigInt type in Stainless does not.
+which Stainless does not support, such as: inheritance by objects,
+abstract type members, and inner classes in case objects. Also,
+Stainless has its own library of some core data types and functions
+which are mapped to corresponding data types and functions inside of
+the SMT solver that Stainless ultimately relies on. Those data types
+in general do not have all the methods of the Scala data types. For
+example, the \texttt{BigInt} type in Scala has methods for bitwise
+operations while the \texttt{BigInt} type in Stainless does not.
 
@@
 \section{The No-Inflation Property}
+
+The \texttt{checkTransaction} function shown in
+Figure~\ref{fig:checktrans} is crucial for the verification of the
+no-inflation property. Given a transaction it returns true if some
+basic checks succeed, otherwise false. For example, one of those
+checks is that both the list of inputs and list of outputs need to be
+non-empty.
+
+To better understand the validation of a transaction in bitcoin-s, it
+is useful to review how transactions are represented and created.
+
+\paragraph{Creating a Transaction.} To create a transaction, we first
+need some coins -- an unspent transaction output. We could load an
+actual unspent transaction output from the bitcoin network, but we
+create one manually in order to see this process. So we first create
+an (invalid) transaction with one output in Figure~\ref{fig:prevtx}.
 
@@
 
-A crucial function for the verification of the no-inflation property
-is the \texttt{checkTransaction} function shown in
-Figure~\ref{fig:checktrans}. Given a transaction it returns true if
-some basic checks succeed, otherwise false. For example, one of those
-checks is that both the list of inputs and list of outputs need to be
-non-empty.
-
-To better understand the validation of a transaction in bitcoin-s, it
-is useful to review how transactions are represented and created.
-
-\paragraph{Creating a Transaction.} The code in this subsection is
-adapted from the bitcoin-s documentation. To create a transaction, we
-first need some coins -- an unspent transaction output. We could load
-an actual unspent transaction output from the bitcoin network, but we
-create one manually in order to see this process. So we first create
-an (invalid) transaction with one output in Figure~\ref{fig:prevtx}.
-
 \begin{figure}
@@
 We first create a keypair, then a lock script with its public key,
-then the amount of satoshis, then a transaction output (utxo) for that
-amount and locked with that script. Finally we create a transaction
-with that output and no inputs. Of course, that is not a valid
-transaction, because it creates coins out of nothing. In particular,
-\texttt{checkTransaction(prevTx)} returns false, simply because the list of
-inputs is empty.
+then the amount of satoshis, then a transaction output (\texttt{utxo})
+for that amount and locked with that script. Finally we create a
+transaction with that output and no inputs. Of course, that is not a
+valid transaction, because it creates coins out of nothing. In
+particular, \texttt{checkTransaction(prevTx)} returns false, simply
+because the list of inputs is empty.
 
@@
 that output, in particular, how to sign the transaction. Now we
-assemble the list of unspent transaction outputs (utxos), in our case
+assemble the list of unspent transaction outputs (\texttt{utxos}), in our case
 just one.
 
-We then set the amount of satoshis that we want to spend. The Int64
+We then set the amount of satoshis that we want to spend. The \texttt{Int64}
 class aims to emulate a C data type in Scala, and we will look at it
@@
 
-We then create a lock script (\textit{destinationSPK}) to receive the
+We then create a lock script (\texttt{destinationSPK}) to receive the
 coins, create our list of transaction outputs (\texttt{destinations}),
@@
 \paragraph{A Bug in the checkTransaction Function.} Note lines 15-17
-of the checkTransaction function in Figure~\ref{fig:checktrans}. Here,
-the value prevOutputTxIds gathers a list of all transaction identifiers
+in Figure~\ref{fig:checktrans}. Here, the value
+\texttt{prevOutputTxIds} gathers a list of all transaction identifiers
 referenced by the inputs of the current transaction. If the size of
@@
 prevents a transaction from spending two different outputs of the same
-previous transaction. The check is too strict: checkTransaction
-returns false for valid transactions.
+previous transaction. The check is too strict:
+\texttt{checkTransaction} returns false for valid transactions.
 
 The fix is simple: we perform the duplicate check on the
-TransactionOutPoint instances instead of on their transaction
-identifiers. Note that TransactionOutPoint is a case class and thus
-its notion of equality is just what we need: equality of of both the
-txId and the output index.
+\texttt{TransactionOutPoint} instances instead of on their transaction
+identifiers. Note that \texttt{TransactionOutPoint} is a case class
+and thus its notion of equality is just what we need: equality of of
+both the transaction identifier and the output index.
 
@@
 code has more than 1500 lines and liberally uses Scala features
-outside of the supported fragment. We tried to transform the code
-into the supported fragment, but quickly realized that a better
-approach is to first verify a simpler property with less code involved
-and later come back to the no-inflation property with more
-experience. So we now turn to the addition-with-zero property.
+outside of the supported fragment. We tried to transform the code into
+the supported fragment, but quickly realized that a better approach is
+to first verify a simpler property depending on less code and later
+come back to the no-inflation property with more experience. So we now
+turn to the addition-with-zero property.
 
@@
 C-like data types. For example, the Satoshis class, which represents
-an amount of satoshis, is implemented using the class Int64 which
+an amount of satoshis, is implemented using the class \texttt{Int64} which
 aims to represent the C-type \texttt{int64\_t}.
@@
 because it is not needed for the verification of our property. For
-example, we removed all number types except for Int64 (so Int32,
-UInt64, etc.) because they are not used. We also removed the
-superclasses Factory and NetworkElement of CurrencyUnit and Number,
+example, we removed all number types except for \texttt{Int64} (so
+\texttt{Int32}, \texttt{UInt64}, etc.) because they are not used. We
+also removed the superclasses \texttt{Factory} and
+\texttt{NetworkElement} of \texttt{CurrencyUnit} and \texttt{Number},
 respectively, because the inherited members are not used. We further
-removed all binary operations on Number that are not used, like
-subtraction and multiplication. The extracted code is shown in
+removed all binary operations on \texttt{Number} that are not used,
+like subtraction and multiplication. The extracted code is shown in
 Figure~\ref{fig:numbertype} and Figure~\ref{fig:currencyunits}.
 
-\paragraph{A Bug in the checkResult Function.} Note the checkResult
-function on line 12 and the value andMask on line 23 of
-NumberType.scala. The function is intended to catch overflows by
-performing a bitwise conjunction of its argument with andMask and
-comparing the result with the argument. However, because of the way
-Java BigIntegers are represented \cite{wikipedia:twocomp} and because
-bitwise operations implicitly perform a sign extension
-\cite{java:bigint} on the shorter operand, the function does not
-actually catch overflows.
+\paragraph{A Bug in the checkResult Function.} Note the
+\texttt{checkResult} function on line 12 and the value
+\texttt{andMask} on line 23 of NumberType.scala. The function is
+intended to catch overflows by performing a bitwise conjunction of its
+argument with \texttt{andMask} and comparing the result with the
+argument. However, because of the way Java BigIntegers are represented
+\cite{wikipedia:twocomp} and because bitwise operations implicitly
+perform a sign extension \cite{java:bigint} on the shorter operand,
+the function does not actually catch overflows.
 
-While this is a potentially serious bug, it turns out that checkResult
-is only ever called inside a constructor call for a number type (like
-Int64) which contains the intended range check, see lines 32-35. The
-checkResult function thus can, and should, be removed entirely. The
-bitcoin-s developers have acknowledged the bug and we submitted a pull
-request to fix it \cite{BitcoinS:pull565}.
+While this is a potentially serious bug, it turns out that
+\texttt{checkResult} is only ever called inside a constructor call for
+a number type which contains the intended range check, see lines
+32-35. The \texttt{checkResult} function thus can, and should, be
+removed entirely. The bitcoin-s developers have acknowledged the bug
+and we submitted a pull request to fix it \cite{BitcoinS:pull565}.
+
+% For further development, this raises the question: if the goal of the
+% \texttt{Int64} type is to emulate \texttt{int64\_t} then why does it
+% prevent overflows? To achieve strict agreement with Bitcoin Core, a
+% better approach might be to remove overflow checking from the data
+% type and to add it in exactly those places where it happens in Bitcoin
+% Core.
 
@@
 Stainless and how to transform the code into the supported fragment.
-All transformations are equivalent in the sense that if the
+All transformations are \emph{equivalent} in the sense that if the
 addition-with-zero property holds for the transformed code, then it
@@
 which Stainless does not support. We simply turn those objects into
-case objects. That transformation is equivalent. Case objects have
+case objects. That transformation is equivalent: case objects have
 various additional properties (for example, being serializable) but
@@
 there is an abstract type that is not supported. Note that we can not
-simply replace it by a (supported) type parameter since the
+simply replace it with a (supported) type parameter since the
 CurrencyUnit class uses one of its implementing classes:
-Satoshis. Since the Satoshis class overrides A with Int64 anyway, we
-just remove the abstract type declaration and replace A by Int64
+Satoshis. Since the Satoshis class overrides \texttt{A} with \texttt{Int64} anyway, we
+just remove the abstract type declaration and replace \texttt{A} by \texttt{Int64}
 everywhere.
@@
 \texttt{underlying.toBigInt}: instead of converting the underlying
-Int64 (which in turn has an underlying BigInt) to Long and then back
-to BigInt we simply directly return the BigInt. This is an equivalent
-transformation: the only thing that might go wrong in the detour via
-\texttt{Long} is that the underlying BigInt does not fit into a
-Long. However, the only constructor of Int64Impl ensures exactly that
-and all functions producing Int64 do so via this constructor.
+\texttt{Int64} (which in turn has an underlying \texttt{BigInt}) to
+\texttt{Long} and then back to \texttt{BigInt} we simply directly
+return the \texttt{BigInt}. This is an equivalent transformation: the
+only thing that might go wrong in the detour via \texttt{Long} is that
+the underlying \texttt{BigInt} does not fit into a
+\texttt{Long}. However, the only constructor of \texttt{Int64Impl}
+ensures exactly that and all functions producing \texttt{Int64} do so
+via this constructor.
 
@@
 support self-referential type boundaries. Since our code only uses
-Number with type parameter T instantiated to Int64, we just remove the
-type parameter declaration and replace all its occurrences by Int64.
+Number with type parameter \texttt{T} instantiated to \texttt{Int64}, we just remove the
+type parameter declaration and replace all its occurrences by \texttt{Int64}.
 
@@
 NumberType on line 6 there is a reference to \texttt{bigInteger}. The
-Scala BigInt class is essentially a wrapper around
-java.math.BigInteger. BigInt has a member bigInteger which is the
-underlying instance of the Java class. The Java class has a method
-longValueExact which returns a long only if the BigInteger fits into a
-long, otherwise throws exception. Stainless does not support Java
-classes and in particular its BigInt has no member
-bigInteger. However, our code does not call toLong anymore, so we just
-remove it.
+Scala \texttt{BigInt} class is essentially a wrapper around
+\texttt{java.math.BigInteger}. \texttt{BigInt} has a member
+\texttt{bigInteger} which is the underlying instance of the Java
+class. The Java class has a method \texttt{longValueExact} which
+returns a \texttt{long} only if the \texttt{BigInteger} fits into a
+\texttt{long}, otherwise throws exception. Stainless does not support
+Java classes and in particular its \texttt{BigInt} has no member
+\texttt{bigInteger}. However, our code does not call \texttt{toLong}
+anymore, so we just remove it.
 
@@
 members. We just remove the declaration and replace all occurrences of
-A with BigInt, since A is never overwritten in an implementing class.
+\texttt{A} with \texttt{BigInt}, since \texttt{A} is never overwritten in an implementing class.
 Note that in the meantime Stainless has implemented support for type
@@
 \paragraph{Missing Bitwise-And Method on BigInt.} Contrary to Scala
-BigInt, the Stainless BigInt class does not support bitwise
+\texttt{BigInt}, the Stainless \texttt{BigInt} class does not support bitwise
 operations, in particular not the \&-method used in NumberType.scala
-on line 13. However, as described above, the checkResult function is
+on line 13. However, as described above, the \texttt{checkResult} function is
 both broken and redundant, so we remove it and all calls to it.
@@
 \paragraph{Missing Implicit Long to BigInt Conversion.} The Scala
-BigInt class has implict conversions from Long which NumberType.scala
-uses on lines 32 and 34 and they are missing in the Stainless
-BigInt. Since a BigInt constructor with a Long argument is also
-missing, we replace the Long literals by an explicit call to the
-BigInt constructor with a literal string argument,
+\texttt{BigInt} class has implict conversions from \texttt{Long} which
+NumberType.scala uses on lines 32 and 34. They are missing in the
+Stainless \texttt{BigInt}. A \texttt{BigInt} constructor with a
+\texttt{Long} argument is also missing. We thus replace the
+\texttt{Long} literals by an explicit call to the \texttt{BigInt}
+constructor with a literal string argument,
 e.g. \texttt{BigInt("-9223...5808")}.
@@
 Stainless, as the output in Figure~\ref{fig:result} shows.
+
+The original bitcoin-s code we started from, the extracted code, and
+the finally verified code are available in our GitHub repository \cite{verification:github}.
 
@@
 
-\todo{reference to verified code}
-
-\todo{say which version of stainless
-  we used, which version of bitcoin-s:
-  91633375a37a170ba260ff4b443582aca4aa0c77 }
-
-\todo{code in text: italic or courier?}
-
 \clearpage

### 2020-07-15 paper-short/main.tex
add short paper

--- 2019-07-06 paper/paper.tex
@@
-\documentclass[runningheads]{llncs}
-
-% to have bookmarks in pdf
-% remove for final version
-\usepackage[open,openlevel=5]{bookmark}
-\setcounter{tocdepth}{5}
-
-\usepackage{graphicx}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
+\documentclass[hyphens, a4paper,USenglish,cleveref, autoref, thm-restate]{oasics-v2019}
+%for section-numbered lemmas etc., use "numberwithinsect"
+%for enabling cleveref support, use "cleveref"
+%for enabling autoref support, use "autoref"
+%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
+%for enabling thm-restate support, use "thm-restate"
 
@@
 
-\usepackage{hyperref}
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-\renewcommand\UrlFont{\color{blue}\rmfamily}
+\bibliographystyle{plainurl}% the mandatory bibstyle
 
-%\usepackage{courier}
+\title{Towards Verifying the Bitcoin-S Library} %TODO Please add
+
+%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
+
+\author{Ramon Boss}{Bern University of Applied Sciences, Switzerland}{ramon.boss@outlook.com}{}{}
+\author{Kai Brünnler}{Bern University of Applied Sciences, Switzerland}{kai.bruennler@bfh.ch}{}{}
+\author{Anna Doukmak}{Bern University of Applied Sciences, Switzerland}{anna.doukmak@gmail.com}{}{}
+
+\authorrunning{R. Boss and K. Brünnler and A. Doukmak} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'
+
+\Copyright{Ramon Boss and Kai Brünnler and Anna Doukmak} %TODO mandatory, please use full first names. OASIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10003752.10003790.10002990</concept_id>
+<concept_desc>Theory of computation~Logic and verification</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Theory of computation~Logic and verification}
+
+\keywords{Bitcoin, Scala, Bitcoin-S, Stainless} %TODO mandatory; please add comma-separated list of keywords
+
+%\category{} %optional, e.g. invited paper
+
+\relatedversion{A full version of the paper is available at
+  \url{https://github.com/kaibr/bitcoin-s-verification/blob/master/paper/paper.pdf}.}
+
+%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
+
+%\funding{(Optional) general funding statement
+% \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.
+
+%\acknowledgements{I want to thank \dots}%optional
+
+\nolinenumbers %uncomment to disable line numbering
+
+%\hideOASIcs  %uncomment to remove references to OASIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
+
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\EventEditors{John Q. Open and Joan R. Access}
+\EventNoEds{2}
+\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
+\EventShortTitle{CVIT 2016}
+\EventAcronym{CVIT}
+\EventYear{2016}
+\EventDate{December 24--27, 2016}
+\EventLocation{Little Whinging, United Kingdom}
+\EventLogo{}
+\SeriesVolume{42}
+\ArticleNo{23}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \usepackage{listings, color}
-\hypersetup{
-  colorlinks   = true,
-  linkcolor  = blue,
-  citecolor  = blue,
-  urlcolor   = blue
-}
-
-% latin modern, no more bitmap fonts in pdf
-\usepackage{lmodern}
-%\usepackage{newtxtext}
-%\usepackage{newtxmath}
-\usepackage[zerostyle=b,scaled=.88]{newtxtt}
-
-% nicer bold
-\renewcommand{\bfdefault}{b}%
-
-% make paragraph headings bold instead of italics
-\renewcommand{\paragraph}{\textbf}%
 
@@
 
-\newcommand{\todo}[1]{{\par \color{red}#1}}
-
 \begin{document}
-
-\title{Towards Verifying the Bitcoin-S Library}
-
-\author{Ramon Boss \and Kai Brünnler \and Anna Doukmak}
-
-\authorrunning{R. Boss et al.}
-\institute{Bern University of Applied Sciences, CH-2501 Biel, Switzerland
-\email{ramon.boss@outlook.com, kai.bruennler@bfh.ch,\\anna.doukmak@gmail.com}}
 
@@
 \begin{abstract}
-  We try to verify properties of the bitcoin-s library, a Scala
+  We try to verify properties of the Bitcoin-S library, a Scala
   implementation of parts of the Bitcoin protocol. We use the
   Stainless verifier which supports programs in a fragment of Scala
-  called \emph{Pure Scala}. Since bitcoin-s is not written in this
+  called \emph{Pure Scala}. Since Bitcoin-S is not written in this
   fragment, we extract the relevant code from it and perform a series
@@
   successfully verify. In that process we find and fix two bugs in
-  bitcoin-s.
-
-\keywords{Bitcoin \and Scala \and bitcoin-s \and Stainless.}
+  Bitcoin-S.
 \end{abstract}
@@
 
-\paragraph{Extracting the Relevant Code} The relevant code for the
+\paragraph{Extracting the Relevant Code.} The relevant code for the
 addition of satoshis is in two files: CurrencyUnits.scala and
@@
 
-\bibliographystyle{splncs04}
+\appendix
 
@@
 
-\clearpage
-\appendix
-
 \end{document}

### 2020-07-22 paper-short/main.tex
remove images

@@
 
-\graphicspath{{./images/}}
+\graphicspath{{./images/}{../images/}}
 
-\bibliographystyle{plainurl}% the mandatory bibstyle
+\bibliographystyle{plainurl}
+% the mandatory bibstyle
 
-\title{Towards Verifying the Bitcoin-S Library} %TODO Please add
+\title{Towards Verifying the Bitcoin-S Library}
 
-%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
+% \titlerunning{Dummy short title}
+% optional, please use if title is longer than one line
 
@@
 
-\authorrunning{R. Boss and K. Brünnler and A. Doukmak} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'
+\authorrunning{R. Boss and K. Brünnler and A. Doukmak}
+% mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'
 
-\Copyright{Ramon Boss and Kai Brünnler and Anna Doukmak} %TODO mandatory, please use full first names. OASIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+\Copyright{Ramon Boss and Kai Brünnler and Anna Doukmak}
+% mandatory, please use full first names. OASIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
 

### 2020-07-22 paper-short/main.tex
work on short version

@@
 
-\keywords{Bitcoin, Scala, Bitcoin-S, Stainless} %TODO mandatory; please add comma-separated list of keywords
+\keywords{Bitcoin, Scala, Bitcoin-S, Stainless}
+% mandatory; please add comma-separated list of keywords
 
@@
 
-%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
+% \supplement{}
+% optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
 
@@
 
+% make paragraph headings bold instead of italics
+\renewcommand{\paragraph}{\textbf}%
+
+% The default bold(bx) is too wide
+% Choose a nicer narrower bold
+\DeclareFontSeriesDefault[rm]{bf}{b}
+
 \begin{document}
@@
   called \emph{Pure Scala}. Since Bitcoin-S is not written in this
-  fragment, we extract the relevant code from it and perform a series
-  of equivalent transformations until we arrive at code that we
-  successfully verify. In that process we find and fix two bugs in
-  Bitcoin-S.
+  fragment, we extract the relevant code from it and rewrite it until
+  we arrive at code that we successfully verify. In that process we
+  find and fix two bugs in Bitcoin-S.
 \end{abstract}
@@
 target for formal verification. In this work, we set out to verify
-properties of the bitcoin-s library with the Stainless verifier.
+properties of the Bitcoin-S library with the Stainless verifier. So
+this is a case study in applying the Stainless verifier to existing
+real-world code.
 
-\paragraph{The Bitcoin-S Library.} The bitcoin-s library is an
+\paragraph{The Bitcoin-S Library.} The Bitcoin-S library is an
 implementation of parts of the Bitcoin protocol in Scala
@@
 Leon verifier and is developed at EPF Lausanne
-\cite{DBLP:conf/ecoop/BlancKKS13,DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}. It
-is intended to be used by programmers without training in formal
-verification. To facilitate that, it accepts specifications written in
-the programming language itself (Scala). Also, it focusses on
-counterexample finding in addition to proving
-correctness. Counterexamples are useful to programmers while
-correctness proofs are not -- correctness is obvious or does not hold,
-and often both at the same time.
+\cite{DBLP:conf/ecoop/BlancKKS13,DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}.
+A distinguishing feature of Stainless is that it accepts
+specifications written in the programming language itself
+(Scala). Also, it focusses on counterexample finding in addition to
+proving correctness. Counterexamples are immediately useful to
+programmers, which can not be said about correctness proofs.
 
@@
   property}. Trying to verify it, we uncover and fix a bug in the
-bitcoin-s library. We then find that there is too much code involved
+Bitcoin-S library. We then find that there is too much code involved
 that lies outside of the supported fragment to currently make this
@@
 to verify it in Section~3. Here as well we see that a significant part
-of the code lies outside of the supported fragment. We perform a
-  series of equivalent transformations on it until we arrive at code that we
-  successfully verify. In that process we find and fix a second bug in
-  bitcoin-s.
+of the code lies outside of the supported fragment. We rewrite it
+until we arrive at code that we successfully verify. In that process
+we find and fix a second bug in Bitcoin-S.
 
@@
 non-empty.
-
-To better understand the validation of a transaction in bitcoin-s, it
-is useful to review how transactions are represented and created.
-
-\paragraph{Creating a Transaction.} To create a transaction, we first
-need some coins -- an unspent transaction output. We could load an
-actual unspent transaction output from the bitcoin network, but we
-create one manually in order to see this process. So we first create
-an (invalid) transaction with one output in Figure~\ref{fig:prevtx}.
 
@@
 
-\begin{figure}
-\begin{lstlisting}[style=scala]
-  val privKey = ECPrivateKey.freshPrivateKey
-  val creditingSPK = P2PKHScriptPubKey(pubKey = privKey.publicKey)
-
-  val amount = Satoshis(Int64(10000))
-
-  val utxo = TransactionOutput(currencyUnit = amount, scriptPubKey = creditingSPK)
-
-  val prevTx = BaseTransaction(
-    version = Int32.one,
-    inputs = List.empty,
-    outputs = List(utxo),
-    lockTime = UInt32.zero
-  )
-\end{lstlisting}
-
-  \caption{Creating a transaction output to spend}
-  \label{fig:prevtx}
-\end{figure}
-
-We first create a keypair, then a lock script with its public key,
-then the amount of satoshis, then a transaction output (\texttt{utxo})
-for that amount and locked with that script. Finally we create a
-transaction with that output and no inputs. Of course, that is not a
-valid transaction, because it creates coins out of nothing. In
-particular, \texttt{checkTransaction(prevTx)} returns false, simply
-because the list of inputs is empty.
-
-\begin{figure}
-\begin{lstlisting}[style=scala]
-  val outPoint = TransactionOutPoint(prevTx.txId, UInt32.zero)
-
-  val utxoSpendingInfo = BitcoinUTXOSpendingInfo(
-    outPoint = outPoint,
-    output = utxo,
-    signers = List(privKey),
-    redeemScriptOpt = None,
-    scriptWitnessOpt = None,
-    hashType = HashType.sigHashAll
-  )
-
-  val utxos = List(utxoSpendingInfo)
-
-  val destinationAmount = Satoshis(Int64(5000))
-
-  val destinationSPK = P2PKHScriptPubKey(pubKey = ECPrivateKey.freshPrivateKey.publicKey)
-
-  val destinations = List(
-    TransactionOutput(currencyUnit = destinationAmount, scriptPubKey = destinationSPK)
-  )
-
-  val feeRate = SatoshisPerByte(Satoshis.one)
-
-  val networkParams = RegTest // some static values for testing
-
-  val txBuilderF: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
-    destinations = destinations, 
-    utxos = utxos,               
-    feeRate = feeRate,           
-    changeSPK = creditingSPK,    // where to send the change
-    network = networkParams      
-  )
-
-  val txF: Future[Transaction] = txBuilderF.flatMap(_.sign)
-
-  val tx: Transaction = Await.result(txF, 1 second)
-\end{lstlisting}
-  \caption{Creating a transaction}
-  \label{fig:tx}
-\end{figure}
-
-Now that we have a transaction output, we create a transaction to
-spend it in Figure~\ref{fig:tx}. First, we need a reference to an
-output of a previous transaction, here called
-\texttt{outPoint}. Second, we add some information on how to spend
-that output, in particular, how to sign the transaction. Now we
-assemble the list of unspent transaction outputs (\texttt{utxos}), in our case
-just one.
-
-We then set the amount of satoshis that we want to spend. The \texttt{Int64}
-class aims to emulate a C data type in Scala, and we will look at it
-more closely in the next section.
-
-We then create a lock script (\texttt{destinationSPK}) to receive the
-coins, create our list of transaction outputs (\texttt{destinations}),
-define the fee rate and set some bitcoin network parameters.
-
-Now we create a transaction builder with those data and we tell it to
-start signing the transaction in line 34.
-
-Finally, we get the actual signed transaction. We could serialize it
-and send it to the Bitcoin network. We can also pass it to the
-checkTransaction function, which will return true.
-
 \paragraph{A Bug in the checkTransaction Function.} Note lines 15-17
@@
 We submitted this fix together with a corresponding unit test to the
-bitcoin-s project in a pull request, which has been merged
+Bitcoin-S project in a pull request, which has been merged
 \cite{BitcoinS:pull435}.
@@
 \paragraph{An Attempt at Verification.} Naively trying Stainless on
-the entire bitcoin-s codebase results in many errors -- as was to be
+the entire Bitcoin-S codebase results in many errors -- as was to be
 expected. We tried to extract only the code relevant to the
@@
 verify that we can think of. However, the code involved in performing
-the addition of two satoshi amounts in bitcoin-s is non-trivial. The
+the addition of two satoshi amounts in Bitcoin-S is non-trivial. The
 reason for that is a peculiarity of consensus code: agreement with the
-majority is more important than correctness, whatever correctness
-might mean. The most widely used bitcoin implementation by far is the
-reference implementation Bitcoin Core, written in C++. For consensus
-code, bitcoin-s has little choice but to be in strict agreement with
-the reference implementation. To achieve that, it implements C-like
-data types in Scala and then implements functionality using those
-C-like data types. For example, the Satoshis class, which represents
-an amount of satoshis, is implemented using the class \texttt{Int64} which
-aims to represent the C-type \texttt{int64\_t}.
+majority is the only relevant notion of correctness. The most widely
+used bitcoin implementation by far is the reference implementation
+Bitcoin Core, written in C++. For consensus code, Bitcoin-S has thus
+little choice but to be in strict agreement with the reference
+implementation. To achieve that, it implements C-like data types in
+Scala and then implements functionality using those C-like data
+types. For example, the Satoshis class, which represents an amount of
+satoshis, is implemented using the class \texttt{Int64} which aims to
+represent the C-type \texttt{int64\_t}.
 
@@
 32-35. The \texttt{checkResult} function thus can, and should, be
-removed entirely. The bitcoin-s developers have acknowledged the bug
+removed entirely. The Bitcoin-S developers have acknowledged the bug
 and we submitted a pull request to fix it \cite{BitcoinS:pull565}.
@@
 
-\paragraph{Transforming the Code.} We now turn to the list of Scala
+\paragraph{Rewriting the Code.} We now turn to the list of Scala
 features used by the extracted code which are not supported by
-Stainless and how to transform the code into the supported fragment.
+Stainless and how to rewrite the code in the supported fragment.
+
 All transformations are \emph{equivalent} in the sense that if the
@@
 
-The original bitcoin-s code we started from, the extracted code, and
+The original Bitcoin-S code we started from, the extracted code, and
 the finally verified code are available in our GitHub repository \cite{verification:github}.
@@
 the code in order to verify it. At the moment, it is unrealistic to
-routinely formally verify properties as part of the bitcoin-s
+routinely formally verify properties as part of the Bitcoin-S
 development process. However, Stainless development has already
@@
 
-On the other hand, bitcoin-s uses features that might not be supported
+On the other hand, Bitcoin-S uses features that might not be supported
 even by future Stainless versions, such as calls to Java code. Here
-the bitcoin-s code can hopefully be adapted to accommodate formal
+the Bitcoin-S code can hopefully be adapted to accommodate formal
 verification.
@@
 From our results we conclude that formal verification of bitcoin
-libraries in general and bitcoin-s in particular is a worthwhile
+libraries in general and Bitcoin-S in particular is a worthwhile
 endeavour. We are looking forward to verifying more substantial parts

### 2020-07-24 paper-short/main.tex
work on final version

@@
 
-\relatedversion{A full version of the paper is available at
-  \url{https://github.com/kaibr/bitcoin-s-verification/blob/master/paper/paper.pdf}.}
+% \relatedversion{A full version of the paper is available at
+%   \url{https://github.com/kaibr/bitcoin-s-verification/blob/master/paper/paper.pdf}.}
 
@@
 
-%\funding{(Optional) general funding statement
-% \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.
+% \funding{(Optional) general funding statement
+% \dots}%optional, to capture a funding statement, which applies to all authors.
+%Please enter author specific funding statements as fifth argument of the \author macro.
 
@@
 
-%\hideOASIcs  %uncomment to remove references to OASIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
+% \hideOASIcs %uncomment to remove references to OASIcs series (logo, DOI, ...),
+% e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
 
-%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \EventEditors{John Q. Open and Joan R. Access}
@@
 \ArticleNo{23}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-The example in Figure~\ref{fig:factorial} adapted from the Stainless
-documentation \cite{Stainless:documentation} shows how the verifier is
+The example in Figure~\ref{fig:factorial} is adapted from the Stainless
+documentation \cite{Stainless:documentation} and shows how the verifier is
 used. Note how a precondition is specified using \texttt{require} and
@@
 anonymous classes and more.  In addition to Pure Scala Stainless also
-supports some imperative features, such as using a (mutable) variable
-in a local scope of a function and while loops. They turn out not to
-be relevant for our current work.
+supports some imperative features, such as while loops and using a
+(mutable) variable in a local scope of a function. They turn out not
+to be relevant for our current work.
 
@@
 try to verify the property that a regular (non-coinbase) transaction
-can not generate new coins. We call it the \emph{no-inflation
-  property}. Trying to verify it, we uncover and fix a bug in the
+can not generate new coins. We call it the \emph{No-Inflation
+  Property}. Trying to verify it, we uncover and fix a bug in the
 Bitcoin-S library. We then find that there is too much code involved
@@
 zero satoshis to a given amount of satoshis yields the given amount of
-satoshis. We call it the \emph{addition-with-zero property} and we try
+satoshis. We call it the \emph{Addition-With-Zero Property} and we try
 to verify it in Section~3. Here as well we see that a significant part
@@
 
-The \texttt{checkTransaction} function shown in
-Figure~\ref{fig:checktrans} is crucial for the verification of the
-no-inflation property. Given a transaction it returns true if some
-basic checks succeed, otherwise false. For example, one of those
-checks is that both the list of inputs and list of outputs need to be
-non-empty.
+\paragraph{An Attempt at Verification.} Naively trying Stainless on
+the entire Bitcoin-S codebase results in many errors -- as was to be
+expected. We tried to extract only the code relevant to the
+No-Inflation Property and to verify that. However, even the extracted
+code has more than 1500 lines and liberally uses Scala features
+outside of the supported fragment. We started to rewrite the code in
+the supported fragment, but quickly realized that a better approach is
+to first verify a simpler property depending on less code and later
+come back to the No-Inflation Property with more experience. However,
+during the process of trying to rewrite the code, we found a bug in
+the \texttt{checkTransaction} function shown in
+Figure~\ref{fig:checktrans}.
+
+\paragraph{A Bug in the checkTransaction Function.} Given a
+transaction the function returns true if some basic checks succeed,
+otherwise false. For example, one of those checks is that both the
+list of inputs and list of outputs need to be non-empty.
 
@@
 
-\paragraph{A Bug in the checkTransaction Function.} Note lines 15-17
-in Figure~\ref{fig:checktrans}. Here, the value
+\begin{figure}
+\begin{lstlisting}[style=scala, firstnumber=15]
+  val prevOutputs = transaction.inputs.map(_.previousOutput)
+  val noDuplicateInputs = 
+    prevOutputs.distinct.size == prevOutputs.size
+\end{lstlisting}
+  \caption{Bug Fix}
+  \label{fig:bugfix1}
+\end{figure}
+
+Note particularly lines 15-17.  Here, the value
 \texttt{prevOutputTxIds} gathers a list of all transaction identifiers
@@
 
-Specifically, we replace lines 15-17 as follows:
-\begin{lstlisting}[style=scala, firstnumber=15]
-  val prevOutputs = transaction.inputs.map(_.previousOutput)
-  val noDuplicateInputs = 
-    prevOutputs.distinct.size == prevOutputs.size
-\end{lstlisting}
+Specifically, we replace lines 15-17 as shown in
+Figure~\ref{fig:bugfix1}. We submitted this fix together with a
+corresponding unit test to the Bitcoin-S project in a pull request,
+which has been merged \cite{BitcoinS:pull435}.
 
-We submitted this fix together with a corresponding unit test to the
-Bitcoin-S project in a pull request, which has been merged
-\cite{BitcoinS:pull435}.
+We now turn to the much simpler Addition-With-Zero Property.
 
-\paragraph{An Attempt at Verification.} Naively trying Stainless on
-the entire Bitcoin-S codebase results in many errors -- as was to be
-expected. We tried to extract only the code relevant to the
-no-inflation-property and to verify that. However, even the extracted
-code has more than 1500 lines and liberally uses Scala features
-outside of the supported fragment. We tried to transform the code into
-the supported fragment, but quickly realized that a better approach is
-to first verify a simpler property depending on less code and later
-come back to the no-inflation property with more experience. So we now
-turn to the addition-with-zero property.
-
-\section{The Addition-with-Zero Property}
+\section{The Addition-With-Zero Property}
 
@@
 used bitcoin implementation by far is the reference implementation
-Bitcoin Core, written in C++. For consensus code, Bitcoin-S has thus
+Bitcoin Core, written in C++. For consensus code, Bitcoin-S thus has
 little choice but to be in strict agreement with the reference
@@
 removed entirely. The Bitcoin-S developers have acknowledged the bug
-and we submitted a pull request to fix it \cite{BitcoinS:pull565}.
+and we submitted a pull request to fix it, which has been merged
+\cite{BitcoinS:pull565}.
 
-% For further development, this raises the question: if the goal of the
-% \texttt{Int64} type is to emulate \texttt{int64\_t} then why does it
-% prevent overflows? To achieve strict agreement with Bitcoin Core, a
-% better approach might be to remove overflow checking from the data
-% type and to add it in exactly those places where it happens in Bitcoin
-% Core.
+For further development of Bitcoin-S, this raises a question. If the
+goal of the \texttt{Int64} type is to emulate \texttt{int64\_t} then
+why does it prevent overflows? To achieve strict agreement with
+Bitcoin Core, a better approach might be to remove overflow checking
+from the data type and to add it in exactly those places where it
+happens in Bitcoin Core.
 
@@
 
-All transformations are \emph{equivalent} in the sense that if the
-addition-with-zero property holds for the transformed code, then it
-also holds for the code before the transformation.
+All code changes are \emph{equivalent} in the (admittedly narrow)
+sense that if the Addition-With-Zero Property holds for the rewritten
+code, then it also holds for the original code.
 
@@
 which Stainless does not support. We simply turn those objects into
-case objects. That transformation is equivalent: case objects have
-various additional properties (for example, being serializable) but
-none of our code depends on the absence of those.
+case objects. That code is equivalent: case objects have various
+additional properties (for example, being serializable) but none of
+our code depends on the absence of those.
 
@@
 CurrencyUnit class uses one of its implementing classes:
-Satoshis. Since the Satoshis class overrides \texttt{A} with \texttt{Int64} anyway, we
-just remove the abstract type declaration and replace \texttt{A} by \texttt{Int64}
-everywhere.
+Satoshis. Since the Satoshis class overrides \texttt{A} with
+\texttt{Int64} anyway, we just remove the abstract type declaration
+and replace \texttt{A} by \texttt{Int64} everywhere.
 
@@
 support self-referential type boundaries. Since our code only uses
-Number with type parameter \texttt{T} instantiated to \texttt{Int64}, we just remove the
-type parameter declaration and replace all its occurrences by \texttt{Int64}.
+Number with type parameter \texttt{T} instantiated to \texttt{Int64},
+we just remove the type parameter declaration and replace all its
+occurrences by \texttt{Int64}.
 
@@
 members. We just remove the declaration and replace all occurrences of
-\texttt{A} with \texttt{BigInt}, since \texttt{A} is never overwritten in an implementing class.
-Note that in the meantime Stainless has implemented support for type
-members \cite{Stainless:pull470}. Since version 0.2 verification
-should succeed without this change.
+\texttt{A} with \texttt{BigInt}, since \texttt{A} is never overwritten
+in an implementing class.  Note that in the meantime Stainless has
+implemented support for type members \cite{Stainless:pull470}. Since
+version 0.2 verification should succeed without this change.
 
 \paragraph{Missing Bitwise-And Method on BigInt.} Contrary to Scala
-\texttt{BigInt}, the Stainless \texttt{BigInt} class does not support bitwise
-operations, in particular not the \&-method used in NumberType.scala
-on line 13. However, as described above, the \texttt{checkResult} function is
-both broken and redundant, so we remove it and all calls to it.
+\texttt{BigInt}, the Stainless \texttt{BigInt} class does not support
+bitwise operations, in particular not the \&-method used in
+NumberType.scala on line 13. However, as described above, the
+\texttt{checkResult} function is both broken and redundant, so we
+remove it and all calls to it.
 
@@
 
-\paragraph{The Specification.} Now that all our code has been
-transformed into the supported fragment, we can finally write our
-specification, shown in Figure~\ref{fig:spec}, and verify it with
-Stainless, as the output in Figure~\ref{fig:result} shows.
+\paragraph{The Specification.} Now that all our code is in the
+supported fragment, we can finally write our specification. We add a
+postcondition to the +-method of the \texttt{CurrencyUnit}-class
+(Figure~\ref{fig:currencyunits}, lines 9-11) resulting in
+Figure~\ref{fig:spec}. We successfully verify it with Stainless, as
+the output in Figure~\ref{fig:result} shows.
 
@@
 \begin{figure}
-\lstinputlisting[style=scala, firstline=9, lastline=12, firstnumber=9]{../code/addition/src/main/scala/verified/currency/CurrencyUnits.scala}
+  \lstinputlisting[style=scala, firstline=9, lastline=12,
+  firstnumber=9]{../code/addition/src/main/scala/verified/currency/CurrencyUnits.scala}
 \caption{Addition function with specification}
@@
 		\includegraphics[width=\textwidth]{result_output}
-	\caption{Stainless output for the transformed code}
+	\caption{Stainless output for the rewritten code}
   \label{fig:result}

### 2020-07-24 paper-short/main.tex
work on final version

@@
 We are happy to see some friendly green verifier output. However,
-apart from the bugs we found, the main conclusion from this work is
-that we had to non-trivially transform even a very small portion of
-the code in order to verify it. At the moment, it is unrealistic to
-routinely formally verify properties as part of the Bitcoin-S
-development process. However, Stainless development has already
-progressed (e.g.\ type members are supported in recent versions) and
-continues to do so (e.g. self-referential type bounds are on the
-roadmap). Some missing features that we identified are presumably very
-easy to support, like the message parameter in the require
-function. Some other features presumably require more substantial
-work, like bitwise operations on integer types.
+apart from the bugs we found, the main conclusion of this work is that
+we had to non-trivially transform even a very small portion of the
+code (70 lines) in order to verify it. And that was true even though
+the code was purely functional to begin with. At the moment, it is
+probably unrealistic to routinely formally verify properties as part
+of the Bitcoin-S development process. However, Stainless development
+has already progressed (e.g.\ type members are supported in recent
+versions) and continues to do so (e.g. self-referential type bounds
+are on the roadmap). Some missing features that we identified are
+presumably very easy to support, like the message parameter in the
+require function. Some other features presumably require more
+substantial work, like bitwise operations on integer types.
 
 On the other hand, Bitcoin-S uses features that might not be supported
-even by future Stainless versions, such as calls to Java code. Here
-the Bitcoin-S code can hopefully be adapted to accommodate formal
-verification.
+even by future Stainless versions, such as calls to Java code.
 
-From our results we conclude that formal verification of bitcoin
-libraries in general and Bitcoin-S in particular is a worthwhile
-endeavour. We are looking forward to verifying more substantial parts
-of the code in future work.
+Given our experience, the best route towards integrating verification
+into the Bitcoin-S development process would be to re-implement parts
+of the library in Pure Scala. We would split the library into a
+\emph{verified} and \emph{non-verified} part, and use Stainless only
+on the \emph{verified} part. It is then both a technical but also a
+political question how much code, if any, can be moved to the
+\emph{verified} part. That is an interesting direction for future
+work.
 

### 2022-12-03 final/main.tex
prepare for final oasics submission
Final version

--- 2020-07-24 paper-short/main.tex
@@
 
-\graphicspath{{./images/}{../images/}}
+\graphicspath{{./images/}}
 
@@
 
-%\category{} %optional, e.g. invited paper
+\category{Short Paper}
 
@@
 
-%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\EventEditors{John Q. Open and Joan R. Access}
+\EventEditors{Bruno Bernardo and Diego Marmsoler}
 \EventNoEds{2}
-\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
-\EventShortTitle{CVIT 2016}
-\EventAcronym{CVIT}
-\EventYear{2016}
-\EventDate{December 24--27, 2016}
-\EventLocation{Little Whinging, United Kingdom}
+\EventLongTitle{2nd Workshop on Formal Methods for Blockchains (FMBC 2020)}
+\EventShortTitle{FMBC 2020}
+\EventAcronym{FMBC}
+\EventYear{2020}
+\EventDate{July 20--21, 2020}
+\EventLocation{Los Angeles, California, USA (Virtual Conference)}
 \EventLogo{}
-\SeriesVolume{42}
-\ArticleNo{23}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\SeriesVolume{84}
+\ArticleNo{7}
 
@@
 % Choose a nicer narrower bold
-\DeclareFontSeriesDefault[rm]{bf}{b}
+%\DeclareFontSeriesDefault[rm]{bf}{b}
 
@@
 \begin{figure}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/extracted/number/NumberType.scala}
+\lstinputlisting[style=scala]{code/extracted/number/NumberType.scala}
   \caption{Extracted Code from NumberType.scala}
@@
 \begin{figure}
-\lstinputlisting[style=scala]{../code/addition/src/main/scala/extracted/currency/CurrencyUnits.scala}
+\lstinputlisting[style=scala]{code/extracted/currency/CurrencyUnits.scala}
   \caption{Extracted Code from CurrencyUnits.scala}
@@
   \lstinputlisting[style=scala, firstline=9, lastline=12,
-  firstnumber=9]{../code/addition/src/main/scala/verified/currency/CurrencyUnits.scala}
+  firstnumber=9]{code/verified/currency/CurrencyUnits.scala}
 \caption{Addition function with specification}

### 2022-12-03 final/code/extracted/currency/CurrencyUnits.scala

@@ -1,0 +1,27 @@
+package extracted.currency
+
+import extracted.number.{BaseNumbers, Int64}
+
+sealed abstract class CurrencyUnit {
+  type A
+  def satoshis: Satoshis
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+  def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  }
+  protected def underlying: A
+}
+
+sealed abstract class Satoshis extends CurrencyUnit {
+  override type A = Int64
+  override def satoshis: Satoshis = this
+  def toBigInt: BigInt = BigInt(toLong)
+  def toLong: Long = underlying.toLong
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}
+
+object Satoshis extends BaseNumbers[Satoshis] {
+  val zero = Satoshis(Int64.zero)
+  def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
+  private case class SatoshisImpl(underlying: Int64) extends Satoshis
+}

### 2022-12-03 final/code/extracted/number/NumberType.scala

@@ -1,0 +1,41 @@
+package extracted.number
+
+sealed abstract class Number[T <: Number[T]] {
+  type A = BigInt
+  protected def underlying: A
+  def toLong: Long = toBigInt.bigInteger.longValueExact()
+  def toBigInt: BigInt = underlying
+  def andMask: BigInt
+  def apply: A => T
+  def +(num: T): T = apply(checkResult(underlying + num.underlying))
+
+  private def checkResult(result: BigInt): A = {
+    require((result & andMask) == result,
+      "Result was out of bounds, got: " + result)
+    result
+  }
+}
+
+sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]
+
+sealed abstract class Int64 extends SignedNumber[Int64] {
+  override def apply: A => Int64 = Int64(_)
+  override def andMask = 0xffffffffffffffffL
+}
+
+trait BaseNumbers[T] {
+  def zero: T
+}
+
+object Int64 extends BaseNumbers[Int64] {
+  private case class Int64Impl(underlying: BigInt) extends Int64 {
+    require(underlying >= -9223372036854775808L,
+      "Number was too small for a int64, got: " + underlying)
+    require(underlying <= 9223372036854775807L,
+      "Number was too big for a int64, got: " + underlying)
+  }
+
+  lazy val zero = Int64(0)
+  def apply(long: Long): Int64 = Int64(BigInt(long))
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}

### 2022-12-03 final/code/verified/currency/CurrencyUnits.scala

@@ -1,0 +1,27 @@
+package verified.currency
+
+import verified.number.{BaseNumbers, Int64}
+import stainless.lang._
+
+sealed abstract class CurrencyUnit {
+  def satoshis: Satoshis
+  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis
+  def +(c: CurrencyUnit): CurrencyUnit = {
+    Satoshis(satoshis.underlying + c.satoshis.underlying)
+  } ensuring (res =>
+    (c == Satoshis.zero) ==> (res == this))
+  protected def underlying: Int64
+}
+
+sealed abstract class Satoshis extends CurrencyUnit {
+  override def satoshis: Satoshis = this
+  def toBigInt: BigInt = underlying.toBigInt
+  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
+}
+
+case object Satoshis extends BaseNumbers[Satoshis] {
+  val zero = Satoshis(Int64.zero)
+  def apply(int64: Int64): Satoshis = SatoshisImpl(int64)
+}
+
+private case class SatoshisImpl(underlying: Int64) extends Satoshis

### 2022-12-03 final/code/verified/number/NumberType.scala

@@ -1,0 +1,28 @@
+package verified.number
+
+sealed abstract class Number {
+  protected def underlying: BigInt
+  def toBigInt: BigInt = underlying
+  def apply: BigInt => Int64
+  def +(num: Int64): Int64 = apply(underlying + num.underlying)
+}
+
+sealed abstract class SignedNumber extends Number
+
+sealed abstract class Int64 extends SignedNumber {
+  override def apply: BigInt => Int64 = Int64(_)
+}
+
+trait BaseNumbers[T] {
+  def zero: T
+}
+
+case object Int64 extends BaseNumbers[Int64] {
+  lazy val zero = Int64(0)
+  def apply(bigInt: BigInt): Int64 = Int64Impl(bigInt)
+}
+
+private case class Int64Impl(underlying: BigInt) extends Int64 {
+  require(underlying >= BigInt("-9223372036854775808"))
+  require(underlying <= BigInt("9223372036854775807"))
+}
